### 2019-06-07 notes/weekly_notes/5.15.19.NotesForWeek.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbb{I}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 5/15/19 - 5/17/19}
\author{Alden Green}
\date{\today}
\maketitle

Consider absolutely continuous distributions $\Pbb$ and $\Qbb$ with density functions $f$ and $g$, respectively. For fixed $n \geq 0$, let $\Z = (z_1, \ldots,z_n)$, where for $i = 1,\ldots,n$, $z_i \sim \frac{\Pbb + \Qbb}{2}$ are independent. Given $\Z$, for $i = 1,...,n$ let
\begin{equation*}
\ell_i = 
\begin{cases}
1~ \text{with probability $\frac{f(z_i)}{f(z_i) + g(z_i)}$} \\
-1~ \text{with probability $\frac{g(z_i)}{f(z_i) + g(z_i)}$}
\end{cases}
\end{equation*} 
be conditionally independent labels, and write
\begin{equation*}
1_X = 
\begin{cases}
1,~ l_i = 1\\
0,~ \text{otherwise}
\end{cases}
1_Y = 
\begin{cases}
1,~ l_i = -1 \\
0,~ \text{otherwise.}
\end{cases}
\end{equation*}
We will write $\X = \set{x_1, \ldots,x_{N_X}} := \set{z_i: \ell_i = 1}$ and similarly $\Y = \set{y_1, \ldots,y_{N_Y}} := \set{z_i: \ell_i = -1}$, where $N_X$ and $N_Y$ are of course random but $N_X + N_Y = n$.

Our statistical goal is hypothesis testing: that is, we wish to construct a test function $\phi$ which differentiates between
\begin{equation*}
\mathbb{H}_0: f = g \text{ and } \mathbb{H}_1: f \neq g.
\end{equation*}

For a given function class $\Hclass$, some $\epsilon > 0$, and test function $\phi$ a Borel measurable function of the data with range $\set{0,1}$, we evaluate the quality of the test using \emph{worst-case risk}
\begin{equation*}
R_{\epsilon}^{(t)}(\phi; \Hclass) = \sup_{f \in \Hclass} \Ebb_{f,f}^{(t)}(\phi) + \sup_{ \substack{f,g \in \Hclass \\ \delta(f,g) \geq \epsilon } } \Ebb_{f,g}^{(t)}(1 - \phi)
\end{equation*} 
where 
\begin{equation*}
\delta^2(f,g) = \int_{\D} (f - g)^2 dx.
\end{equation*}

\paragraph{Test statistic.}

For $r \geq 0$, define the \emph{$r$-graph} $G_r = (V,E_r)$ to have vertex set $V = \set{1,\ldots,t}$ and edge set $E_r$ which contains the pair $(i,j)$ if and only if $\norm{z_i - z_j}_2 \leq r$. Let $D_{r}$ denote the incidence matrix of $G_r$.

Define the \emph{Laplacian Smooth} test statistic over the neighborhood graph to be
\begin{equation*}
T_{LS} = \sup_{\theta: \norm{D_{r}\theta}_2 \leq C(n,r)} \dotp{\theta }{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
\end{equation*}
where we note that the test statistic is implicitly a function of $r$ and $C(n,r)$.

\paragraph{General assumptions.}
We will need to make the following assumptions in order to obtain meaningful theoretical results.

\begin{assumption}[General assumptions.]
	\label{asmp: general}
	For each $f \in \mathcal{H}$, assume $\mathrm{supp}(f)$ is an open connected, bounded subset of $\Rd$ with $l$-Lipschitz boundary. Further assume $\mathcal{H}$ is uniformly bounded above and below: that is, there exists $f_{\min} < f_{\max}$ such that for each $f \in \mathcal{H}$, 
	\begin{equation*}
	0 < f_{\min} < f(x) < f_{\max} < \infty, \quad \textrm{for all $x \in \mathrm{supp}(f)$}
	\end{equation*}
\end{assumption}

We motivate these assumptions through the following theorem relating the $r$-neighborhood graph to the $d$-dimensional grid, a useful result as the spectrum of grid graphs is well known.

Formally, for $\kappa = n^{1/d}$, consider the \emph{grid graph}
\begin{equation*}
G_{grid} = (V_{grid},E_{grid}),~~ V_{grid} = \set{k: k \in [\kappa]^d},~~ E_{grid} = \set{(k,k'): k, k' \in V_{grid}, \norm{k - k'}_1 = 1}
\end{equation*}
with associated incidence matrix $D_{grid}$.

\begin{lemma}
	Let $z_1, \ldots,z_n \sim p$ for some density function $p \in \mathcal{H}$ which satisfies Assumption \ref{asmp: general}.

	Fix $a > 2$. There exists $c_1 > 0$ (potentially depending on $d, l, f_{\min}, f_{\max}$ and $a$) such that for any $r \geq c_1 \left(\frac{\log n}{n}\right)^{1/d}$, the following statement holds with probability at least $1 - n^{-a}$: there exists a bijection $T: [n] \to [\kappa]^d$ such that for all $\theta = (\theta_1, \ldots, \theta_n) \in \Reals^n$, letting
	\begin{equation*}
	(\theta_{T})_{k} = \theta_{T^{-1}(k)}, \textrm{for all $k \in [\kappa]^d$}
	\end{equation*}
	we have
	\begin{equation*}
	\norm{D_{\grid}\theta_{T}}_2 \leq \norm{D_r \theta}_2.
	\end{equation*}
\end{lemma}
\textcolor{red}{Prove this.}

We use this to control the empirical process $T_{LS}$ under the null hypothesis.
\begin{lemma}
	Assume $\mathcal{H}$ satisfies Assumption \ref{asmp: general}, and fix $a \geq 2$. Then, there exists $c_1 > 0$ (potentially depending on $d, l, f_{\min}$, $f_{\max}$, and $a$) and $c_2$ (potentially depending on $d$) such that for any $r \geq c_1 \left(\frac{\log n}{n}\right)^{1/d}$, and any $C_{n,r} > 0$,
	\begin{equation*}
	\Pbb_{\mathbb{H}_0}\left(T_{LS} \leq \frac{c_2 C(n,r)\sqrt{\log n} }{a \sqrt{n}} \right) \leq 1 - 2\exp(2a) - a
	\end{equation*}
\end{lemma}
\begin{proof}
	We write
	\begin{align*}
	T_{LS} & \leq \abs{T_{LS} - \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y}} + \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y} \\
	& \leq \abs{T_{LS} - \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y}} + \sup_{\substack{\theta: \norm{D_{\grid} \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y} \tag{Lemma \ref{lem: closed_form_ls_statistic} } \\
	& \leq \abs{T_{LS} - \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y}} + C(n,r) \frac{1}{n/2} \sqrt{\sigma^T L_{\grid}^{\dagger} \sigma}
	\end{align*}
	whence the statement follows by Lemmas \ref{lem: normalization_error} and \ref{lem: noise_quadratic_form_on_grid}.
\end{proof}

\section{Additional Theory}

The Laplacian smooth test statistic has a closed form solution. Let $L = D^T D$ be the Laplacian of $G$, and write $\Linv$ for the pseudoinverse of $L$.

\begin{lemma}
	\label{lem: closed_form_ls_statistic}
	For any unweighted, undirected, connected graph $G = (V,E)$ with incidence matrix $D$, number $C > 0$, and any vector $v \in \Reals^n$ with $\sum_{i = 1}^{n} v_i = 0$,
	\begin{equation*}
	\sup_{\theta: \norm{D\theta}_2 \leq C} \dotp{\theta}{v} = C \sqrt{v^T \Linv v}
	\end{equation*}
	Additionally, for any vector $v \in \Reals^n$ (not necessarily $\sum_{i = 1}^{n} v_i = 0$), under the additional constraint $\theta^T \mathbf{1} = 0$, the same statement holds. That is,
	\begin{equation*}
	\sup_{ \substack{\theta: \norm{D\theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \dotp{\theta}{v} = C \sqrt{v^T \Linv v}
	\end{equation*}
\end{lemma}
\begin{proof}
	Note that the condition $\norm{D \theta}_2 \leq C$ is equivalent to $\theta^T L \theta \leq C$. The solution then follows from the KKT conditions.
\end{proof}

\begin{lemma}
	\label{lem: normalization_error}
	Under the general assumptions,
	\begin{equation*}
	\abs{T_{LS} - \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C(n,r), \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y}} \leq \frac{8C(n,r)}{a n^{1 - 1/d}\pi}
	\end{equation*}
	with probability at least $1 - 2 \exp(2a)$. 
\end{lemma}
\begin{proof}
	Note that as $(1_X/N_X - 1_Y/X_Y)^T \mathbf{1} = 0$, we may write
	\begin{align*}
	T_{LS} & = \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq 1, \\ \theta^T \mathbf{1} = 0} } \dotp{\theta}{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
	\end{align*}
	and therefore by Cauchy-Schwarz,
	\begin{align*}
	\abs{T_{LS} - \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \frac{1}{n/2}\dotp{\theta}{1_X - 1_Y}}  & \leq \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \cdot \norm{1_X(2/n - 1/N_X) - 1_Y(1/N_Y - 2/n)} \\
	& \leq \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \cdot \norm{1_X(2/n - 1/N_X) - 1_Y(1/N_Y - 2/n)} \\
	& \leq \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \cdot \left(\norm{1_X} \frac{\abs{n/2 - N_X}}{n/2 N_X} + \norm{1_Y} \frac{\abs{n/2 - N_Y}}{n/2 N_Y}\right) \\
	& \leq \sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \cdot \left(\frac{2\abs{n/2 - N_X}}{\sqrt{N_X} n} + \frac{2\abs{n/2 - N_Y}}{\sqrt{N_Y} n} \right).
	\end{align*}
	The reasoning in the proof of Lemma \ref{lem: noise_quadratic_form_on_grid} leads to the bound
	\begin{equation*}
	\sup_{\substack{\theta: \norm{D_r \theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \leq \frac{C n^{1/d}}{\pi},
	\end{equation*}
	and a standard application of Hoeffding's inequality to the quantities $N_X$ and $N_Y$ yields, for any $a > 0$
	\begin{equation*}
	\abs{N_X - n/2}, \abs{N_Y - n/2} \leq  a \sqrt{n}
	\end{equation*}
	with probability at least $1 - 2 \exp(-2a)$. As a result,
	\begin{equation*}
	\sup_{\substack{\theta: \norm{D_r \theta}_2 \leq 1, \\ \theta^T \mathbf{1} = 0} } \set{\norm{\theta}} \cdot \left(\frac{2\abs{n/2 - N_X}}{\sqrt{N_X} n} + \frac{2\abs{n/2 - N_Y}}{\sqrt{N_Y} n} \right) \leq \frac{8a C n^{1/d}}{n\pi} 
	\end{equation*}
	with probability at least $1 - 2 \exp(-2a)$, which proves the claim.
\end{proof}

\subsection{Type I error.}

Denote the eigenvalues and eigenvectors of $\Lgrid = \Dgrid^T \Dgrid$ as $\set{\lambda_{k}: k \in [\kappa]^d}$ and $\set{u_{k}: k \in [\kappa]^d}$, respectively.

\begin{lemma}
	\label{lem: noise_quadratic_form_on_grid}
	Let $\sigma_1, \ldots, \sigma_n \sim \mathrm{Rademacher}(1/2)$ be independent random variables, and write $\sigma = (\sigma_1, \ldots, \sigma_n)$. Then, 
	\begin{equation*}
	\sigma^T \Linvgrid \sigma \leq \frac{(2d)^d\kappa^2}{a\pi^2} \left(8 + 2\pi \log(\sqrt{2} \kappa) \left(\sum_{p = 2}^{d}\kappa^{p}\right) \right)
	\end{equation*}
	with probability at least $1  - a$.
\end{lemma}
\begin{proof}
	Taking the eigendecomposition $\Lgrid = U \Lambda U^T$, we can then write $\Linvgrid = U \Lambda^{\dagger} U^T$. Therefore
	\begin{equation*}
	\Ebb \bigl(\sigma^T \Linvgrid \sigma\bigr) = \sum_{k \in [\kappa]^d} \frac{\Ebb \bigl(u_{k}^T \sigma\bigr)^2}{\lambda_k}
	\end{equation*}
	and as 
	\begin{equation*}
	\Ebb \bigl(u_{k}^T \sigma\bigr)^2 = \sum_{i,j = 1}^{n} \Ebb(u_{k_i} u_{k_j} \sigma_i \sigma_j) = \sum_{i = 1}^{n} u_{k_i}^2 = 1
	\end{equation*}
	we are left with
	\begin{equation}
	\label{eqn: noise_quadratic_form_on_grid_1}
	\Ebb \bigl(\sigma^T \Linvgrid \sigma\bigr) = \sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \frac{1}{\lambda_k}.
	\end{equation}

	It is well known \textcolor{red}{HR, CB}, that the $d$-dimensional grid can be written as the Kronecker product of $d$ path graphs, and exploiting this fact as in \textcolor{red}{SWT} we obtain
	\begin{align*}
	\lambda_{k} = 4 \sin^2\left(\frac{\pi(k_1 - 1)}{2 \kappa}\right) + \cdots + 4 \sin^2\left(\frac{\pi(k_d - 1)}{2 \kappa}\right).
	\end{align*}
	The inequality $\sin(x) \geq x/2$ holds for $x \in [0,\pi/2]$, and so we have
	\begin{align*}
	\sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \frac{1}{\lambda_k} & \leq \frac{\kappa^2}{\pi^2} \sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d} (k_j - 1)^2\right)^{-1} \\
	& \leq \frac{(2d)^d\kappa^2}{\pi^2} \left(8 + 2\pi \log(\sqrt{2} \kappa) \left(\sum_{p = 2}^{d}\kappa^{p - 2}\right) \right) \\
	& = \frac{(2d)^d\kappa^2}{\pi^2} \left(8 + 2\pi \log(\sqrt{2} \kappa) \left(\sum_{p = 2}^{d}\kappa^{p}\right) \right) \tag{Lemma \ref{lem: grid_eigenvalues}}
	\end{align*}
	whence the claim follows by Markov's inequality.
\end{proof}

\begin{lemma}
	\label{lem: grid_eigenvalues}
	There exists a universal constant $c_1 > 0$ such that for any $\kappa \geq 1$ and for all integers $d \geq 3$,
	\begin{equation*}
	\sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d} (k_j - 1)^2\right)^{-1} \leq (2d)^d \Biggl(8 + 2 \pi \log (\sqrt{2} \kappa) \biggl(\sum_{p = 2}^{d} \kappa^{p - 2}\biggr) \Biggr)
	\end{equation*}
\end{lemma}
\begin{proof}
	We will prove by induction on $d$. As a base case, consider $d  = 2$. Rewrite
	\begin{equation*}
	\sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d} (k_j - 1)^2\right)^{-1} \leq 4 \sum_{k_1 = 1}^{\kappa - 1} \frac{1}{k_1^2}  + \sum_{k_1 =2}^{\kappa - 1} \sum_{k_2 = 2}^{\kappa - 1} (k_1^2 + k_2^2)^{-1}.
	\end{equation*}
	Then, upper bounding sums by integrals, we obtain
	\begin{equation*}
	\sum_{k_1 = 1}^{\kappa - 1} \frac{1}{k_1^2} \leq 1 +  \int_{x = 1}^{\kappa} \frac{1}{x^2} \dif x \leq 2,
	\end{equation*}
	and 
	\begin{align*}
	\sum_{k_1 = 2}^{\kappa - 1} \sum_{k_2 = 2}^{\kappa - 1} (k_1^2 + k_2^2)^{-1} & \leq \int_{x = 1}^{\kappa} \int_{y = 1}^{\kappa} \frac{1}{x^2 + y^2} \dif x \dif y \\
	& \leq 2 \pi \log(\sqrt{2} \kappa),
	\end{align*}
	so that
	\begin{equation*}
	\sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d} (k_j - 1)^2\right)^{-1} \leq 8 + 2 \pi \log(\sqrt{2} \kappa).
	\end{equation*}
	and the base case is shown.

	We proceed by induction. For $d \geq 3$ we write
	\begin{align*}
	\sum_{\substack{k \in [\kappa]^d, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d} (k_j - 1)^2\right)^{-1} & \leq 2d \sum_{\substack{k \in [\kappa]^{d - 1}, \\ k \neq \bf{0}}} \left(\sum_{j = 1}^{d - 1} (k_j - 1)^2\right)^{-1} + \sum_{k_1 = 2}^{\kappa - 1} \cdots \sum_{k_d = 2}^{\kappa - 1} \left(\sum_{j = 1}^{d} k_j^2\right)^{-1} \\
	& \leq 2d \left( (2d)^{d - 1}\Biggl(8 + 2 \pi \log \sqrt{2} \kappa \biggl(\sum_{p = 2}^{d - 1} \kappa^{p - 2}\biggr) \Biggr)\right) + \sum_{k_1 = 2}^{\kappa - 1} \cdots \sum_{k_d = 2}^{\kappa - 1} \left(\sum_{j = 1}^{d} k_j^2\right)^{-1} \\
	& = \left( (2d)^{d}\Biggl(8 + 2 \pi \log \sqrt{2} \kappa \biggl(\sum_{p = 2}^{d - 1} \kappa^{p - 2}\biggr) \Biggr)\right) + \sum_{k_1 = 2}^{\kappa - 1} \cdots \sum_{k_d = 2}^{\kappa - 1} \left(\sum_{j = 1}^{d} k_j^2\right)^{-1} 
	\end{align*}
	and again upper bounding sums by integrals, we obtain
	\begin{align*}
	\sum_{k_1 = 2}^{\kappa - 1} \cdots \sum_{k_d = 2}^{\kappa - 1} \left(\sum_{j = 1}^{d} (k_j)^2\right)^{-1} & \leq \int_{1}^{\kappa} \cdots \int_{1}^{\kappa} \frac{1}{x_1^2 + \cdots + x_d^2} \dif x_d \ldots \dif x_1 \\
	& \leq \kappa^{d-2} \int_{1}^{\kappa} \int_{1}^{\kappa} \frac{1}{x^2 + y^2} \dif x \dif y \\
	& \leq \kappa^{d-2} 2 \pi \log(\sqrt{2} \kappa)
	\end{align*}
	and the proof is complete.
\end{proof}

\paragraph{Type II error.}

Lemma \ref{lem: approx_error_lb_1} will be useful in lower bounding the \textcolor{red}{approximation error} of the $T_{LS}$ test statistic. Write $P_1^{\perp}$ for the projection operator onto the subspace of $\Reals^n$ orthogonal to the constant vector $\mathbf{1}$.

\begin{lemma}
	\label{lem: approx_error_lb_1}
	For any vector $v \in \Reals^n$ and Laplacian matrix $L$ of a connected graph $G$,
	\begin{equation*}
	v^T \Linv v \geq \frac{(v^T P_{1}^{\perp}v)^2}{v^T L v}
	\end{equation*}
\end{lemma}
\begin{proof}
	We can expand $v^T P_1^{\perp} v  = v^T (L)^{1/2}(\Linv)^{1/2} v$, for any matrix square root of $L$ and $\Linv$. The statement follows by Cauchy-Schwarz.
\end{proof}

\end{document}