### 2019-06-07 notes/weekly_notes/5.29.19.NotesForWeek.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbb{I}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 5/29/19 - 5/31/19}
\author{Alden Green}
\date{\today}
\maketitle

Consider absolutely continuous distributions $\Pbb$ and $\Qbb$ with density functions $f$ and $g$, respectively. For fixed $n \geq 0$, let $\Z = (z_1, \ldots,z_n)$, where for $i = 1,\ldots,n$, $z_i \sim \frac{\Pbb + \Qbb}{2}$ are independent. Given $\Z$, for $i = 1,...,n$ let
\begin{equation*}
\ell_i = 
\begin{cases}
1~ \text{with probability $\frac{f(z_i)}{f(z_i) + g(z_i)}$} \\
-1~ \text{with probability $\frac{g(z_i)}{f(z_i) + g(z_i)}$}
\end{cases}
\end{equation*} 
be conditionally independent labels, and write
\begin{equation*}
1_X = 
\begin{cases}
1,~ l_i = 1\\
0,~ \text{otherwise}
\end{cases}
1_Y = 
\begin{cases}
1,~ l_i = -1 \\
0,~ \text{otherwise.}
\end{cases}
\end{equation*}
We will write $\X = \set{x_1, \ldots,x_{N_X}} := \set{z_i: \ell_i = 1}$ and similarly $\Y = \set{y_1, \ldots,y_{N_Y}} := \set{z_i: \ell_i = -1}$, where $N_X$ and $N_Y$ are of course random but $N_X + N_Y = n$.

Our statistical goal is hypothesis testing: that is, we wish to construct a test function $\phi$ which differentiates between
\begin{equation*}
\mathbb{H}_0: f = g \text{ and } \mathbb{H}_1: f \neq g.
\end{equation*}

For a given function class $\Hclass$, some $\epsilon > 0$, and test function $\phi$ a Borel measurable function of the data with range $\set{0,1}$, we evaluate the quality of the test using \emph{worst-case risk}
\begin{equation*}
R_{\epsilon}^{(t)}(\phi; \Hclass) = \sup_{f \in \Hclass} \Ebb_{f,f}^{(t)}(\phi) + \sup_{ \substack{f,g \in \Hclass \\ \delta(f,g) \geq \epsilon } } \Ebb_{f,g}^{(t)}(1 - \phi)
\end{equation*} 
where 
\begin{equation*}
\delta^2(f,g) = \int_{\D} (f - g)^2 dx.
\end{equation*}

\paragraph{Test statistic.}

For $r \geq 0$, define the \emph{$r$-graph} $G_r = (V,E_r)$ to have vertex set $V = \set{1,\ldots,t}$ and edge set $E_r$ which contains the pair $(i,j)$ if and only if $\norm{z_i - z_j}_2 \leq r$. Let $D_{r}$ denote the incidence matrix of $G_r$.

Define the \emph{Laplacian Smooth} test statistic over the neighborhood graph to be
\begin{equation*}
T_{LS} = \sup_{\theta: \norm{D\theta}_2 \leq C(n,r)} \dotp{\theta }{\frac{1_X}{N_X} - \frac{1_Y}{N_Y}}
\end{equation*}
where we note that the test statistic is implicitly a function of $r$ and $C(n,r)$.

\section{Additional Theory}
Let $\theta^{\star} = (\thetast_i)_{i = 1}^n$, with $\thetast_i := f(z_i) - g(z_i)$. Write $L = D^T D$ for the Laplacian matrix of the $r$-neighborhood graph, $\gradient f$ for the gradient of a function $f$, and $\Hessian f$ for the Hessian of a function $f$. We write $C^{2}(L)$ for the set of functions $f$ twice continuously differentiable over $\Rd$, with bounded Hessian $\norm{\Hessian f}_{\infty} \leq L$.
\begin{lemma}
	\label{lem: approx_error}
	For all density functions $f,g \in C^2(L)$ with $\int (f - g)^2 \geq \epsilon^2$, if $r \geq c_1 (\log n / n)^{1/d}$ and $\epsilon \geq c_2 n r^{d + 1} $ then
	\begin{equation*}
	\sup_{\norm{D\theta}_2 \leq 1} \dotp{\theta}{\thetast} \geq \frac{c_2}{2} (\log n)^{1 + 2/d} n^{-2/d}
	\end{equation*}
	with probability at least $1 - \delta$.
\end{lemma}
\begin{proof}
	By Lemmas \ref{lem: closed_form_ls_statistic} and \ref{lem: lin_alg_1}, we have
	\begin{align}
	\label{eqn: approx_error_1}
	\sup_{\norm{D\theta}_2 \leq 1} \dotp{\theta}{\thetast} & = \frac{1}{n} \sqrt{(\thetast)^T \Linv \thetast} \nonumber \\
	& \geq \frac{1}{\lambda_k}\left(\dotp{\thetast}{\thetast} - \dotp{P_k^{\perp} \thetast}{P_k^{\perp}\thetast}\right)^2 
	\end{align}
	where the latter inequality holds for any $k = 1, \ldots, n - 1$. We upper bound $\dotp{P_k^{\perp} \thetast}{P_k^{\perp}\thetast}$ using the following relations
	\begin{equation*}
	(\thetast)^T L \thetast \geq (P_k^{\perp}\thetast)^T \Linv ( P_k^{\perp}\thetast) \geq \lambda_k \dotp{P_k^{\perp} \thetast}{P_k^{\perp} \thetast}
	\end{equation*}
	and obtain
	\begin{align*}
	\frac{1}{\lambda_k}\left(\dotp{\thetast}{\thetast} - \dotp{P_k^{\perp} \thetast}{P_k^{\perp}\thetast}\right)^2  & \geq \frac{1}{\lambda_k}\left(\dotp{\thetast}{\thetast} - \frac{(\thetast)^T L \thetast}{\lambda_k} \right)^2 \\
	& \geq \frac{1}{\lambda_k}\left(c_1 n \epsilon^2 - \frac{c_2 n^2 r^{d + 2}}{\lambda_k} \right)^2 \tag{Lemmas \ref{lem: witness_inner_product} and \ref{lem: witness_discrete_energy}}
	\end{align*}
	where the latter inequality occurs with probability at least $1 - 2\delta$. Choose $k = n - 1$. As $r \geq c_1 (\log n / n)^{1/d}$, we have that $G \preceq \mathrm{Grid}$, and as a result \textcolor{red}{$4 \geq \lambda_{n - 1} \geq 1$}. Therefore,
	\begin{align*}
	\frac{1}{\lambda_k}\left(c_1 n \epsilon^2 - \frac{c_2 n^2 r^{d + 2}}{\lambda_k} \right)^2 & \geq \frac{1}{4} \left(c_1 n \epsilon^2 - c_2 n^2 r^{d + 2} \right)^2 \\
	& \geq \frac{1}{4} \left(c_2 n^2 r^{d + 2}\right)^2 \\
	& \geq \frac{1}{4} \left(c_2 n^{1 - 2/d} \log(n)^{1 + 2/d}\right)^2 = \frac{c_2^2}{4} n^{2 - 4/d} \log(n)^{2 + 4/d},
	\end{align*}
	and the proof is complete.
\end{proof}

\begin{lemma}
	\label{lem: witness_inner_product}
	For any $f, g \in L^2(\Rd)$ satisfying the \textcolor{red}{regularity conditions}, such that $\int_{\Rd} (f - g)^2 dx \geq \epsilon^2$ there exists constant $c_1$ such that 
	\begin{equation*}
	\dotp{\thetast}{\thetast} \geq c_1 n \epsilon^2
	\end{equation*}
	with probability at least $1 - \delta$.
\end{lemma}

\begin{lemma}
	\label{lem: witness_discrete_energy}
	For any $f, g \in C^2(L)$ satisfying the \textcolor{red}{regularity conditions}, there exists a constant $c_2$ such that
	\begin{equation*}
	(\thetast)^T L \thetast \leq c_2 n^2 r^{d + 2}
	\end{equation*}
	with probability at least $1 - \delta$.
\end{lemma}

\section{Linear Algebra}

\begin{lemma}
	\label{lem: closed_form_ls_statistic}
	For any unweighted, undirected, connected graph $G = (V,E)$ with incidence matrix $D$, and any vector $v \in \Reals^n$ with $\sum_{i = 1}^{n} v_i = 0$,
	\begin{equation*}
	\sup_{\theta: \norm{D\theta}_2 \leq C} \dotp{\theta}{v} = C \sqrt{v^T \Linv v}
	\end{equation*}
	Additionally, for any vector $v \in \Reals^n$ (not necessarily $\sum_{i = 1}^{n} v_i = 0$), under the additional constraint $\theta^T \mathbf{1} = 0$, the same statement holds. That is,
	\begin{equation*}
	\sup_{ \substack{\theta: \norm{D\theta}_2 \leq C, \\ \theta^T \mathbf{1} = 0} } \dotp{\theta}{v} = C \sqrt{v^T \Linv v}
	\end{equation*}
\end{lemma}
\begin{proof}
	Note that the condition $\norm{D \theta}_2 \leq C$ is equivalent to $\theta^T L \theta \leq C$. The solution then follows from the KKT conditions.
\end{proof}

Let $L$ be the Laplacian matrix of a connected graph, and $\Linv$ be the pseudo-inverse $L$. Write the eigendecomposition of $\Linv = U\Lambda^{\dagger}U^{T}$, where $\Lambda$ is an $n \times n$ diagonal matrix with entries $0 = \lambda_0 < \lambda_1 \leq \lambda_2 \leq \ldots \leq \lambda_{n-1}$ and $U$ is an orthogonal matrix with columns $U =\left(u_0 \ldots u_{n-1}\right)$. For each $k = 0, \ldots, n - 1$, write $U_{k} = (u_0 \ldots u_{k-1})$ for the first $k$ columns of $U$, $P_k$ as the projection operator onto the span of $U_k$, and $P_k^{\perp}$ for the projection operator onto the subspace orthogonal to the span of $U_k$.
\begin{lemma}
	\label{lem: lin_alg_1}
	For any $k = 1, \ldots, n$, 
	\begin{equation*}
	(\thetast)^T \Linv \thetast \geq \frac{1}{\lambda_k}\left(\dotp{\thetast}{\thetast} - \dotp{P_k^{\perp} \thetast}{P_k^{\perp}\thetast}\right)^2
	\end{equation*}
\end{lemma}
\begin{proof}
	Note that $\Lambda^{\dagger}$ is a diagonal matrix, with entries $\rho_{1,1} = 0$ and $\rho_{k,k} = \frac{1}{\lambda_k}$ for $k = 1, \ldots, n$. Therefore,
	\begin{equation*}
	(\thetast)^T \Linv \thetast = \sum_{k = 1}^{n} \frac{\dotp{\thetast}{u_k}^2}{\lambda_k}.
	\end{equation*}
	Clearly, for any $k = 1, \ldots, n$,
	\begin{equation*}
	\sum_{k = 1}^{n} \frac{\dotp{\thetast}{u_k}^2}{\lambda_k} \geq \frac{\dotp{P_k \thetast}{P_k \thetast}^2}{\lambda_k}
	\end{equation*}
	and as $\dotp{P_k \thetast}{P_k \thetast} + \dotp{P_k^{\perp} \thetast}{P_k^{\perp} \thetast} = \dotp{\thetast}{\thetast}$, we obtain
	\begin{align*}
	\frac{\dotp{P_k \thetast}{P_k \thetast}^2}{\lambda_k} & =  \frac{1}{\lambda_k}\left(\dotp{\thetast}{\thetast} - \dotp{P_k^{\perp} \thetast}{P_k^{\perp}\thetast}\right)^2.
	\end{align*}
\end{proof}

\end{document}