### 2019-06-19 notes/weekly_notes/6.18.19.NotesForWeek.tex
added weekly notes

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbb{I}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 6/13/19 - 6/18/19}
\author{Alden Green}
\date{\today}
\maketitle

Let $f,g$ be $L^2$ density functions supported on the $d$-dimensional unit cube $[0,1]^d$, meaning
\begin{equation*}
\int_{[0,1]^d} f(x)^2 dx, \int_{[0,1]^d} g(x)^2 dx < \infty, \quad \text{and} \quad \int_{[0,1]^d} f(x) dx, \int_{[0,1]^d} g(x) dx = 1.
\end{equation*}
with $f$,$g$ bounded above and below, $0 < p_{\min} < f(x), g(x) < p_{\max} < \infty$ for all $x \in [0,1]^d$.

We observe data $(X,\ell)$, a design matrix and associated labels, specified as follows. We let $x_1, \ldots, x_n$ be the rows of $X$, each sampled independently from $\mu = (f + g)/2$. For each $i = 1,\ldots,n$, we then sample $\ell_i$ according to
\begin{equation*}
\ell_i =
\begin{cases}
1,~ \textrm{with probability} \frac{f(x_i)}{f(x_i) + g(x_i)} \\
-1,~ \textrm{with probability} \frac{g(x_i)}{f(x_i) + g(x_i)}
\end{cases}
\end{equation*}
and let $\ell = (\ell_i)$.

Our statistical goal is hypothesis testing: that is, we wish to construct a test function $\phi$ which differentiates between
\begin{equation*}
\mathbb{H}_0: f = g \text{ and } \mathbb{H}_1: f \neq g.
\end{equation*}

For a given function class $\Hclass$, some $\epsilon > 0$, and test function $\phi$ a Borel measurable function of the data with range $\set{0,1}$, we evaluate the quality of the test using \emph{worst-case risk}
\begin{equation*}
R_{\epsilon}^{(t)}(\phi; \Hclass) = \sup_{f \in \Hclass} \Ebb_{f,f}^{(t)}(\phi) + \sup_{ \substack{f,g \in \Hclass \\ \delta(f,g) \geq \epsilon } } \Ebb_{f,g}^{(t)}(1 - \phi)
\end{equation*} 
where 
\begin{equation*}
\delta^2(f,g) = \int_{\D} (f - g)^2 dx.
\end{equation*}
We will consider the function class $\mathcal{H} := \mathcal{H}_{1}^d(L)$, the $d$-dimensional Lipschitz ball with norm $L$. Formally, a function $f \in \mathcal{H}$ if for all $x,y \in [0,1]^d$, $\abs{f(x) - f(y)} \leq L \norm{x - y}$.\footnote{Note that the Lipschitz requirement, along with the restriction that $f$ be a density function over a bounded domain, imply that $f$ has finite $L^2$ norm.}

\section{Test Statistic}

Let $r > 0$ be a bandwidth parameter, let $\eta_r: \mathcal{X} \times \mathcal{X} \to [0,\infty)$ be a Mercer kernel given by
\begin{equation*}
\eta_r(x,y) =
\begin{cases}
1, \quad & \textrm{$\norm{x - y} \leq r$} \\
0, \quad & \textrm{$\norm{x - y} > r$},
\end{cases}
\end{equation*}
and let $\widetilde{A}$ be an $n \times n$ matrix with entries $\widetilde{A}_{ij} = n^{-1}\eta(x_i,x_j)$. We obtain $A$ from $\widetilde{A}$ be setting all diagonal entries of $\widetilde{A}$ to be zero. Our test statistic is then
\begin{equation*}
T_{\eta} := \ell^T A \ell.
\end{equation*}

\section{Theory}

\begin{theorem}

\end{theorem}

\section{Supporting Theory}

We begin by recalling Chebyshev's inequality, which we will use to bound deviations of $T_{\eta}$ from its expectation.
\begin{lemma}[Chebyshev]
	\label{lem: chebyshev}
	Let $X$ be a random variable with $\Ebb X < \infty$ and $0 < \Var(X) < \infty$. Then,
	\begin{equation*}
	\Pbb(\abs{X - \Ebb X} \geq a) \leq \frac{\Var(X)}{a^2}.
	\end{equation*}
\end{lemma}

In order to employ Lemma \ref{lem: chebyshev}, we must bound the expectation and variance of $T_{\eta}$ under the null and alternative hypotheses. We begin with the expectation $\Ebb T_{\eta}$. 
\begin{lemma}
	\label{lem: mmd_expectation}
	Under the null hypothesis $h = g$, for any $h, g \in \mathcal{H}_{1}^{d}(L)$,
	\begin{equation*}
	\Ebb T_{\eta} = 0.
	\end{equation*}

	Under the alternative hypothesis $h \neq g$, for any $h, g \in \mathcal{H}_{1}^{d}(L)$, 
	\begin{equation*}
	\Ebb T_{\eta} \geq \frac{(n - 1)p_{\min}^2}{ p_{\max}^2} \left(\delta^2(f,g)\frac{r^d}{2^d} - L\delta(f,g)r^{d + 1}\right)
	\end{equation*}
\end{lemma}
\begin{proof}
	Let $\thetast$ be a vector in $\Reals^{n}$, with entries $ \thetast_i = \Ebb(\ell_i|x_i)$. Letting $w = \ell - \thetast$, we note that conditional on $X$, the noise component $w$ consists of $n$ independent and zero mean terms.

	We expand the test statistic
	\begin{equation*}
	T_{\eta} = (\thetast)^T A \thetast + w^T A w + 2 w^T A \thetast
	\end{equation*}
	and note that by the law of iterated expectation, 
	\begin{equation*}
	\Ebb(T_{\eta}) = \Ebb(T_{\eta}|X) = \Ebb\left( (\thetast)^T A \thetast \right)
	\end{equation*}

	Now, let $X$ and $Y$ be independent random variables with distribution $\mu$, and let $\Delta(x) := f(x) - g(x)$. By the linearity of expectation,
	\begin{align*}
	\Ebb (\thetast)^T A \thetast & = \sum_{i \neq j = 1}^{n} \Ebb( \thetast_i \thetast_j A_{ij}) \\
	& = (n - 1) \Ebb \bigl(\thetast_1 \thetast_2 \eta_r(x_1, x_2) \bigr) \\
	& \geq \frac{4(n - 1)}{p_{\max}^2} \Ebb(\Delta(x) \Delta(y) \eta_r(X,Y))
	\end{align*}

	We can lower bound $\Ebb(\Delta(x) \Delta(y) \eta_r(X,Y))$ by $\delta^2(f,g)$, minus a remainder term, as follows
	\begin{align*}
	\Ebb(\Delta(x) \Delta(y) \eta_r(X,Y)) & = \iint \Delta(x) \Delta(y) \eta_r(x,y) d\mu(x) d\mu(x) \\
	& \geq \frac{p_{\min}^2}{4} \iint \Delta(x) \Delta(y) \eta_r(x,y) dx dy \\
	& \geq \frac{p_{\min}^2}{4} \left( \iint \Delta^2(x) \eta_r(x,y) dx dy - \iint \abs{\Delta(x)} \abs{\Delta(x) - \Delta(y)} \eta(x,y) dx dy \right)
	\end{align*}
	To lower bound the first term, observe that for any $x \in [0,1]^d$,
	\begin{equation*}
	\int_{[0,1]^d} \eta_r(x,y) dx \geq \frac{r^d}{2^d}
	\end{equation*}
	and therefore
	\begin{equation*}
	\iint \Delta^2(x) \eta_r(x,y) dx dy \geq \frac{r^d}{2^d} \int_{[0,1]^d} \Delta^2(x) dx \geq \frac{r^d}{2^d} \delta^2(f,g).
	\end{equation*}

	To upper bound the remainder term, observe that $\Delta \in \mathcal{H}_{1}^{d}(2L)$. Using this fact, we obtain the upper bound
	\begin{align*}
	\iint \abs{\Delta(x)} \abs{\Delta(x) - \Delta(y)} \eta(x,y) dx dy & \leq : \iint \abs{\Delta(x)} \norm{x - y} \eta(x,y) dx dy \\
	& \leq L r^{d + 1} \int_{[0,1]^d} \abs{\Delta(x)} dx \\
	& \leq L r^{d + 1} \delta(f,g).
	\end{align*}
	where the last inequality follows from Jensen's inequality.
\end{proof}

\begin{lemma}
	\label{lem: var_mmd}
	For any $h,g \in \mathcal{H}_1^{d}(L)$,
	\begin{equation*}
	\Var(T_{\eta}) \leq 2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + 2r^d
	\end{equation*}
\end{lemma}
\begin{proof}
	We begin by computing the second moment $\Ebb T_{\eta}^2$. Expanding the quadratic form, we obtain
	\begin{align}
	\Ebb T_{\eta}^2 =  \sum_{i,i' = 1}^{n} \sum_{j,j' = 1}^{n} \Ebb(\ell_i \ell_{i'} \ell_{j} \ell_{j'} A_{ii'} A_{jj'})
	\end{align}

	We next compute this expectation, based on the number of distinct indices $i,i',j,j'$. Let $X, X', Y, Y'$ be four independent random variables, each with distribution $\mu$, and write $\widetilde{\Delta} := 2 \Delta / \mu$.  Then, applying the law of conditional expectation, we obtain
	\begin{equation*}
	n^2 \Ebb(\ell_i \ell_{i'} \ell_{j} \ell_{j'} A_{ii'} A_{jj'}) = 
	\begin{cases}
	\Ebb(\widetilde{\Delta}(X) \wt{\Delta}(X') \eta_r(X,X'))^2, \quad & \textrm{if}~ \set{i,i',j,j'} ~\textrm{contains four distinct elements}, \\
	\Ebb(\widetilde{\Delta}(X') \wt{\Delta}(Y') \eta_r(X,X') \eta_r(X,Y')), \quad & \textrm{if}~ \set{i,i',j,j'} ~\textrm{contains three distinct elements and}~ i = j, \\
	\Ebb\biggl(\eta_r(X,Y)^2\biggr), \quad  & \textrm{if}~ \set{i,i',j,j'} ~\textrm{contains two distinct elements and}~ i = j. \\
	0, \quad & \textrm{if}~ i = i' ~\textrm{or}~ j = j'.
	\end{cases}
	\end{equation*}
	and as a result we obtain
	\begin{align}
	\label{eqn: var_mmd_1}
	n^2 \Ebb T_{\eta}^2 = & n(n-1)(n-2)(n-3)\Ebb\biggl(\widetilde{\Delta}(X) \wt{\Delta}(X') \eta_r(X,X')\biggr) + \nonumber \\
	& \quad 2n(n-1)(n-2) \Ebb \biggl(\widetilde{\Delta}(X') \wt{\Delta}(Y') \eta_r(X,X') \eta_r(X,Y') \biggr) + \nonumber \\
	& \quad \quad 2n(n-1) \Ebb\biggl(\eta_r(X,Y)^2\biggr).
	\end{align}

	The first term -- covering the case when $\set{i,i',j,j'}$ contains four distinct elements --  will be canceled out by the mean, as we will see, and we do not simplify it further.

	The second term we upper bound by methods similar to those used to obtain the upper bound in Lemma \ref{lem: mmd_expectation}. We will need the following upper bound,
	\begin{align*}
	\int \eta(x,x') \eta(x,y') \dx & \leq r^d \1(\norm{x' - y'} \leq 2r) \\
	& = r^d \eta_{2r}(x',y').
	\end{align*}
	We proceed,
	\begin{align}
	\label{eqn: var_mmd_2}
	\Ebb \biggl(\widetilde{\Delta}(X') \wt{\Delta}(Y') \eta_r(X,X') \eta_r(X,Y') \biggr) & \leq \frac{p_{\max}^2}{p_{\min}^2} \iiint \Delta(x') \Delta(y') \eta(x,x') \eta(x,y') \dx \dypr \dxpr \nonumber\\
	& \leq \frac{p_{\max}^2}{p_{\min}^2} r^d \iint \Delta(x') \Delta(y') \eta_{2r}(x',y') \dypr \dxpr \nonumber \\
	& \leq \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right).
	\end{align}

	The following upper bound on the third term is a straightforward consequence of the fact $\eta_r^2 = \eta_r$, and therefore
	\begin{equation}
	\label{eqn: var_mmd_3}
	\Ebb(\eta_r(X,Y)^2) = \Ebb(\eta_r(X,Y)) = r^d. 
	\end{equation}

	Finally, we have that
	\begin{equation}
	\label{eqn: var_mmd_4}
	n^2 \Ebb(T_{\eta}^2) = \left(n(n-1) \Ebb(\widetilde{\Delta}(X) \wt{\Delta}(X') \eta_r) \right)^2 .
	\end{equation}

	Combining \eqref{eqn: var_mmd_1}-\eqref{eqn: var_mmd_4}, we obtain
	\begin{equation*}
	\Var(T_\eta) \leq 2n \frac{p_{\max}^2}{p_{\min}^2} r^d \left(r^d \delta^2(f,g) + Lr^{d+1}\delta(f,g)\right) + 2r^d.
	\end{equation*}
\end{proof}

\end{document}