### 2019-08-02 notes/weekly_notes/7.17.19.NotesForWeek.tex
updated weekly notes; began drafting master document

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbb{I}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 7/17/19 - 7/31/19}
\author{Alden Green}
\date{\today}
\maketitle

\section{Goodness-of-Fit Testing in the Sampling Model.}

Let $\mathcal{D} = [0,1]^d$. Suppose we observe the random design $x_1,\ldots,x_n$ independently sampled from the uniform distribution over $[0,1]^d$. Additionally, for $i \in [n]$, we observe
\begin{equation*}
z_i = f(x_i) + \varepsilon_i, \quad \varepsilon_i \overset{\textrm{i.i.d}}{\sim} \mathcal{N}(0,1), \quad \varepsilon \perp x
\end{equation*}
where $f \in \mathcal{H} \subseteq L^2(\mathcal{D})$. Our statistical goal is hypothesis testing. We wish to distinguish:
\begin{equation*}
\mathbf{H}_0: \norm{f}_2 = 0 \quad \textrm{vs} \quad \mathbf{H}_a: \norm{f}_2 > 0. 
\end{equation*}
We will evaluate our performance using worst-case error: for a given function class $\mathcal{H}$, test function $\phi: \Reals^n \to \set{0,1}$ and $\epsilon> 0$, let
\begin{equation*}
\mathcal{R}_{\epsilon}(\phi; \mathcal{H}) = \Ebb_{f = 0}(\phi) + \sup_{f \in \mathcal{H}: \norm{f}_2 > \epsilon} \Ebb_f(1 - \phi)
\end{equation*}

An example function class $\mathcal{H}$ we will consider will be the unit ball in a Sobolev space. Let $s,d$ be known, fixed positive integers. For $f: \mathcal{D} \to \Reals$ locally summable, we use the multiindex notation $D^{\alpha}f$ to denote the $\alpha$th-weak partial derivative of $f$ (if one exists). Then, the Sobolev norm is 
\begin{equation*}
\norm{f}_{W^{s,2}(\D)}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{D}} \norm{D^{\alpha}f}_2^2 \,dx
\end{equation*}
and the corresponding unit ball is $W^{s,2}(\D; L) = \set{f: \norm{f}_{W^{s,2}(\mathcal{D})} \leq L}$.

\subsection{Test statistic.}

For $x,y \in \mathcal{D}$ and radius $r > 0$ to be specified later, let $\eta_r(x,y) = \mathbf{1}(\norm{x - y}/ r \leq 1)$. and let $A$ be the $n \times n$ adjacency matrix with entires $A_{ij} = \eta_r(x_i,x_j)$. Let $B$ be the incidence matrix associated with $A$, and $L = B^TB$ be the associated Laplacian matrix. Write $B = U \Lambda^{1/2}V^T$ for the singular value decomposition of $B$. Then $L = V \Lambda V^T$ is the eigendecomposition of $L$ , where $\Lambda$ is a diagonal matrix of eigenvalues with diagonal entries $\lambda_1 \leq \lambda_2 \leq \ldots \leq \lambda_n$, and $V$ is an orthonormal matrix of eigenvectors.

Our test statistic will be the norm of a projection of $z$ onto the subspace spanned by the first few eigenvectors of $V$. In particular, for $C > 0$ to be specified later, our test statistic will be
\begin{equation*}
T_C := \sum_{k: \lambda_k^s \leq C^2} y_k^2, \quad y_k = \frac{1}{\sqrt{n}} z^T v_k.
\end{equation*}
For $b \geq 1$, let $\tau(b) = b\sqrt{2N(C)/n^2}$. Then our test will be $\phi_{C} = \mathbf{1}\set{T_C \leq N(C)/n + \tau(b)}$.

\subsection{Fixed design testing.}

Let $\beta$ be the length-$n$ random vector with $j$th entry $\beta_j = f(x_j)$. We introduce the scaling $C_{n,r} > 0$, defined by $C_{n,r}^2 = n^2r^{d + 2s}$.  We will apply results about testing over balls in discrete Sobolev classes. Let $b \geq 1$ denote $N(C) := \sharp\set{k: \lambda_k^s \leq C^2}$, and consider the following events:
\begin{enumerate}[label=(E\arabic*)]
	\item 
	\label{event:discrete_sobolev_norm}
	Discrete Sobolev norm of $\beta$:
	\begin{equation*}
	\norm{B^{(s)}\beta}_2 \leq C_{n,r}
	\end{equation*}
	\item Eigenvalue tail decay:
	\label{event:tail_decay}
	For the choice $C = C_r^{\star}$, where
	\begin{equation*}
	C_r^{\star} = \frac{(C_{n,r}n^{s/d})^{4s/(4s + d)}}{n^{s/d}},
	\end{equation*}
	the following inequality is satisfied:
	\begin{equation*}
	N(C_r^{\star}) \leq n(C_r^{\star})^{d/s}.
	\end{equation*}
	\item 
	\label{event:l2_norm}
	$L_2$ norm of $\beta$:
	\begin{equation*}
	\frac{\norm{\beta}_2^2}{n} \geq \frac{(2\sqrt{2}b + 1)}{n} \left(C_{n,r}n^{s/d}\right)^{2d/(4s+d)}.
	\end{equation*}
\end{enumerate}

Suppose for a given $x = x_1, \ldots, x_n$, and for a particular choice of radius $r$, the events \ref{event:l2_norm}-\ref{event:tail_decay} hold. The worst-case error conditional on $x$ can then be upper bounded.
\begin{lemma}
	\label{lem:discrete_test_error}
	For any $x$ and $r$ such that \ref{event:discrete_sobolev_norm} and \ref{event:tail_decay} hold, we have that if $f = 0$,
	\begin{equation*}
	\Ebb_{\beta}(\phi|x) \leq \frac{1}{b^2}
	\end{equation*}
	If $f \neq 0$ is such that \ref{event:l2_norm} is additionally satisfied, we have that
	\begin{equation*}
	\Ebb_{\beta}(1 -\phi|x) \leq \frac{1}{2b^2} + o(1)
	\end{equation*}
\end{lemma}

We now turn to showing that each of \ref{event:discrete_sobolev_norm} and \ref{event:tail_decay} (and, under appropriate conditions on $\norm{f}_2$, \ref{event:l2_norm}) hold with probability at least $1 - \frac{1}{b^2} - o(1)$.

\section{Bounding discrete Sobolev norm.}
Recall that our goal is to show \ref{event:discrete_sobolev_norm} occurs with high probability. We will build slowly to this goal.

\subsection{$s = 1$, $f$ Lipschitz.}
To start, we provide a bound in the case when $s = 1$, and $f$ has bounded Lipschitz norm. Precisely, we define the space of $1$-Holder (Lipschitz) functions $C_{\mathcal{X}}^{0,1}(1)$ to consist of all continuous functions $g: \D \to \Reals$ such that
\begin{equation*}
\norm{g}_{C_{\D}^{0,1}} := \norm{g}_{C(\D)} + [g]_{C_{\D}^{0,1}} \leq 1
\end{equation*}
where $\norm{g}_{C(\D)} = \sup_{x \in \D} \abs{g(x)}$, and
\begin{equation*}
[g]_{C_{\D}^{0,1}} = \sup_{x,y \in \D} \frac{\abs{g(x) - g(y)}}{\norm{x - y}}.
\end{equation*}

Let $P_n$ be the empirical distribution associated with $x_1,\ldots,x_n$, i.e. $P_n = \frac{1}{n}\sum_{i = 1}^{n} \delta_{x_i}$. We begin by providing a deterministic bound involving the distance between the measures $P_n$ and $P$. As these are measures over different spaces, it is not obvious how to relate them. We will use transportation distance to do so. Recall that a mapping $T: \D \to \D$ is a \emph{transportation map} between $P$ and $P_n$ if for all measurable $A \subseteq \D$,
\begin{equation*}
P_n(A) = P(T^{-1}(A))
\end{equation*}

\begin{lemma}
	\label{lem:holder_1_bound}
	For any $f \in C_{\mathcal{D}}^{0,1}$, the following bound holds on the Sobolev discrete norm:
	\begin{equation*}
	\norm{B\beta}_2^2 \leq n^2r^2(r + \norm{\mathrm{Id} - T}_{L^{\infty}(P)})^d
	\end{equation*}
\end{lemma}
\begin{proof}
	We write
	\begin{align*}
	\frac{1}{n^2}\norm{B^{(s)}\beta}_2^2 & = \frac{1}{n^2} \beta^T L \beta \\
	& = \frac{1}{2n^2} \sum_{i,j = 1}^{n} (\beta_i - \beta_j)^2 A_{ij} \\
	& = \int_{\D} \int_{\D} (f(x) - f(y))^2 \eta_r(x,y) \,dP_n(y) \,dP_n(x)
	\end{align*}
	We examine the inner integral. By the Holder property of $f$, and a change of variables, we obtain
	\begin{align*}
	\int_{\D} (f(x) - f(y))^2 \eta_r(x,y) \,dP_n(y) & \leq \int_{\D} \norm{x - y}^2 \eta_r(x,y) \,dP_n(y) \\
	& = \int_{\D} \norm{x - T(y)}^2 \eta_r(x,T(y)) \,dP(y) \\
	& \leq r^2 \int_{\D} \eta_r(x,T(y)) \,dP(y) \\
	& \leq r^2 \bigl( r + \norm{\mathrm{Id} - T}_{L^{\infty}(P)} \bigr)^d
	\end{align*}
	and the desired result is shown.
\end{proof}

\subsection{$s = 1$, $f \in \mathcal{W}^{1,2}(\D; 1)$.}

Let $P$ be an absolutely continuous probability measure over $\D$ with density function $p$ bounded above and below by constants, i.e
\begin{equation*}
0 < p_{\textrm{min}} < p(x) < p_{\textrm{max}} < \infty, \quad \textrm{for all $x \in \mathcal{D}$}
\end{equation*}

\begin{lemma}
	\label{lem:sobolev_1_bound}
	For any $f \in W^{1,2}(\mathcal{D};1)$, and any $b \geq 1$, we have that there exists a constant $C > 0$ which depends only on $d$ and $p_{\max}$ such that
	\begin{equation}
	\label{eqn:sobolev_1_bound}
	\norm{B\beta}_2^2 \leq b C n^2 r^{d + 2}
	\end{equation}
	with probability at least $1 - \frac{1}{b}$. 
\end{lemma}
\begin{proof}
	Observe that
	\begin{align*}
	\frac{1}{n^2}\Ebb(\beta^T L \beta) = \Ebb\left(\frac{1}{n^2} \sum_{i,j = 1}^{n} (\beta_i - \beta_j)^2 A_{ij}\right) & = \frac{(n - 1)}{n} \int_{\D} \int_{\D} (f(x) - f(y))^2\eta_r(x,y) \,dP(y) \,dP(x) \\
	& \leq p_{\max}^2 \frac{(n - 1)}{n} \int_{\D} \int_{\D} (f(x) - f(y))^2\eta_r(x,y) \,dy \,dx.
	\end{align*}
	We will show that for any $f \in \mathcal{W}^{1,2}(\D;1)$, there exists a constant $C$ which depends only on dimension $d$ such that
	\begin{equation}
	\label{eqn:mean_bound}
	\int_{\D} \int_{\D} (f(x) - f(y))^2\eta_r(x,y) \,dy \,dx \leq C r^{d + 2}
	\end{equation}
	whence the desired result of \eqref{eqn:sobolev_1_bound} follows by Markov's inequality.

	We begin by dealing with complications due to the boundary of $\mathcal{D}$. Let $V$ be any bounded open set such that $\D \subset \subset V$. Note that as $\partial \D$ is $C^1$, by Theorem \ref{thm:evans_extension} there exists $g \in W^{1,2}(\Reals^d)$ such that
	\begin{enumerate}
		\item
		\label{eqn:evans_extension_1}
		$g = f$, $P$-almost-everywhere in $\D$
		\item 
		$g$ has support within $V$, and  
		\item 
		\label{eqn:sobolev_1_bound_2}
		$\norm{g}_{W^{1,2}(\Rd)} \leq C \norm{f}_{W^{1,2}(\D)}$ for a constant $C$ which depends only on $\mathcal{D}$.
	\end{enumerate}
	As a result of the first point,
	\begin{equation}
	\label{eqn:sobolev_1_bound_1}
	\int_{\D} \int_{\D} (f(x) - f(y))^2\eta_r(x,y) \,dy \,dx \leq \int_{\Rd} \int_{\Rd} (g(x) - g(y))^2\eta_r(x,y) \,dy \,dx.
	\end{equation}

	Next, we smooth $g$, so that we may work with ordinary partial derivatives.
	We let $\kappa \in C^{\infty}(\Rd)$ be given by
	\begin{equation*}
	\kappa(x) :=
	\begin{cases}
	C \exp \left\{\frac{1}{\norm{x}^2 - 1}\right\}& \quad \textrm{if $\norm{x}_2 \leq 1$} \\
	0 & \quad \textrm{if $\norm{x}_2 \geq 1$}
	\end{cases}
	\end{equation*}
	where the normalizing constant $C > 0$ is chosen so that $\int_{\Rd} \eta dx = 1$. Let $\kappa_r(x) := (1/r^d) \kappa(x/r)$. Then, the mollification of $g$ by $\kappa_r$ is given by
	\begin{align*}
	g^r & := g \ast \eta_r \\
	& = \int_{\Rd} \eta_r(x - y) g(y) dy
	\end{align*}
	(Refer to \citep{evans10}, Appendix C, Theorem 7 for a proof that $g^r \in C^{\infty}(\Rd)$.)
	Adding and subtracting within \eqref{eqn:sobolev_1_bound_1}, we have
	\begin{align}
	\int_{\Rd} & \int_{\Rd} (g(x) - g(y))^2\eta_r(x,y) \,dy \,dx \\
	& \leq \int_{\Rd} \int_{\Rd} \bigl((g(x) - g^r(x))^2 + (g^r(x) - g^r(y))^2 + (g^r(y) - g(y))^2\bigr)\eta_r(x,y) \,dy \,dx \nonumber \\
	& = 2 \int_{\Rd} \int_{\Rd} \bigl((g(y) - g^r(y))^2 \eta_r(x,y) \,dy \,dx + \int_{\Rd} \int_{\Rd} \bigl((g^r(x) - g^r(y))^2 \eta_r(x,y) \,dy \,dx \label{eqn:sobolev_1_bound_5}
	\end{align}
	We deal with each summand individually, beginning with the first one. We have
	\begin{align}
	\int_{\Rd} \int_{\Rd} \bigl((g(y) - g^r(y))^2 \eta_r(x,y) \,dy \,dx & = \int_{\Rd} \int_{B(x,r)} \bigl((g(y) - g^r(y))^2\,dy \,dx \nonumber \\
	& = \int_{\Rd} \int_{B(y,r)} \bigl((g(y) - g^r(y))^2\,dx \,dy \tag{Tonelli's Theorem} \nonumber \\
	& = r^d \int_{\Rd}\bigl((g(y) - g^r(y))^2 \,dy \\
	& \leq r^{d + 2} \int_{\Rd} \norm{\nabla g(y)}^2 \,dy \label{eqn:sobolev_1_bound_4}
	\end{align}
	where the last line follows from Lemma \ref{lem:poincare_mollify}.

	We now turn our attention to the second summand. Note that as $g^r \in C^{\infty}(\Rd)$, we may apply Theorem \ref{thm:taylor_expansion} and obtain
	\begin{align*}
	(g^r(x) - g^r(y))^2 & = \left(\int_{0}^{1} \nabla g^r(x + t(y - x)) \cdot (y - x) \,dt\right)^2 \\
	& \leq \int_{0}^{1} \bigl(\nabla g^r(x + t(y - x)) \cdot (y - x)\bigr)^2 \,dt  \tag{Jensen's inequality} \\
	& \leq \norm{y - x}^2 \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 \,dt. \tag{Cauchy-Schwarz inequality}
	\end{align*}
	As a result, we have
	\begin{align*}
	\int_{\Rd} \int_{\Rd} \bigl((g^r(x) - g^r(y))^2 \eta_r(x,y) \,dy) \,dx & \leq \int_{\Rd} \int_{\Rd} \norm{y - x}^2  \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 \eta_r(x,y) \,dt \,dy \,dx \\
	& \leq r^{2} \int_{\Rd} \int_{\Rd} \int_{0}^{1} \norm{\nabla g^r(x + t(y - x))}^2 \eta_r(x,y) \,dt \,dy \,dx \\
	& = r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + t(y - x))}^2 \eta_r(x,y) \,dx \,dt \,dy \\
	& = r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + tz)}^2 \eta_r(z) \,dx \,dt \,dz \tag{$z = y - x$}
	\end{align*}

	where we write $\eta_r(z) = \mathbf{1}(\norm{z} \leq r)$ in an abuse of notation. Next, we note that
	\begin{equation*}
	\int_{\Rd} \norm{\nabla g^r(x + tz)}^2 \,dx = \int_{\Rd} \norm{\nabla g^r(x)}^2 \,dx \leq \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
	\end{equation*}
	with the inequality following from Lemma \ref{lem:gradient_mollify_commute}. Therefore,
	\begin{align}
	r^{2} \int_{\Rd} \int_{0}^{1} \int_{\Rd} \norm{\nabla g^r(x + tz)}^2 \eta_r(z) \,dx \,dt \,dz & \leq r^{2} \int_{\Rd} \eta_r(z) \int_{0}^{1} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx  \,dt \,dz \nonumber \\
	& = r^{2 + d} \int_{\Rd} \norm{\nabla g(x)}^2 \,dx. \label{eqn:sobolev_1_bound_3}
	\end{align}
	By \eqref{eqn:sobolev_1_bound_5}, \eqref{eqn:sobolev_1_bound_4} and \eqref{eqn:sobolev_1_bound_3}, we have that 
	\begin{equation*}
	\int_{\Rd} \int_{\Rd} (g(x) - g(y))^2 \eta_r(x,y) \,dx \,dy \leq 3 \int_{\Rd} \norm{\nabla g(x)}_2^2 \,dx
	\end{equation*}
	Then by \eqref{eqn:sobolev_1_bound_2}, $\int_{\Rd} \norm{\nabla g(x)}_2^2 \,dx \leq C\int_{\D} \norm{\nabla f(x)}_2^2 \,dx \leq C$ where $C$ is a constant depending only on $\D$. So the desired result of \eqref{eqn:sobolev_1_bound_1} follows.
\end{proof}

\section{Supporting Results.}

\begin{theorem}[\citep{evans10} Chapter 5.4, Theorem 1]
	\label{thm:evans_extension}
	Assume $U$ is bounded and $\partial U$ is $C^1$. Select a bounded open set $V$ such that $U \subset \subset V$ ($U$ is compactly contained in $V$). Then there exists a bounded linear operator $E: W^{1,2}(U) \to W^{1,2}(\Rd)$ such that for each $u \in W^{1,2}(U)$:
	\begin{enumerate}
		\item $Eu = u$ a.e. in $U$,
		\item $Eu$ has support within $V$, and 
		\item 
		\begin{equation*}
		\norm{Eu}_{W^{1,2}(\Rd)} \leq C \norm{u}_{W^{1,2}(\Rd)}
		\end{equation*}
		the constant $C$ depending only on $U$ and $V$.
	\end{enumerate}
\end{theorem}

For $u \in W^{1,2}(\Rd)$ and $x \in \Rd$, write $\nabla u(x) = (D^{e_1}(x),\ldots,D^{e_d}(x)$, for the gradient of $u$.

\begin{theorem}[\citep{evans10} Chapter 5.8.1, Theorem 2]
	\label{thm:evans_poincare}
	There exists a constant $C$, depending only on $d$, such that
	\begin{equation*}
	\norm{u - u^r}_{L^2(B(x,r))} \leq Cr\norm{\nabla u}_{L^2(B(x,r))}
	\end{equation*}
\end{theorem}

\begin{theorem}[Taylor expansion.]
	\label{thm:taylor_expansion}
	For any function $u \in C^1$, and any $x,y \in \Rd$, 
	\begin{equation*}
	u(y) - u(x) = \int_{0}^{1} \nabla(u(x + t(y - x))) \cdot (y - x) \,dt
	\end{equation*}
\end{theorem}

\begin{lemma}
	\label{lem:poincare_mollify}
	For any function $g \in W^{1,2}(\Rd)$ compactly supported in a bounded open set $V \subset \Rd$, we have
	\begin{equation}
	\label{eqn:poincare_mollify}
	\int_{\Rd} (g(x) - g^r(x))^2 \,dx \leq r^2 \int_{\Rd} \norm{\nabla g(x)}^2 \,dx
	\end{equation}
\end{lemma}
\begin{proof}
	This Lemma is essentially a reproduction of part of the proof of the Rellich-Kondrachov Compactness Theorem from \citep{evans10}. Note that it is sufficient to prove in the case when $g$ is smooth. To see this, for the moment assume \eqref{eqn:poincare_mollify} holds for all $u \in C^{\infty}(V)$, and let $g \in W^{1,2}(V)$. By Theorem \ref{thm:local_approx_smooth_functions}, we may take a sequence $g_m \in C^{\infty}(V)$ such that
	\begin{equation*}
	\norm{g - g_m}_{L^2(V)} \to 0, \quad \textrm{and} \quad \norm{\nabla g - \nabla g_m}_{L^2(V)} \to 0,
	\end{equation*}
	Then we have
	\begin{align*}
	\int_{\Rd} (g(x) - g^r(x))^2 \,dx & \leq \int_{\Rd} (g(x) - g_m(x))^2 \,dx + \int_{\Rd}(g_m(x) - g_m^r(x))^2 \,dx + \int_{\Rd}(g_m^r(x) - g^r(x))^2 \,dx \\
	& \leq \int_{\Rd} (g(x) - g_m(x))^2 \,dx + r^2 \int_{\Rd} \norm{\nabla g_m(x)}^2 \,dx + \int_{\Rd}(g_m^r(x) - g^r(x))^2 \,dx
	\end{align*}
	and taking the limit as $m$ goes to infinity, the right hand side converges to $\int_{\Rd} \norm{\nabla g(x)}^2 \,dx$.

	It remains to show \eqref{eqn:poincare_mollify} in the case where $g$ is smooth. In this case, we have
	\begin{align*}
	g^{r}(x) -  g(x) & = \frac{1}{r^d} \int_{B(x,r)} \kappa \left(\frac{x - z}{r}\right)\bigl(g(z) - g(x)\bigr) \,dz \\
	& = \int_{B(0,1)} \kappa(y) \bigl(g(x - ry) - g(x)\bigr) \,dy \\
	& = \int_{B(0,1)} \kappa(y) \int_{0}^{1} \frac{d}{dt} \bigl(g(x - try) \bigr) \,dt \,dy \\
	& = -r \int_{B(0,1)} \kappa(y) \int_{0}^{1} \bigl(\nabla g (x - try) \bigr) \cdot y \,dt \,dy.
	\end{align*} 
	Therefore, by Jensen's and Cauchy-Schwarz inequalities, we have
	\begin{align*}
	\int_{\Rd} (g(x) - g^r(x))^2 \,dx & \leq r^2 \int_{\Rd} \int_{B(0,1)} \kappa(y) \int_{0}^{1} \norm{\nabla g (x - try)}^2 \norm{y}^2 \,dt \,dy \,dx \\
	& \leq r^2 \int_{B(0,1)} \kappa(y) \int_{0}^{1} \int_{\Rd} \norm{\nabla g (x - try)}^2 \,dx \,dt \,dy \\
	& = r^2 \int_{\Rd} \norm{\nabla g(z)}^2 \,dz
	\end{align*}
\end{proof}

The following theorem is Theorem 1 in Section 5.3 of \citep{evans10}.
\begin{theorem}[Local approximation by smooth functions.]
	\label{thm:local_approx_smooth_functions}
	Assume $U$ is bounded, and $u \in W^{k,p}(U)$ for some $1 \leq p < \infty$. Then there exists functions $u_m \in C^{\infty}(U) \cap W^{k,p}(U)$ such that
	\begin{equation*}
	\norm{u_m - u}_{W^{k,p}(U)} \overset{m}{\to} 0.
	\end{equation*}
\end{theorem}

\begin{lemma}
	\label{lem:gradient_mollify_commute}
	For any $u \in W^{1,2}(\Rd)$, we have
	\begin{equation*}
	\int_{\Reals^d} \norm{\nabla u^r(x)}_2^2 \,dx \leq \int_{\Reals^d} \norm{\nabla u(x)}_2^2 \,dx
	\end{equation*}
\end{lemma}
\begin{proof}
	Observe that $\norm{\nabla u(x)}_2^2 = \sum_{j = 1}^{d} (D^{e_j}u(x))^2 $. Therefore,
	\begin{align*}
	\int_{\Rd} \norm{\nabla u^r(x)}_2^2 \,dP(x) & = \sum_{j = 1}^{d} \int_{\Rd} (D^{e_j}u^r(x))^2 \,dx \\
	& = \sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u)^r(x))^2 \,dx
	\end{align*}
	where the second equality follows from equation (1) in Section 5.3 of \citep{evans10}. Then, for any $v \in L^2(\Rd)$, we have that
	\begin{align*}
	\abs{v^r(x)} & = \int_{\Rd} \kappa_r^{1/2}(x - y) \kappa_r^{1/2}(x - y) v(y) \,dy \\
	& \leq \left(\int_{\Rd} \kappa_r^(x - y) \,dy\right)^{1/2} \left(\int_{\Rd} \kappa_r(x - y) v^2(y) \right)^{1/2} \\
	& = \left(\int_{\Rd} \kappa_r(x - y) v^2(y) \right)^{1/2}
	\end{align*}
	and therefore
	\begin{align*}
	\int_{\Rd} (v^r(x))^2 \,dx & \leq \int_{\Rd} \int_{\Rd } \kappa_r(x - y) v^2(y) \,dy \,dx \\
	& = \int_{\Rd} \int_{\Rd } \kappa_r(x - y) v^2(y) \,dy \,dx \\
	& = \int_{\Rd} v^2(y) \,dy
	\end{align*}

	Applying this to $D^{e_j}u \in L^2(\Rd)$, we have that
	\begin{equation*}
	\sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u)^r(x))^2 \,dx \leq \sum_{j = 1}^{d} \int_{\Rd} ((D^{e_j}u(x))^2 \,dx = \int_{\Rd} \norm{\nabla u(x)}^2 \,dx.
	\end{equation*}
\end{proof}

\bibliographystyle{plainnat}
\bibliography{../../graph_testing_bibliography}

\end{document}