### 2019-11-23 notes/weekly_notes/11.8.19.NotesForWeek.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\distiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\domain}{\mathcal{X}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\Partial}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 11/1/19 - 11/8/19}
\author{Alden Green}
\date{\today}
\maketitle

Let $P$ and $Q$ be distributions on $\Rd$, absolutely continuous with respect to Lebesgue measure with densities $p$ and $q$, and let $\mu = (P + Q)/2$. For fixed points $Z = \set{z_1,\ldots,z_N} \subset \mathcal{Z}$ for some subset $\mathcal{Z}$ of the support of $\mu$, suppose we observe $\ell = (\ell_1,\ldots,\ell_N) \in \Reals^n$, where $\ell_1,\ldots,\ell_N$ are independent Rademacher random variables with means
\begin{equation*}
\mathbb{E}[\ell_i] = \frac{p(z_i) - q(z_i)}{p(z_i) + q(z_i)}.
\end{equation*}

Let $\theta_p = (p(z_1),\ldots,p(z_N))$ and similarly $\theta_q = (q(z_1),\ldots,q(z_N))$; let $\theta_{\mu} = \frac{1}{2}\theta_p + \frac{1}{2}\theta_q$. Our goal: use the data $\ell,Z$ to test the following hypothesis:
\begin{equation*}
\textbf{H}_0: \theta_p = \theta_q ~~\textrm{versus}~~ \textbf{H}_a: d(\theta_p,\theta_q) > \epsilon 
\end{equation*}
for some metric $d$ on the space $\Reals^n$, and some $\epsilon > 0$. For a given test $\phi: \{-1,1\}^N \times \mathcal{Z} \to \{0,1\}$, we will evaluate our test error using the sum of Type I and Type II error
\begin{equation*}
\mathcal{R}(\phi;\theta_p,\theta_q) = \mathbb{E}_{\theta_\mu,\theta_\mu}(\phi) + \mathbb{E}_{\theta_p,\theta_q}(1 - \phi),
\end{equation*}
and the worst case risk
\begin{equation*}
\mathcal{R}(\phi,\epsilon,\Theta) = \sup_{\theta_p,\theta_q \in \Theta} \mathcal{R}(\phi; \theta_p, \theta_q)
\end{equation*}
where $\Theta$ is a subset of $\Reals^n$ and the supremum is over all vectors $\theta_p,\theta_q$ such that $d(\theta_p,\theta_q) \geq \epsilon$.

In particular, we propose and analyze the following projection-based test. Letting $G_{N,r}$ be the $r$-neighborhood graph over $Z$ with combinatorial Laplacian $L$, we write the spectral decomposition of $L$ as $L = V S V^T$, where $V = (v_1,\ldots,v_N)$ is $N \times N$ orthonormal matrix, and $S$ is a diagonal matrix with entries $s_1 \leq s_2 \leq \ldots \leq s_N$. Our test statistic $T$ depends on a tuning parameter $C > 0$ in the following manner:
\begin{equation*}
T = T(\ell,Z) = \frac{1}{N}\sum_{k: s_k \leq C} \left(\dotp{\ell}{v_k}\right)^2
\end{equation*}
and our test $\phi := \phi(\ell,Z)$ will simply be $\phi = \1\{T \geq \frac{\kappa}{N} + b\tau\}$ for a threshold $\tau > 0$ to be specified later.

\subsection{Moments of $T$.}
To upper bound the test error we must show that the fluctuations of $T$ (under null or alternative)  are small compared to the difference in means $\Ebb_{\theta_{p},\theta_q}(T) - \Ebb_{\theta_{\mu},\theta_{\mu}}(T)$. We therefore must compute the first and second moments of $T$, under null and alternative. To compute the moments of $T$, we decompose the vector $\ell$ into a mean term and a noise term. Letting $\Delta_{P,Q} = \frac{\theta_P - \theta_q}{2\theta_{\mu}}$, we have $\ell = \Delta_{P,Q} + \varepsilon$ where $\varepsilon = (\varepsilon_1,\ldots,\varepsilon_n)$ consists of $n$ independent mean-zero random variables, distributed
\begin{equation*}
\varepsilon_i = 
\begin{cases}
\frac{\theta_{Q,i}}{\theta_{\mu,i}}, ~ & \textrm{with probability $\frac{\theta_{P,i}}{\theta_{\mu,i}}$} \\
-\frac{\theta_{P,i}}{\theta_{\mu,i}}, ~ & \textrm{with probability $\frac{\theta_{Q,i}}{\theta_{\mu,i}}$}.
\end{cases}
\end{equation*}

\subsubsection{Mean of $T$.}
Let $\kappa$ be the largest integer $k \in [N]$ such that $s_k \leq C$. (Here $\kappa$ is a fixed quantity.) Then,
\begin{align*}
\Ebb(T) & = \frac{1}{N}\sum_{k = 1}^{\kappa} \Ebb\left[\dotp{v_k}{\ell}^2\right] \\
& = \frac{1}{N}\sum_{k = 1}^{\kappa} \Ebb\left[\dotp{\Delta_{P,Q} + \epsilon}{v_k}^2\right] \\
& = \frac{1}{N}\sum_{k = 1}^{\kappa} \bigl\{\dotp{\Delta_{P,Q}}{v_k}^2 + \Ebb\left[\dotp{\epsilon}{v_k}^2\right]\bigr\}
\end{align*}
For a given $k \in [\kappa]$, we have
\begin{align*}
\Ebb\left[\dotp{\epsilon}{v_k}_2^2\right] & = \sum_{i = 1}^{N} \sum_{j = 1}^N \Ebb\left[\varepsilon_i \varepsilon_j v_{k,i} v_{k,j}\right] \\
& = \sum_{i = 1}^{N} v_{k,i}^2 \Ebb\left[\varepsilon_i^2 \right]
\end{align*}
and by a direct computation of $\Ebb\left[\varepsilon_i^2 \right]$ we obtain
\begin{align*}
\mathbb{E}(T) & = \frac{1}{N}\left(\sum_{k = 1}^{\kappa} \dotp{\Delta_{P,Q}}{v_k}^2 + \sum_{k = 1}^{\kappa} \sum_{i = 1}^{N} v_{k,i}^2 \frac{\theta_{P,i} \theta_{Q,i}}{\theta_{\mu,i}^2}\right) \\
& = \frac{1}{N}\left(\sum_{k = 1}^{\kappa} \dotp{\Delta_{P,Q}}{v_k}^2 + \kappa - \sum_{k = 1}^{\kappa} \sum_{i = 1}^{N} v_{k,i}^2 \Delta_{P,Q,i}^2\right) \\
& = \frac{1}{N}\left(\sum_{k = 1}^{\kappa} \dotp{\Delta_{P,Q}}{v_k}^2 + \kappa - \sum_{i = 1}^{N} \Pi_{\kappa,ii} \Delta_{P,Q,i}^2\right) \\
& \geq \frac{1}{N}\left(\sum_{k = 1}^{\kappa} \dotp{\Delta_{P,Q}}{v_k}^2 + \kappa - \Pi_{\max} \norm{\Delta_{P,Q}}^2\right)
\end{align*}
Under the null, $\Delta_{P,Q} = 0$ and therefore
\begin{equation}
\label{eqn:mean_null}
\mathbb{E}_{\theta_{\mu},\theta_{\mu}}(T) = \frac{\kappa}{N}.
\end{equation}
Under the alternative, we have the following lower bound on the expectation of 
\begin{align*}
\sum_{k = 1}^{\kappa} \dotp{\Delta_{P,Q}}{v_k}^2 & = \sum_{k = 1}^{N} \dotp{\Delta_{P,Q}}{v_k}^2  - \sum_{k = \kappa + 1}^{N} \dotp{\Delta_{P,Q}}{v_k}^2 \\
& = \norm{\Delta_{P,Q}}_2^2 - \sum_{k = \kappa + 1}^{N} \frac{1}{s_k}\dotp{\Delta_{P,Q}}{v_k}^2 s_k \\
& \geq \norm{\Delta_{P,Q}}_2^2 - \frac{1}{C} \dotp{L\Delta_{P,Q}}{\Delta_{P,Q}}.
\end{align*}
leading to the lower bound
\begin{equation}
\label{eqn:mean_alternative}
\mathbb{E}_{\theta_p,\theta_q}(T) \geq \frac{\kappa}{N} + \frac{(1 - \Pi_{\max})}{N}\norm{\Delta_{P,Q}}^2 - \frac{1}{CN} \dotp{L\Delta_{P,Q}}{\Delta_{P,Q}}.
\end{equation}

\subsubsection{Variance of $T$.}
Let $V_{\kappa} = (v_1\ldots v_{\kappa})$ be the $n \times \kappa$ matrix containing the first $\kappa$ eigenvectors of $L$, and let $\Pi_{\kappa} = V_{\kappa} V_{\kappa}^T$. Observing that $T = \frac{1}{N}\dotp{\Pi_{\kappa}(\Delta_{P,Q} + \varepsilon)}{\Delta_{P,Q} + \varepsilon}$, we have
\begin{align}
\Var(T) & = \frac{1}{N^2}\Var(2\dotp{\Pi_{\kappa}\Delta_{P,Q}}{\varepsilon} + \dotp{\Pi_{\kappa}\varepsilon}{\varepsilon}) \nonumber \\
& = \frac{1}{N^2}\Bigl\{4\underbrace{\Var(\dotp{\Pi_{\kappa}\Delta_{P,Q}}{\varepsilon})}_{=: V_1} + 2\underbrace{\Cov(\dotp{\Pi_{\kappa}\Delta_{P,Q}}{\varepsilon},\dotp{\Pi_{\kappa}\varepsilon}{\varepsilon})}_{=: K_1} + \underbrace{\Var(\dotp{\Pi_{\kappa}\varepsilon}{\varepsilon})}_{=:V_2}\Bigr\}, \label{eqn:var}
\end{align}
and we now upper bound each of the three terms on the right hand side of the previous display.
\paragraph{Upper bound on $V_1$:}
Let $\Sigma := \Cov(\varepsilon)$ be the covariance matrix of $\varepsilon$. Noting that $\Sigma \preceq I$, we have
\begin{equation}
\label{eqn:var1}
\Var(\dotp{\Pi_{\kappa}\Delta_{P,Q}}{\varepsilon}) =\Delta_{P,Q}^T \Pi_{\kappa}^T \Sigma \Pi_{\kappa} \Delta_{P,Q} \leq \norm{\Pi_{\kappa}\Delta_{P,Q}}^2.
\end{equation}

\paragraph{Upper bound on $K_1$:}
Noting that $\Ebb(\dotp{\Pi_{\kappa} \Delta_{P,Q}}{\varepsilon}) = 0$, we have that
\begin{align*}
K_1 & = \Ebb\left[\dotp{\Pi_{\kappa} \Delta_{P,Q}}{\varepsilon} \dotp{\Pi_{\kappa} \varepsilon}{\varepsilon}\right] \\
& = \sum_{i = 1}^{N} \sum_{j = 1}^{N} \sum_{i' = 1}^{N} \Ebb\left[\varepsilon_i \varepsilon_j \Pi_{\kappa,ij} \varepsilon_{i'} (\Pi_{\kappa}\Delta_{P,Q})_i\right] \\
& = \sum_{i = 1}^{N} \Ebb\left[\varepsilon_i^3\right] \Pi_{\kappa,ii} (\Pi_{\kappa} \Delta_{P,Q})_i.
\end{align*}
Standard computations yield $\Ebb\left[\varepsilon_i^3\right]  = \frac{1}{2}(1 - \Delta_{P,Q,i}^2)(\Delta_{Q,P,i})$, and plugging in to the previous expression by Cauchy-Schwarz we have
\begin{align}
K_1 = \sum_{i = 1}^{N} \Ebb\left[\varepsilon_i^3\right] \Pi_{\kappa,ii}  (\Pi_{\kappa}\Delta_{P,Q})_i & = \frac{1}{2}\sum_{i = 1}^{N} (1 - \Delta_{P,Q,i}^2)(\Delta_{Q,P,i}) \Pi_{\kappa,ii}  (\Pi_{\kappa}\Delta_{P,Q})_i \nonumber \\
& \leq \norm{\Pi_{\kappa} \Delta_{P,Q}} \cdot \left(\sum_{i = 1}^{N} \Delta_{P,Q,i}^2 \Pi_{\kappa,ii}^2\right)^{1/2} \nonumber \\
& \leq \Pi_{\max} \cdot \norm{\Pi_{\kappa} \Delta_{P,Q}} \cdot \norm{\Delta_{P,Q}}  \label{eqn:var2} 
\end{align}

\paragraph{Upper bound on $V_2$:}
$V_2$ is a variance of a sum, which we re-express as the sum of covariances:
\begin{align*}
V_2 & = \Var(\dotp{\Pi_{\kappa} \varepsilon}{\varepsilon}) \\
& = \sum_{i,j,i',j' = 1}^{N} (\Pi_{\kappa})_{ij} (\Pi_{\kappa})_{i'j'} \Cov(\varepsilon_i \varepsilon_j, \varepsilon_{i'} \varepsilon_{j'}).
\end{align*}
This covariance will be non-zero only when $i = i' \neq j = j'$, $i = j' \neq j = i'$, or $i = i' = j = j'$, and therefore
\begin{align}
V_2 & = 2\sum_{i = 1}^{N}\sum_{j = 1}^{N} (\Pi_{\kappa})_{ij}^2 \Var(\varepsilon_i\varepsilon_j) + \sum_{i = 1}^{N} (\Pi_{\kappa})_{ii}^2 \Var(\varepsilon_i^2) \nonumber\\
& \leq 2\sum_{i = 1}^{N}\sum_{j = 1}^{N} (\Pi_{\kappa})_{ij}^2 + \sum_{i = 1}^{N} (\Pi_{\kappa})_{ii}^2 \leq 3 \mathrm{tr}(\Pi_{\kappa}^2) = 3\kappa. \label{eqn:var3}
\end{align}

\paragraph{Putting the pieces together.}
Plugging \eqref{eqn:var1}, \eqref{eqn:var2}, and \eqref{eqn:var3} back into \eqref{eqn:var}, we obtain
\begin{equation}
\label{eqn:var_5}
\Var(T) \leq \frac{1}{N^2}\left\{4\norm{\Pi_{\kappa} \Delta_{P,Q}}^2 + \Pi_{\max} \cdot \norm{\Pi_{\kappa} \Delta_{P,Q}} \cdot \norm{\Delta_{P,Q}} + 3\kappa \right\}.
\end{equation}

\subsection{Type I and Type II error.}

We translate our bounds on the mean and variance of $T$ to bounds on the Type I and Type II error of our test $\phi$ using Chebyshev's inequality. Recall that our test statistic is of the form $\phi = \1\{T \geq \frac{\kappa}{N} + b\tau\}$, and let $\tau = \frac{\sqrt{3\kappa}}{N}$.

\paragraph{Type I error.}
We apply Chebyshev's inequality, and obtain
\begin{align*}
\Pbb_{\theta_{\mu},\theta_{\mu}}\left(T \geq \frac{\kappa}{N} + b\tau\right) & = \Pbb_{\theta_{\mu},\theta_{\mu}}\left(T - \frac{\kappa}{N} \geq b\tau - \Ebb(T)\right) \\
& = \Pbb_{\theta_{\mu},\theta_{\mu}}\left( (T - \frac{\kappa}{N})^2 \geq b^2\tau^2 \right) \\
& \leq \frac{\Var_{\theta_{\mu},\theta_{\mu}}(T)}{b^2\tau^2} = \frac{1}{b^2}.
\end{align*}

\paragraph{Type II error.}
To obtain meaningful bounds on Type I error, we must assume some separation between $\theta_P$ and $\theta_Q$. In particular, letting $d(\theta_P,\theta_Q) := \norm{\Delta_{P,Q}}^2$ we assume
\begin{equation}
\label{eqn:dist_thetaP_thetaQ}
\frac{1}{N}d(\theta_P,\theta_Q) \geq \frac{1}{(1 - \Pi_{\max})} \left(\frac{1}{CN}\norm{D \Delta_{P,Q}}^2 + 2b\tau\right).
\end{equation}
Note that this implies
\begin{equation*}
\mathbb{E}_{\theta_P,\theta_Q}(T) - \left(\frac{\kappa}{N} + b\tau\right) \geq \frac{(1 - \Pi_{\max})}{N} \norm{\Delta_{P,Q}}^2 - \frac{1}{CN}\norm{D \Delta_{P,Q}}^2 - b\tau \geq b\tau \vee \frac{(1 - \Pi_{\max})}{2N} \norm{\Delta_{P,Q}}^2.
\end{equation*}
Now, we apply Chebyshev's inequality:
\begin{align*}
\Pbb\left(T \leq \frac{\kappa}{N} + b\tau\right) & = \Pbb\bigl(T - \Ebb[T] \leq \frac{\kappa}{N} + b\tau - \Ebb[T]\bigr) \\
& = \Pbb\bigl( (T - \Ebb[T])^2 \leq (\Ebb[T] - \frac{\kappa}{N} - b\tau)^2 \bigr) \\
& \leq \frac{\Var(T)}{(\Ebb[T] - \frac{\kappa}{N} - b\tau)^2} \\
& \leq \frac{1}{N^2}\frac{4\norm{\Pi_{\kappa} \Delta_{P,Q}}^2 + \Pi_{\max} \cdot \norm{\Pi_{\kappa} \Delta_{P,Q}} \cdot \norm{\Delta_{P,Q}} + 3\kappa}{b^2\tau^2 \vee \frac{(1 - \Pi_{\max})^2}{4N^2} \norm{\Delta_{P,Q}}^4} \\
& \leq \frac{16}{(1 - \Pi_{\max})^2 \norm{\Delta_{P,Q}}^2} \vee \frac{4\Pi_{\max}}{(1 - \Pi_{\max})^2 \norm{\Delta_{P,Q}}^2} \vee \frac{1}{b^2} \\
& \leq \frac{3}{b^2}
\end{align*}
where the last line follows by \eqref{eqn:dist_thetaP_thetaQ} whenever $N \geq \sqrt{48}$.

\section{Upper bounds on $\Pi_{\max}$.}
Clearly, the critical radius established by inequality~\eqref{eqn:dist_thetaP_thetaQ} is only meaningful if $1 - \Pi_{\max} > 0$. In fact, we will want to show that $\Pi_{\max} < \frac{1}{2}$ so the factor of $(1 - \Pi_{\max})^{-1}$ does not affect the rate at which the critical radius shrinks. We review what $\Pi_{\max}$ looks like on some common graphs.

\subsection{Chain.}
Suppose $G := G_{\mathrm{1N}}$ is a length-$N$ chain, i.e. a 1d grid graph on vertices $V = 1:N$. The eigenvectors of $G$ are given by the discrete Fourier transform;
\begin{equation*}
v_{1,i} = \frac{1}{\sqrt{N}},~~ v_{k,i} =\sqrt{\frac{2}{N}} \cos\left(\frac{(i - .5)(k - 1)\pi}{N}\right)~ \textrm{for $i = 1,\ldots,N$, k = $2,\ldots,N$.}
\end{equation*}
Clearly $\Pi_{\max} \leq \frac{2\kappa}{N}$, so that whenever $\kappa < \frac{N}{4}$ we have $\Pi_{\max} < \frac{1}{2}$ as desired.

\subsection{Grid.}
Now suppose $G := G_{\mathrm{dN}}$ is a $d$-dimensional grid graph on $N$ nodes, i.e. $G = G_{1M} \otimes \cdots \otimes G_{1M}$ for $M = \frac{N}{d}$ (assume $M$ is an integer for simplicity). The product structure of $G$ yields an expression for its eigenvectors from the eigenvectors of $G_{\mathrm{1M}}$. In particular, for a given $i = (i_1,\ldots,i_d) \in [M]^d$ and $k = (k_1,\ldots,k_d) \in [M]^d$, we have
\begin{equation*}
v_{k,i} = \prod_{j = 1}^{d} \sqrt{\frac{2}{M}} \cos\left(\frac{(i_j - .5)(k_j - 1)\pi}{M}\right)
\end{equation*}
and therefore $v_{k,i}^2 \leq \frac{2}{M^{d}} = \frac{2}{N}$ for all $i,k \in [M]^d$. So we arrive at the same bound $\Pi_{\max} \leq \frac{2\kappa}{N}$ and again $\kappa < \frac{N}{4}$ implies $\Pi_{\max} < \frac{1}{2}$ as desired.

\subsection{Spectral similarity.}
Suppose we have graphs $G$ and $H$ with Laplacians $L_G$ and $L_H$, such that $G$ and $H$ are $\epsilon$ spectrally-similar, meaning
\begin{equation}
\label{eqn:spectral_similarity}
(1 - \epsilon) x^T L_H x \leq x^T L_G x \leq (1 + \epsilon) x^T L_H x,~~ \textrm{for all $x \in \Reals^N$.}
\end{equation}
Write $L_G = V \Lambda V^T$ and $L_H = U S U^T$ for the spectral decompositions of $G$ and $H$, respectively, and assume that we have
\begin{equation*}
\max_{(i,l) \in [N]^2} u_{l,i}^2 \leq \frac{2}{N},
\end{equation*}
which we have already shown holds for $H = G_{dN}$. For each $k$, and for any $L(k), R(k) \in [N]$, write
\begin{equation*}
v_{k} = \sum_{l = R(k)}^{L(k)}\alpha_{k,l} u_l +  \textrm{proj}_{U_{L(k)}^{\perp}}(v_k);
\end{equation*}
as a result, we have that
\begin{align*}
v_{k,i}^2 & \leq 2\sum_{h = R(k)}^{L(k)} \sum_{l = R(k)}^{L(k)}\alpha_{k,h}\alpha_{k,l} u_{h,i} u_{l,i} + 2(\textrm{proj}_{U_{R(k):L(k)}^{\perp}}(v_k))_i^2 \\
& \leq \frac{4}{N} \norm{\alpha}_1^2 + 2\norm{ \textrm{proj}_{U_{L(k)}^{\perp}}(v_k)}^2 \\
& \leq \frac{4}{N} (L(k) - R(k)) + 2\norm{ \textrm{proj}_{U_{R(k):L(k)}^{\perp}}(v_k)}^2
\end{align*}
where the last inequality follows since $\norm{\alpha}_2 \leq 1$, and the inequality $\norm{\alpha}_1 \leq \sqrt{L - R} \norm{\alpha}_2$, since $\alpha$ is a length $L - R$-vector.

\textcolor{red}{TODO:} Upper bound $\norm{ \textrm{proj}_{U_{L(k)}^{\perp}}(v_k)}^2$ using~\eqref{eqn:spectral_similarity}, and choose $L(k),R(k)$ to make the sum in the previous display as small as possible.
\section{Notation.}

We write $\Pi_{\max} = \max_{i = 1,\ldots,n} \Pi_{\kappa,ii}$ where we recall $\Pi_{\kappa} = V_{\kappa} V_{\kappa}^T$ and therefore $\Pi_{\kappa,ii} = \sum_{k = 1}^{\kappa} v_{k,i}^2$.

We write $D$ for the incidence matrix of $L$.

\end{document}