### 2020-01-09 notes/weekly_notes/1.3.20.NotesForWeek.tex
added weekly notes for break

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Xsetiam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{\mathcal{L}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\Unq}{\mathrm{Unq}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 12/30/19 - 1/3/19}
\author{Alden Green}
\date{\today}
\maketitle

Suppose we observe independent samples $X = \set{x_1,\ldots,x_n}$ from a distribution $P$ with density $p$ supported on $\mathcal{X} = [0,1]^d$. Let $G_{n,r}$ be the random geometric graph of radius $r$ formed over the samples $X$, that is, the neighborhood graph with kernel $K(x) = \1(\norm{x} \leq r)$, and let $L$ denote the corresponding graph Laplacian matrix. In this week's notes, we will show that as long as $r$ is chosen to be sufficiently small, the eigenvalues of $L$ we care about scale at an appropriate rate. More precisely, we will prove the following Lemma.
\begin{lemma}
	\label{lem:higher_order_eigenvalue_tail_bound}
	Let $\kappa = n^{2d/(4s + d)}$. Suppose $p$ is uniformly bounded away from $0$ and $\infty$ on its support,
	\begin{equation*}
	0 < p_{\min} < p(x) < p_{\max} < \infty,~~\textrm{for all $x \in \Xset$.}
	\end{equation*}
	Then there exists a constant $c$ such that for any $r \leq n^{-4/((2+d)(4s + d))}$, the eigenvalue $\lambda_{\kappa}$ is lower bounded,
	\begin{equation}
	\label{eqn:higher_order_eigenvalue_tail_bound}
	\lambda_{\kappa} \geq c\kappa^{2/d}nr^{2 + d}
	\end{equation}
	with probability $1 - o(n^{-1})$ as $n \to \infty$.
\end{lemma}
Lemma~\ref{lem:higher_order_eigenvalue_tail_bound} establishes that eigenvalues of the graph Laplacian, suitably normalized, scale comparably to the eigenvalues of the continuum Laplacian operator $\mathcal{D}f := \textrm{div}(\nabla f)$. Along with an upper bound on the graph Sobolev semi-norm $R_{s,n}(f)$ (established in the \textit{12.20.19 Notes}), this allows us to argue that if $f$ is a smooth function with large $\mathcal{L}_2$ norm, then the projection of a function $f$ onto the eigenvectors $v_1,\ldots,v_{\kappa}$ is still quite large (measured in empirical norm); or, equivalently, the bias induced by projection onto these eigenvectors is small.

\section{Proofs}

\subsection{Proof of Lemma~\ref{lem:higher_order_eigenvalue_tail_bound}}

We prove Lemma~\ref{lem:higher_order_eigenvalue_tail_bound} by comparing $G_{n,r}$ to the tensor product of a $d$-dimensional lattice and a complete graph. The latter is a highly structured graph with known eigenvalues, which as we will see are sufficiently lower bounded for our purposes.

Let $\wt{r} = r/(3(\sqrt{d} + 1)), M = (1/\wt{r})^d, N = n\wt{r}^d$. Assume without loss of generality that $M$ and $N$ are integers. Additionally, for $t = n^{1/d}$ and $m = M^{1/d}$ let 
\begin{equation*}
\overline{X} = \set{\frac{1}{t}(k_1,\ldots,k_d): k \in [t]^d},~~ \overline{Z} = \set{\frac{1}{m}(j_1,\ldots,j_d): j \in [m]^d}.
\end{equation*}
For a given $\overline{z}_j \in \overline{Z}$, we write $Q(z_j) = m^{-1}[j_1 - 1,j_1] \times \cdots \times m^{-1}[j_d - 1,j_d]$ for the cube of side length $1/m$ with $z_j$ at one corner.

Consider the graph $H = (\overline{X}, E_H)$, where $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$ if
\begin{equation*}
\textrm{there exists}~\ol{z}_i, \ol{z}_j \in \ol{Z}~\textrm{such that}~\ol{x}_k \in Q(\ol{z}_i),~ \ol{x}_\ell \in Q(\ol{z}_j),~\textrm{and}~\norm{i - j}_1 \leq 1.
\end{equation*}
On the one hand $H \cong \ol{G}^M_d \otimes K_N$ where $\ol{G}^M_d$ is the $d$-dimensional lattice on $M$ nodes, and $K_N$ is the complete graph on $N$ nodes. On the other hand, we now show that with high probability $G_{n,r} \succeq H$. If $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$, then there exist $\ol{z}_i, \ol{z}_j$ such that
\begin{equation*}
\norm{\ol{x}_k - \ol{x}_{\ell}}_2 \leq m^{-1} + \norm{\ol{x}_k - \ol{z}_{i}}_2 + \norm{\ol{x}_{\ell} - \ol{z}_{j}}_2 \leq \wt{r}(1 + \sqrt{d}) = r/3.
\end{equation*}

By Theorem 1.1 of \textcolor{red}{Garcia Trillos and Slepcev}, with probability at least $1 - n^{-1}$ there exists a bijection $\pi: \overline{X} \to X$ such that
\begin{equation}
\label{eqn:transport_distance}
\max_{k \in [t]^d} \abs{\overline{x}_k - \pi(\overline{x}_k)} \leq c \left(\frac{\log n}{n}\right)^{1/d}
\end{equation}
Assuming~\eqref{eqn:transport_distance} holds, if $(\ol{x}_k, \ol{x}_{\ell}) \in E_H$, then for sufficiently large $n$
\begin{equation*}
\norm{\pi(\ol{x}_k) - \pi(\ol{x}_\ell)}_2 \leq 2 c \left(\frac{\log n}{n}\right)^{1/d} + \frac{r}{3} \leq r,
\end{equation*}
implying that $(\pi(\ol{x}_k), \pi(\ol{x}_{\ell})) \in E$. Therefore, $G_{n,r} \succeq \ol{G}^M_d \otimes K_N$ whenever~\eqref{eqn:transport_distance} holds.

The eigenvalues of lattices and complete graphs are known to satisfy, respectively
\begin{equation*}
\lambda_k(\ol{G}^{M}_d) \geq \frac{k^{2/d}}{M^{2/d}}~\textrm{for $k = 0,\ldots,M - 1$},~~ \textrm{and}~\lambda_{j}(K_N) \geq N\1\{j > 0\}~\textrm{for $j = 0,\ldots,N-1$.}
\end{equation*}
and by standard facts regarding the eigenvalues of tensor product graphs (\textcolor{red}{reference}), we have that the spectrum $\Lambda(H)$ satisfies
\begin{equation*}
\Lambda(H) = \set{N\lambda_k(\ol{G}^{M}_d) + M\lambda_j(K_N): \textrm{for $k = 0,\ldots,M - 1$ and $j = 0,\ldots,N-1$}}
\end{equation*}
For all $j = 1,\ldots,N-1$, we have that $M\lambda_j(K_N) = MN = n$. Therefore,
\begin{align*}
\lambda_{\kappa}(H) & \geq \{n \wedge N\lambda_{\kappa}(\ol{G}^{M}_d)\} \\
& \geq \{n \wedge n\wt{r}^d\frac{\kappa^{2/d}}{M^{2/d}}\} \\
& \geq \{n \wedge (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d}\} \\
& \geq (3\sqrt{d} + 3)^{-(2+d)}nr^{d+2}\kappa^{2/d},
\end{align*}
where the last inequality can be verified by a quick calculation in light of $\kappa = n^{2d/(4s + d)}$ and $r \leq n^{-4/((2+d)(4s + d))}$. Since we've already shown $\lambda_{\kappa}(G_{n,r}) \geq \lambda_{\kappa}(H)$ with probability $1 - o(n^{-1})$, this completes the proof of Lemma~\ref{lem:higher_order_eigenvalue_tail_bound}. 
\end{document}