### 2020-08-09 notes/graph_regression2.tex
Added section on minimax testing with graph regression methods.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage{tikz}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Xsetiam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}
\newcommand{\Lap}{{\bf L}}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Minimax-optimal nonparametric regression with graph Laplacians}
\author{Alden Green}
\date{\today}
\maketitle

\section{Introduction}

In the random design nonparametric regression setup, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ on a domain $\Xset \subset \Rd$ and 
\begin{equation}
\label{eqn:random_design_regression}
Y_i = f_0(X_i) + \varepsilon_i
\end{equation}
with independent Gaussian noise $\varepsilon_i \sim N(0,1)$. The goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether or not $f_0 = 0$, i.e whether or not there is a signal present.

In graph based nonparametric regression, the idea is to perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$. For a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$. The weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ is an $n \times n$ matrix with entries
\begin{equation*}
\label{eqn:neighborhood_graph}
{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr),
\end{equation*}
and the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$.  Our star operator, the neighborhood graph Laplacian, can then be written as
\begin{equation}
\label{eqn:graph_Laplacian}
\Lap_{n,r} = \bf{D} - \bf{W}
\end{equation}
The two graph-based regression methods we will consider are \emph{Laplacian eigenmaps} and \emph{Laplacian smoothing}. Each use $\Lap_{n,r}$ to construct an estimator $\wh{f}$ of $f_0$. Letting $v_1(G_{n,r}),\ldots,v_n(G_{n,r})$ denote eigenvectors of $\Lap_{n,r}$, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ \citep{belkin2003} is given by 
\begin{equation}
\label{eqn:laplacian_eigenmaps}
\wh{f}_{\textrm{LE}} = \sum_{k = 1}^{\kappa} \Dotp{Y}{v_k(G_{n,r})}_n v_{k}(G_{n,r})
\end{equation}
where $\kappa \in \{1,...,n\}$ is a tuning parameter. (Here $\dotp{u}{v}_n = \frac{1}{n}\sum_{i = 1}^{n} u_i v_i$ for vectors $u,v \in \Reals^n$.) The Laplacian smoothing estimator $\wt{f}_{\LS}$ \citep{smola2003} is given by
\begin{equation}
\label{eqn:laplacian_smoothing}
\wt{f}_{\LS}= \min_{f \in \Reals^n} \Bigl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T L^s f \Bigr\}
\end{equation}
where $\rho > 0$ acts a tuning parameter. Then, assuming~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are reasonable estimators of $f_0$, the size of the statistics
\begin{equation}
T_{\LE} = \Bigl\|\wh{f}_{\LE}\Bigr\|_n^2 \label{eqn:laplacian_eigenmaps_test}
\end{equation}
and
\begin{equation}
T_{\LS} = \Bigl\|\wt{f}_{\LS}\Bigr\|_n^2 \label{eqn:laplacian_smoothing_test}
\end{equation}
can be used to test whether or not $f_0 = 0$.

\begin{itemize}
	\item \textbf{Motivate graph Laplacians.} There already exist a vast number of methods for both the nonparametric testing and, especially, the nonparametric estimation problems. Nevertheless, methods such as~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} deserve a place in the nonparametric regression toolbox. Their advantages include:
	\begin{itemize}
		\item Simplicity.
		\item Computational efficiency.
		\item Domain and density adaptivity.
		\item Practical success.
	\end{itemize}
	\item \textbf{Our contributions, short version.} For these reasons, there is a  substantial body of work assessing the statistical properties of graph Laplacian based methods. \textcolor{red}{(TODO): Include citations.} In this paper, we adopt a different perspective, and analyze~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} through the minimax lens. We show that for a variety of function classes $\mc{F}$, distributions $P$, and choices of domain $\mc{X}$ for which minimax rates are known, both~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are minimax optimal estimators (with loss measured in a suitable norm) and both~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} are minimax optimal test statistics. This lends additional theoretical support to the case for using graph Laplacian methods.
	\item \textbf{Summary}
\end{itemize}

\section{Problem Setup and Background}
\label{sec:problem_setup_and_background}

Before we get to our main results, we will review some of the relevant literature on nonparametric estimation and regression, and recall some relevant minimax analysis.  We will also briefly recall some (of the many) estimators and tests which achieve these rates. In particular, we will focus on the \emph{truncated series} and \emph{smoothing spline} methods. As we will see, Laplacian eigenmaps and Laplacian smoothing can be seen as graph-based analogues to these ``continuuum'' methods.

\begin{itemize}
	\item \textbf{Minimax estimation.} For given integers $s > 0$ and $p > 0$, the Sobolev space $W^{s,p}(P)$ consists of all functions $f \in \Leb^p(P)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d)$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ belongs to $\mathcal{L}^p(P)$. The Sobolev $\{s,p\}$ norm is then 
	\begin{equation*}
	\norm{f}_{W^{s,p}(P)}^2 = \sum_{\abs{\alpha} \leq s} \int_{\mathcal{X}} \abs{D^{\alpha}f}^p \,dP
	\end{equation*}
	and for a given $M > 0$, the corresponding ball is $W^{s,p}(P, M) = \set{f: \norm{f}_{W^{s,p}(P)} \leq M}$. In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space; we adopt the usual convention of writing $H^s(P) = W^{s,p}(P)$ and $H^s(P,M) = W^{s,2}(P,M)$. We will confine our attention to this special case hereafter. Let us also make the following assumption:
	\begin{enumerate}[(P1)]
		\item 
		\label{asmp:bounded_density}
		The distribution $P$ is defined on domain $\Xset = [0,1]^d$, and admits a density $p$ with respect to the uniform (Lebesgue) measure on $[0,1]^d$. The density $p$ is bounded away from $0$ and $\infty$,
		\begin{equation*}
		0 < p_{\min} < p(x) < p_{\max} < \infty
		\end{equation*}
		for all $x \in [0,1]^d$.
	\end{enumerate}
	The assumption~\ref{asmp:bounded_density} is a standard but significant assumption; we will return to discuss it further at \textcolor{red}{a later point}. Assuming~\ref{asmp:bounded_density}, the minimax estimation rate over the Sobolev ellipsoids $H^{s}(P,M)$ is
	\begin{equation*}
	\inf_{\wh{f}} \sup_{f \in H^s(P,M)} \Ebb\Bigl[\norm{\wh{f} - f}_{L^2(P)}^2\Bigr] \asymp n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d \geq 0$,}
	\end{equation*}
	where the infimum is over all measurable maps $\wh{f}(X_1,Y_1,\ldots,X_n,Y_n)$
	\item \textbf{Minimax hypothesis testing.}
	\item \textbf{Our contributions, long version.}
	\begin{itemize}
		\item \textbf{Minimax optimal estimation.}
		We show that with high probability,  $\norm{\wh{f}_{\LE} - f_0}_{n}^2 \lesssim n^{-2s/(2s + d)}$ whenever $f_0 \in C^s(\mc{X})$ for any $s \leq 1$, or $f_0 \in H^s(P)$ for $s = 1$. The same conclusions hold with respect to $\wt{f}_{\LS}$ when $d \leq 4$. 
		\item \textbf{Minimax optimal testing.}
		We show that with high probability, a level-$\alpha$ test constructed using $T_{\LE}$ has non-trivial power whenever $f_0 \in C^s(\mc{X})$ for any $s \leq 1$---or $f_0 \in H^s(P)$ for $s = 1$---additionally satisfies $\norm{f_0}_{\Leb^2(P)} \gtrsim \max\{n^{-4s/(4s + d)},n^{-1/4}\}$. \textcolor{red}{The same conclusions hold with respect to $\wh{f}_{LS}$.}
		\item \textbf{Manifold adaptivity.}
		If $\mc{X} \subset \Rd$ is a submanifold of dimension $m < d$, each of the aforementioned rates hold with $d$ replaced by $m$.
		\item \textbf{Taking advantage of higher-order smoothness conditions.} Under appropriate boundary conditions on the function classes $C^s$ and $H^s$ and additional smoothness conditions on $P$, when $s > 1$  each of the aforementioned conclusions hold for certain values of $d$.
	\end{itemize}
	\item \textbf{Notation.}
\end{itemize}

\section{Minimax optimal graph Laplacian methods}
\label{sec:minimax_optimal_graph_Laplacian_methods}

Let us start by assuming $f_0 \in H^1(P,M)$. In this section, we will establish that the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and the resulting test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test}, achieve the standard minimax optimal estimation and hypothesis testing rates, respectively.

\subsection{Minimax optimal estimation}

Under appropriate conditions, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves minimax rates over all $f_0 \in H^1(P,M)$ with high probability. The conditions we impose will be on the neighborhood graph $G_{n,r}$. As the construction of the graph is in the user's control, these assumptions are therefore mild.
\begin{enumerate}[label=(K\arabic*)]
	\item 
	\label{asmp:kernel_form}
	The kernel function $K(z) = \1\Bigl\{\norm{z}_2/r \leq 1 \Bigr\}$.
	\item 
	\label{asmp:kernel_radius_estimation}
	The radii $r = r(n)$ satisfies $r(n) \to 0$ as $n \to \infty$, and additionally
	\begin{equation*}
	n^{-1/d}\leq r(n) \leq M^{2/(2 + d)}n^{-1/(2 + d)}
	\end{equation*}
	for all $n$ sufficiently large.
\end{enumerate}

We use~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_estimation} to ensure that the approximation error (bias) of $\wh{f}_{\LE}$ is of the correct order. On the other hand, bounding the estimation error of $\wh{f}_{\LE}$ -- both in expectation and in probability-- is straightforward. Together, bounds on these two sources of error imply Theorem~\ref{thm:laplacian_eigenmaps_estimation1}.
\begin{theorem}
	\label{thm:laplacian_eigenmaps_estimation1}
	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}. Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_estimation}. Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as given by~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
	\begin{equation*}
	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4 + d)} n^{-2/(2 + d)}
	\end{equation*}
	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
\end{theorem}

\begin{itemize}
	\item Discuss the choice of radius $r$ imposed by~\ref{asmp:kernel_radius_estimation}.
	\item Comment on the distinction between error rates as measured by the norm in $L^2(P_n)$ vs. $L^2(P)$.
	\item \textbf{Proof techniques}: to prove results we rely on both (a) classical bias-variance decomposition of orthogonal series estimators (see e.g. \textcolor{red}{Tsybakov}), and (b) pointwise and variational results on neighborhood graph Laplacian semi-norms and eigenvalues (see e.g. \textcolor{red}{Belkin04, Belkin08, GarciaTrillos18, Calder19})
\end{itemize}

Theorem~\ref{thm:laplacian_smoothing_estimation1} is analogous to Theorem~\ref{thm:laplacian_eigenmaps_estimation1}, except it deals with the Laplacian smoothing estimator $\wt{f}_{\LS}$ rather than the Laplacian eigenmaps estimator $\wh{f}_{\LE}$. The key difference is that it holds only when the dimension $d < 4$.  
\begin{theorem}
	\label{thm:laplacian_smoothing_estimation1}
	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}. Suppose that $f_0 \in H^1(P,M)$, that $P$ satisfies~\ref{asmp:bounded_density}, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_estimation}. Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as given by~\eqref{eqn:laplacian_eigenmaps}, when we set $\kappa = M^{2/(2 + d)}n^{d/(2 + d)}$ the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
	\begin{equation*}
	\Bigl\|\wh{f}_{\LS} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4 + d)} n^{-2/(2 + d)}
	\end{equation*}
	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
\end{theorem}

\begin{itemize}
	\item \textbf{Proof techniques}: Because the solution to~\eqref{eqn:laplacian_smoothing} has an analytic form, we can rely on a bias-variance decomposition---as opposed to empirical process theory---to derive~Theorem~\ref{thm:laplacian_smoothing_estimation1}.
	\item The restriction in Theorem~\ref{thm:laplacian_smoothing_estimation1} that the dimension $d$ be less than $4$ is unsurprising. In a related context (estimation over a $d$-dimensional lattice graph), \cite{sadhanala2016} require the same restriction. The underlying phenomenon is similar to that elucidated by the seminal papers~\cite{birge1993} and~\cite{birge1998}: when the dimension $d$ is at least $4$, the parameter space is too large, and (penalized) least squares approaches are no longer minimax optimal.
\end{itemize}

\subsection{Minimax optimal hypothesis testing}

Under appropriate conditions, the Laplacian eigenmaps test statistic $T_{\LE}$ has non-trivial power over all $f_0 \in H^1(P,M)$ that are separated from $0$ in $\Leb^2(P)$ norm by at least the minimax critical radius. These conditions include~\ref{asmp:kernel_form} and~\ref{asmp:smooth_density}, but not~\ref{asmp:kernel_radius_estimation}. Instead, we will require the following scaling of the neighborhood graph radius.
\begin{enumerate}[label=(K\arabic*)]
	\setcounter{enumi}{2}
	\item 
	\label{asmp:kernel_radius_testing}
	The radii $r = r(n)$ satisfy $r(n) \to 0$ as $n \to \infty$, and additionally
	\begin{equation*}
	n^{-1/d} \leq r(n) \leq n^{-2/(4 + d)}
	\end{equation*}
	for all $n$ sufficiently large.
\end{enumerate}

\begin{theorem}
	\label{thm:laplacian_eigenmaps_testing1}
	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}. Suppose that $f_0 \in H^1(P,M)$, that
	$P$ satisfies~\ref{asmp:bounded_density}, and that the neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_testing}.
	Then for the Laplacian eigenmaps test statistic $T_{\LE}$ as in~\eqref{eqn:laplacian_eigenmaps_test} with $\kappa = n^{2d/(4 + d)} M^{4d/(4 + d)}$, the following statement holds true for all $n$ sufficiently large: there exist constants $C_1$ and $C_2$ which do not depend on $f_0$ such that if
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_testing}
	\norm{f_0}_{\Leb^2(P)}^2 \geq \frac{C_1}{\delta^2} M^{2d/(4 + d)} n^{-4/(4 + d)}
	\end{equation}
	then
	\begin{equation*}
	\Pbb \biggl(T_{\LE} \geq \frac{\kappa}{n} + \frac{1}{\delta}\sqrt{\frac{2\kappa}{n^2}}\biggr) \geq 1 - C_2\Bigl(\delta + r^{-d} \exp\{- c n r^d\}\Bigr).
	\end{equation*}
\end{theorem}
\begin{itemize}
	\item Explain why Theorem~\ref{thm:laplacian_eigenmaps_testing1} implies minimax optimality.
\end{itemize}

The estimation and testing rates derived in Section~\ref{sec:minimax_optimal_graph_Laplacian_methods} exhibit the curse of dimensionality: as $d$ gets large, inference becomes hard. As covered in Section~\ref{sec:problem_setup_and_background}, it is well known that by making additional structural assumptions---either on the regression function $f_0$, the distribution $P$, or both---one can derive improved minimax rates. In the following two sections, we show that under such assumptions, the estimators~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing}, and test statistics~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} can in certain situations obtain the faster (known) minimax rates in the nonparametric estimation and testing problems.

\section{Higher-order smoothness classes}
\label{sec:higher_order_smoothness_classes}
In this section, we will assume that $f_0 \in H^s(P,M)$ for some integer $s > 1$. In other words, we will assume that $f_0$ has at least two smooth derivatives.

\subsection{Minimax optimal estimation}
Under the following three conditions, the estimation error of the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ achieves the minimax optimal rate for any $f_0 \in H^s(P,M)$.
\begin{enumerate}[label=(K\arabic*)]
	\setcounter{enumi}{2}
	\item 
	\label{asmp:kernel_radius_higher_order}
	\textbf{Neighborhood graph radius}:
	The sequence of radii $r = r(n)$ satisfy $r(n) \to 0$ as $n \to \infty$, and additionally
	\begin{equation*}
	n^{-1/(2(s - 1) + d)}\leq r(n) \leq M^{2/(2s + d)}n^{-1/(2s + d)}
	\end{equation*}
	for all $n$ sufficiently large.
\end{enumerate}
\begin{enumerate}[label=(F\arabic*)]
	\setcounter{enumi}{0}
	\item 
	\label{asmp:boundary} \textbf{Boundary condition}; The regression function $f_0$ is zero-trace. In other words, $f_0 \in H_0^s(P,M)$. 
\end{enumerate}
\begin{enumerate}[label=(P\arabic*)]
	\setcounter{enumi}{1}
	\item 
	\label{asmp:smooth_density}\textbf{$\beta$-Holder density}: The design distribution $P$ has a density $p \in C^{\beta}([0,1]^d;p_{\max})$ for some $\beta \geq 0$. 
\end{enumerate}

We note that the assumption~\ref{asmp:kernel_radius_higher_order} is rather mild, since the construction of the neighborhood graph $G_{n,r}$ is in the user's control. whereas the latter two assumptions are more stringent. That being said, some restrictions on the boundary behavior of the regression function $f_0$ are typical in the analysis of orthogonal series estimators (see e.g. \textcolor{red}{Tsybakov}), as are assumptions on the smoothness of the density $p$ under a random design setup (see e.g. \textcolor{red}{Mukherjee16}).

Together, conditions~\ref{asmp:boundary} and~\ref{asmp:smooth_density} are used to ensure that the approximation error (bias) of $\wh{f}_{\LE}$ is of the correct order. Bounds on the estimation error of $\wh{f}_{\LE}$ -- both in expectation and in probability-- are straightforward. Together, bounds on these two sources of error imply Theorem~\ref{thm:laplacian_eigenmaps_estimation}.
\begin{theorem}
	\label{thm:laplacian_eigenmaps_estimation}
	Let $(X_1,Y_1),\ldots,(X_n,Y_n)$ be random samples drawn according to the model~\eqref{eqn:random_design_regression}. Suppose that $f_0 \in H_0^s(P,M)$, that $P$ satisfies~\ref{asmp:smooth_density} with $\beta = s - 1$, and that neighborhood graph $G_{n,r}$ is computed with kernel $K$ and radius $r$ which satisfy~\ref{asmp:kernel_form} and~\ref{asmp:kernel_radius_higher_order}. Then, for the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ as in~\eqref{eqn:laplacian_eigenmaps}, when $\kappa = M^{2/(2s + d)}n^{d/(2s + d)}$, the following statement holds: for all $n$ sufficiently large, there exist constants $c$ and $C$ such that
	\begin{equation*}
	\Bigl\|\wh{f}_{\LE} - f_0\Bigr\|_n^2 \leq \frac{C}{\delta} M^{2d/(4s + d)} n^{-2s/(2s + d)}
	\end{equation*}
	with probability at least $1 - \delta -  Cr^{-d}\exp\{-cnr^d\} - \exp(-\kappa)$.
\end{theorem}

\begin{enumerate}
	\item \textcolor{red}{(TODO)}: Comment on the difference between this and minimax optimality. Cite others \textcolor{red}{(Wainwright, Liu)} who do the same thing.
	\item \textcolor{red}{(TODO)}: Comment on why we insist that $f_0$ satisfy a boundary condition.
	\item \textcolor{red}{(TODO)}: Comment a bit on the proof strategy; i.e., how we don't convergence properties of the eigenvectors $v_k$ towards eigenfunctions of the weighted Laplace-Beltrami operator, why we think that's advantageous, and how it might be limiting.
\end{enumerate}

\section{Manifold Assumption}
\label{sec:manifold_assumption}

\section{Discussion}

\begin{itemize}
	\item Comment on the different types of relationships between $s$ and $d$. (1) Those due to fundamental lower bounds on testing; (2) those due to largeness of the parameter space for Laplacian smoothing, and (3) those which (potentially) could be improved by different proof techniques.
\end{itemize}

\section{Simulations}
\begin{itemize}
	\item \textbf{Estimation MSE 1.} Plot $L^2(P_n)$ error as a function of $n$ and choice $f_0^{(n)}$ when $P$ is (close to) uniform, showing that Laplacian methods are competitive with their non-graph-based brethren.
	\item \textbf{Estimation fitted values.} Show fits for two regression problems, one where the distribution $P$ is assumed to be close to uniform, one where $P$ is dramatically non-uniform.
	\item \textbf{Estimation MSE 2.} Plot $L^2(P_n)$ error as a function of the distribution $P$, as we vary $P$ from being uniform to being very non-uniform.

	\item \textbf{Testing power.} 
\end{itemize}

\begin{figure}
	\centering
	\includegraphics[width = .48\textwidth]{plots/graph_regression/mean_squared_errors/sobolev_MSE_by_density}
	\caption{}
	\label{fig:piecewise_cosine}
\end{figure}

\clearpage

\appendix

Many of our results follow the same general two-part proof strategy. First, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to functionals of the graph $G$. We then analyze the behavior of these functionals with respect to the particular neighborhood graph $G_{n,r}$ and giving high probability (upper or lower )bounds on these functionals. It is in this second step where we invoke our various assumptions on the distribution $P$ and regression function $f_0$.

\section{Fixed graph error bounds}
\label{sec:fixed_graph_error_bounds}

Suppose we observe a graph $G = ([n],E)$, and responses
\begin{equation}
\label{eqn:fixed_graph_regression_model}
Y_i = f_{0,i} + \varepsilon_i
\end{equation}
where $f_0 \in \Reals^n$ and $\varepsilon_i$ are independent $N(0,1)$ noise. The Laplacian eigenmaps estimator of $f_0$ on $G$ is then
\begin{equation}
\label{eqn:le_G}
\wh{f}_{\LE}(G) := \sum_{k = 1}^{\kappa} \biggl(\frac{1}{n}\sum_{i = 1}^{n} Y_i v_{k,i}(G) \biggr) v_k(G),
\end{equation}
the Laplacian smoothing estimator of $f_0$ on $G$ is
\begin{equation}
\label{eqn:ls_G}
\wh{f}_{\LS}(G) = \argmin_{f \in \Reals^n} \biggl\{ \sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_G^{s}  f \biggr\} = (\rho \Lap_G^s + I)^{-1}Y
\end{equation}
and their associated test statistics are
\begin{equation}
\label{eqn:le_ts_G}
T_{\LE}(G) := \Bigl\|\wh{f}_{\LE}(G)\Bigr\|_n^2
\end{equation}
and
\begin{equation}
\label{eqn:ls_ts_G}
T_{\LS}(G) := \Bigl\|\wt{f}_{\LS}(G)\Bigr\|_n^2
\end{equation}
We note that~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing_test} specialize~\eqref{eqn:le_G}-\eqref{eqn:ls_ts_G} to the case where $G$ = $G_{n,r}$.

\subsection{Estimation: Laplacian eigenmaps on a fixed graph}

In Lemma~\ref{lem:le_fixed_graph_estimation}, we upper bound the mean squared error of $\wh{f}_{\LE}(G)$. Our bound will be stated as a function of $f_0^T \Lap_G^s f_0$--a measure of the smoothness the signal $f_0$ displays over the graph $G$--as well as the $\kappa$th eigenvalue $\lambda_{\kappa}(G)$.

\begin{lemma}
	\label{lem:le_fixed_graph_estimation}
	Let $G$ be a fixed (non-random) graph. Suppose we observe $G$ and responses $Y$ according to the model~\eqref{eqn:fixed_graph_regression_model}. Then for any $1 \leq \kappa \leq n$, 
	\begin{equation}
	\label{eqn:le_fixed_graph_estimation}
	\Ebb\Bigl[\norm{\wh{f}_{\LE}(G) - f_0}_n^2\Bigr] \leq 2\biggl(\frac{f_0^T \Lap_{G}^s f_0}{n \bigl(\lambda_{\kappa}(G)\bigr)^s} + \frac{\kappa}{n}\biggr).
	\end{equation}
	Additionally, 
	\begin{equation}
	\label{eqn:le_fixed_graph_estimation_1}
	\norm{\wh{f}_{\LE}(G) - f_0}_n^2 \leq 2\biggl(\frac{f_0^T \Lap_{G}^s f_0}{n \bigl(\lambda_{\kappa}(G)\bigr)^s} + \frac{4\kappa}{n}\biggr)
	\end{equation}
	with probability at least $1 - \exp(-\kappa)$.
\end{lemma}
The two terms on the right hand side of~\eqref{eqn:le_fixed_graph_estimation} and~\eqref{eqn:le_fixed_graph_estimation_1} represent the squared bias and variance of the estimator $\wh{f}_{\LE}(G)$. Note that the squared bias--- and therefore the overall mean squared error---depends on the quadratic form $f_0^T \Lap_{G}^s f_0$ and the eigenvalue $\lambda_{\kappa}(G)$, whereas the variance depends instead on the dimension $\kappa$ of the subspace $V_{\kappa} = \textrm{span}\{v_1,\ldots,v_{\kappa}\}$ to which $\wh{f}_{\LE}(G)$ belongs. In this respect, the error bound on the Laplacian eigenmaps estimator resembles the error bound on the \textcolor{red}{orthogonal series} estimator (see e.g. \textcolor{red}{Tsybakov}).

\paragraph{Proof of Lemma~\ref{lem:le_fixed_graph_estimation}:}

\textcolor{red}{(TODO)}

\subsection{Estimation: Laplacian smoothing on a fixed graph.}

\begin{lemma}
	\label{lem:ls_fixed_graph_estimation}
	Suppose we observe a graph $G$ and responses $Y$ according to the model~\eqref{eqn:fixed_graph_regression_model}. Then for any $\rho > 0$ and positive integer $s$,
	\begin{equation}
	\label{eqn:ls_fixed_graph_estimation}
	\Ebb\Bigl[\norm{\wt{f}_{\LS}(G) - f_0}_n^2\Bigr] \leq 2\biggl(\frac{\rho \cdot f_0^T \Lap_{G}^s f_0}{n} + \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{(\rho \cdot \lambda_{k}(G)^s + 1)^2}\biggr)
	\end{equation}
	Additionally,
	\begin{equation}
	\label{eqn:ls_fixed_graph_estimation_prob}
	\Pbb\Biggl[\norm{\wt{f}_{\LS}(G) - f_0}_n^2 \leq 2\biggl(\frac{\rho \cdot f_0^T \Lap_{G}^s f_0}{n} + \frac{4}{n}\sum_{k = 1}^{n} \frac{1}{(\rho \cdot \lambda_{k}(G)^s + 1)^2}\biggr)\Biggr] \geq 1 - \exp\biggl(-\sum_{k = 1}^{n}\frac{1}{(\rho \lambda_k(G)^s + 1)^2}\biggr)
	\end{equation}
\end{lemma}

\paragraph{Proof of Lemma~\ref{lem:ls_fixed_graph_estimation}:}

\textcolor{red}{(TODO)}

\subsection{Testing: Laplacian eigenmaps on a fixed graph.}

\begin{lemma}
	\label{lem:fixed_graph_testing}
	Suppose we observe a graph $G$ and responses $Y$ according to the model~\eqref{eqn:fixed_graph_regression_model}. Then for any $1 \leq \kappa \leq n$, each of the following holds for any $b \geq 1$,
	\begin{enumerate}
		\item \textbf{Type I error:} Under the null hypothesis $f_0 = 0$,
		\begin{equation}
		\label{eqn:graph_spectral_type_I_error}
		\Pbb\biggl(T_{\LE}(G) \geq \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}\biggr) \leq \frac{2}{b^2}.
		\end{equation}
		\item \textbf{Type II error:} Assume that
		\begin{equation}
		\label{eqn:fixed_graph_testing_critical_radius_1}
		\Bigl\|\sum_{k = 1}^{\kappa} \dotp{f_0}{n}_n v_k\Bigr\|_n^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}}.
		\end{equation}
		Then,
		\begin{equation}
		\label{eqn:graph_spectral_type_II_error}
		\Pbb\biggl(T_{\LE}(G) \leq \frac{\kappa}{n} + b\sqrt{\frac{2\kappa}{n^2}}\biggr) \leq \frac{2}{b^2} + \frac{8}{b\sqrt{2\kappa}}.
		\end{equation}
		In particular if
		\begin{equation}
		\label{eqn:fixed_graph_testing_critical_radius}
		\frac{1}{n} \sum_{i = 1}^{n} f_{0,i}^2 \geq 2b\sqrt{\frac{2\kappa}{n^2}} + \frac{f_0^T L_G^s f_0}{n\lambda_{\kappa}(G)^s}
		\end{equation}
		then~\eqref{eqn:fixed_graph_testing_critical_radius_1} and thus~\eqref{eqn:graph_spectral_type_II_error} follow.
	\end{enumerate}
\end{lemma}

\subsection{Testing: Laplacian smoothing on a fixed graph.}

\clearpage
To prove Theorems~\ref{thm:laplacian_eigenmaps_estimation},\textcolor{red}{...}, we would like to apply Lemmas~\ref{lem:fixed_graph_estimation},\textcolor{red}{...} in the random design context, i.e. to the random geometric graph $G_{n,r}$. To do so, we must establish the following (high probability) bounds: (i) an upper bound on the graph Sobolev seminorm of the regression function $f_0 \Lap_{n,r}^s f_0$, (ii) and lower boud on the Laplacian eigenvalue $\lambda_{\kappa}(G_{n,r})$ as a function of $\kappa$, and (iii) a lower bound on the empirical norm $\norm{f}_{n}$ as a function of $\norm{f}_{L^2(P)}$.

\section{Bounds on the graph Sobolev semi-norm}
\label{sec:graph_sobolev_seminorm}
We first bound $f_0^T \Lap_{n,r}^s f_0$ in expectation.
\begin{lemma}
	\label{lem:roughness_functional_expectation_sobolev}
	Let $\Xset$ be a Lipschitz domain and let $s \geq 1$ be an integer. Suppose that \textcolor{red}{$f \in H_0^{s}(P,M)$}, and that $P$ admits a density $p \in C^{s-1}(\Xset,p_{\max})$ for some constant $p_{\max} < \infty$. Let $\{r(n)\}$ be a sequence of connectivity radii such that $n^{-1/(2(s - 1) + d)} \leq r(n)$ for all $n$ sufficiently large, and $r(n) \to 0$ as $n \to \infty$. Then \textcolor{red}{for any $2$nd-order kernel $K$}, there exists a constant $c$ depending only on $p_{\max}, K_{\max}$, $s$, $d$ and $\Xset$ such that the resulting neighborhood graph Laplacian satisfies
	\begin{equation}
	\label{eqn:roughness_functional_expectation_sobolev}
	\Ebb\bigl[f^T L_{n,r}^s f\bigr] \leq c \cdot n \cdot (nr^{d+2})^s \cdot M^2
	\end{equation}
	for all $n$ sufficiently large.
\end{lemma}

\section{Bounds on graph eigenvalues}
\label{sec:graph_eigenvalues}
Lemma~\ref{lem:neighborhood_eigenvalue} establishes a scaling rate which holds for all eigenvalues $\lambda_{k}(G_{n,r})$, assuming appropriate regularity conditions on $P$.
\begin{lemma}
	\label{lem:neighborhood_eigenvalue}
	Suppose the distribution $P$ satisfies~\ref{asmp:bounded_density}. Then there exists constants $c$ and $C$ which depend only on $p_{\min}, p_{\max}$ and $d$, such that with probability at least $1 - C r^{-d} \exp\bigl\{- c n r^d\bigr\}$, for all $2 \leq k \leq n$
	\begin{equation}
	\label{eqn:neighborhood_eigenvalue}
	\lambda_k(G_{n,r}) \geq c \cdot \min\Bigl\{nr^{d + 2}k^{2/d}, nr^d\Bigr\}.
	\end{equation}	
\end{lemma}
The first term on the right hand side of~\eqref{eqn:neighborhood_eigenvalue} is the scaling of the eigenvalues we desire. The second term reflects the fact that the maximum eigenvalue of the Laplacian can never be more than twice the maximum degree of the graph. It is a bit of simple algebra to derive the conditions under which the first term is the smaller of the two when $k = \kappa = M^{2/(2s + d)} n^{d/(2s + d)}$.
\begin{corollary}
	\label{cor:neighborhood_eigenvalue}
	Let $\kappa = M^{2/(2s + d)} n^{d/(2s + d)}$. For any $r \leq M^{-2/(d(2s +d))} n^{-1/(2s + d)}$, 
	\begin{equation}
	\label{eqn:neighborhood_eigenvalue_1}
	\lambda_{\kappa}(G_{n,r}) \geq c \cdot nr^{d+2}{\kappa}^{2/d}
	\end{equation}
	with probability at least $1 - C r^{-d} \exp\{- c n r^d\}$. The constants $c$ and $C$ are the same as in Lemma~\ref{lem:neighborhood_eigenvalue}.
\end{corollary}

\section{Proofs of Theorems}
\label{sec:proofs_theorems}
With the results of Sections~\ref{sec:fixed_graph_error_bounds}-\ref{sec:graph_eigenvalues} established, the proofs of our main theorems are quite straightforward.

\subsection{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation}}
\label{subsec:laplacian_eigenmaps_estimation_pf}

\paragraph{\eqref{eqn:fixed_graph_estimation_sobolev_rate_2} holds with high probability.}

Since $\Lap_{n,r}$ is a positive semi-definite matrix, $f_0^T \Lap_{{n,r}}^s f_0$ must be non-negative. We may therefore apply Markov's inequality and conclude
\begin{equation}
\label{eqn:roughness_functional_probability_sobolev}
f_0^T \Lap_{{n,r}}^s f_0 \leq  \frac{1}{\delta} \cdot c \cdot n \cdot (nr^{d+2})^s \cdot M^2
\end{equation}
with probability at least $1 - \delta$.

\paragraph{\eqref{eqn:fixed_graph_estimation_sobolev_rate_1} holds with high probability.}

\paragraph{Putting the pieces together.}

Let $n$ be sufficiently large so that $n^{-1/(2(s - 1)+ d)} \leq r(n) \leq M^{-2/(d(2s + d))} n^{-1/(2s + d)}$. Then by~\eqref{eqn:roughness_functional_probability_sobolev} and~\eqref{eqn:neighborhood_eigenvalue_1}, there exists a set $\mathcal{E}(f_0) \subseteq \Xset^n$ of probability at least $1 - \delta - Cr^{-d}\exp\{-cnr^d\}$ such that whenever ${\bf X} \in \mathcal{E}(f_0)$,
\begin{equation*}
f_0^T \Lap_{n,r}^s f_0 \leq C \cdot n \cdot C_n^s \cdot M^2
\end{equation*}
and
\begin{equation*}
\lambda_{\kappa}(G_{n,r}) \geq c \cdot nr^{d+2}\kappa^{2/d}
\end{equation*}
Theorem~\ref{thm:laplacian_eigenmaps_estimation} then follows by Corollary~\ref{cor:fixed_graph_estimation_sobolev_rate}.

\section{Concentration Inequalities}

\bibliographystyle{plainnat}
\bibliography{../graph_testing_bibliography}

\end{document}