### 2020-08-14 notes/weekly_notes/8.12.20.NotesForWeek.tex
8.12.20.NotesForWeek. Added fixed graph testing results for Laplacian smoothing test.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage{tikz}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\Xsetotp}[2]{\langle #1, #2 \rangle}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Xsetiam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}
\newcommand{\Lap}{L}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}
\newcommand{\Id}{I}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 8/12/2020 - 8/19/2020}
\author{Alden Green}
\date{\today}
\maketitle

\section{Fixed graph regression}

Suppose we observe a graph $G = \bigl([n],W\bigr)$, and responses
\begin{equation}
\label{eqn:fixed_graph_regression_model}
Y_i = \theta_{0,i} + \varepsilon_i
\end{equation}
with signal vector $\theta_0 = (\theta_{0,1},\ldots,\theta_{0,n}) \in \Reals^n$, and noise vector $\varepsilon = (\varepsilon_1,\ldots,\varepsilon_n) \sim N(0,\Id_{n \times n})$. Letting $L$ denote the Laplacian matrix of graph $G$, the \emph{Laplacian smoothing} estimator $\wt{\theta}(G) \in \Reals^n$ is given by
\begin{equation}
\label{eqn:ls_G}
\wt{\theta}(G) := \argmin_{\theta \in \Reals^n} \biggl\{ \sum_{i = 1}^{n}(Y_i - \theta_i)^2 + \rho \cdot \theta^T \Lap_G^{s}  \theta \biggr\} = (\rho \Lap_G^s + I)^{-1}Y.
\end{equation}
Now suppose we wish to test 
\begin{equation}
\mathbf{H}_0: \theta_0 = 0 ~~\textrm{vs.}~~ \mathbf{H}_a: \theta_0 \neq 0
\end{equation}

 A natural candidate is the quadratic form
\begin{equation}
\label{eqn:ls_ts_G}
\wt{T}(G) := \frac{1}{n}\sum_{i = 1}^{n} \bigl(\wt{\theta}_i(G)\bigr)^2
\end{equation}
In Lemma~\ref{lem:ls_fixed_graph_testing}, we provide concentration bounds on the test statistic $\wt{T}(G)$, under both the null and alternative hypotheses. These statements are a function of the spectral decomposition $\Lap_G = \sum_{k = 1}^{n} \lambda_k(G) \cdot v_k(G) v_k(G)^T$. 
\begin{lemma}
	\label{lem:ls_fixed_graph_testing}
	Define the threshold $\wt{t}_b$ to be
	\begin{equation*}
	\wt{t}_b := \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k^s + 1\bigr)^2} + \frac{2b}{n}\sqrt{\sum_{k = 1}^{n} \frac{1}{\bigl(\rho \lambda_k^s + 1\bigr)^4}}
	\end{equation*}
	Then,
	\begin{itemize}
		\item \textbf{Type I error.}
		\begin{equation}
		\label{eqn:ls_fixed_graph_testing_typeI}
		\Pbb_0\Bigl(\wt{T}(G) > \wt{t}_b\Bigr) \leq \frac{1}{b^2}
		\end{equation}
		\item \textbf{Type II error.} If
		\begin{equation}
		\label{eqn:ls_fixed_graph_testing_critical_radius}
		\frac{1}{n}\norm{\theta_0}_2^2 \geq \frac{2 \rho}{n} \bigl(\theta_0^T \Lap^s \theta_0\bigr) + \frac{4b}{n} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k^s + 1)^4} \Biggr)^{1/2}
		\end{equation}
		then
		\begin{equation}
		\label{eqn:ls_fixed_graph_testing_typeII}
		\Pbb_{\theta_0}\Bigl(\wt{T}(G) \leq \wt{t}_b\Bigr) \leq \frac{4}{b^2} + \frac{8}{b} \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k^s + 1)^4} \Biggr)^{-1/2}
		\end{equation}
	\end{itemize}
\end{lemma}

\subsection{Notation}
\begin{itemize}
	\item When the graph $G$ is obvious from context, we may drop the notational dependence on $G$ and simply write the Laplacian of $G$ as $\Lap$, and the $k$th eigenvalue of the Laplacian as $\lambda_k$.
\end{itemize}

\section{Proofs}

\subsection{Proof of Lemma~\ref{lem:ls_fixed_graph_testing}}

Let $\wt{S} := (\rho \Lap^s + \Id)^{-1}$. The matrix $\wt{S} \in \Reals^{n \times n}$ is symmetric and positive semidefinite, and $\wt{T}(G) = \frac{1}{n}Y^T \wt{S}^2 Y$. The desired result thus follows from Lemma~\ref{lem:linear_smoother_fixed_graph_testing}. To see that the conditions of this Lemma are satisfied, we first note that since
\begin{equation*}
\lambda_k(S) = \frac{1}{(\rho\lambda_k^s + 1)}
\end{equation*}
and $\rho, \lambda_k > 0$, it is evident that $\lambda_{\max}(\wt{S}) \leq 1$.  Then, by assumption~\eqref{eqn:ls_fixed_graph_testing_critical_radius}
\begin{equation*}
\theta_0^{T} \wt{S}^2 \theta_0 = \norm{\theta_0}_2^2 - \theta_0^T(I - \wt{S}^2)\theta_0 \geq 2 \rho (\theta_0^T \Lap^s \theta_0) + \theta_0^T(I - \wt{S}^2)\theta_0 + 4b \Biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k^s + 1)^4} \Biggr)^{-1/2},
\end{equation*}
and along with the following calculations,
\begin{equation*}
\begin{aligned}
\theta_0^T \Bigl(\Id - \wt{S}^2\Bigr) \theta_0  & \overset{(i)}{=} \theta_0^T L^{s/2} L^{-s/2}\Bigl(\Id - \wt{S}^2\Bigr) L^{-s/2} L^{s/2} \theta_0 \\ 
& \leq \theta_0^T L^{s} \theta_0 \cdot  \lambda_{\max}\biggl(L^{-s/2}\Bigl(\Id - \wt{S}^2\Bigr) L^{-s/2}\biggr) \\ 
& \overset{(ii)}{=}  \theta_0^T L^{s} \theta_0 \cdot \max_{k} \biggl\{ \frac{1}{\lambda_k^s} \Bigl(1 - \frac{1}{(\rho \lambda_k^s + 1)^2}\Bigr) \biggr\} \\
& \overset{(iii)}{\leq} \theta_0^T L^{s} \theta_0 \cdot 2\rho,
\end{aligned}
\end{equation*}
we have that
\begin{equation*}
\theta_0^{T} \wt{S}^2 \theta_0 \geq 2b \biggl(\sum_{k = 1}^{n} \frac{1}{(\rho\lambda_k^s + 1)^4} \biggr)^{-1/2}.
\end{equation*} 
In other words condition~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} in Lemma~\ref{lem:linear_smoother_fixed_graph_testing} is met, and applying that Lemma completes the proof.

(In the previous derivation: in $(i)$ when we write $\Lap^{-s/2}$ we are referring to the pseudoinverse of $\Lap$, since the Laplacian always has at least one eigenvalue equal to $0$; in $(ii)$ the maximum is over all indices $k$ such that the eigenvalue $\lambda_k$ is strictly positive; and $(iii)$ follows from the basic algebraic identity $1 - 1/(1 + \rho x)^2 \leq 2 \rho x$ for any $x, \rho > 0$.

\section{Technical Lemmas}

\subsection{Type I and Type II error of quadratic forms.}
Let $S \in \Reals^{n \times n}$ be a square symmetric matrix. The quadratic form
\begin{equation}
T = Y^T S^2 Y
\end{equation}
can be used as a test statistic to distinguish $\mathbf{H}_0$ from $\mathbf{H}_a$. In Lemma~\ref{lem:linear_smoother_fixed_graph_testing}, we establish conditions under which a test based on $T$ has small Type I and Type II error.
\begin{lemma}
	\label{lem:linear_smoother_fixed_graph_testing}
	Define the threshold $t_b$ to be 
	\begin{equation}
	t_{b} := \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^2 + 2b \sqrt{\sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^4}
	\end{equation}
	Suppose the eigenvalues $0 \leq \lambda_{\min}(S) \leq \lambda_{\max}(S) \leq 1$. Then,
	\begin{itemize}
		\item \textbf{Type I error.}
		\begin{equation}
		\label{eqn:linear_smoother_fixed_graph_testing_typeI}
		\Pbb_0\bigl(T > t_b\bigr) \leq \frac{1}{b^2}
		\end{equation}
		\item \textbf{Type II error.} Assuming that
		\begin{equation}
		\label{eqn:linear_smoother_fixed_graph_testing_critical_radius}
		\theta_0^T S^2 \theta_0 \geq 4b \sqrt{\sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^4}
		\end{equation}
		then
		\begin{equation}
		\label{eqn:linear_smoother_fixed_graph_testing_typeII}
		\Pbb_{\theta_0}\bigl(T \leq t_b\bigr) \leq \frac{4}{b^2} + \frac{8}{b} \Biggl(\sum_{k = 1}^{n}\bigl(\lambda_k(S)\bigr)^4\Biggr)^{-1/2}
		\end{equation}
	\end{itemize}
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:linear_smoother_fixed_graph_testing}]
	We compute the mean and variance of $T$ as a function of $\theta_0$, then apply Chebyshev's inequality.
	\paragraph{Mean.}
	Writing $Y = \theta_0 + \varepsilon$, we make use of the eigendecomposition $S = \sum_{k = 1}^{n} \lambda_k(S) \cdot v_k(S) v_k(S)^T$---where in this case we fix the eigenvectors $v_1(S),\ldots,v_n(S)$ to be unit-norm---and obtain
	\begin{equation}
	\label{pf:linear_smoother_fixed_graph_testing1}
	\begin{aligned}
	T & = \theta_0^T S^2 \theta_0 + 2 \theta_0^T S^2 \varepsilon + \varepsilon^T S^2 \varepsilon \\
	& = \theta_0^T S^2 \theta_0 + 2 \theta_0^T S^2 \varepsilon + \sum_{k = 1}^{n}  \bigl(\lambda_k(S)\bigr)^2 (\varepsilon^T v_k(S))^2 \\
	& = \theta_0^T S^2 \theta_0 + 2 \theta_0^T S^2 \varepsilon + \sum_{k = 1}^{n}  \bigl(\lambda_k(S)\bigr)^2 Z_k^2
	\end{aligned}
	\end{equation}
	where in the last line $Z_k = (\varepsilon^T v_k(S))$, and $Z = (Z_1,\ldots,Z_k) \sim N(0,\Id)$ follows from the rotational invariance of the Gaussian distribution. Thus
	\begin{equation}
	\label{pf:linear_smoother_fixed_graph_testing_mean}
	\Ebb_{\theta_0}[T] = \theta_0^T S^2 \theta_0 + \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^2.
	\end{equation}
	\paragraph{Variance.}  
	Starting from~\eqref{pf:linear_smoother_fixed_graph_testing1} and recalling the basic fact $\Var(Z_k^2) = 2$, we derive
	\begin{equation}
	\label{pf:linear_smoother_fixed_graph_testing_var}
	\Var_{\theta_0}[T] \leq 8 \theta_0^T S^4 \theta_0 + 4 \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^4 \leq 8 \theta_0^T S^2 \theta_0 + 4 \sum_{k = 1}^{n} \bigl(\lambda_k(S)\bigr)^4
	\end{equation}
	where the second inequality follows since by assumption $\lambda_{\max}(S) \leq 1$.

	\paragraph{Bounding Type I and Type II error.}
	The bound~\eqref{eqn:linear_smoother_fixed_graph_testing_typeI} follows directly from Chebyshev's inequality, along with our above calculations on the mean and variance of $T$.

	The bound~\eqref{eqn:linear_smoother_fixed_graph_testing_typeII} also follows from Chebyshev's inequality, as can be seen by the following manipulations,
	\begin{equation*}
	\begin{aligned}
	\Pbb_{\theta_0}\bigl(T \leq t_b\bigr) & = \Pbb_{\theta_0}\bigl(T - \Ebb_{\theta_0}[T] \leq t_b - \Ebb_{\theta_0}[T]\bigr) \\
	& \overset{(i)}{\leq} \Pbb_{\theta_0}\bigl(\abs{T - \Ebb_{\theta_0}[T]} \geq \abs{t_b - \Ebb_{\theta_0}[T]}\bigr) \\ 
	& \overset{(ii)}{\leq} 4 \frac{\Var_{\theta_0}[T]}{(\theta_0^T S^2 \theta_0)^2} \\
	& \overset{(iii)}{\leq} \frac{32}{\theta_0^T S^2 \theta_0} + \frac{4}{b^2} \\
	& \overset{(iv)}{\leq} \frac{8}{b} \Biggl(\sum_{k = 1}^{n}\bigl(\lambda_k(S)\bigr)^4\Biggr)^{-1/2} + \frac{4}{b^2}
	\end{aligned}
	\end{equation*}
	In the previous expression, $(i)$ and $(ii)$ follow since assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and equation~\eqref{pf:linear_smoother_fixed_graph_testing_mean} together imply $\Ebb_{\theta_0}(T) - \frac{1}{2}\theta_0^T S^2\theta_0 \geq t_b$, $(iii)$ follows from assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius} and the inequality~\eqref{pf:linear_smoother_fixed_graph_testing_var}, and $(iv)$ follows assumption~\eqref{eqn:linear_smoother_fixed_graph_testing_critical_radius}.
\end{proof}

\end{document}