### 2020-09-30 notes/weekly_notes/9.23.20.NotesForWeek.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage{tikz}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\iid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}
\newcommand{\Lap}{L}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}
\newcommand{\Id}{I}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Vset}{\mathcal{V}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Notes for Week 9/16/2020 - 9/23/2020}
\author{Alden Green}
\date{\today}
\maketitle

In these notes, I will carefully examine the results of \textcolor{red}{(GarciaTrillos18, Calder19)} on the convergence of neighborhood-graph eigenvalues towards eigenvalues of a continuum Laplace-Beltrami operator. The idea is to: (a) prove results in the ``flat Euclidean'' setting (where the manifold is full-dimensional and has boundary), and (b) to ensure there are no hidden dependencies in the stated constants that I am missing.

\section{Main results}
Let $X_1,\ldots,X_n$ be sampled iid from a distribution $P$, supported on a compact set $\Xset \subset \Reals^d$.

\begin{enumerate}[(P1)]
	\item 
	\label{asmp:bounded_lipschitz_density} We assume the distribution $P$ admits density $p$ that is Lipschitz continuous with Lipschitz parameter $L_p$ and satisfies
	\begin{equation*}
	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty,~~\textrm{for all $x \in \Xset$.}
	\end{equation*}
\end{enumerate}

For a given radius $r > 0$, and kernel function $K: [0,\infty) \to [0,\infty)$, define the neighborhood graph $G_{n,r} = ([n],\mathbf{W})$ to be the graph with weighted edges $\mathbf{W}_{ij} = K(\|X_i - X_j\|/r)$. The neighborhood graph Laplacian is $\Lap_{n,r} = \mathbf{D} - \mathbf{W}$, where $\mathbf{D}$ is the degree matrix, and $0 = \lambda_1(G_{n,r}) \leq \cdots \leq \lambda_n(G_{n,r})$ are the eigenvalues of $\Lap_{n,r}$. Our goal is to relate these eigenvalues to the eigenvalues $\lambda_k(\Delta_P)$ of the Laplace-Beltrami operator
\begin{equation*}
\Delta_{P}f  := -\frac{1}{2p} \dive(p^2 \nabla f)
\end{equation*}
We would like to show the following result.
\begin{theorem}
	\label{thm:neighborhood_eigenvalue}
	There constants $A_1$ and $a_1$ which do not depend on $n$ or $r$ such that the following statement holds with probability at least $1 - A_1n\exp(-a_1n\theta^2\wt{\delta}^{m})$: for any $\ell \in \mathbb{N}$ such that
	\begin{equation}
	\label{eqn:neighborhood_eigenvalue_1}
	1 - A_1\Biggl(r \sqrt{\lambda_{\ell}(\Delta_P)} + (\theta + \wt{\delta})\Biggr)  \geq \frac{1}{2}
	\end{equation}
	we have that
	\begin{equation}
	\label{eqn:eigenvalue_bound}
	\frac{1}{A_1} \lambda_k(G_{n,r}) \leq nr^{d+2} \lambda_k(\Delta_P) \leq A_1 \lambda_k(G_{n,r}),~~\textrm{for all $1 \leq k \leq \ell$}
	\end{equation}
\end{theorem}
Some results along these lines (see \textcolor{red}{(Burago13, GarciaTrillos18, Calder19)}) have been shown when $\Xset$ is an $m$-dimensional submanifold of $\Reals^d$---in this case, replace $r^{d + 2}$ by $r^{m+2}$ in~\eqref{eqn:eigenvalue_bound}. Indeed in these results the constant $A = (1 + e)$ for an error term $e$ which goes to $0$ as $n \to \infty, r(n) \to 0$. However, (a) the manifold is assumed to be boundaryless, and (b) the right hand side in (their equivalent of)~\eqref{eqn:neighborhood_eigenvalue_1} depends on $\ell$ which is a bit strange and unsuitable for our purposes. We would like to ``fix'' both (a) and (b).

\section{Ancillary results.}
Following the lead of~\textcolor{red}{(Burago13, GarciaTrillos18, Calder19)}, we relate $\lambda_k(\Delta_P)$ and $\lambda_k(G_{n,r})$ by means of the Sobolev semi-norms (Dirichlet energies)
\begin{equation*}
b_r(f) := \frac{1}{n^2 r^{d+ 2}}f^T \Lap_{n,r} f~~\textrm{for any $f \in \Leb^2(P_n)$}
\end{equation*}
and
\begin{equation*}
D_2(f) :=
\begin{cases*}
\int_{\Xset} \|\nabla f(x)\|^2 p^2(x) \,dx~~ &\textrm{if $f \in H^1(\Xset)$} \\
\infty~~ & \textrm{otherwise,}
\end{cases*}
\end{equation*}
Suppose one could upper bound $b_r(u)$ by $D_2\bigl(G(u)\bigr)$ for an appropriate choice of mapping $G: \Leb^2(P_n) \to \Leb^2(P)$, and vice versa upper bound $D_2(f)$ by $b_r(V(f))$ for some $V: \Leb^2(P) \to \Leb^2(P_n)$. Suppose also that $G$ and $V$ are near-isometries, meaning $\|G(u)\|_{\Leb^2(P)} \approx \|u\|_{\Leb^2(P_n)}$ and $\|V(f)\|_{\Leb^2(P_n)} \approx \|f\|_{\Leb^2(P)}$. Then using the variational characterization of eigenvalues $\lambda_k(\Delta_P)$ and $\lambda_k(G_{n,r})$---i.e. the Courant-Fischer Theorem---one can obtain estimates on the error of $|\lambda_k(\Delta_P) - \lambda_k(G_{n,r})|$.

We now move to define maps $V$ and $G$ which provably satisfy such bounds. Let $\wt{P}_n$ be a probability measure with density $\wt{p}_n$, that satisfy the conclusions of Theorem~\ref{thm:calder19_2}. Denote by $\wt{T}_n$ an \emph{optimal transport map} (in sup-norm) between $\wt{P}_n$ and $P_n$, and let $U_1,\ldots,U_n$ be the preimages of $X_1,\ldots,X_n$ under $\wt{T}_n$. Define the discretization map  $\wt{\mathcal{P}}: \Leb^2(P) \to \Leb^2(P_n)$ by
\begin{equation*}
\bigl(\wt{\mathcal{P}}f\bigr)(X_i) := n \cdot \int_{U_i} f(x) \wt{p}_n(x) \,dx
\end{equation*}
Define the interpolation map $\wt{\mc{I}}: \Leb^2(P_n) \to \Leb^2(\Xset)$ as $\wt{\mc{I}}u := \Lambda_{r - 2\wt{\delta}}(\wt{\mc{P}}^{\star}u)$. Here, $\wt{\mc{P}}^{\star} = u \circ \wt{T}$ is the adjoint of $\wt{\mc{P}}$, i.e.
\begin{equation*}
\bigl(\wt{\mc{P}}^{\star}u\bigr)(x) = \sum_{j = 1}^{n} u(x_i) \1\{x \in U_i\} 
\end{equation*} 
and $\Lambda_r$ is a smoothening of this extension, formally
\begin{equation*}
\Lambda_r(f) := \frac{1}{r^d\tau(x)}\int_{\Xset} \eta_r(x',x) f(x') \,dx',~~ \eta_r(x',x) := \psi\biggl(\frac{\|x' - x\|}{r}\biggr)
\end{equation*}
where $\psi(t) := (1/\sigma_K)\int_{t}^{\infty} s K(s) \,ds$ and $\tau(x) := (1/r^d)\int_{\Xset} \eta_r(x',x) \,dx'$ is a normalizing constant.

Assume the following.
\begin{enumerate}[label=(K\arabic*)]
	\item
	\label{asmp:kernel}
	$K:[0,\infty) \to [0,\infty)$ is a non-increasing function supported on $[0,1]$, and its restriction to $[0,1]$ is Lipschitz with Lipschitz constant $L_K$. Additionally, it is normalized so that
	\begin{equation*}
	\int_{\Reals^d} K\bigl(\norm{z}\bigr) \,dz = 1,
	\end{equation*}
	and $K(1/2) > 0$. Let $\sigma_{K} := \frac{1}{d}\int_{\Reals^d} \|z\|^2 K(\|z\|) \,dy$.
\end{enumerate}
\begin{enumerate}[label=(OT\arabic*)]
	\item 
	\label{asmp:calder19_1}
	There exists $n^{-1/m} < \wt{\delta} < \frac{1}{4}r$.
	\item 
	\label{asmp:calder19_2} There exists $\theta$ such that $C(\theta + \wt{\delta}) \leq p_{\min}/2$.
	\item
	\label{asmp:r_small_1} For all $x,x' \in \Xset$ such that $\|x - x'\| \leq r$, we have that for $W = B(x,r) \cap B(x',r) \cap \Xset$,
	\begin{equation*}
	\vol(W) \geq a_1(\Xset,d) \cdot \nu_d \biggl(\frac{r}{2}\biggr)^d
	\end{equation*}
	for some constant $A_1(\Xset,d)$. 
	\item
	\label{asmp:r_small_2} For all $x,x' \in \Xset$, we have that $z_{xx'} = \frac{1}{2}x + \frac{1}{2}x' \in \Xset$.
	\item 
	\label{asmp:r_small_3} For any $x \in \Xset$, letting $\mc{B}_x:= (\Xset - x)/r \cap B(0,1)$, we have that $\int_{\mc{B}_x} \psi(\|z\|) \,dz \geq a_1(\Xset,d) \cdot \int_{B(0,1)} \psi(\|z\|) \,dz$.
\end{enumerate}
Very roughly speaking, one can view $\theta + \wt{\delta}$ as an upper bound on the optimal transport distance between $P_n$ and $\wt{P}_n$ and the $\Leb_{\infty}$ distance between $\wt{p}_n$ and $p$.

\textcolor{red}{(TODO)}: Eliminate these assumptions, by showing that each of them is a conseequence of the regularity assumption that $\Xset$ is a Lipschitz domain with positive reach.

\begin{proposition}[\textbf{c.f. Proposition 4.1 of \textcolor{red}{Calder19}}]
	\label{prop:dirichlet_energies}
	With probability at least $1 - A_1n\exp(-a_1n\theta^2\wt{\delta}^{m})$, we have the following.
	\begin{enumerate}[(1)]
		\item For any $u \in \Leb^2(P_n)$,
		\begin{equation*}
		\sigma_{K} D_2(\wt{\mc{I}}u) \leq \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \biggl(1 + A_3(d,K,p)\frac{\wt{\delta}}{r}\biggr) b_r(u)
		\end{equation*}
		\item For any $f \in \Leb^2(P)$,
		\begin{equation*}
		b_r(\wt{\mc{P}}f) \leq \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + L_p(2r + 4\wt{\delta})\Bigr) \cdot \Bigl(1 + A_1(K)\frac{\wt{\delta}}{r}\Bigr) \sigma_{K} D_2(f).
		\end{equation*}
	\end{enumerate}
\end{proposition}

\begin{proposition}[\textbf{c.f. Proposition 4.2 of \textcolor{red}{Calder19}}]
	\label{prop:isometry}
	With probability at least $1 - A_1n\exp(-a_1n\theta^2\wt{\delta}^{m})$, we have the following.
	\begin{enumerate}[(1)]
		\item For every $f \in \Leb^2(\Xset)$,
		\begin{equation*}
		\Bigl|\|f\|_{\Leb^2(P)}^2 - \|\wt{P}f\|_{\Leb^2(P_n)}^2\Bigr| \leq A_2(\Xset,d,K,p_{\min}) \cdot \wt{\delta} \|f\|_{\Leb^2(P)} \sqrt{D_2(f)} + \frac{A_1\bigl(\theta + \wt{\delta}\bigr)}{p_{\min}} \|f\|_{\Leb^2(P)}^2
		\end{equation*}
		\item For every $u \in \Leb^2(P_n)$,
		\begin{equation*}
		\Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(P)}^2 - \|u\|_{\Leb^2(P_n)}^2\Bigr| \leq C r \|u\|_{\Leb^2(P_n)} \sqrt{b_r(u)} + C\bigl(\theta + \wt{\delta}\bigr) \|u\|_{\Leb^2(P_n)}^2
		\end{equation*}
	\end{enumerate}
\end{proposition}

To prove these inequalities, it will help to introduce the intermediate energies
\begin{equation*}
\wt{E}_r(f,\eta,V) := \frac{1}{r^{d + 2}}\int_{V} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{r}\biggr) \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx
\end{equation*}
and
\begin{equation*}
{E}_r(f,\eta,V) := \frac{1}{r^{d + 2}}\int_{V} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx
\end{equation*}
where $\eta: [0,\infty) \to [0,\infty)$ is an arbitrary kernel, and $V \subseteq \Xset$ is a measurable set. We will abbreviate $\wt{E}_r(f,\eta,\Xset)$ as $\wt{E}_r(f,\eta)$ and $\wt{E}_r(f,K) = \wt{E}_r(f)$ (and likewise with $E_r$.)

Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized} relates the graph Sobolev semi-norm $b_r(\wt{\mc{P}}f)$ to the non-local energy $\wt{E}_r(f)$. It follows exactly from the derivations in the proof of Lemma 13 in \cite{trillos2019}; indeed, in our flat Euclidean setting, the proof is simplified, and we include this simpler proof for completeness purposes only.
\begin{lemma}[\textbf{c.f. Lemma 13 of \cite{trillos2019}, Lemma 4.3 of \cite{burago2014}}]
	\label{lem:first_order_graph_sobolev_seminorm_discretized}
	Suppose~\ref{asmp:kernel} is satisfied, and that the conclusions of Theorem~\ref{thm:calder19_2} are met. Then
	\begin{equation*}
	b_r(\wt{\mc{P}}f) \leq \wt{E}_{r + 2\wt{\delta}}(f) + 2\frac{L_K\wt{\delta}}{K(1/2)r} \wt{E}_{2(r + 2\wt{\delta})}(f)
	\end{equation*}
	for any $f \in \Leb^2(\Xset)$.
\end{lemma}

Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected} relates the non-local energy $E_r(f)$ to $D_2(f)$. It follows in large part from the proofs of Lemma 13 of \cite{trillos2019} and Lemma 6 of \cite{burago2014}, with minor modifications to account for the boundary of $\Xset$.
\begin{lemma}[c.f. Lemma 13 of \cite{trillos2019}, Lemma 6 of \cite{burago2014}]
	\label{lem:first_order_graph_sobolev_seminorm_expected}
	For any $f \in \Leb^2(\Xset)$,
	\begin{equation*}
	E_r(f) \leq (1 + L_pr)^2 \cdot \sigma_K D_2(f)
	\end{equation*}
\end{lemma}

In Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, we establish the reverse of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected}. 
\begin{lemma}[c.f. Lemma 9 of \cite{trillos2019}, Lemma 5.5 of \cite{burago2014}]
	\label{lem:first_order_graph_sobolev_seminorm_expected_lb}
	For any $f \in \Leb^2(\Xset)$,
	\begin{equation*}
	\sigma_KD_2(\Lambda_rf) \leq \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K}E_r(f)
	\end{equation*}
\end{lemma}

In Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}, we establish the reverse of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized}. 
\begin{lemma}[c.f. Lemma 14 of \cite{trillos2019}]
	\label{lem:first_order_graph_sobolev_seminorm_discretized_lb}
	Suppose $r$ satisfies~\ref{asmp:r_small_2}. For any $u \in \Leb^2(P_n)$, 
	\begin{equation*}
	\wt{E}_{r - 2\wt{\delta}}\bigl(\wt{\mc{P}}^{\star}u\bigr) \leq \biggl(1 + A_3(d,K,P)\frac{\wt{\delta}}{r}\biggr) b_{r}(u)
	\end{equation*}
\end{lemma}

To prove Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, we require upper and lower bounds on $\tau(x)$, as well as an upper bound on the gradient of $\tau$. The lower bound here---$\tau(x) \geq 1/2$---is quite a bit a looser than in the case where $\Xset$ has no boundary---$\tau(x) \geq (1 + Cr^2)^{-1}$. The same is the case regarding the upper bound of the size of the gradient $\|\nabla \tau(x)\|$. However, the bounds as stated here will be sufficient for our purposes.
\begin{lemma}
	\label{lem:tau_bound}
	Suppose $r$ satisfies~\ref{asmp:r_small_3}. For all $x \in \Xset$,
	\begin{equation*}
	\frac{1}{2} \leq \tau(x) \leq 1,
	\end{equation*}
	and additionally $\|\nabla \tau(x)\| \leq \frac{1}{\sigma_K r}$.
\end{lemma}

To prove part (2) of Proposition~\ref{prop:isometry}, we require Lemma~\ref{lem:smoothening_error}, which gives an estimate on the error $\Lambda_h f - f$ in $\Leb^2(P)$ norm.
\begin{lemma}[c.f Lemma 8 of \cite{trillos2019}, Lemma 5.4 of \cite{burago2014}]
	\label{lem:smoothening_error}
	For any $h > 0$, 
	\begin{equation}
	\label{eqn:smoothening_error_norm}
	\bigl\|\Lambda_hf\bigr\|_{\Leb^2(P)}^2 \leq \frac{2p_{\max}}{p_{\min}} \bigl\|f\bigr\|_{\Leb^2(P)}^2
	\end{equation}
	and
	\begin{equation}
	\label{eqn:smoothening_error_energy}
	\bigl\|\Lambda_hf - f\bigr\|_{\Leb^2(P)}^2 \leq \frac{2}{\sigma_Kp_{\min}} r E_r(f)
	\end{equation}
	for all $f \in \Leb^2(\Xset)$.
\end{lemma}

\section{Other's results}

For a transport map $T: \Xset \to \mathbf{X}$ and a measure $\mu$ supported on $\Xset$, let $T_{\sharp}\mu$ denote the push-forward of $\mu$ by $T$, i.e the measure for which
\begin{equation*}
\bigl(T_{\sharp}\mu\bigr)(\mathbf{Z}) = \mu\bigl(T^{-1}(\mathbf{Z}))
\end{equation*}
for any $\mathbf{Z} \subseteq \mathbf{X}$.

\begin{theorem}[Proposition~2.11 of \textcolor{red}{(Calder2019)}]
	\label{thm:calder19_2}
	Suppose the assumptions~\ref{asmp:calder19_1} and~\ref{asmp:calder19_2} are satisfied. Then with probability at least $1 - A_1n\exp(-a_1n\theta^2\wt{\delta}^d)$, there exists a probability measure $\wt{P}_n$ with density function $\wt{p}_n: \Xset \to \Reals$ such that
	\begin{equation*}
	\min_{T_{\sharp}\wt{P}_n = P_n} d_{\mc{M}}(x,T(x)) \leq \wt{\delta}
	\end{equation*}
	and
	\begin{equation*}
	\bigl\|\wt{p}_n - p\bigr\|_{\infty} \leq A_1\bigl(\theta + \wt{\delta}\bigr)
	\end{equation*}
\end{theorem}
\textcolor{red}{(TODO)}: Replace this Theorem with it's ``flat'' predecessor.

By Theorem~\ref{thm:calder19_2}, the two energies can be related,
\begin{equation}
\label{eqn:calder19_1}
\bigl(1 - A_1(\theta + \wt{\delta})\bigr) E_r(f) \leq \wt{E}_r(f) \leq \bigl(1 + A_1(\theta + \wt{\delta})\bigr) E_r(f),
\end{equation}
and so can the norms
\begin{equation}
\label{eqn:calder19_2}
\bigl(1 - A_1(\theta + \wt{\delta})\bigr) \|f\|_{\Leb^2(P)}^2 \leq \|f\|_{\Leb^2(\wt{P}_n)}^2 \leq \bigl(1 + A_1(\theta + \wt{\delta})\bigr) \|f\|_{\Leb^2(P)}^2.
\end{equation}

\section{Additional proofs}

\paragraph{Proof of Theorem~\ref{thm:neighborhood_eigenvalue}.}
\mbox{} \\
\mbox{} \\
We start with the upper bound, proceeding exactly as in \citep{burago2014} (Proposition 4.4). Let $W$ be the span of $f_1,\ldots,f_{\ell}$, the first ${\ell}$ eigenfunctions of $-\Delta_P$, so that for every $f \in W$ we have $D_2(f) \leq \lambda_k(\Delta_P) \|f\|_{\Leb^2(P)}$. As a result, by Part (1) of Proposition~\ref{prop:isometry} we have that for any $f \in W$
\begin{equation*}
\bigl\|\wt{\mc{P}}f\bigr\|_{\Leb^2(P_n)}^2 \geq \biggl(1 - A_2(\Xset,d,K,p_{\min}) \wt{\delta} \sqrt{\lambda_{\ell}(\Delta_P)} - \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr)\|f\|_{\Leb^2(P)}^2  \geq \frac{1}{2} \|f\|_{\Leb^2(P)}^2,
\end{equation*}
where the second inequality follows by assumption~\eqref{eqn:neighborhood_eigenvalue_1}.

Therefore $\wt{\mc{P}}$ is injective over $W$, and $\wt{\mc{P}}W$ has dimension $\ell$. This means we can invoke the Courant-Fischer Theorem, along with Proposition \ref{prop:dirichlet_energies}, and conclude that
\begin{align*}
\lambda_k(G_{n,r}) & \leq \max_{\substack{u \in \wt{\mc{P}}W \\ u \neq 0} } \frac{b_r(u)}{\|u\|_{\Leb^2(P_n)}^2} \\
& = \max_{\substack{f \in W \\ f \neq 0} } \frac{b_r(\wt{\mc{P}}f)}{\bigl\|\wt{\mc{P}}f\bigr\|_{\Leb^2(P_n)}^2} \\
& \leq \frac{\Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + L_p(2r + 4\wt{\delta})\Bigr) \cdot \Bigl(1 + A_1(K)\frac{\wt{\delta}}{r}\Bigr)}{1 - A_2(\Xset,d,K,p_{\min}) \wt{\delta} \sqrt{\lambda_{\ell}(\Delta_P)} - 2\frac{A_1(\theta + \wt{\delta})}{p_{\min}}} \sigma_K \lambda_{\ell}(\Delta_P),
\end{align*}
establishing the upper bound in~\eqref{eqn:eigenvalue_bound} upon proper choice of $A$.

\textcolor{red}{(TODO)}: Ask Siva how to `finish this proof off'.

\paragraph{Proof of Proposition~\ref{prop:dirichlet_energies}.}
\mbox{} \\
\mbox{} \\
Part (1) follows from
\begin{align*}
\sigma_K D_2(\Lambda_{r - 2\wt{\delta}} \wt{\mc{P}}^{\star}u) & \overset{(i)}{\leq} \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} E_{r - 2\wt{\delta}}(\wt{\mc{P}}^{\star}u) \\
& \overset{(ii)}{\leq} \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \wt{E}_{r - 2\wt{\delta}}(\wt{\mc{P}}^{\star}u) \\
& \overset{(iii)}{\leq} \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \biggl(1 + A_3(d,K,p)\frac{\wt{\delta}}{r}\biggr) b_r(u)
\end{align*}
where $(i)$ follows from Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}, $(ii)$ follows from~\eqref{eqn:calder19_1}, and $(iii)$ follows from~Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}.

Part (2) follows from
\begin{align*}
b_r(\wt{\mc{P}}f) & \overset{(iv)}{\leq} \wt{E}_{r + 2\wt{\delta}}(f) + A_1(K)\frac{\wt{\delta}}{r}\wt{E}_{2(r + 2\wt{\delta})}(f)\\
& \overset{(v)}{\leq} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl({E}_{r + 2\wt{\delta}}(f) + A_1(K)\cdot\frac{\wt{\delta}}{r}{E}_{2(r + 2\wt{\delta})}(f)\Bigr) \\
& \overset{(vi)}{\leq} \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \cdot \Bigl(1 + L_p(2r + 4\wt{\delta})\Bigr) \cdot \Bigl(1 + A_1(K)\frac{\wt{\delta}}{r}\Bigr) \cdot \sigma_{\eta} D_2(f)
\end{align*}
where $(iv)$ follows from Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized}, $(v)$ follows from~\eqref{eqn:calder19_1}, and $(vi)$ follows from Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected}.

\paragraph{Proof of Proposition~\ref{prop:isometry}.}
\mbox{}\\
\mbox{}\\
\textit{Proof of (1).}
\mbox{}\\
\mbox{}\\
We begin by upper bounding $\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}$. By the Cauchy-Schwarz inequality and the bound on $\|\wt{p}_n - p\|_{\infty}$ in~\eqref{eqn:calder19_1},
\begin{align*}
\Bigl|\wt{P}f(X_i)\Bigr|^2 & = n^2 \Bigl|\int_{U_i} f(x) \wt{p}_n(x) \,dx\Bigr|^2 \\
& \leq n \int_{U_i} \bigl|f(x)\bigr|^2 \wt{p}_n(x) \,dx \\
& \leq n \cdot \biggl[\int_{U_i} \bigl|f(x)\bigr|^2 p(x) \,dx + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \int_{U_i} \bigl|f(x)\bigr|^2 p(x) \,dx\biggr]
\end{align*}
and summing over $i = 1,\ldots,n$, we obtain
\begin{equation}
\label{pf:prop_isometry_1}
\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}^2 \leq \biggl(1 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \bigl\|f\bigr\|_{\Leb^2(P)}^2.
\end{equation}
Now, noticing that $\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)} = \bigl\|\wt{P}^{\star}\wt{P}f\bigr\|_{\Leb^2(\wt{P}_n)}$, we can use the upper bound~\eqref{pf:prop_isometry_1} to show that
\begin{align}
\Bigl|\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}^2 - \bigl\|f\bigr\|_{\Leb^2(P)}^2\Bigr| & \leq \Bigl|\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}^2 - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2\Bigr| + \Bigl|\bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2 - \bigl\|f\bigr\|_{\Leb^2(P)}^2\Bigr| \nonumber \\
& \overset{(i)}{\leq} \Bigl|\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)}^2 - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}^2\Bigr|  + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{\Leb^2(P)}^2 \\
& \overset{(ii)}{\leq} \biggl(2 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \Bigl|\bigl\|\wt{P}f\bigr\|_{\Leb^2(P_n)} - \bigl\|f\bigr\|_{\Leb^2(\wt{P}_n)}\Bigr| \cdot \bigl\|f\bigr\|_{\Leb^2(P)} + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{\Leb^2(P)}^2 \nonumber \\
& \leq \biggl(2 + \frac{A_1(\theta + \wt{\delta})}{p_{\min}}\biggr) \bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)} \cdot \bigl\|f\bigr\|_{\Leb^2(P)} + \frac{A_1(\theta + \wt{\delta})}{p_{\min}} \bigl\|f\bigr\|_{\Leb^2(P)}^2 \label{pf:prop_isometry_2}
\end{align}
where $(i)$ follows from~\eqref{eqn:calder19_2} and $(ii)$ follows from~\eqref{pf:prop_isometry_1}.

It remains to upper bound $\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)}^2$. Noting that $\wt{P}^{\star}\wt{P}f$ is piecewise constant over the cells $U_i$, we have
\begin{equation*}
\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)}^2 = \sum_{i = 1}^{n} \int_{U_i} \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx.
\end{equation*}
We will now establish the following Poincare-type inequality, 
\begin{equation}
\label{pf:prop_isometry_3}
\int_{U_i} \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx \leq \frac{2^{d + 2}}{a_1(\Xset,d) \nu_d K(1/2)} \wt{\delta}^2 \wt{E}_{2\wt{\delta}}(f,U_i)
\end{equation}
which holds for all $i = 1,\ldots,n$. Summing up over $i$ implies
\begin{equation*}
\bigl\|\wt{P}^{\star}\wt{P}f - f\bigr\|_{\Leb^2(\wt{P}_n)}^2 \leq \frac{2^{d+2}}{a_1(\Xset,d)K(1/2)p_{\min}} \cdot \wt{E}_{2\wt{\delta}}(f) \leq \frac{2^{d+3}}{a_1(\Xset,d)K(1/2)p_{\min}} A_1(\Xset,d,K,p_{\min}) D_2(f)
\end{equation*}
with the latter inequality following just as in the proof of Proposition~\ref{prop:dirichlet_energies}; the Lemma then follows by plugging this inequality into~\eqref{pf:prop_isometry_2}.

Now it remains to show~\eqref{pf:prop_isometry_3}. The key point is that for any pair of points $x,x' \in \Xset$ such that $\|x - x'\| \leq r$  the triangle inequality
\begin{equation*}
\bigl|f(x') - f(x)\bigr|^2 \leq 2\bigl|f(x') - f(z)\bigr|^2 + 2\bigl|f(z) - f(x)\bigr|^2
\end{equation*}
holds for each $z \in B\bigl(x,\wt{\delta}\bigr) \cap B\bigl(x',\wt{\delta}\bigr) \cap \Xset =: W$, and we have that
\begin{align*}
\bigl|f(x') - f(x)\bigr|^2 & \leq \frac{2}{\vol(W)}\biggl[\int_{W} \bigl|f(x') - f(z)\bigr|^2 \,dz + \int_{W} \bigl|f(z) - f(x)\bigr|^2 \,dz\biggr] \\
& \leq \frac{2}{\vol(W)K(1/2)}\biggl[\int_{\Xset} K\biggl(\frac{x' - z}{2\wt{\delta}}\biggr)\bigl|f(x') - f(z)\bigr|^2 \,dz + \int_{\Xset} K\biggl(\frac{z - x}{2\wt{\delta}}\biggr)\bigl|f(z) - f(x)\bigr|^2 \,dz\biggr] \\
& \leq \frac{2^{d + 1}}{a_1(\Xset,d) \nu_d \wt{\delta}^d K(1/2)}\biggl[\int_{\Xset} K\biggl(\frac{x' - z}{2\wt{\delta}}\biggr)\bigl|f(x') - f(z)\bigr|^2 \,dz + \int_{\Xset} K\biggl(\frac{z - x}{2\wt{\delta}}\biggr)\bigl|f(z) - f(x)\bigr|^2 \,dz\biggr]
\end{align*}
where the last inequality follows from Assumption~\ref{asmp:r_small_1}. We use this along with the Cauchy-Schwarz inequality to prove~\eqref{pf:prop_isometry_3}:
\begin{align*}
\int_{U_i} & \biggl(f(x) - n\cdot\int_{U_i} f(x') \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx \\
 & =\int_{U_i} \biggl(n\cdot\int_{U_i} \bigl(f(x) - f(x')\bigr) \wt{p}_n(x') \,dx'\biggr)^2 \wt{p}_n(x) \,dx \\
& \leq n \cdot \int_{U_i} \int_{U_i} \bigl(f(x) - f(x')\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx \\
& \leq \frac{2^{d + 2}}{a_1(\Xset,d) \nu_d \wt{\delta}^d K(1/2)}n\int_{U_i} \int_{U_i} \int_{\Xset} K\biggl(\frac{x - z}{2\wt{\delta}}\biggr)\bigl|f(x) - f(z)\bigr|^2 \wt{\rho}_n(x') \wt{\rho}_n(x)  \,dx' \,dz \,dx \\
& \leq \frac{2^{d + 2}}{a_1(\Xset,d) \nu_d \wt{\delta}^d K(1/2) p_{\min}} \int_{U_i} \int_{U_i} K\biggl(\frac{x - z}{2\wt{\delta}}\biggr)\bigl|f(x) - f(z)\bigr|^2 \wt{\rho}_n(z) \wt{\rho}_n(x)  \,dz \,dx \\
& =  \frac{2^{d + 2}}{a_1(\Xset,d) \nu_d K(1/2) p_{\min}} \wt{\delta}^2 \wt{E}_{2\wt{\delta}}(f,U_i).
\end{align*}

\textit{Proof of (2).}
\mbox{}\\
\mbox{}\\

By the triangle inequality,
\begin{align}
\Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(P)}^2 - \|u\|_{\Leb^2(P_n)}^2\Bigr| & \leq \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(P)}^2 - \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2\Bigr| + \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 - \|u\|_{\Leb^2(P_n)}^2\Bigr| \nonumber \\
& \leq A_1(\theta + \wt{\delta}) \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 + \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 - \|u\|_{\Leb^2(P_n)}^2\Bigr| \nonumber\\
& = A_1(\theta + \wt{\delta}) \|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 + \Bigl(\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} + \|u\|_{\Leb^2(P_n)}\Bigr) \cdot \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|u\|_{\Leb^2(P_n)}\Bigr| \label{pf:prop_isometry_4}
\end{align}
To upper bound the second term in the above expression, we first note that~$\|u\|_{\Leb^2(P_n)} = \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)}$, and thus
\begin{align}
\Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|u\|_{\Leb^2(P_n)} \Bigr| & = \Bigl|\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)} - \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)}\Bigr| \nonumber \\
& \overset{(iii)}{\leq} \|\Lambda_{\wt{r}}\wt{\mc{P}}^{\star}u - \wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)} \nonumber \\
& \overset{(iv)}{\leq} r \sqrt{\frac{2}{\sigma_K p_{\min}} E_{\wt{r}}(\wt{\mc{P}}^{\star}u)} \nonumber \\
& \overset{(v)}{\leq} r \sqrt{\frac{2}{\sigma_K p_{\min}} \Bigl(1 + A_3(d,K,p)\frac{\wt{\delta}}{r}\Bigr) b_r(u)} \label{pf:prop_isometry_5}
\end{align}
where $(iii)$ follows by the triangle inequality, $(iv)$ follows from Lemma~\ref{lem:smoothening_error}, and $(v)$ follows from~\eqref{eqn:calder19_1} and Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}. On the other hand, by~\eqref{eqn:calder19_2} and Lemma~\ref{lem:smoothening_error},
\begin{align*}
\|\wt{\mc{I}}u\|_{\Leb^2(\wt{P}_n)}^2 & \leq \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \|\wt{\mc{I}}u\|_{\Leb^2(P)}^2 \\
& \leq \frac{2p_{\max}}{p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr) \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(P)}^2 \\
& \leq \frac{2p_{\max}}{p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)^2 \|\wt{\mc{P}}^{\star}u\|_{\Leb^2(\wt{P}_n)}^2 \\
& = \frac{2p_{\max}}{p_{\min}} \cdot \Bigl(1 + A_1(\theta + \wt{\delta})\Bigr)^2 \|u\|_{\Leb^2(P_n)}^2
\end{align*}
and plugging this estimate along with~\eqref{pf:prop_isometry_5} back into~\eqref{pf:prop_isometry_4}, we obtain part (2) of Proposition~\ref{prop:isometry}.

\paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized}.} \mbox{} \\
\mbox{} \\
Recalling that $\bigl(\wt{P}f\bigr)(X_i) = n \cdot \int_{U_i} f(x) \wt{p}_n(x)$, the key point is that
\begin{equation*}
\biggl(\bigl(\wt{P}f\bigr)(X_i) - \bigl(\wt{P}f\bigr)(X_j)\biggr)^2 \leq n^2 \cdot \int_{U_i} \int_{U_j} \bigl(f(x') - f(x)\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx.
\end{equation*}
Additionally, the non-increasing and Lipschitz properties of $K$ imply that for any $x \in U_i$ and $x' \in U_j$, 
\begin{equation*}
K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) \leq K\biggl(\frac{\bigl(\|x' - x\| - 2\wt{\delta}\bigr)_{+}}{r}\biggr) \leq K\biggl(\frac{\|x' - x\|}{r + 2\wt{\delta}}\biggr) + \frac{2L_K\wt{\delta}}{r}\1\Bigl\{\|x' - x\| \leq r + 2\wt{\delta}\Bigr\}
\end{equation*}
Using this along with the Lipschitz property of $K$, we get
\begin{align*}
b_r(f) & = \frac{1}{n^2r^{d + 2}} \sum_{i,j = 1}^n \Bigl(\bigl(\wt{P}f\bigr)(X_i) - \bigl(\wt{P}f\bigr)(X_j)\Bigr)^2 K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) \\
& \leq \frac{1}{r^{d + 2}} \sum_{i,j = 1}^n \int_{U_i} \int_{U_j}  \bigl(f(x') - f(x)\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) K\biggl(\frac{\|X_i - X_j\|}{r}\biggr) \,dx' \,dx \\
& \leq \frac{1}{r^{d + 2}} \sum_{i,j = 1}^n \int_{U_i} \int_{U_j}  \bigl(f(x') - f(x)\bigr)^2 \wt{p}_n(x') \wt{p}_n(x) \biggl[K\biggl(\frac{\|x' - x\|}{r + 2\wt{\delta}}\biggr) + \frac{2L_K\wt{\delta}}{r}\1\Bigl\{\|x' - x\| \leq r + 2\wt{\delta}\Bigr\}\biggr] \,dx' \,dx \\
& = \wt{E}_{r + 2\wt{\delta}}(f) + \frac{2L_K\wt{\delta}}{r}\wt{E}_{r + 2\wt{\delta}}(f; \1_{[0,1]})
\end{align*}
for $\1_{[0,1]}(t) = \1\{0 \leq t \leq 1\}$. But clearly $\wt{E}_{r + 2\wt\delta}(f; \1_{[0,1]}) \leq 1/(K(1/2))\wt{E}_{2r + 4\wt{\delta}}(f)$, and the Lemma is shown.

\paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected}.}
\mbox{}\\
\mbox{}\\
Suppose $f \in C^{\infty}(\wb{\Xset})$, which we may do without loss of generality due to the density of $C^{\infty}(\wb{\Xset})$ in  $\Leb^2(\Xset)$.

Taylor expanding $f(x')$ about $x' = x$, applying first Cauchy-Schwarz' and then Jensen's inequality, and then using the Lipschitz property of $p$, we have
\begin{align*}
E_r(f) & = \frac{1}{r^{d + 2}}\int_{\Xset} \int_{\Xset} (f(x') - f(x))^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \\
& = \frac{1}{r^{d + 2}}\int_{\Xset} \int_{\Xset} \biggl(\sum_{i = 1}^{d} (x' - x)^{e_i} \int_0^1 f^{(e_i)}\bigl(x + t(x' - x)\bigr) \,dt\biggr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \\
& \leq \frac{1}{r^{d + 2}}\int_{\Xset} \int_{\Xset} \|x' - x\|^2 \sum_{i = 1}^{d}\biggl(\int_0^1 f^{(e_i)}\bigl(x + t(x' - x)\bigr) \,dt\biggr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \\
& \leq \frac{1}{r^{d + 2}}\int_{\Xset} \int_{\Xset} \int_0^1   \|x' - x\|^2 \biggl\{\sum_{i = 1}^{d} \Bigl(f^{(e_i)}\bigl(x + t(x' - x)\bigr)\Bigr)^2\biggr\} K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x)   \,dt   \,dx' \,dx \\
& \leq \frac{(1 + L_pr)^2}{r^{d + 2}}\int_{\Xset} \int_{\Xset} \int_0^1  \|x' - x\|^2 \biggl\{\sum_{i = 1}^{d} \Bigl(f^{(e_i)}\bigl(x + t(x' - x)\bigr)\Bigr)^2\biggr\} K\biggl(\frac{\|x' - x\|}{r}\biggr) p\bigl(x + t(x' - x)\bigr)^2   \,dt   \,dx' \,dx.
\end{align*} 
Then letting $z = (x' - x)/r$ and $\wt{x} = x + trz$, we obtain
\begin{align*}
E_r(f) & \leq (1 + L_pr)^2 \int_{\Xset} \int_{\mathcal{Z}_r(x)} \int_{0}^{1} \|z\|^2 \biggl\{\sum_{i = 1}^{d} \Bigl(f^{(e_i)}\bigl(x + trz\bigr)\Bigr)^2\biggr\} K\bigl(\|z\|\bigr) p(x + trz)^2 \,dt \,dz \,dx \\
& = (1 + L_pr)^2 \int_{B(0,1)} \int_{0}^{1}  \int_{\Xset_r(z)} \|z\|^2 \biggl\{\sum_{i = 1}^{d} \Bigl(f^{(e_i)}\bigl(x + trz\bigr)\Bigr)^2\biggr\} K\bigl(\|z\|\bigr) p(x + trz)^2 \,dx \,dt \,dz \\
& \leq (1 + L_pr)^2 \int_{B(0,1)} \int_{0}^{1} \|z\|^2 K\bigl(\|z\|\bigr) \biggl\{\int_{\wt{\Xset}_r(t,z)} \|\nabla f(\wt{x})\|^2 p(\wt{x})^2 \,d\wt{x}\biggr\} \,dt \,dz  \\
& \leq (1 + L_pr)^2 \int_{B(0,1)} \int_{0}^{1} \|z\|^2 K\bigl(\|z\|\bigr) \biggl\{\int_{\Xset} \|\nabla f(\wt{x})\|^2 p(\wt{x})^2 \,d\wt{x}\biggr\} \,dt \,dz \\
& = (1 + L_pr)^2 \int_{B(0,1)} \|z\|^2 K\bigl(\|z\|\bigr) \,dz 
\end{align*}
where $\mathcal{Z}_r(x) = B(0,1) \cap (\Xset - x)/r$, $\Xset_r(z) = \Xset \cap( \Xset - zr)$, $\wt{\Xset}_r(t,z) = \Xset_r(z) + trz$, and we note that since
\begin{align*}
\wt{x} \in \wt{\Xset}_r(t,z) & \Longrightarrow \wt{x} = x + trz ~~ \textrm{for some $x \in \Xset - zr$ and $t \in [0,1]$} \\
& \Longrightarrow \wt{x} \in \Xset,~~ \textrm{(since $x \in \Xset$, $x + rz \in \Xset$ and $\Xset$ is connected)}
\end{align*}
it must be that $\wt{X}_r(t,z) \subseteq \Xset$ for all $z \in B(0,1)$ and $t \in [0,1]$.

\paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}.}
\mbox{} \\
\mbox{} \\
For any $a \in \Reals$, $\Lambda_rf$ satisfies the identity
\begin{equation*}
\Lambda_rf(x) = a + \frac{1}{r^d\tau(x)}\int_{\Xset} \eta_r(x',x)\bigl(f(x') - a\bigr)\,dx'
\end{equation*}
and by differentiating we obtain
\begin{equation*}
\bigl(\nabla \Lambda_rf\bigr)(x)= \frac{1}{r^d\tau(x)}\int_{\Xset} \bigl(\nabla \eta_r(x',\cdot)\bigr)(x)\bigl(f(x') - a\bigr)\,dx' + \nabla\bigl(\tau^{-1}\bigr)(x)\cdot \frac{1}{r^d}\int_{\Xset} \eta_r(x',x)\bigl(f(x') - a\bigr)\,dx'
\end{equation*} 
Plugging in $a = f(x)$, we get $\nabla\Lambda_rf(x) = J_1(x)/\tau(x) + J_2(x)$ for
\begin{equation*}
J_1(x) := \frac{1}{r^d}\int_{\Xset} \bigl(\nabla \eta_r(x',\cdot)\bigr)(x)\bigl(f(x') - f(x)\bigr)\,dx',~~ J_2(x) := \nabla\bigl(\tau^{-1}\bigr)(x)\cdot \frac{1}{r^d}\int_{\Xset} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)\,dx'.
\end{equation*}
To upper bound $\bigl\|J_1(x)\bigr\|^2$, we first compute the gradient of $\eta_r(x',\cdot)$,
\begin{align*}
\bigl(\nabla\eta_r(x',\cdot)\bigr)(x) & = \frac{1}{r} \psi'\biggl(\frac{\|x'  - x\|}{r}\biggr) \frac{(x - x')}{\|x' - x\|} \\
& = \frac{1}{\sigma_Kr^2} K\biggl(\frac{\|x' - x\|}{r}\biggr) (x' - x),
\end{align*}
so that by the Cauchy-Schwarz inequality,
\begin{align*}
\bigl\|J_1(x)\bigr\|^2 & = \frac{1}{\sigma_K^2 r^{4 + 2d}} \Biggl[\int_{\Xset} \bigl(f(x') - f(x)\bigr)K\biggl(\frac{\|x' - x\|}{r}\biggr)(x' - x)\Biggr]^2 \,dx' \\
& \leq \frac{1}{\sigma_K^2r^{4 + 2d}} \biggl[\int_{\Xset}K\biggl(\frac{\|x' - x\|}{r}\biggr)\|x' - x\|^2\,dx'\biggr] \biggl[\int_{\Xset}K\biggl(\frac{\|x' - x\|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2\,dx'\biggr] \\
& \leq \frac{1}{\sigma_K r^{2 + d}}\int_{\Xset}K\biggl(\frac{\|x' - x\|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2\,dx'.
\end{align*}
To upper bound $\bigl\|J_2(x)\bigr\|^2$, we use the Cauchy-Schwarz inequality along with the observation $\eta_r(x',x) \leq \frac{1}{\sigma_K} K\bigl(\|x' - x\|/r\bigr)$ to deduce
\begin{align*}
\bigl\|J_2(x)\bigr\|^2 & \leq \Bigl\|\nabla\bigl(\tau^{-1}\bigr)(x)\Bigr\|^2\frac{1}{r^{2d}} \biggl[\int_{\Xset}\eta_r(x',x) \,dx'\biggr] \cdot \biggl[\int_{\Xset} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr] \\
& = \Bigl\|\nabla\bigl(\tau^{-1}\bigr)(x)\Bigr\|^2\frac{\tau(x)}{r^d} \int_{\Xset} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)^2 \,dx'\\ 
& \leq \Bigl\|\nabla\bigl(\tau^{-1}\bigr)(x)\Bigr\|^2\frac{\tau(x)}{\sigma_K r^d}\int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 \,dx' \\
& \leq \frac{16}{\sigma_K^2r^{2 + d}} \biggl[\int_{\Xset} K\biggl(\frac{|x' - x|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr].
\end{align*}
where the last inequality follows by the estimates on $\tau$ and $\nabla \tau$ given in Lemma~\ref{lem:tau_bound}. Combining our bounds on $\bigl\|J_1(x)\bigr\|^2$ and $\bigl\|J_2(x)\bigr\|^2$ along with the lower bound on $\tau(x)$ in Lemma~\ref{lem:tau_bound} and integrating over $\Xset$, we have
\begin{align*}
\sigma_K D_2(\Lambda_r f) & = \sigma_K\int_{\Xset} \biggl\|\Bigl(\nabla \Lambda_rf)(x)\biggr\|^2 p^2(x) \,dx \\
& \leq 2 \sigma_K \int_{\Xset} \Biggl(\frac{\bigl\|J_1(x)\bigr\|^2}{\tau^2(x)} + \bigl\|J_2(x)\bigr\|^2\Biggr) p^2(x) \,dx \\
& \leq \frac{40}{\sigma_K r^{2 + d}} \int_{\Xset} \int_{\Xset} K\biggl(\frac{|x' - x|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 p^2(x) \,dx' \,dx \\
& \leq \frac{40 \bigl(1 + L_p r\bigr)}{\sigma_K} E_r(f),
\end{align*}
completing the proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}.

\paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_discretized_lb}.}
\mbox{} \\
\mbox{} \\
For brevity, we write $\wt{r} := r - 2\wt{\delta}$. We begin by expanding the energy $\wt{E}_{r - 2\wt{\delta}}\bigl(\wt{\mc{P}}^{\star}u\bigr)$ as a double sum of double integrals,
\begin{align*}
\wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr) & = \frac{1}{\wt{r}^{d + 2}} \sum_{i = 1}^{n} \sum_{j = 1}^{n} \int_{U_i} \int_{U_j} \Bigl(u(X_i) - u(X_j)\Bigr)^2 K\biggl(\frac{\|x' - x\|}{\wt{r}}\biggr) \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx
\end{align*}
We next use the Lipschitz property of the kernel $K$---in particular that for $x \in U_i$ and $x' \in U_j$,
\begin{equation*}
K\biggl(\frac{\|x' - x\|}{\wt{r}}\biggr) \leq K\biggl(\frac{\|x_i - x_j\|}{r}\biggr) + \frac{2L_K\wt{\delta}}{\wt{r}} \cdot \1\biggl\{\frac{\|x' - x\|}{\wt{r}}\biggr\}
\end{equation*}
---to conclude that
\begin{align}
\wt{E}_{\wt{r}}\bigl(\wt{\mc{P}}^{\star}u\bigr) & \leq \frac{1}{n^2\wt{r}^{d + 2}} \sum_{i = 1}^{n} \sum_{j = 1}^{n} \Bigl(u(X_i) - U(X_j)\Bigr)^2 K\biggl(\frac{\|x_i - x_j\|}{r}\biggr) + \frac{2L_K\wt{\delta}}{\wt{r}}\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \nonumber \\
& \leq \biggl(1 + 4\bigl(2^{d + 2} - 1\bigr)\frac{\wt{\delta}}{r}\biggr) b_r(u) + \frac{2L_K\wt{\delta}}{\wt{r}}\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \label{pf:lem_first_order_graph_sobolev_seminorm_discretized_1}.
\end{align}
We focus now on the factor of $\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,1_{[0,1]})$. For any $x$ and $x' \in \Xset$ such that $\|x - x'\| \leq r$, note that by assumption $z_{xx'}:= \frac{1}{2}x + \frac{1}{2}x' \in \Xset \cap B(x,\wt{r}/2) \cap B(x',\wt{r}/2)$. Moreover by the triangle inequality
\begin{equation*}
\bigl|f(x) - f(x')\bigr|^2 \leq 2\bigl|f(x) - f(z_{xx'})\bigr|^2 + 2\bigl|f(z_{xx'}) - f(x')\bigr|^2
\end{equation*}
so that
\begin{align*}
\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) & \leq \int_{\Xset} \int_{\Xset \cap B(x,r)} \Bigl[ 2\bigl|f(x) - f(z_{xx'})\bigr|^2 +  2\bigl|f(z_{xx'}) - f(x')\bigr|^2\Bigr] \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx.
\end{align*}
Now, by a change of variables along with the Lipschitz property of $p$, we can deduce that
\begin{align*}
 \int_{\Xset} \int_{\Xset \cap B(x,r)} \bigl|f(x) - f(z_{xx'})\bigr|^2 \wt{p}_n(x') \wt{p}_n(x) \,dx' \,dx & \leq 2^d \bigl(1 + L_p\wt{r}\bigr)\int_{\Xset} \int_{\Xset \cap B(x,r/2)} \bigl|f(x) - f(z)\bigr|^2 \wt{p}_n(z) \wt{p}_n(x) \,dz \,dx \\
 & = 2^d \bigl(1 + L_p\wt{r}\bigr) \wt{E}_{\wt{r}/2}\bigl(f; \1_{[0,1]}\bigr),
\end{align*}
and therefore by symmetry
\begin{equation*}
\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr) \leq 2^{d + 2} \bigl(1 + L_p\wt{r}\bigr) \wt{E}_{\wt{r}/2}\bigl(f; \1_{[0,1]}\bigr).
\end{equation*}
Finally, noting that $\wt{E}_{\wt{r}/2}\bigl(f; \1_{[0,1]}) \leq \frac{1}{K(1/2)} \wt{E}_{\wt{r}}\bigl(f)$, the proof is completed by plugging our upper bound on $\wt{E}_{\wt{r}}(\wt{\mc{P}}^{\star}u,\1_{[0,1]}\bigr)$ back into~\eqref{pf:lem_first_order_graph_sobolev_seminorm_discretized_1}.
\paragraph{Proof of Lemma~\ref{lem:tau_bound}.}
\mbox{}\\
\mbox{}\\
First prove the estimates of $\tau(x)$, and then upper bound on $\|\nabla\tau(x)\|$. Substituting $z = (x' - x)/r$ and using~\ref{asmp:r_small_3}, we have that
\begin{align*}
\tau(x) & = \frac{1}{r^d} \int_{\Xset} \psi\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx' \\
& =  \int_{\mc{B}_x} \psi\bigl(\|z\|\bigr) \,dz \\
& \geq a_1(\Xset,d) \cdot \int_{B(0,1)} \psi(\|z\|) \,dz
\end{align*}
We will now show that $\int_{B(0,1)} \psi(\|z\|) \,dz = 1$, which will imply our estimate of $\tau(x)$. To see this identity, note that on the one hand, by converting to polar coordinates and integrating by parts we obtain
\begin{align*}
\int_{B(0,1)} \psi\bigl(\|z\|\bigr) \,dz = d \nu_d \int_{0}^{1} \psi(t) t^{d - 1} \,dt = \nu_d \int_{0}^{1} \psi'(t) t^{d} \,dt = \frac{\nu_d}{\sigma_K} \int_{0}^{1} t^{d + 1} K(t) \,dt;
\end{align*}
on the other hand, again converting to polar coordinate, we have
\begin{equation*}
\sigma_K = \frac{1}{d} \int_{\Reals^d} \|x\|^2 K(\|x\|) \,dx = \nu_d \int_{0}^{1}t^{d + 1} K(t) \,dt.
\end{equation*}

Finally, we upper bound $\|\nabla\tau(x)\|^2$. Exchanging derivative and integral, we have
\begin{align*}
\nabla\tau(x) = \frac{1}{r^d} \int_{\Xset} \bigl(\nabla \eta_r(x',\cdot)\bigr)(x) \,dx' = \frac{1}{\sigma_K r^{d + 2}} \int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr)(x' - x)\,dx',
\end{align*}
and by the Cauchy-Schwarz inequality,
\begin{equation*}
\|\nabla\tau(x)\|^2 \leq \frac{1}{\sigma_K^2 r^{2d + 4}} \biggl[\int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr)\,dx'\biggr] \int_{\Xset} K\biggl(\frac{\|x' - x\|}{r}\biggr)\|x' - x\|^2\,dx', \leq \frac{1}{\sigma_K r^{2}}
\end{equation*}
concluding the proof of Lemma~\ref{lem:tau_bound}. One note: while $\nabla\tau(x) = 0$ when $B(x,r) \in \Xset$, near the boundary the upper bound we derived by using Cauchy-Schwarz appears tight.

\paragraph{Proof of Lemma~\ref{lem:smoothening_error}.}
\mbox{}\\
\mbox{}\\
By Jensen's inequality and Lemma~\ref{lem:tau_bound},
\begin{align*}
\Bigl[\Lambda_rf(x)\Bigr]^2 & \leq \frac{1}{r^d \tau(x)}\int_{\Xset} \eta_r(x',x) \bigl[f(x')\bigr]^2 \,dx' \\
& \leq \frac{2}{r^d p_{\min}}\int_{\Xset} \eta_r(x',x) \bigl[f(x')\bigr]^2 p(x') \,dx'
\end{align*}
Then integrating over $x$ and swapping the order of integraton, we have
\begin{align*}
\bigl\|\Lambda_rf\bigr\|_{\Leb^2(P)}^2 & \leq \frac{2}{r^d p_{\min}} \int_{\Xset} \int_{\Xset} \eta_r(x',x) \bigl[f(x')\bigr]^2 p(x') p(x) \,dx' \,dx \\ 
& \leq \frac{2p_{\max}}{p_{\min}} \int_{\Xset} \bigl[f(x')\bigr]^2 p(x') \,dx' \\
& = \frac{2p_{\max}}{p_{\min}} \|f\|_{\Leb^2(P)}^2.
\end{align*}

Noting that $\Lambda_ra = a$ for any $a \in \Reals$, by the Cauchy-Schwarz inequality we have that for all $x \in \Xset$,
\begin{align*}
\bigl|\Lambda_rf(x) - f(x)\bigr|^2 & = \biggl[\frac{1}{r^d\tau(x)} \int_{\Xset} \eta_r(x',x) \bigl(f(x') - f(x)\bigr) \,dx'\biggr]^2 \\
& \leq \frac{1}{r^{2d} \tau^2(x)} \biggl[\int_{\Xset} \eta_r(x',x) \,dx'\biggr] \cdot \biggl[\int_{\Xset} \eta_r(x',x) \bigl(f(x') - f(x)\bigr)^2 \,dx'\biggr] \\
& = \frac{1}{r^d \tau(x)} \int_{\Xset} \eta_r(x',x) \bigl(f(x') - f(x)\bigr)^2 \,dx'. \\
& \leq \frac{1}{r^d \tau(x) p_{\min}} \int_{\Xset} \eta_r(x',x) \bigl(f(x') - f(x)\bigr)^2 p(x') \,dx'.
\end{align*}
Then integrating over $\Xset$ with respect to $p$ yields~\eqref{eqn:smoothening_error_energy}.

\clearpage

\section{Miscellaneous}

\paragraph{Proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb} (simplified setting).}
\mbox{} \\
\mbox{} \\
Assume $d = 1$ and $\Xset = [0,1]$. For any $a \in \Reals$,
\begin{equation*}
\Lambda_rf(x) = a + \frac{1}{r\tau(x)}\int_{0}^{1} \eta_r(x',x)\bigl(f(x') - a\bigr)\,dx'
\end{equation*}
Differentiating with respect to $x$, we obtain
\begin{equation*}
\frac{d}{dx}\Lambda_rf = \frac{1}{r\tau(x)}\int_{0}^{1} \frac{\partial}{\partial x}\eta_r(x',x)\bigl(f(x') - a\bigr)\,dx' + \frac{d}{dx}\bigl(\tau^{-1}(x)\bigr) \frac{1}{r}\int_{0}^{1} \eta_r(x',x)\bigl(f(x') - a\bigr)\,dx'
\end{equation*} 
and plugging in $a = f(x)$, we get $\frac{d}{dx}\Lambda_rf(x) = J_1(x)/\tau(x) + J_2(x)$ for
\begin{equation*}
J_1(x) := \frac{1}{r}\int_{0}^{1} \frac{\partial}{\partial x}\eta_r(x',x)\bigl(f(x') - f(x)\bigr)\,dx',~~ J_2(x) := \frac{d}{dx}\bigl(\tau^{-1}(x)\bigr)\frac{1}{r}\int_{0}^{1} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)\,dx'.
\end{equation*}
To upper bound $\bigl|J_1(x)\bigr|^2$, we first compute the partial derivative
\begin{align*}
\frac{\partial}{\partial x}\eta_r(x',x) & = \frac{1}{r} \psi'\biggl(\frac{|x'  - x|}{r}\biggr) \cdot \frac{(x - x')}{\|x' - x\|} \\
& = \frac{1}{\sigma_Kr^2} K\biggl(\frac{\|x' - x\|}{r}\biggr) \cdot (x' - x),
\end{align*}
so that by the Cauchy-Schwarz inequality,
\begin{align*}
\bigl|J_1(x)\bigr|^2 & = \frac{1}{\sigma_K^2 r^6} \Biggl[\int_{0}^{1} K\biggl(\frac{\|x' - x\|}{r}\biggr) \cdot (x' - x) \cdot \bigl(f(x') - f(x)\bigr)\Biggr]^2 \,dx' \\
& \leq \frac{1}{\sigma_K^2r^6} \biggl[\int_{0}^{1}K\biggl(\frac{\|x' - x\|}{r}\biggr)(x' - x)^2\,dx'\biggr] \cdot \biggl[\int_{0}^{1}K\biggl(\frac{\|x' - x\|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2\,dx'\biggr] \\
& \leq \frac{1}{\sigma_K r^3}\int_{0}^{1}K\biggl(\frac{\|x' - x\|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2\,dx'.
\end{align*}
To upper bound $\bigl|J_2(x)\bigr|^2$, we note that $\eta_r(x',x) \leq \frac{1}{\sigma_K} K_r(x',x)$. Then by the Cauchy-Schwarz inequality again,
\begin{align*}
\bigl|J_2(x)\bigr|^2 & \leq \biggl[\frac{d}{dx}\bigl(\tau^{-1}(x)\bigr)\biggr]^2\frac{1}{r^2} \biggl[\int_{0}^1\eta_r(x',x) \,dx'\biggr] \cdot \biggl[\int_{0}^{1} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr] \\
& \leq \biggl[\frac{d}{dx}\bigl(\tau^{-1}(x)\bigr)\biggr]^2\frac{\tau(x)}{r}\biggl[\int_{0}^{1} \eta_r(x',x)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr]\\ 
& \leq \biggl[\frac{d}{dx}\bigl(\tau^{-1}(x)\bigr)\biggr]^2\frac{\tau(x)}{\sigma_K r}\biggl[\int_{0}^{1} K\biggl(\frac{|x' - x|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr] \\
& \leq \frac{8}{\sigma_Kr^3} \biggl[\int_{0}^{1} K\biggl(\frac{|x' - x|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 \,dx' \biggr].
\end{align*}
where the last inequality follows by Lemma~\ref{lem:tau_bound}. Combining our bounds on $\bigl|J_1(x)\bigr|^2$ and $\bigl|J_2(x)\bigr|^2$, we have
\begin{align*}
\sigma_K D_2(\Lambda_r f) & = \sigma_K\int_{0}^{1} \biggl|\frac{d}{dx} \Lambda_rf(x)\biggr|^2 p^2(x) \,dx \\
& \leq \frac{18}{r^3} \int_{0}^{1} \int_{0}^{1} K\biggl(\frac{|x' - x|}{r}\biggr)\bigl(f(x') - f(x)\bigr)^2 p^2(x) \,dx' \,dx \\
& = 18 E_r(f),
\end{align*}
completing the proof of Lemma~\ref{lem:first_order_graph_sobolev_seminorm_expected_lb}. 
\paragraph{Proof of Lemma~\ref{lem:tau_bound} (simplified setting).} 
\mbox{} \\
\mbox{} \\
Assume $d = 1$ and $\Xset = [0,1]$. First we show the lower bound on $\tau(x)$, then the upper bound on $\tau(x)$, and finally the upper bound on $|\tau'(x)|$. Substituting $z = (x' - x)/r$ and integrating by parts, we have
\begin{align*}
\tau(x) & = \frac{1}{r} \int_{0}^{1} \psi\biggl(\frac{|x' - x|}{r}\biggr) \,dx' \\
& =  \int_{-x/r}^{(1 - x)/r}  \psi\bigl(\abs{z}\bigr) \,dz \\
& = z \psi\bigl(|z|\bigr) \bigg|_{-x/r}^{(1 - x)/r} - \int_{-x/r}^{(1 - x)/r}  z \psi'\bigl(\abs{z}\bigr) \,dz \\
& \geq \frac{1}{\sigma_K}\int_{-x/r}^{(1 - x)/r}  z^2 K\bigl(|z|\bigr) \,dz. \\
\end{align*}
By our smallness assumption on $r$, one of $x > r$ or $(1 - x) > r$ must be true; we will assume without loss of generality that $(1 - x) > r$. In that case,
\begin{equation*}
\int_{-x/r}^{(1 - x)/r}  z^2 K\bigl(|z|\bigr) \,dz \geq \int_{0}^{1} z^2K\bigl(|z|\bigr) = \frac{1}{2} \int_{-1}^{1} z^2K\bigl(|z|\bigr) \,dz = \frac{1}{2}\sigma_K,
\end{equation*}
so that $\tau(x) \geq 1/2$.

The upper bound on $\tau(x)$ follows along similar lines. Again substituting $z = (x' - x)/r$ and integrating by parts, we have
\begin{align*}
\tau(x) & = \frac{1}{r} \int_{0}^{1} \psi\biggl(\frac{|x' - x|}{r}\biggr) \,dx' \\
& \leq \int_{-\infty}^{\infty}  \psi\bigl(\abs{z}\bigr) \,dz \\
& = \frac{1}{\sigma_K}\int_{-\infty}^{\infty} z^2 K(z) \,dz = 1.
\end{align*}

Finally, we upper bound $|\tau'(x)|$:
\begin{align*}
\abs{\tau'(x)} & = \frac{1}{r} \int_{0}^{1} \frac{\partial}{\partial x} \eta_r(x',x) \,dx' \\
& = \frac{1}{\sigma_Kr^{3}} \int_{0}^{1} K\biggl(\frac{\|x' - x\|}{r}\biggr) \cdot (x' - x) \,dx' \\
& = \frac{1}{\sigma_Kr} \int_{-x/r}^{(1 - x)/r} z K\bigl(|z|\bigr) \,dz \\
& \leq \frac{1}{\sigma_Kr} \biggl[\int_{-\infty}^{\infty} K\bigl(|z|\bigr) \,dz\biggr] \cdot \biggl[\int_{-\infty}^{\infty}z^2 K\bigl(|z|\bigr) \,dz \biggr] \\
& \leq \frac{1}{r}.
\end{align*}
The first upper bound in the previous display follows from the Cauchy-Schwarz inequality. While much tighter upper bounds are possible when $B(x,r) \subseteq \Xset$---indeed $\tau'(x) = 0$ for such $x$---near the boundary this inequality is tight.

\clearpage

\bibliographystyle{plainnat}
\bibliography{../../graph_testing_bibliography}

\end{document}