### 2020-10-07 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing3.tex
Begun a 2nd attempt at AISTATS paper on Laplacian smoothing.

\documentclass[twoside]{article}
\usepackage{aistats2021}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{mathtools}
\usepackage{tikz}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
\DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\iid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}
\newcommand{\lambdavec}{\boldsymbol{\lambda}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}
\newcommand{\Lap}{L}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}
\newcommand{\Id}{I}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Vset}{\mathcal{V}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\SM}{\mathrm{SM}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

%%% Theorem environments
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}

\twocolumn[

	\aistatstitle{Minimax optimal Laplacian smoothing}

	\aistatsauthor{Alden Green, Sivaraman Balakrishnan, and Ryan J. Tibshirani}

	\aistatsaddress{Carnegie Mellon University}]

\begin{abstract}
	In this paper we study the statistical properties of Laplacian smoothing, a graph-based approach to non-parametric regression. We show that under standard regularity conditions, the Laplacian smoothing estimator $\wh{f}$, and a test using the statistic $T = \frac{1}{n}\|\wh{f}\|_2^2$, attain the minimax optimal estimation and testing rates of convergence over the first-order Sobolev class $H^1(\Xset)$, where $\Xset \subset \Rd$ and $1 \leq d \leq 4$. Additionally, we prove that Laplacian smoothing is manifold adaptive: if $\Xset$ is an $m$ dimensional manifold embedded in $\Reals^d$ with $m \ll d$, then the error rate of Laplacian smoothing methods depends only on the intrinsic dimension $m$ and not on the ambient dimension $d$. We support our theory with simulations.
\end{abstract}

\section{INTRODUCTION}

Suppose we observe responses according to the signal plus noise model:
\begin{equation}
\label{eqn:signal_plus_noise_model}
Y_i = f_{0,i} + \varepsilon_i
\end{equation}
with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise. Our goal is to perform statistical inference on the unknown regression function $f_0$, by which we mean either \emph{estimating} $f_0$ or more simply \emph{testing} whether $f_0 = 0$, i.e whether there is any signal present.

The Laplacian smoothing estimator $\wh{f}$ \citep{smola2003} is a penalized least squares estimator, defined on a graph. Letting $G = \bigl(\{1,\ldots,n\},\mathbf{W}\bigr)$ be a weighted graph over the vertices $\{1,\ldots,n\}$, 
\begin{equation}
\label{eqn:laplacian_smoothing}
\wh{f} :=  \min_{f \in \Reals^n} \biggl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap f \biggr\},
\end{equation}
where $\Lap$ is the graph Laplacian matrix (defined in Section 2), and the penalty
\begin{equation*}
f^T \Lap f = \frac{1}{2} \sum_{i,j = 1}^{n} {\bf W}_{ij}\bigl(Y_i - Y_j\bigr)^2
\end{equation*}
encourages $\wh{f}_i \approx \wh{f}_j$ when ${\bf W}_{ij}$ is large. Assuming~\eqref{eqn:laplacian_smoothing} is a reasonable estimator of $f_0$, the squared empirical norm
\begin{equation}
\label{eqn:laplacian_smoothing_test}
\wh{T} = \frac{1}{n}\bigl\|\wh{f}\bigr\|_2^2 
\end{equation}
is in turn a reasonable statistic to assess whether or not $f_0 = 0$.

Of course there exist many methods for nonparametric regression, but Laplacian smoothing has its own set of advantages. For instance:
\begin{itemize}
	\item \emph{Computational ease.} Laplacian smoothing is fast, easy, and stable to compute. The estimate $\wh{f}$ can be computed by solving a symmetric, diagonally dominant linear system. There are known extremely fast and reliable solvers for exactly this problem \textcolor{red}{(TODO)}: references.
	\item \emph{Generalizability.} Laplacian smoothing is well defined whenever one can associate a graph with some observed responses. This generality lends itself to many different data modalities: for instance sentiment analysis and image classification, to name only two. \textcolor{red}{(TODO)}: references
	\item \emph{Supervision.} Although we study Laplacian smoothing in the supervised learning setup~\eqref{eqn:signal_plus_noise_model}, it can be easily adapted to the semi-supervised or unsupervised settings.
\end{itemize}

For these reasons, a body of work has emerged analyzing the statistical properties of Laplacian smoothing, and graph-based methods more generally. Roughly speaking, these works can be divided into two categories, based on the perspective they adopt. 
\begin{itemize}
	\item In the \emph{discrete (fixed design) perspective}, both $G$ and $f_0 \in \Reals^n$ are treated as fixed, and the regression ``function'' is really a vector. In this context, tight upper bounds have been derived on the error of graph-based estimators, which certify that such estimators are \emph{optimal} over ``function'' classes $\mc{F} \subseteq \Reals^n$. The downside of this work is that, depending on the context, it may be unnatural to assume a priori that the graph $G$ is fixed, and that the signal $f_0$ belongs to some discrete ``function'' class.
	\item In the arguably more natural \emph{continuum (random design) perspective}, in addition to the responses $Y_i$ one observes randomly sampled design points $\{X_1,\ldots,X_n\}$ belonging to a domain $\Xset \subseteq \Rd$, and assumes that $f_{0,i} \equiv f_0(X_i)$ are $n$ evaluations of a regression function $f_0: \Xset \to \Reals$. The user then builds a neighborhood graph over the design points, so that ${\bf W}_{ij}$ is large iff $X_i$ and $X_j$ are close, say in Euclidean distance. In this context, various graph-based estimators have been shown to be \emph{consistent}: as $n \to \infty$, the estimator converges to $f_0$ in an appropriate norm, assuming $f_0$ satisfies some notion of smoothness such as possessing a bounded derivative. However, until recently such statements were not accompanied by error rates, and such error rates as have been proved are not optimal.
\end{itemize}
Viewed as a whole, this body of work leaves a key question unanswered: what are the optimality properties of graph Laplacian smoothing, viewed as an estimator of a (continuum) regression function? This is no small point---rather, it is arguably the fundamental question of nonparametric regression, and without an answer one cannot fully compare the statistical properties of Laplacian smoothing to its competitors.

Our work addresses this gap, and our main contributions can all be summarized as follows: Laplacian smoothing over a random design is minimax optimal over certain first-order \emph{continuum} Sobolev balls. We will show that this is true for the estimator $\wh{f}$ and test statistc $\wh{T}$, when $\Xset \subset \Rd$ is a full-dimensional domain and $1 \leq d \leq 4$. Additionally, we will consider the \emph{manifold assumption}---roughly, that $\Xset$ is a manifold embedded in $\Rd$, with intrinsic dimension $m \ll d$---and derive upper bounds which depend only on the intrinsic dimension $m$, and are independent of the ambient dimension $d$; these bounds too will be optimal. For ease of reading, we provide the upper bounds derived in our main theorems in Table~\textcolor{red}{(?)}.

\paragraph{Organization.}
We now outline the structure of the rest of this paper. In Section~\ref{sec:problem_setup_and_background}, we formalize our model of nonparametric regression with random design, recall the minimax rates for the estimation and goodness-of-fit testing over Sobolev spaces, and review some related work. In Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we give our main results on the optimality of Laplacian smoothing methods, offering some comparison with other nonparametric regression approaches as well as giving insight into our proof techniques. In Section~\ref{sec:simulations} we support our theory with some experiments, before concluding in Section~\ref{sec:discussion}.

\textcolor{red}{(TODO)}:

\section{PROBLEM SETUP AND BACKGROUND}
\label{sec:problem_setup_and_background}
Formally speaking, In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Reals^d$, and 
\begin{equation}
\label{eqn:random_design_regression}
Y_i = f_0(X_i) + \varepsilon_i
\end{equation}
with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise. We will assume the following throughout:
\begin{enumerate}[label=(P\arabic*)]
	\item
	\label{asmp:bounded_lipschitz_density} 
	$P$ admits a density $p$ bounded away from $0$ and $\infty$, i.e.
	\begin{equation*}
	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty,~~\textrm{for all $x \in \Xset$.}
	\end{equation*}
	Additionally, $p$ is Lipschitz, with Lipschitz constant $L_p$.
\end{enumerate}

\paragraph{Neighborhood graph Laplacians.}
As mentioned, in the graph-based approach to nonparametric regression, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense. The neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$. The $n \times n$ weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ encodes proximity between pairs of samples; for a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the entries $\mathbf{W}_{ij}$ are given by
\begin{equation*}
\label{eqn:neighborhood_graph}
{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}}{r}\Biggr).
\end{equation*}
where $\norm{\cdot} = \norm{\cdot}_{\Rd}$ is the Euclidean norm. Then the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$, and the graph Laplacian can be written as
\begin{equation}
\label{eqn:graph_Laplacian}
\Lap_{n,r} = \bf{D} - \bf{W};
\end{equation}
we add the subscripts $n$ and $r$ to $\Lap$ and $G$ to emphasize that they depend on the random design points $\bf{X}$ and the radius $r$.

\paragraph{Minimax rates over Sobolev spaces.}
Now that our estimator $\wh{f}$ and test statistic $\wh{T}$ are defined in the random design setup, we step away from these methods---indeed, away from graphs entirely--- for a moment, to briefly recall some very classical results regarding minimax rates over Sobolev classes.

\paragraph{Previous work.}

\textcolor{red}{(TODO)}

\section{MINIMAX OPTIMALITY OF LAPLACIAN SMOOTHING}
\label{sec:minimax_optimal_laplacian_smoothing}

\textcolor{red}{(TODO)}: Section 4 of the previous draft.

\paragraph{Comparison with the continuous-time approach.}

\textcolor{red}{(TODO)}: Take some of section 4, and a little of section 2, to do this paragraph.

\paragraph{Overview of analysis.}
\begin{itemize}
	\item Speaking in gross generality, there are two ways in which one can analyze nonparametric regression methods. 
	\item Way (1) --- by far the more common approach to analyzing estimators defined as solutions to variational problems---is to use \emph{empirical process theory} as the workhorse. In other words, one bounds the supremum of the empirical process, i.e.
	\begin{equation}
	\label{eqn:empirical_process}
	\sup_{f \in H^1(\Xset)} \abs{\langle \varepsilon, f \rangle} \lesssim \delta_n
	\end{equation}
	and then uses~\eqref{eqn:empirical_process} to guarantee that the estimator in question will not overfit to the noise $\varepsilon$. While this can give the right rate when $d = 1$, it will totally fail when $d \geq 2$. 
	\item Way (2) is the bias-variance tradeoff. Typically this is not tractable for an estimators such as $\wh{f}_{\LS}$ that is defined as the solution to a variational problem. However the problem~\textcolor{red}{(?)} is special: it has a nice closed-form solution that is amenable to a bias-variance decomposition. We therefore do not need to rely on empirical process theory, and so we will be able to analyze $\wh{f}_{\LS}$ and $T_{\LS}$ beyond the univariate setting.
\end{itemize}

\section{SIMULATIONS}
\label{sec:simulations}

\section{DISCUSSION}
\label{sec:discussion}

\bibliographystyle{plainnat}
\bibliography{../../../graph_testing_bibliography}

\end{document}