### 2020-10-26 notes/weekly_notes/10.28.20.NotesForWeek.tex
Added a note outlining where we get stuck in our proof of a lower bound for Laplacian smoothing.

\documentclass{article}
\usepackage{amsmath, amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{verbatim}
\usepackage{fullpage}

\usepackage{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
\DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}

% Make a widecheck symbol (thanks, Stack Exchange!)
\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{
	<5> <6> <7> <8> <9> <10>
	<10.95> <12> <14.4> <17.28> <20.74> <24.88>
	mathx10
}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareFontSubstitution{U}{mathx}{m}{n}
\DeclareMathAccent{\widecheck}{0}{mathx}{"71}
% widecheck made

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\iid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\dagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}
\newcommand{\lambdavec}{\boldsymbol{\lambda}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Dgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\dagger}}
\newcommand{\Lap}{L}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}
\newcommand{\Id}{I}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Vset}{\mathcal{V}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\dive}{\mathrm{div}}
\newcommand{\dif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\dx}{\,dx}
\newcommand{\dy}{\,dy}
\newcommand{\dr}{\,dr}
\newcommand{\dxpr}{\,dx'}
\newcommand{\dypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\SM}{\mathrm{SM}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

%%% Theorem environments
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Note on: Lower bound for Laplacian smoothing on Geometric Graphs}
\author{Alden Green}
\date{\today}
\maketitle

Suppose we observe design points $X_1,\ldots,X_n$, with associated responses
\begin{equation*}
Y_i = f_0(X_i) + \varepsilon_i;
\end{equation*}
here each $X_i$ belongs to a domain $\Xset$, the regression function $f_0: \Xset \to \Reals$, and $\varepsilon_i \sim N(0,1)$ are independent Gaussian noise. Let $G = (\{1,\ldots,n\},W)$ be a geometric graph defined over the vertices $\{1,\ldots,n\}$, which we associate with the design points $X_1,\ldots,X_n$, where $W_{ij} = 0$ unless $\|X_i - X_j\| \leq r$.  Let $\Lap$ be the (combinatorial) Laplacian matrix associated with $G$. The Laplacian smoothing estimator $\wh{f}$ is defined as
\begin{equation*}
\wh{f} := \argmin_{f \in \Reals^n} \Bigl\{\sum_{i = 1}^{n} (Y_i - f_0(X_i))^2 + \rho f^T \Lap f\Bigr\}.
\end{equation*}
The accuracy of $\wh{f}$ is measured by the in-sample mean squared error
\begin{equation*}
\|\wh{f} - f_0\|_n^2 := \frac{1}{n} \sum_{i = 1}^{n} \bigl(\wh{f}_i - f_0(X_i)\bigr)^2.
\end{equation*}

Suppose we make the following~\emph{random design} assumptions.
\begin{enumerate}[label=(P\arabic*)]
	\item
	\label{asmp:domain}
	 The domain $\Xset$ is an open, connected set, with Lipschitz boundary.
	\item 
	\label{asmp:density}
	The design points $X_1,\ldots,X_n$ are independently drawn by a distribution $P$ defined on $\Xset$. The distribution $P$ admits a density $p$ with respect to Lebesgue measure, which is bounded away from $0$ and $\infty$,
	\begin{equation*}
	0 < \frac{1}{\alpha} \leq p(x) \leq \alpha < \infty,~~\textrm{for all $x \in \Xset$.}
	\end{equation*}
\end{enumerate}
\begin{enumerate}[label=(G\arabic*)]
	\item 
	\label{asmp:geometric_graph}
	The graph $G$ is an unweighted \emph{random geometric graph}: $W_{ij} = 1$ if $\|X_i - X_j\| \leq r$ and $0$ otherwise.
\end{enumerate}
\begin{enumerate}[label=(F\arabic*)]
	\item 
	\label{asmp:sobolev_norm}
	The regression function $f_0$ belongs to the unit-norm first-order Sobolev ball $H^1(\Xset,1)$, meaning
	\begin{equation}
	\label{eqn:sobolev_norm}
	\int_{\Xset} \|\nabla f_0(x)\|^2 \,dx \leq 1
	\end{equation}
\end{enumerate}
We have derived the following upper bounds on the in-sample mean squared error of $\wh{f}$: with probability at least $1 - \delta$,
\begin{equation*}
\|\wh{f} - f_0\|_n^2 \leq C_{\delta}
\begin{cases*}
n^{-2/(2 + d)},& ~~\textrm{if $d = 1,2,3$} \\
\Bigl(\log n/n\Bigr)^{1/3},&~~\textrm{if $d = 4$} \\
n^{-4/(3d)},& ~~\textrm{if $d \geq 5$.}
\end{cases*}
\end{equation*}
These upper bounds fail to match the known minimax rate $n^{-2/(2 + d)}$ when $d \geq 4$. Our question is: is it our estimator, or our theory, that is deficient when $d \geq 4$? Simulations, and theory from the related fixed design problem, suggest it is the estimator. In this note, I detail this fixed design theory, and show the difficulty in applying similar reasoning to the random design case.

\section{Lower bounds in the fixed design setup}
In the fixed design setting, we assume $X_1,\ldots,X_n$ are a priori fixed, so that the only randomness comes from the noise variables $\varepsilon_i$ (one can do a better job of formalizing this assumption, but for the moment I won't bother.) In the fixed design setup, we make different assumptions.
\begin{enumerate}[label=(P\arabic*)]
	\setcounter{enumi}{2}	
	\item  
	\label{asmp:grid}
	The design points form a grid over the $d$-dimensional unit cube $[0,1]^d$. Adopting multiindex notation for conciseness, we have that for $N = n^{1/d}$, the design points are the set
	\begin{equation*}
	\Bigl\{X_j := \frac{1}{N}(j_1,\ldots,j_d): j \in [N]^d\Bigr\}
	\end{equation*}
\end{enumerate}

\begin{enumerate}[label=(G\arabic*)]
	\setcounter{enumi}{1}
	\item 
	\label{asmp:lattice}
	The graph $G$ is an unweighted lattice graph, with $W_{ij} = 1$ if $\|i - j\|_1 = 1$, and $0$ otherwise.
\end{enumerate}

\begin{enumerate}[label=(F\arabic*)]
	\setcounter{enumi}{1}
	\item 
	\label{asmp:graph_sobolev_norm}
	The regression function $f_0: \{X_1,\ldots,X_n\} \to \Reals$ satisfies
	\begin{equation*}
	\frac{1}{n^2\wb{r}^{d + 2}} f_0^{\top} \Lap f_0 \leq 1
	\end{equation*}
	for $\wb{r} := n^{-1/d}$.
\end{enumerate}
(NB: the grid and lattice assumption are not essential. We could likely replace the lattice by a general geometric graph of radius $r$, replace $\wb{r}$ by $r$, and replace the grid assumption by a Weyl's-Law-type assumption on the resulting eigenvalues of $G$. But the above is simpler, and gets the main idea across.)

We shall prove the following when $d > 4$: For any $\rho > 0$, there exists a function $f_0 \in \Reals^n$ satisfying~\ref{asmp:graph_sobolev_norm} such that
\begin{equation}
\label{eqn:lb_mse}
\mathbb{E}\bigl[\|\wh{f} - f_0\|_n^2\bigr] \gtrsim n^{-4/(3d)} \gg n^{-2/(d + 2)}.
\end{equation}

\paragraph{Proof of~\eqref{eqn:lb_mse}.}
Our strategy to prove this claim parallels the development of the corresponding upper bound. In the fixed design setup, the expected in-sample mean squared error of $\wh{f}$ can be decomposed into squared bias and variance:
\begin{align*}
\Ebb\bigl[\|\wh{f} - f_0\|_n^2\bigr] & = \underbrace{\|\Ebb[\wh{f}] - f_0]\|_n^2}_{\mathbb{B}(\wh{f},f_0)} + \underbrace{\Ebb\bigl[\|\wh{f} - \Ebb[\wh{f}]\|_n^2\bigr]}_{\mathbb{V}(\wh{f})}
\end{align*}
and the task reduces to lower bounding of lower bounding mean squared error reduces to lower bounding the bias and variance terms.

In the following two subsections, we give lower bounds on $\mathbb{B}(\wh{f},f_0)$ and $\mathbb{V}(\wh{f})$. The bounds take different forms, depending on which of the following situations we are in: (i) there exists $k \in \{2,\ldots,n/2 - 1\}$ such that $\rho \lambda_k = 1$, (ii) $\rho \lambda_2 > 1$, or (iii) $\rho \lambda_n < 1$. If we are in the first situation, then by Lemmas~\ref{lem:lb_bias} and~\ref{lem:lb_variance} there exists a choice of $f_0$ (which we make explicit in the statement of Lemma~\ref{lem:lb_bias}) such that
\begin{equation*}
\Ebb\bigl[\|\wh{f} - f_0\|_n^2\bigr] \geq c\biggl(\wb{\rho} + \frac{1}{\wb{\rho}^2}n^{-4/d}\biggr)
\end{equation*}
and a simple computation tells us that the right hand side is always greater than $cn^{-4/(3d)}$. 
If we are in the second situation, then by Lemma~\ref{lem:lb_bias} $\Ebb\bigl[\|\wh{f} - f_0\|_n^2\bigr] \gtrsim 1$; if we are in the third situation, then by Lemma~\ref{lem:lb_variance} the same thing is true, and so we have proved the claim.

(Of course, we have not handled the case where $\rho\lambda_k < 1 < \rho\lambda_{k + 1}$, but in spirit this is basically situation (i)---it should only introduce some negligible rounding error which I am not going to bother calculating right now.)

\subsection{Lower bound on Bias}
Let $S := (I + \rho \Lap)^{-1}$, and note that Laplacian smoothing can be written as $\wh{f} = S Y_{1:n}$ (writing $Y_{1:n} = (Y_1,\ldots,Y_n)$), and the bias of Laplacian smoothing as
\begin{equation}
\label{eqn:bias}
\mathbb{B}(\wh{f},f_0) = \frac{1}{n}f_0^{\top}(I - S)^2 f_0
\end{equation}
For $\rho > 0$, set $\wb{\rho} := \rho \cdot n\wb{r}^{d + 2}$.

\begin{lemma}
	\label{lem:lb_bias}
	Fix tuning parameter $\rho > 0$. 
	\begin{itemize}
		\item Suppose there exists a number $k \in \{1,\ldots,n\}$ such that $\rho \lambda_k = 1$. Then, setting $f_0$ equal to $\sqrt{\frac{n^2 \wb{r}^{d + 2}}{\lambda_k}}v_k$, we have that $(n^2\wb{r}^{d + 2})^{-1}f_0^{\top}\Lap f_0 = 1$, and 
		\begin{equation}
		\label{eqn:lb_bias_1}
		\mathbb{B}(\wh{f},f_0) = \frac{1}{4}\wb{\rho}
		\end{equation}
		\item Suppose $\lambda_1 \rho > 1$. Then, setting $f_0$ equal to $\sqrt{\frac{n^2 \wb{r}^{d + 2}}{\lambda_k}}v_1$, we have that $(n^2\wb{r}^{d + 2})^{-1}f_0^{\top}\Lap f_0 = 1$, and 
		\begin{equation}
		\label{eqn:lb_bias_2}
		\mathbb{B}(\wh{f},f_0) \geq c.
		\end{equation}
	\end{itemize}
\end{lemma}
Note: to properly trade off bias and variance when $d < 4$, we set $\wb{\rho} = n^{-2/(2 + d)}$. In this case, $\rho \lambda_k = 1$ implies that $k \asymp n^{d/(d + 2)}$, so that our choice of $f_0$ which saturates the upper bound $f_0^{\top} \Lap f_0 \lesssim \wb{\rho}$ (i.e the choice of $f_0$ for which~\eqref{eqn:lb_bias_1} holds) is a harmonic function of frequency $n^{1/(d + 2)}$. In plain English: its very wiggly.

\subsection{Lower bound on Variance}
The variance term $\mathbb{V}(\wh{f})$ is simply
\begin{equation*}
\mathbb{V}(\wh{f}) = \Ebb\bigl[\varepsilon^{\top} S^2 \varepsilon\bigr] = \frac{1}{n}\sum_{j = 1}^{n} \frac{1}{(\rho \lambda_j + 1)^2}
\end{equation*}

\begin{lemma}
	\label{lem:lb_variance}
	Fix tuning parameter $\rho > 0$. Assume $d > 4$.
	\begin{itemize}
		\item Suppose there exists a number $k \in \{1,\ldots,n/2 - 1\}$ such that $\rho \lambda_k = 1$. Then
		\begin{equation}
		\label{eqn:lb_variance_1}
		\mathbb{V}(\wh{f}) \geq c\frac{1}{\wb{\rho}^2}n^{-4/d}
		\end{equation}
		for $\wb{\rho} = (nr^{d + 2})\rho$. 
		\item Suppose instead that $\rho \lambda_{n/2 - 1} < 1$. Then
		\begin{equation}
		\label{eqn:lb_variance_2}
		\mathbb{V}(\wh{f}) \geq \frac{1}{4}.
		\end{equation}
	\end{itemize}
\end{lemma}

\section{What's tricky about a lower bound in the random design setup?}
The simple answer is: the bias. That is, it should not be hard to prove a rough analog to Lemma~\ref{lem:lb_variance} under the random design assumptions~\ref{asmp:density},~\ref{asmp:domain} and~\ref{asmp:geometric_graph}. However, to prove Lemma~\ref{lem:lb_bias} we set $f_0$ as a function of the graph $G$ (more specifically, as proportional to an eigenvector $v_k$ of $L$). In the random design setup, this is an impermissible operation, as we must choose the ``bad'' function $f_0$ before the design points $X_1,\ldots,X_n$ are drawn. Now, we could choose $f_0$ proportional to an eigenfunction $f_k$ of an appropriately weighted continuum Laplace-Beltrami operator. However, we would then need to argue that $f_k \approx v_k$, for a value of $k$ which is growing (quite rapidly) with $n$.

\section{Proofs}

\paragraph{Proof of Lemma~\ref{lem:lb_bias}.}
Take the representation~\eqref{eqn:bias}, and note that $(I - S)^2$ admits the eigendecomposition
\begin{equation*}
(I - S)^2 = \sum_{j = 1}^{n} \biggl(1 - \frac{1}{1 + \rho\lambda_j}\biggr)^2 v_j v_j^{\top} = \rho^2 \sum_{j = 1}^{n}\frac{\lambda_j^2}{(1 + \rho \lambda_j)^2} v_j v_j^{\top}
\end{equation*}
As a result, if $\lambda_k\rho = 1$, then 
\begin{equation*}
v_k^{\top} (I - S)^2 v_k = \rho^2 \frac{\lambda_k^2}{(1 + \rho \lambda_k)^2} = \frac{1}{4}
\end{equation*}
and setting $f_0 = \sqrt{\frac{n\wb{r}^{d+2}}{\lambda_k}}v_k$, we obtain
\begin{equation*}
\frac{1}{n}f_0^{\top} (I - S)^2 f_0 = \frac{n\wb{r}^{d + 2}}{\lambda_k} v_k^{\top} (I - S)^2 v_k =  \frac{1}{4}n\wb{r}^{d + 2}\rho.
\end{equation*}
On the other hand, if $\lambda_1\rho > 1$, then
\begin{equation*}
v_1^{\top} (I - S)^2 v_1  = \rho^2 \frac{\lambda_1^2}{(1 + \rho \lambda_1)^2} \geq \frac{1}{4}
\end{equation*}
and setting $f_0 = \sqrt{\frac{n\wb{r}^{d+2}}{\lambda_1}}v_1$ we obtain
\begin{equation*}
\frac{1}{n} f_0^{\top} (I - S)^2 f_0 \geq \frac{n\wb{r}^{d + 2}}{4\lambda_1} \geq c.
\end{equation*}

\paragraph{Proof of Lemma~\ref{lem:lb_variance}.}
The inequality~\eqref{eqn:lb_variance_2} is immediate.

To prove the inequality~\eqref{eqn:lb_variance_1}, we note that by assumption
\begin{equation*}
1 + \lambda_j\rho \leq 
\begin{cases*}
2,& ~~\textrm{for all $j < k$} \\
2\lambda_j\rho, & ~~\textrm{for all $j \geq k$.}
\end{cases*}
\end{equation*}
and therefore
\begin{equation}
\label{pf:lb_variance_1}
\mathbb{V}(\wh{f}) = \frac{1}{n}\sum_{j = 1}^{n} \frac{1}{(\rho \lambda_j + 1)^2} \geq \frac{1}{4n} \biggl( k - 1 + \sum_{j = k}^{n} \frac{1}{\rho^2 \lambda_j^2} \biggr)
\end{equation}
To lower bound the tail sum, we rely on the fact that for all $j = 1,\ldots,n$,
\begin{equation*}
\lambda_j \leq C j^{-2/d} n\wb{r}^{d + 2}
\end{equation*}
and so $\rho \lambda_j \leq C j^{-2/d} \wb{\rho}$. Therefore,
\begin{align*}
\sum_{j = k}^{n} \frac{1}{\rho^2 \lambda_j^2} & \geq \frac{c}{\wb{\rho}^2} \sum_{j = k}^{n} j^{-4/d} \\
& \geq  \frac{c}{\wb{\rho}^2} \int_{k + 1}^{n} x^{-4/d} \,dx \\
& = \frac{c}{\wb{\rho}^2}\biggl[n^{1 - 4/d} - (k + 1)^{4/d - 1}\biggr] \\
& \geq \frac{c}{\wb{\rho}^2}n^{1 - 4/d}
\end{align*}
and by plugging back in to~\eqref{pf:lb_variance_1} we obtain~\eqref{eqn:lb_variance_1}.

\end{document}