### 2021-04-27 notes/misc/splines_graph_laplacians_connections.tex
Began a short note regarding connections between Laplacian smoothing splines and graph Laplacian smoothing.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	linkcolor=blue,
	citecolor=blue
}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{bm}
\usepackage{multirow}
\usepackage[font={small,it}]{caption}

\usepackage{natbib}
\renewcommand{\bibname}{References}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\DeclareFontFamily{U}{mathx}{\hyphenchar\font45}
\DeclareFontShape{U}{mathx}{m}{n}{<-> mathx10}{}
\DeclareSymbolFont{mathx}{U}{mathx}{m}{n}
\DeclareMathAccent{\wb}{0}{mathx}{"73}
\DeclareMathAccent{\wc}{0}{mathx}{"71}

\DeclarePairedDelimiterX{\norm}[1]{\lVert}{\rVert}{#1}
\DeclarePairedDelimiterX{\seminorm}[1]{\lvert}{\rvert}{#1}

\newcommand{\eqdist}{\ensuremath{\stackrel{d}{=}}}
\newcommand{\Graph}{\mathcal{G}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Identity}{\mathbb{I}}
\newcommand{\Xsetistiid}{\overset{\text{i.i.d}}{\sim}}
\newcommand{\convprob}{\overset{p}{\to}}
\newcommand{\convdist}{\overset{w}{\to}}
\newcommand{\Expect}[1]{\mathbb{E}\left[ #1 \right]}
\newcommand{\Risk}[2][P]{\mathcal{R}_{#1}\left[ #2 \right]}
\newcommand{\Prob}[1]{\mathbb{P}\left( #1 \right)}
\newcommand{\iset}{\mathbf{i}}
\newcommand{\jset}{\mathbf{j}}
\newcommand{\myexp}[1]{\exp \{ #1 \}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\restr}[2]{\ensuremath{\left.#1\right|_{#2}}}
\newcommand{\ext}[1]{\widetilde{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\set{#1}_{n \in \N}}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\diam}{\mathrm{diam}}

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

\newcommand{\Linv}{L^{\Xsetagger}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\newcommand{\emF}{\mathbb{F}_n}
\newcommand{\emG}{\mathbb{G}_n}
\newcommand{\emP}{\mathbb{P}_n}
\newcommand{\F}{\mathcal{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\RD}{\Reals^D}
\newcommand{\Nbb}{\mathbb{N}}

%%% Vectors
\newcommand{\thetast}{\theta^{\star}}
\newcommand{\betap}{\beta^{(p)}}
\newcommand{\betaq}{\beta^{(q)}}
\newcommand{\vardeltapq}{\varDelta^{(p,q)}}
\newcommand{\lambdavec}{\boldsymbol{\lambda}}
\newcommand{\bj}{{\bf j}}

%%% Matrices
\newcommand{\X}{X} % no bold
\newcommand{\Y}{Y} % no bold
\newcommand{\Z}{Z} % no bold
\newcommand{\Lgrid}{L_{\grid}}
\newcommand{\Xsetgrid}{D_{\grid}}
\newcommand{\Linvgrid}{L_{\grid}^{\Xsetagger}}
\newcommand{\Lap}{{\bf L}}
\newcommand{\NLap}{{\bf N}}
\newcommand{\PLap}{{\bf P}}
\newcommand{\Id}{I}

%%% Sets and classes
\newcommand{\Xset}{\mathcal{X}}
\newcommand{\Sset}{\mathcal{S}}
\newcommand{\Hclass}{\mathcal{H}}
\newcommand{\Pclass}{\mathcal{P}}
\newcommand{\Leb}{L}
\newcommand{\mc}[1]{\mathcal{#1}}

%%% Distributions and related quantities
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\Qbb}{\mathbb{Q}}
\newcommand{\Ibb}{\mathbb{I}}

%%% Operators
\newcommand{\Tadj}{T^{\star}}
\newcommand{\Xsetive}{\mathrm{div}}
\newcommand{\Xsetif}{\mathop{}\!\mathrm{d}}
\newcommand{\gradient}{\mathcal{D}}
\newcommand{\Hessian}{\mathcal{D}^2}
\newcommand{\dotp}[2]{\langle #1, #2 \rangle}
\newcommand{\Dotp}[2]{\Bigl\langle #1, #2 \Bigr\rangle}

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\SM}{\mathrm{SM}}
\newcommand{\OS}{\mathrm{FS}}
\newcommand{\PLS}{\mathrm{PLS}}

%%% Order of magnitude
\newcommand{\soom}{\sim}

% \newcommand{\span}{\textrm{span}}

\newtheoremstyle{alden}
{6pt} % Space above
{6pt} % Space below
{} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{alden}

\newtheoremstyle{aldenthm}
{6pt} % Space above
{6pt} % Space below
{\itshape} % Body font
{} % Indent amount
{\bfseries} % Theorem head font
{.} % Punctuation after theorem head
{.5em} % Space after theorem head
{} % Theorem head spec (can be left empty, meaning `normal')

\theoremstyle{aldenthm}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{remark}

\begin{document}
\title{Note on: Connections between Smoothing Splines and Laplacian smoothing}
\author{Alden Green}
\date{\today}
\maketitle

Recall the definition of an \emph{order-$m$ smoothing spline}. For observations $(x_1,y_1),\ldots,(x_n,y_n)$, design points $x_i$ each belonging to a domain $\Omega \subset \Rd$, 
\begin{equation}
\label{eqn:smoothing_spline}
\wh{g} = \argmin_{g \in G} \frac{1}{n} \sum_{i = 1}^{n} (y_i - g(x_i))^2 + \lambda |g|_{m}^2;
\end{equation}
here we use $|g|_{m}$ to denote the $H^m(\Omega)$, and $G$ to be an appropriately defined function class. The standard choice is $G = H^m(\Omega)$, in which case we must also choose $m > d/2$ to make~\eqref{eqn:smoothing_spline} be well-defined.

The goal of this note is to review properties of order-$m$ smoothing splines, and see if/when graph Laplacian smoothing possesses analogous properties.

\section{Well-posedness of smoothing splines}
\textcolor{red}{(TODO)}

\section{Optimality of smoothing splines}
Suppose $x_1,\ldots,x_n$ are sampled independently from a \textcolor{red}{sufficiently regular} density $p$ on $\Omega$, and that
\begin{equation}
\label{model:regression}
y_i = f_0(x_i) + \varepsilon_i,
\end{equation}
where $\varepsilon_i \sim N(0,1)$ are i.i.d. Gaussian noise that are also independent of $x_i$. Assume that $f_0 \in H^s(\Omega)$. What can we say about mean squared error $\|\wh{g} - g\|_P^2$?

\textcolor{red}{(TODO)}:
\begin{itemize}
	\item Smoothing splines are optimal when $2m > s$. Is there a lower bound on $s$ relative to $m$ or $d$?
	\item When $m < s < 2s$, the optimality requires some boundary conditions on $f_0$. Detail these boundary conditions and explain intuition behind why they are required. Compare with the boundary conditions required for orthogonal series/non-parametric least squares estimators to be optimal.
 	\item Detail various techniques for establishing optimality. Refer to \textcolor{red}{(van der Geer, Nussbaum, Cox)}. Review pros and cons of each technique.
\end{itemize}

\section{Low-smoothness regime}
We now introduce a \emph{restricted smoothing spline}, in order to compute estimates for $m < d/2$. Suppose we take $\Phi_N \subseteq G$ to be a finite set of $L^2(P)$-orthogonal low-frequency functions; formally $\Phi_N = \{\phi_1,\ldots,\phi_N\}$, where each $\phi_k$ satisfies the Laplacian eigenfunction equation,
\begin{equation}
\label{eqn:laplacian_eigen}
\Delta_P \phi_k = \lambda_k \phi_k,\quad \frac{\partial}{\partial {\bf n}}\phi_k = 0.
\end{equation}
Let $F_N = \mathrm{span}\{\Phi_N\}$, and consider
\begin{equation}
\label{eqn:restricted_smoothing_spline}
\wh{g}_{\Phi} = \argmin_{g \in F_N} \frac{1}{n} \sum_{i = 1}^{n} (y_i - g(x_i))^2 + \lambda |g|_{m}^2.
\end{equation}
The restricted smoothing spline~\eqref{eqn:smoothing_spline} is well-posed, and its solution $\wh{g}_{\Phi}$ possesses a relatively simple analytic solution. It may be totally impossible to exactly compute because it is difficult to exactly solve~\eqref{eqn:laplacian_eigen}, but that is besides the point. Now we analyze its statistical properties. To do so, we reduce the analysis to that of a related estimator in the Normal means/Gaussian sequence model.

\textcolor{red}{(TODO)}
\begin{itemize}
	\item Refer to the results of \textcolor{blue}{(Birge and Massart)} for empirical risk minimization (ERM) over normed balls in H\"{o}lder spaces. Resolve why Birge and Massart find a phase transition at $m = 1, d = 2$, which is not at the $4m = d$ boundary; I suspect it is because Birge and Massart find the point at which the empirical process diverges.
\end{itemize}

\subsection{Gaussian sequence model}
In the Gaussian sequence model, for each $i \in \mathbb{N}$ one observes
\begin{equation}
\label{model:gaussian_sequence_model}
Y_i = \theta_i^{\ast} + \frac{1}{\sqrt{n}}\varepsilon_i,\quad \varepsilon_i \sim N(0,1).
\end{equation}
Supposes the sequence $\theta^{\ast}$ possesses some ``smoothness''. In particular, suppose $\theta^{\ast}$ belongs to the \emph{order-$s$ Sobolev ellipsoid} of radius $M$, $\Theta^s = \{\theta: |\theta|_s \leq M\}$, where for $s \in \mathbb{N}$ the Sobolev seminorm $|\theta|_s$ is defined by
\begin{equation*}
|\theta|_s^2 := \sum_{j = 1}^{\infty} \theta_j^2 j^{2s/d}.
\end{equation*}
Consider the \emph{sequence smoothing spline} estimator, defined for $N \in \mathbb{N}$ as
\begin{equation}
\label{eqn:sequence_smoothing_spline}
\wh{\theta}_{N} := \argmin_{\theta \in \ell^2(N)} |Y - \theta|_{0} + \lambda |\theta|_m^2, \quad \wh{\theta} := \lim_{N \to \infty} \wh{\theta}_{N}.
\end{equation}
Note that the sequence space estimator $\wh{\theta}$ is perfectly well-posed even though its analogous function space estimator is not. In fact, the sequence space estimator has the explicit solution
\begin{equation}
\wh{\theta}_{j} = Y_j[1 + \lambda j^{2m/d}]^{-1},
\end{equation}
showing that it is a shrinkage estimator.

\textcolor{red}{(TODO)}
\begin{itemize}
	\item Relate the sampling model~\eqref{model:gaussian_sequence_model} to~\eqref{model:regression}, and likewise relate the estimator~\eqref{eqn:sequence_smoothing_spline} to~\eqref{eqn:restricted_smoothing_spline}. Refer to \textcolor{blue}{(Tsybakov, Johnstone)}.
	\item Show that in the Gaussian sequence model, if $m > \max\{d/4,s/2\}$ then~\eqref{eqn:sequence_smoothing_spline} achieves optimal $\ell_2$ loss when $N = \infty$.
\end{itemize}
\subsection{``Failure'' of restricted smoothing splines in the very-low-smoothness regime}
On the other hand, whenever $m \leq d/4$ meaningful restrictions on the size of $N$ kick in.
\begin{lemma}
	\label{lem:failure_very_low_smoothness}
	Suppose we observe data $Y_1,Y_2,\ldots$ according to Model~\eqref{model:gaussian_sequence_model}. If we compute $\wh{\theta}$ and any $m \leq d/4$, then
	\begin{equation}
	\label{eqn:failure_very_low_smoothness}
	\Ebb[|\wh{\theta} - \theta^{\ast}|_{0}] = \infty.
	\end{equation}
\end{lemma}
The requirement $m \leq d/4$ is at first glance somewhat mysterious. It is nicely demystified by the following Lemma, which describes the expected ``smoothness'' properties of the estimator $\wh{\theta}$, in terms of the various Sobolev seminorms $|\cdot|_r$.
\begin{lemma}
	\label{lem:smoothness_of_sequence_smoothing_splines}
	Suppose we observe data $Y_1,Y_2,\ldots$ according to Model~\eqref{model:gaussian_sequence_model}, and we compute $\theta$ as in~\eqref{eqn:sequence_smoothing_spline}.
	\begin{itemize}
		\item If $4m > 2r + d$, then
		\begin{equation}
		\label{eqn:smoothness_of_sequence_smoothing_splines_1}
		\Ebb[|\wh{\theta}|_{r}^2] \leq |\theta^{\star}|_r^2 + O(\lambda^{-(2s + d)/(2m)}/n).
		\end{equation}
		\item If $4m \leq 2r + d$, then 
		\begin{equation}
		\label{eqn:smoothness_of_sequence_smoothing_splines_2}
		\Ebb[|\wh{\theta}|_{r}^2] = \infty.
		\end{equation}
	\end{itemize}
\end{lemma}
\begin{proof}
	Recall that $\wh{\theta}_i = Y_i \cdot [1 + \lambda i^{2m/d}]^{-1}$, and note that
	\begin{equation*}
	\Ebb[Y_i^2] = (\theta_i^{\ast})^2 + \frac{1}{n}.
	\end{equation*} 
	Therefore
	\begin{equation}
	\label{pf:smoothness_of_sequence_smoothing_splines_-1}
	\Ebb[|\wh{\theta}|_{r}^2] = \Ebb\biggl[\sum_{i = 1}^{\infty} \biggl(\frac{Y_i}{1 + \lambda i^{2m/d}}\biggr)^2 i^{2r/d}\biggr] = \sum_{i = 1}^{\infty} \frac{(\theta_i^{\ast})^2 + 1/n}{(1 + \lambda i^{2m/d})^2} \cdot i^{2r/d}
	\end{equation}
	Clearly,
	\begin{equation}
	\label{pf:smoothness_of_sequence_smoothing_splines_0}
	\sum_{i = 1}^{\infty} \frac{(\theta_i^{\ast})^2}{(1 + \lambda i^{2m/d})^2}  \cdot i^{2r/d} \leq \sum_{i = 1}^{\infty} (\theta_i^{\ast})^2 \cdot i^{2r/d} = |\theta^{\ast}|_{r}^2.
	\end{equation}
	Moreover, when $4m > 2r + d$, then
	\begin{equation}
	\label{pf:smoothness_of_sequence_smoothing_splines_1}
	\frac{1}{n}\sum_{i = 1}^{\infty} \frac{i^{2r/d}}{(1 + \lambda i^{2m/d})^2} \leq \frac{1}{n} \lambda^{-(2r + d)/(2m)}.
	\end{equation}
	On the other hand, when $4m \leq 2r + d$, then setting $i^{\ast}$ such that $(i^{\ast})^{2m/d}\lambda = 1$, we have
	\begin{equation}
	\label{pf:smoothness_of_sequence_smoothing_splines_2}
	\frac{1}{n}\sum_{i = 1}^{\infty} \frac{i^{2r/d}}{(1 + \lambda i^{2m/d})^2} \geq 	\frac{1}{2\lambda^2n}\sum_{i = i^{\ast}}^{\infty} i^{(2r - 4m)/d} = \infty.
	\end{equation}
	Plugging~\eqref{pf:smoothness_of_sequence_smoothing_splines_0}-\eqref{pf:smoothness_of_sequence_smoothing_splines_2} back into~\eqref{pf:smoothness_of_sequence_smoothing_splines_-1} establishes the claim.
\end{proof}
Lemma~\ref{lem:smoothness_of_sequence_smoothing_splines} has a particularly interesting interpretation when $r = 1$ or $r = 0$. 
\begin{itemize}
	\item When $r = 1$---corresponding to the first-order Sobolev norm--- Lemma~\ref{lem:smoothness_of_sequence_smoothing_splines} tells us that the sequence smoothing spline $\wh{\theta}$ has infinite $|\cdot|_1$ norm when $d \geq 2$. Remarkably, when $d = 2$ or $d = 3$, $\wh{\theta}$ continues to be optimal despite being an incredibly ``wiggly'' estimator of a ``smooth'' sequence. Note that this forecloses upon the possibility of analyzing $\wh{\theta}$ using uniform convergence over $\Theta^1(M)$, since $\wh{\theta}$ does not belong to this class for any finite $M$.
	\item When $r = 0$---corresponding to the zeroth-order Sobolev norm, or in other words the $\ell^2$-norm---Lemma~\ref{lem:smoothness_of_sequence_smoothing_splines} tells us that the sequence smoothing spline $\wh{\theta}$ has infinite $\Theta^0$ norm when $d \geq 4$. Clearly, since the regression function $f_0$ has finite $\Theta^0$ norm, and the estimator $\wh{f}$ has infinite $\Theta^0$ norm, it follows from the triangle inequality that $|\wh{f} - f|_0 = \infty$. 
\end{itemize}

\textcolor{red}{(TODO)}
\begin{itemize}
	\item Do an analogous calculation for the $\Theta^{1,1}$ case, with the estimator
	\begin{equation*}
	\wt{\theta}_{N} = \argmin_{\theta \in \ell^2(N)} |Y - \theta|_0^2 + \lambda |\theta|_{1,1}^2.
	\end{equation*}
	 Confirm your expectation: that when error measured in $|\cdot|_0$ norm and $d > 1$, this estimator is optimal when $N = n$, but not when $N = \infty$. This would give some evidence that the graph is saving the TV denoiser when $d > 1$.
	 \item Figure out other smoothness properties of $\wh{\theta}_N$, for instance ``H\"{o}lder'' smoothness.
	 \item Ponder what implication the lack-of-smoothness of sequence smoothing splines has for other problems (besides estimation with error measured in $|\cdot|_0$ norm).
\end{itemize}

\section{Relationship to Graph Laplacian methods}

\begin{itemize}
	\item Review why your analysis of Laplacian smoothing breaks down when $d > 4$, making reference to the previous section. If possible, prove a lower bound to this effect. Compare with the analogous situation for restricted smoothing splines, pointing out that the discretization imposed by the graph improves the situation slightly.
	\item Study whether order-$m$ Laplacian smoothing estimate is optimal over $H^1(\mc{X})$, as long as $m > d/4$. 
	\item Explain why your analysis of Laplacian smoothing fails at estimation over $H^s(\mc{X})$ for $s > 1$. If possible, prove a lower bound.
\end{itemize}

\end{document}