### 2021-08-13 manuscripts/laplacian_eigenmaps/sections/preliminaries.tex

\section{Preliminaries}
\label{sec:setup_main_results}

In this section, we precisely define our modeling assumptions and the Laplacian eigenmaps methods we study.

\subsection{Nonparametric regression over Sobolev spaces}
\label{sec:regression_laplacian_eigenmaps}

We will operate in the usual setting of nonparametric regression with random design, in which we observe independent random samples $(X_1,Y_1),\ldots,(X_n,Y_n)$. The design points $X_1,\ldots,X_n$ are sampled from a distribution $P$ with support $\mc{X} \subseteq \Rd$, and the responses follow the signal plus noise model
\begin{equation}
\label{eqn:model}
Y_i = f_0(X_i) + w_i,
\end{equation}
with regression function $f_0: \mc{X} \to \Reals$, and $w_i \sim N(0,1)$ independent Gaussian noise. 
% (AG): Removed the comment regarding noise with variance = \sigma^2.

We now formulate two models, which differ in the assumed nature of the support $\mc{X}$ of the design distribution $P$: the \emph{flat Euclidean} and \emph{manifold} models.

\begin{definition}[Flat Euclidean model]
	\label{def:model_flat_euclidean}
	The data $(X_1,Y_1),\ldots,(X_n,Y_n)$ are sampled according to~\eqref{eqn:model}. The support $\mc{X}$ of the design distribution $P$ is an open, connected, and bounded subset of $\Rd$, with Lipschitz boundary. The distribution $P$ admits a Lipschitz density $p$ with respect to the $d$-dimensional Lebesgue measure $\nu$, which is bounded away from $0$ and $\infty$,
	\begin{equation*}
	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty, \quad \textrm{for all $x \in \mc{X}$.}
	\end{equation*}
\end{definition}

In the following, we recall that the injectivity radius of an $m$-dimensional Riemannian manifold $\mc{X}$ is the maximum value of $\delta$ such that the exponential map $\exp_x: B_m(0,\delta) \subset T_x(\mc{X}) \to B_{\mc{X}}(x,\delta) \subset \mc{X}$ is a diffeomorphism for all $x \in \mc{X}$.
\begin{definition}[Manifold model]
	\label{def:model_manifold}
	The data $(X_1,Y_1),\ldots,(X_n,Y_n)$ are sampled according to~\eqref{eqn:model}. 
	The support $\mc{X}$ of the design distribution $P$ is a closed, connected, smooth and boundaryless Riemannanian manifold embedded in $\Rd$, of intrinsic dimension $1 \leq m < d$. The injectivity radius of $\mc{X}$ is lower bounded by a positive constant $i_0 > 0$. The design distribution $P$ admits a Lipschitz density $p$ with respect to the volume form $d\mu$ induced by the Riemannian structure of $\mc{X}$, which is bounded away from $0$ and $\infty$,
	\begin{equation*}
	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty, \quad \textrm{for all $x \in \mc{X}$.}
	\end{equation*}
\end{definition}

Finally, at various points we will have to assume that the density $p$ also satisfies different types of higher-order regularity, beyond Lipschitz continuity. When such assumptions are necessary we will state them explicitly.

\subsection{Laplacian eigenmaps}
We now formally define the estimator and test statistic we study. Both are derived from eigenvectors of a graph Laplacian.  For a positive, symmetric kernel $\eta: [0,\infty) \to [0,\infty)$, and a radius parameter $\varepsilon > 0$, let $G = ([n],W)$ be the neighborhood graph formed over the design points $\{X_1,\ldots,X_n\}$, with a weighted edge $W_{ij} = \eta(\|X_i - X_j\|/\varepsilon)$ between vertices $i$ and $j$. Then the 
\emph{neighborhood graph Laplacian} $L_{n,\varepsilon}: \Reals^n \to \Reals$ is defined by its action on vectors $u \in \Reals^n$ as
\begin{equation}
\label{eqn:neighborhood_graph_laplacian}
\bigl(L_{n,\varepsilon}u\bigr)_i := \frac{1}{n\varepsilon^{2 + \mathrm{dim}(\mc{X})}} \sum_{j = 1}^{n} \bigl(u_i - u_j\bigr) \eta\biggl(\frac{\|X_i - X_j\|}{\varepsilon}\biggr).
\end{equation}
(Here $\mathrm{dim}(\mc{X})$ stands for the dimension of $\mc{X}$. It is equal to $d$ under the assumptions of Model~\ref{def:model_flat_euclidean}, and equal to $m$ under the assumptions of Model~\ref{def:model_manifold}. The pre-factor $(n\varepsilon^{2 + \mathrm{dim}(\mc{X})})^{-1}$ is purely for convenience in taking limits as $n \to \infty, \varepsilon \to 0$). Written in standard coordinates we have $(n\varepsilon^{\dim(\mc{X}) + 2}) \cdot L_{n,\varepsilon} = D - W$, where $D \in \Reals^{n \times n}$ is the diagonal degree matrix, $D_{ii} = \sum_{i = 1}^{n} W_{ij}$.

The graph Laplacian is a positive semi-definite matrix, and admits the eigendecomposition $L_{n,\varepsilon} = \sum_{k = 1}^{n} \lambda_k v_k v_k^{\top}$, where for each $k = 1,\ldots,n$ the eigenvalue-eigenvector pair $(\lambda_k,v_k)$ satisfies
\begin{equation*}
L_{n,\varepsilon}v_k = \lambda_k v_k, \quad \|v_k\|_2^2 = 1.
\end{equation*}
We will assume without loss of generality that each eigenvalue $\lambda$ of $L_{n,\varepsilon}$ has algebraic multiplicity $1$, and so we can index the eigenpairs $(\lambda_1,v_1),\ldots,(\lambda_n,v_n)$ in ascending order of eigenvalue, $0 = \lambda_1 < \ldots < \lambda_n$.

The Laplacian eigenmaps estimator $\wh{f}$ simply projects the response vector ${\bf Y}$ onto the first $K$ eigenvectors of $L_{n,\varepsilon}$: letting $V_K \in \Reals^{n \times K}$ be the matrix with columns $v_1,\ldots,v_K$, we have that
\begin{equation}
\label{eqn:laplacian_eigenmaps_estimator}
\wh{f} := \sum_{k = 1}^{K} \dotp{{\bf Y}}{v_k}_{n} v_k = \frac{1}{n} V_K V_K^{\top} {\bf Y}.
\end{equation} 
Thus $\wh{f}$ is equivalently a vector in $\Reals^n$, or a function in $L^2(P_n)$. If $\wh{f}$ is a reasonable estimate of $f_0$, then the Laplacian eigenmaps test statistic
\begin{equation}
\label{eqn:laplacian_eigenmaps_test}
\wh{T} := \|\wh{f}\|_n^2
\end{equation}
is in turn a reasonable estimate of $\|f_0\|_{P}^2$, and can be used to distinguish whether or not $f_0 = 0$.

% AG 8/12/21: Commented out due to overlap with footnote in new introduction.
% It may be helpful to comment briefly on the term ``Laplacian eigenmaps'', which we use a bit differently than is typical in the literature. Laplacian eigenmaps typically refers to an algorithm for embedding, which maps each design point $X_1,\ldots,X_n$ to $\Reals^K$ according to $X_i \mapsto (v_{1,i}, \ldots, v_{K,i})$. Viewing this embedding as a feature map, we can then interpret the estimator $\wh{f}$ as the least-squares solution to a linear regression problem with responses $Y_1,\ldots,Y_n$ and features $v_1,\ldots,v_K$. Often, the Laplacian eigenmaps embedding is viewed as a tool for dimensionality reduction, wherein it is implicitly assumed that $K$ is much smaller than $d$. We will neither explicitly nor implicitly take $K < d$; indeed, the embedding perspective is not particularly illuminating in what follows, and we do not henceforth make reference to it. Instead, we use ``Laplacian eigenmaps'' to directly refer to the estimator $\wh{f}$ or test statistic $\wh{T}$.

\subsection{Sobolev Classes}
\label{sec:sobolev}
We now review the definition of Sobolev classes, dividing our discussion into two cases, the flat Euclidean case (Model~\ref{def:model_flat_euclidean}) and the manifold case (Model~\ref{def:model_manifold}).

\paragraph{Flat Euclidean case.}
Take $\mc{X}$ to be a open, connected and bounded set with Lipschitz boundary, as in Model~\ref{def:model_flat_euclidean}. Recall that for a given multi-index $\alpha \in \mathbb{N}^d$, a function $f$ is \emph{$\alpha$-weakly differentiable} if there exists some $h \in L^1(\mc{X})$ such that
\begin{equation*}
\int_{\mc{X}} h g = (-1)^{|\alpha|} \int_{\mc{X}} f D^{\alpha}g, \quad \textrm{for every $g \in C_c^{\infty}(\mc{X})$.}
\end{equation*}
If such a function $h$ exists, it is the $\alpha$th weak partial derivative of $f$, and denoted by $D^{\alpha}f := h$. For functions $f$ which are $|\alpha|$-times differentiable, this coincides with the classical definition of derivative.

\begin{definition}[Sobolev space on a flat Euclidean domain]
	\label{def:sobolev_space}
	For an integer $s \geq 1$, a function $f \in L^2(\mc{X})$ belongs to the Sobolev space $H^s(\mc{X})$ if for all $|\alpha| \leq s$, the weak derivatives $D^{\alpha}f$ exists and satisfy $D^{\alpha}f \in L^2(\mc{X})$. The $j$th order semi-norm for $f \in H^s(\mc{X})$ is $|f|_{H^j(\mc{X})} := \sum_{|\alpha| = j}\|D^{\alpha}f\|_{\Leb^2(\mc{X})}$, and the corresponding norm
	\begin{equation*}
	\|f\|_{H^s(\mc{X})}^2 := \|f\|_{\Leb^2(\mc{X})}^2 + \sum_{j = 1}^{s} |f|_{H^j(\mc{X})}^2,
	\end{equation*}
	induces the Sobolev ball
	\begin{equation*}
	H^s(\mc{X};M) := \bigl\{f \in H^s(\mc{X}): \|f\|_{H^s(\mc{X})} \leq M\bigr\}.
	\end{equation*} 
\end{definition}
We note that $H^s(\mc{X})$ is the completion of $C^{\infty}(\mc{X})$ with respect to the $\|\cdot\|_{H^s(\mc{X})}$ norm, so that $C^{\infty}(\mc{X})$ is dense in $H^s(\mc{X})$.

\paragraph{Manifold case.}

There are several equivalent ways to define Sobolev spaces on a compact, smooth, $m$-dimensional Riemannian manifold embedded in $\Reals^d$. We will stick with a definition that parallels our setup in the flat Euclidean setting as much as possible. To do so, we first recall the notion of partial derivatives on a manifold, which are defined with respect to a local coordinate system. Letting $r_1,\ldots,r_m$ be the standard basis of $\Reals^m$, for a given chart $(\phi,U)$ (meaning an open set $U \subseteq \mc{X}$, and a smooth mapping $\phi: U \to \Reals^m$) we write $\phi =: (x_1,\ldots,x_m)$ in local coordinates, meaning $x_i = r_i \circ \phi$. Then the partial derivative $\partial f/\partial x_i$ of a function $f$ with respect to $x_i$ at $x \in U$ is
\begin{equation*}
\frac{\partial f}{\partial x_i}(x) := \frac{\partial(f \circ \phi^{-1})}{\partial r_i}\bigl(\phi(x)\bigr).
\end{equation*}
The right hand side should be interpreted in the weak sense of derivative. As before, we use the multi-index notation $D^{\alpha}f := \partial^{|\alpha|}f/\partial^{\alpha_1}x_1\ldots\partial^{\alpha_m}x_m$.

\begin{definition}[Sobolev space on a manifold]
	\label{def:sobolev_space_manifold}
	A function $f \in \Leb^2(\mc{X})$ belongs to the Sobolev space $H^{s}(\mc{X})$ if for all $\abs{\alpha} \leq s$, the weak derivatives $D^{\alpha}f$ exist and satisfy  $D^{\alpha}f \in \Leb^2(\mc{X})$. The $j$th order semi-norm $|f|_{H^j(\mc{X})}$, the norm $\|f\|_{H^s(\mc{X})}$, and the ball $H^s(\mc{X};M)$ are all defined as in Definition~\ref{def:sobolev_space}.
\end{definition}
The partial derivatives $D^{\alpha}f$ will depend on the choice of local coordinates, and so will the resulting Sobolev norm~$\|f\|_{H^s(\mc{X})}$. However, the important point is that regardless of the choice of local coordinates the resulting norms will be equivalent\footnote{Recall that norms $\|\cdot\|_1$ and $\|\cdot\|_2$ on a space $\mc{F}$ are said to be equivalent if there exist constants $c$ and $C$ such that
\begin{equation*}
c \|f\|_1 \leq \|f\|_2 \leq C \|f\|_1 \quad \textrm{for all $f \in \mc{F}$.}
\end{equation*}} 
and so the ultimate Sobolev space $H^s(\mc{X})$ is independent of the choice of local coordinates.

\paragraph{Boundary conditions.}
In the flat Euclidean model (Model~\ref{def:model_flat_euclidean}), in order to show that Laplacian eigenmaps is optimal over $H^{s}(\mc{X})$ for $s > 1$ we will need to assume that the regression function $f_0$ satisfies some boundary conditions. In particular, we will assume that $f_0$ is zero-trace.
\begin{definition}[Zero-trace Sobolev space]
	\label{def:zero_trace_sobolev_space}
	The \emph{order-s zero-trace Sobolev space} $H_0^s(\mc{X})$ is the closure of $C_c^\infty(\mc{X})$ with respect to $\|\cdot\|_{H^s(\mc{X})}$ norm. That is, $f \in H_0^s(\mc{X})$ if $f \in H^s(\mc{X})$ and additionally there exists a sequence $f_1,f_2,\ldots$ of functions in $C_c^{\infty}(\mc{X})$ such that
	\begin{equation*}
	\lim_{k \to \infty}\|f_k - f\|_{H^s(\mc{X})} = 0.
	\end{equation*}
	The normed ball $H_0^{s}(\mc{X};M) := H_0^{s}(\mc{X}) \cap H^{s}(\mc{X};M)$.
\end{definition}
The zero-trace condition can be made more concrete when $f \in C^\infty(\mc{X})$, since we can then speak of the pointwise behavior of $f$ and its derivatives. Letting $\partial/(\partial {\bf n})$ be the partial derivative operator in the direction of the vector $\mathbf{n}$ normal to the boundary of $\mc{X}$, then the zero-trace condition implies that $\partial^{k}f/\partial{\bf n}^k(x) = 0$ for each $k = 0,\ldots,s - 1$, and for all $x \in \partial\mc{X}$.

% (SB via AG) Move following chunk to appendix, and replace with one sentence explaining that when studying spectral projection methods, it is standard to impose boundary conditions.
We now explain why Laplacian eigenmaps should be optimal only when $f_0$ satisfies certain boundary conditions. Let $(\lambda_1(\Delta_P),\psi_1),(\lambda_2(\Delta_P),\psi_2),\ldots$ be the solutions to the weighted Laplacian eigenvector equation with Neumann boundary conditions, 
\begin{equation}
\label{eqn:laplacian_eigenfunction_neumann}
\Delta_P\psi_k = \lambda_{k}(\Delta_P) \psi_k, \quad \frac{\partial}{\partial {\bf n}}\psi_k = 0~~\textrm{on $\partial \mc{X}$}.
\end{equation}
As we have already alluded to, it is known~\citep{garciatrillos18} that each graph Laplacian eigenpair ($\lambda_k$,$v_k$) converges to a corresponding solution ($\lambda_k(\Delta_P)$, $\psi_k$) of~\eqref{eqn:laplacian_eigenfunction_neumann}. Thus it is relevant to consider which Sobolev functions $f \in H^s(M)$ we can reconstruct using the eigenfunctions $\psi_1,\psi_2,\ldots$. To that end, we introduce the spectrally defined Sobolev space
\begin{equation}
\label{eqn:spectral_sobolev_space}
\mc{H}^s(\mc{X}) := \biggl\{f \in L^2(\mc{X}): \sum_{k = 1}^{\infty} \bigl[\dotp{f}{\psi_k}_P\bigr]^2 \cdot\bigl[\lambda_k(\Delta_P)\bigr]^s < \infty \biggr\}.
\end{equation}
Under the conditions $p \in C^{\infty}(\mc{X})$ and $\partial\mc{X} \in C^{1,1}$, \cite{dunlop2020} show the strict inclusion $\mc{H}^{2s}(\mc{X}) \subset H^{2s}(\mc{X})$. More precisely, they show that for any $s > 0$,
\begin{equation}
\label{eqn:spectral_sobolev_space_equivalence}
\mc{H}^{2s}(\mc{X}) = \biggl\{f \in H^{2s}(\mc{X}): \frac{\partial \Delta_P^rf}{\partial {\bf n}} = 0~\textrm{on}~\partial\mc{X},~~\textrm{for all $0 \leq r \leq s - 1$} \biggr\},
\end{equation}
and for any $s \geq 0$, $\mc{H}^{2s + 1}(\mc{X}) = \mc{H}^{2s}(\mc{X}) \cap H^{2s + 1}(\mc{X})$. If $P$ is uniform on $\mc{X}$, then~\eqref{eqn:spectral_sobolev_space_equivalence} means that a Sobolev function $f \in H^{s}(\mc{X})$ additionally belongs to $\mc{H}^{s}(\mc{X})$ only if all its odd lower order derivatives vanish at the boundary $\partial\mc{X}$.

The bottom line is that the eigenvectors $v_k$ of the graph Laplacian accurately approximate only those functions $f \in H^{s}(\mc{X})$ which satisfy some additional boundary conditions. Although the zero-trace boundary condition is more restrictive than~\eqref{eqn:spectral_sobolev_space_equivalence}---since it also requires that derivatives of even-order be equal to $0$---the point is that some kind of boundary condition will be necessary in order to obtain optimal rates. We will stick to the zero-trace condition, since it greatly eases some of the steps in our proofs.
% (SB via AG) Move above chun to appendix.

On the other hand, in the manifold model (Model~\ref{def:model_manifold}) the domain $\mc{X}$ is without boundary---precisely, every point $x \in \mc{X}$ has a neighborhood that is homeomorphic to an open set in $\Reals^m$, for instance the open ball $B(x,\delta)$ for any $\delta$ smaller than the injectivity radius $i_0$---and so boundary conditions are irrelevant.

\subsection{Minimax Rates and Spectral Projection Methods}
\label{subsec:minimax_rates_sobolev}
We now review the minimax estimation and goodness-of-fit testing rates over Sobolev balls. We will pay special attention to certain classical spectral projection methods which achieve these rates. This is because, as we have already discussed, classical spectral projection methods are very related to Laplacian eigenmaps.

\paragraph{Estimation rates over Sobolev balls.}
In the estimation problem, we ask for an estimator---formally, a Borel measurable function $\wh{f}$ that maps from $\mc{X}$ to $\Reals$---which is close to the regression function $f_0$ with respect to the squared norm $\|\wh{f} - f_0\|_{P}^2$. Under Model~\ref{def:model_flat_euclidean}, the minimax estimation rate over the order-$s$ Sobolev ball is
\begin{equation}
\label{eqn:sobolev_space_minimax_estimation_rate}
\inf_{\wh{f}} \sup_{f_0} \Ebb\|\wh{f} - f_0\|_P^2 \asymp M^2(M^2n)^{-2s/(2s + d)};
\end{equation}
here the infimum is taken over all estimators $\wh{f}$, and the supremum over all $f_0 \in H^1(\mc{X};M)$ (first-order case) or $f_0 \in H_0^s(\mc{X};M)$ (higher-order case), and we assume $n^{-1/2} \lesssim M$.
% (SB via AG) Supremum also taken over all densities p? Do the densities have to satisfy certain conditions?

% (SB via AG) 
% 	1. Not clear that $\wt{f}$ is an estimator in Model 2.1.
%   2. My description of the analysis sells the proposition short.
%   3. Replace the proof description with the proof itself.
The lower bound in~\eqref{eqn:sobolev_space_minimax_estimation_rate} is due to~\citep{stone1980} (at least for the case of $M$ constant, as is most typically considered). The upper bound can be certified by a particular spectral projection estimator $\wt{f}$,
\begin{equation}
\label{eqn:spectral_series_estimator}
\wt{f} := \sum_{k = 1}^{K} \dotp{Y}{\psi_k}_n \psi_k,
\end{equation}
where $\psi_k$ are the eigenfunctions of $\Delta_P$ defined in~\eqref{eqn:laplacian_eigenfunction_neumann}. The optimality of spectral projection estimators over Sobolev type spaces is generally well-understood. For instance, see \cite{tsybakov08,johnstone2011,gine16}, who work in the Gaussian sequence model and show that analogous estimators are optimal over Sobolev ellipsoids. However, we have not found an analysis of the specific estimator~\eqref{eqn:spectral_series_estimator} under Model~\ref{def:model_flat_euclidean}, and so for completeness we state the result in the following proposition.
\begin{proposition}
	\label{prop:spectral_series_estimation}
	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $\partial \mc{X} \in C^{1,1}$, that $p \in C^{\infty}(\mc{X})$, and that $f_0 \in H^1(\mc{X};M)$ (first-order case) or $f_0 \in H_0^s(\mc{X};M)$ for some $s > 1$ (higher-order case). Then there exists a constant $C$ which does not depend on $f_0,M$ or $n$ such that the following statement holds: if the spectral projection estimator $\wt{f}$ is computed with parameter $K = \floor{M^2n}^{d/(2s + d)}$, then
	\begin{equation*}
	\Ebb\bigl[\|\wt{f} - f_0\|_P^2\bigr] \leq C \min\bigl\{M^2(M^2n)^{-2s/(2s + d)}, M^2\bigr\}
	\end{equation*}
\end{proposition}
The proof of Proposition~\ref{prop:spectral_series_estimation}, along with proofs of all of our results, can be found in the appendix. It is worth mentioning some aspects of the analysis here, because it sets the stage for the strategy we will use to analyze Laplacian eigenmaps.

In particular, three essential facts are needed to establish Proposition~\ref{prop:spectral_series_estimation}. 
\begin{enumerate}
	\item The continuous embedding of $H_0^s(\mc{X})$ into $\mc{H}^s(\mc{X})$---recall the latter is defined in~\eqref{eqn:spectral_sobolev_space}---which is a consequence of the zero-trace condition, the conditions on $\partial \mc{X}$ and $p$, and~\eqref{eqn:spectral_sobolev_space_equivalence}.
	\item Weyl's Law, which gives the asymptotic scaling of eigenvalues $\lambda_{k}(\Delta_P) \asymp k^{2/d}$, and allows us to properly control the bias induced by spectral projection.
	\item A local version of Weyl's law, which gives an estimate on $\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2$, and allows us to appropriately control the difference $\dotp{f_0}{\psi_k}_n - \dotp{f_0}{\psi_k}_P$  between the empirical and population Fourier coefficients.
\end{enumerate}
With these facts in hand the proof of Proposition~\ref{prop:spectral_series_estimation} follows from calculations standard to analysis of the Gaussian sequence model. Our analysis of Laplacian eigenmaps will depend on analogues of the first and second of these facts, with the space $\mc{H}^s(\mc{X})$ and eigenvalues $\lambda_k(\Delta_P)$ replaced by alternatives suitably defined with respect to the neighborhood graph Laplacian $L_{n,\varepsilon}$.

\paragraph{Goodness-of-fit testing rates over Sobolev balls.}
In the goodness-of-fit testing problem, we ask for a test function---formally, a Borel measurable function $\phi$ that takes values in $\{0,1\}$--- which can distinguish between the hypotheses
\begin{equation}
\mathbf{H}_0: f_0 = f_0^{\ast}, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{f_0^{\ast}\}.
\end{equation} 
Typically, the null hypothesis $f_0 = f_0^{\ast} \in \mc{F}$ reflects the absence of interesting structure, and $\mc{F} \setminus  \{f_0^{\ast}\}$ is a set of smooth departures from this null. To fix ideas, as in \citet{ingster2009} we focus on the problem of \emph{signal detection} in Sobolev spaces, where $f_0^{\ast} = 0$ and $\mc{F} = \mc{H}^s(\mc{X};M)$ is a Sobolev ball. This is without loss of generality since our test statistic and its analysis are easily modified to handle the case when $f_0^{\ast}$ is not $0$, by simply subtracting $f_0^{\ast}(X_i)$ from each observation $Y_i$.

The Type I error of a test $\phi$ is $\mathbb{E}_0[\phi]$, and if $\mathbb{E}_0[\phi] \leq a$ for a given $a \in (0,1)$ we refer to $\phi$ as a level-$a$ test\footnote{We reserve the more common symbol $\alpha$ for multi-indices, so as to avoid confusion}. The worst-case risk of $\phi$ over $\mc{F}$ is
\begin{equation*}
R_n(\phi, \mc{F}, \epsilon) := \sup\Bigl\{\mathbb{E}_{f_0}[1 - \phi]: f_0 \in \mc{F}, \|f_0\|_{P} > \epsilon\Bigr\},
\end{equation*}
and for a given constant $b \in (0,1)$, the minimax critical radius $\epsilon_n(\mc{F})$ is the smallest value of $\epsilon$ such that some level-$a$ test has worst-case risk of at most $b$. Formally,
\begin{equation*}
\epsilon_n(\mc{F}) := \inf\Bigl\{\epsilon > 0: \inf_{\phi} R_n(\phi,\mc{F},\epsilon) \leq b \Bigr\},
\end{equation*} 
where in the above the infimum is over all level-$a$ tests $\phi$, and $\Ebb_{f_0}[\cdot]$ is the expectation operator under the regression function $f_0$.\footnote{Clearly, the minimax critical radius $\epsilon_n(\mc{F})$ depends on $a$ and $b$. However, we adopt the typical convention of treating $\alpha,b \in (0,1)$ and  as small but fixed positive constants; hence they will not affect the testing error rates, and we suppress them notationally.} We will refer to the rate at which the squared critical radius $\epsilon_n^2(\mc{F})$ tends to $0$ as the minimax testing rate.

Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing rate is much smaller than the minimax estimation rate. \citet{ingster2009} give the minimax testing rate in the special case where $\mc{X} = [0,1]^d$ and $P$ is uniform, and we restate their result in the terms and notation of our paper.
\begin{theorem}[Theorem 1 of \citet{ingster2009}]
	\label{thm:spectral_series_testing}
	Suppose Model~\ref{def:model_flat_euclidean}, and additionally that $\mc{X} = [0,1]^d$ and $P$ is the uniform distribution on $[0,1]^d$. Then there exist constants $c,C,n_0$ which do not depend on $f_0$ or $n$ such that for all $n \geq n_0$,
	\begin{equation}
	\label{eqn:sobolev_space_testing_critical_radius}
	c n^{-4s/(4s + d)} \leq \epsilon_n^2\bigl(\mc{H}^s(\mc{X};1)\bigr) \leq C n^{-4s/(4s + d)}~~\textrm{for $1 \leq d < 4s$.}
	\end{equation}
\end{theorem}
The analysis used by \citet{ingster2009} to show the upper bound in~\eqref{eqn:sobolev_space_testing_critical_radius} relies on a similar trio of facts as used in the proof of Proposition~\ref{prop:spectral_series_estimation}. It is otherwise reminiscent of calculations made in the Gaussian sequence model, which can be found in~\cite{ingster2012}. This analysis can be straightforwardly adapted to handle design distributions $P$ that satisfy the conditions of Model~\ref{def:model_flat_euclidean}, or to handle the case where $M$ is not $1$. Finally, the test \citet{ingster2009} use to certify the upper bound in~\eqref{eqn:sobolev_space_testing_critical_radius} is implicitly a spectral projection method: the test statistic is the $L^2(P_n)$ norm of a particular spectral projection estimator.\footnote{\cite{ingster2009} project the responses onto a the span of trigonometric basis functions. Since $P$ is uniform on the unit cube, such functions are eigenfunctions of $\Delta_P$ when the right boundary conditions are imposed.}

A major difference between testing and estimation over Sobolev spaces is the requirement that $4s > d$. When $4s \leq d$, the functions in $H^s(\Xset)$ are very irregular. Crucially, $H^s(\Xset)$ no longer continuously embeds into $\Leb^4(\Xset)$ when $4s \leq d$, and test statistics using the $L^2(P_n)$ norm are no longer guaranteed to have finite variance.\footnote{Note that this will not affect the analysis for estimation, because for estimation we only need to control the first two moments of $f_0$.} In fact, we have not seen any analysis which describes the minimax rate for nonparametric regression testing over Sobolev spaces in the $4s \leq d$ regime. However, if one explicitly assumes that $f_0 \in \Leb^4(\mc{X})$, then the critical radius is characterized by the dimension-free rate $\epsilon_n^2(\Leb^4(\mc{X};M)) \asymp Mn^{-1/2}$.\footnote{As a sanity check, note that when $1 \leq d < 4s$, $n^{-1/2}$ converges to $0$ at a slower rate than $n^{-4s/(4s + d)}$, which is the minimax rate for $M = 1$ given in~\eqref{eqn:sobolev_space_testing_critical_radius}.} As we discuss after our first main theorem regarding testing with Laplacian eigenmaps (Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}), this rate is achievable by a test based on $\wh{T}$.

\paragraph{Manifold setup.}
Under Model~\ref{def:model_manifold}, both~\eqref{eqn:sobolev_space_minimax_estimation_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced everywhere by the intrinsic dimension $m$. The estimator $\wt{f}$ and a test using the statistic $\wt{T}$---with $\psi_1,\psi_2,\ldots$ now the eigenfunctions of the manifold weighted Laplace-Beltrami operator $\Delta_P$---achieve the optimal estimation and testing rates. This is because each of the three facts mentioned after Proposition~\ref{prop:spectral_series_estimation} have analogues when the domain is a smooth manifold (See~\citet{hendriks1990}, who analyzes a spectral projection density estimator, for details.)

\paragraph{In-sample mean squared error.}
As mentioned in our introduction, roughly speaking one of our main conclusions is that the Laplacian eigenmaps estimator $\wh{f}$ is minimax rate-optimal. It is worth being clear about what we do and do not mean by this statement. We do not mean that the estimator $\wh{f}$ will match the upper bound given in~\eqref{eqn:sobolev_space_minimax_estimation_rate}, since such a statement does not make sense when the estimator is defined only at the random design points $X_1,\ldots,X_n$. Instead we will measure loss using the squared $L^2(P_n)$ error. In Section~\ref{sec:out_of_sample} we show that an extension of $\wh{f}$ defined over all $\mc{X}$ has $L^2(P)$ error comparable to the $L^2(P_n)$ error of $\wh{f}$. We also believe that in the random design setting we work in, simple arguments will imply that $L^2(P_n)$ risk has the same minimax rate of convergence as $L^2(P)$ risk. We sketch such an argument in Section~\ref{sec:out_of_sample}, but do not further pursue the details.

% (SB via AG): Move this somewhere later.
Additionally, we will not actually measure accuracy using the expectation of the loss. Rather, we will give a constant probability bound on $\|\cdot\|_n^2$. For instance, when $f_0 \in H^1(\mc{X};1)$, we will show that with probability $1 - \delta$ the loss $\|\wh{f} - f_0\|_n^2 \leq C_{\delta} n^{-2/(2 + d)}$, for a constant $C_{\delta}$ that depends on $\delta$ but not on $f_0$ or $n$. Thus we give an upper bound on the $(1 - \delta)$th quantile of $\|\wh{f} - f_0\|_n^2$, rather than an upper bound on its expectation. We explain the reason for this in Section~\ref{sec:minimax_optimal_laplacian_eigenmaps}. We also show that if $f_0$ is bounded in a larger norm---for instance, if it is H\"{o}lder rather than Sobolev smooth---then we can obtain bounds on the expected $L^2(P_n)$ loss.

There is one other subtlety introduced by the use of in-sample mean squared error. Technically speaking, elements $f \in H^s(\mc{X})$ are equivalence classes, defined only up to a set of measure zero. Thus one cannot speak of the pointwise evaluation $f_0(X_i)$, as we do by defining our target of estimation to be $f_0(X_i)$, $i=1,\ldots,n$, until one selects \emph{representatives}. When $s > d/2$, every element $f$ of $H^s(\mc{X})$ admits a continuous version $f^{\ast}$, and as is standard we set this to be our favored representative. When $s \leq d/2$, some elements in $H^s(\mc{X})$ do not have any continuous version; however they admit a \emph{quasi-continuous} version \citep{evans15} known as the \emph{precise representative}, and we use this representative. To be clear, however, it does not really matter which representative we choose. Since all versions agree except on a set of measure zero, and since $P$ is absolutely continuous with respect to Lebesgue measure (in Model~\ref{def:model_flat_euclidean}) or the volume form $d\mu$ (in Model~\ref{def:model_manifold}), with probability $1$ any two versions $g_0, h_0 \in f_0$ will satisfy $g_0(X_i) = h_0(X_i)$ for all $i = 1,\ldots,n$. The bottom line is that we can use the notation $f_0(X_i)$ without fear of ambiguity or confusion.

Finally, we note that for testing none of these comments are relevant. We will show that our test has small worst-case risk whenever $\epsilon \gtrsim \epsilon_n(H_0^s(\mc{X};M))$, thus establishing that it is a minimax optimal test in the usual sense.