### 2021-10-16 manuscripts/laplacian_eigenmaps/appendix.tex

\noindent

\section{Upper bounds on population-level spectral series regression}

In this section we first give the proof of Proposition~\ref{prop:spectral_series_estimation}, then of Proposition~\ref{prop:spectral_series_testing}. In both cases the structure of the analysis, which is fairly classical and straightforward, can be usefully compared to our analysis of PCR-LE (see Section~\ref{subsec:analysis}).

\paragraph{Proof of Proposition~\ref{prop:spectral_series_estimation}.}
We decompose risk into squared bias and variance,
\begin{equation}
\label{pf:spectral_series_estimation_0}
\Ebb \|\wt{f} - f_0\|_P^2 = \Ebb\| \Ebb[\wt{f}]  - f_0\|_P^2 + \Ebb\| \wt{f} - \Ebb[\wt{f}]\|_P^2.
\end{equation}
Since the eigenfunctions $\{\psi_k\}$ form an orthonormal basis of $L^2(P)$, and $f_0 \in \mc{H}^{s}(\mc{X}) \subseteq L^2(P)$, we can write the squared bias in terms of squared Fourier coefficients of $f_0$, leading to the following upper bound,
\begin{equation*}
\|f_0 - \Ebb[\wt{f}]\|_P^2 = \sum_{k = K + 1}^{\infty}  \dotp{f_0}{\psi_k}_P^2 \leq  \frac{1}{\{\lambda_{K + 1}(\Delta_P)\}^s} \sum_{k = K + 1}^{\infty} \{\lambda_{k + 1}(\Delta_P)\}^s \dotp{f_0}{\psi_k}_P^2 \leq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}}{\{\lambda_{K + 1}(\Delta_P)\}^s}.
\end{equation*}
On the other hand, the variance term can be written as the sum of the variance of each empirical Fourier coefficient, and subsequently by the law of total variance we derive that
\begin{align}
\label{pf:spectral_series_estimation_2}
\Ebb\| \wt{f} - \Ebb[\wt{f}]\|_P^2 = \sum_{k = 1}^{K} \Var\Bigl[\dotp{{\bf Y}}{\psi_k}_n\Bigr] & = \sum_{k = 1}^{K} \Var\Bigl[\Ebb[\dotp{Y}{\psi_k}_n|{\bf X}\Bigr] + \Ebb\Bigl[\Var[\dotp{Y}{\psi_k}_n|{\bf X}\Bigr] \nonumber \\
& = \sum_{k = 1}^{K} \Var\Bigl[\dotp{f_0}{\psi_k}_n\Bigr] + \frac{1}{n}\Ebb\Bigl[\|\psi_k\|_n^2\Bigr] \nonumber \\
& \leq \frac{K}{n} + \frac{1}{n}\sum_{k = 1}^{K}\Ebb\Bigl[\Bigl(f_0(X)\psi_k(X)\Bigr)^2\Bigr].
\end{align}
Consequently,
\begin{equation}
\label{pf:spectral_series_estimation_1}
\Ebb \|\wt{f} - f_0\|_P^2 \leq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\bigl[\lambda_{K + 1}(\Delta_P)\bigr]^s} + \frac{K}{n} + \frac{1}{n}\Ebb\Bigl[(f_0(X))^2 \cdot \sum_{k = 1}^{K} (\psi_k(X))^2\Bigr].
\end{equation}
The claim of the proposition then follows from variants of two classical results in spectral geometry. The first is a Weyl's Law asymptotic scaling of the eigenvalues of $\Delta_P$ due to~\cite{dunlop2020}; formally, there exist constants $c$ and $C$ (which will depend on $P$ and $d$) such that
\begin{equation}
\label{eqn:weyl}
ck^{2/d} \leq \rho_k \leq Ck^{2/d}\quad\textrm{for all $k \in \mathbb{N}$, $k \geq 2$}.
\end{equation}
The second is a local analog to Weyl's Law, which says that there exists a constant $C$ (again depending on $P$ and $d$) such that
\begin{equation}
\label{eqn:local_weyl}
\sup_{x \in \mc{X}}\biggl\{\sum_{k = 1}^{K} \bigl(\psi_k(x)\bigr)^2\biggr\} \leq CK \quad\textrm{for all $K \in \mathbb{N}$}.
\end{equation}
Equation~\eqref{eqn:local_weyl} is a direct implication of~\eqref{eqn:weyl} along with Theorem 17.5.3 of~\cite{hormander1973}. Plugging the upper bounds~\eqref{eqn:weyl} and~\eqref{eqn:local_weyl} back into~\eqref{pf:spectral_series_estimation_1}, we conclude that
\begin{equation}
\label{pf:spectral_series_estimation_3}
\Ebb \|\wt{f} - f_0\|_P^2 \leq C\biggl(\frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{(K + 1)^{2s/d}} + \frac{K}{n}\biggr).
\end{equation}
If $n^{-1/2} \geq M$, then taking $K = 1$ implies $\Ebb \|\wt{f} - f_0\|_P^2 \leq C(M^2 + 1/n)$. Otherwise, setting $K = \floor{M^2n}^{d/(2s + d)}$ balances squared bias and variance, and yields the claim.
\qed.

\paragraph{Proof of Proposition~\ref{prop:spectral_series_testing}.}
We briefly lay out the main ideas needed to prove Proposition~\ref{prop:spectral_series_testing}, following the lead of~\cite{ingster2009} who prove a similar result in the special case where $M = 1$ and $P$ is the uniform distribution over $\mc{X} = [0,1]^d$, and referring to that work for more details.

We begin by computing the first two moments of the test statistic $\wt{T}$. The expectation is
\begin{align*}
\Ebb[\wt{T}] = \frac{(n - 1)}{n} \sum_{k = 1}^{K} \dotp{f_0}{\psi_K}_P^2 + \frac{K}{n} + \Ebb\Bigl[(f_0(X))^2 \dot \sum_{k = 1}^{K} (\psi_k(X))^2\Bigr],
\end{align*}
and from~\eqref{eqn:weyl} (Weyl's Law) we have that under the alternative $f_0 \neq 0$,
\begin{equation*}
\Ebb_{f_0}[\wt{T}] \geq \frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\lambda_{K + 1}^s(\Delta_{P})} + \frac{K}{n}.
\end{equation*}
To compute the variance, we decompose $\wt{T} = \wt{T}_{1,1} + \wt{T}_{1,2} + \wt{T}_{1,3} + \wt{T}_2$ into the sum of 3 U-statistics and the remaining diagonal terms, defined in terms of the equivalent kernel $\kappa(x,x') = \sum_{k = 1}^{K} \psi_k(x) \psi_k(x')$ as,
\begin{align*}
T_{1,1} & := \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} w_i w_j \kappa(X_i,X_j),\quad && T_{1,2} := \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} \bigl(w_i f_0(X_j) + w_jf_0(X_i)\bigr)\kappa(X_i,X_j) \\
T_{1,3} & := \frac{1}{n^2} \sum_{1 \leq i \neq j \leq n} f_0(X_i) f_0(X_j) \kappa(X_i,X_j), \quad && T_2 := \frac{1}{n^2} \sum_{i = 1}^{n} Y_i^2 \kappa(X_i,X_i).
\end{align*}
The variances of each statistic can be found by routine computation (see~\cite{ingster2009}), and in particular satisfy the upper bounds
\begin{align*}
\Var(T_{1,1}) & \leq \frac{2K}{n^2}, \quad && \Var(T_{1,2}) \overset{\mathrm{(i)}}{\leq} \frac{C}{n}\|f_0\|_P^2\\
\Var(T_{1,3}) & \overset{\mathrm{(ii)}}{\leq} C\biggl(\frac{K}{n}\|f_0\|_P^4 + \frac{K}{n^2}\|f_0\|_{L^4(\mc{X})}^4\biggr), \quad && \Var(T_{2}) \overset{\mathrm{(iii)}}{\leq} \frac{CK^2}{n^3}\biggl(1 + \|f_0\|_{L^4(\mc{X})}^4\biggr)
\end{align*}
where $\mathrm{(i)}-\mathrm{(iii)}$ hold due to local Weyl's law, i.e.~\eqref{eqn:local_weyl}. Upper bounds on Type I and Type II error,
\begin{equation*}
\Ebb_0[\wt{\varphi}] \leq \biggl(1 + CK/n^2\biggr)a, \quad \Ebb_{f_0}[1 - \wt{\varphi}] \leq \frac{C(K/n^2 + \|f_0\|_P^2 + K/n \|f_0\|_P^4 + K/n^2 \|f_0\|_{L^4(\mc{X})}^4)}{(\sum_{k = 1}^{K} \dotp{f_0}{\psi_k}_P^2 - \sqrt{2K/an})^2},
\end{equation*}
follow from Chebyshev's inequality. It can be verified that so long as
\begin{equation}
\label{pf:spectral_series_test}
\|f_0\|_P^2 \geq C\Biggl(\frac{\|f_0\|_{\mc{H}^s(\mc{X})}^2}{\lambda_{K + 1}^s(\Delta_{P})} + \frac{\sqrt{K}}{n}\biggl(\sqrt{\frac{1}{a}} + \sqrt{\frac{1}{b}}\biggr)\Biggr)
\end{equation}
for a sufficiently large constant $C$, then $\Ebb_{f_0}[1 - \wt{\varphi}] \leq b$. The two summands in~\eqref{pf:spectral_series_test} are bias and standard deviation terms, respectively. When $M^2 \leq n^{-1}$, setting $K = 1$ gives the desired result. Otherwise, choosing $K = \floor{M^2n}^{2d/(4s + d)}$ balances these two terms, and leads to~\eqref{eqn:spectral_series_testing}. \qed
% (AG 8/19/21): Same comment as above, regarding the role of M in random design regression testing.

\section{Graph-dependent error bounds}
\label{sec:fixed_graph_error_bounds}
In this section, we adopt the fixed design perspective; or equivalently, condition on $X_i = x_i$ for $i = 1,\ldots,n$. Let $G = \bigl([n],W\bigr)$ be a fixed graph on $\{1,\ldots,n\}$ with Laplacian matrix $L = \sum_{k = 1}^{n}\lambda_k v_k v_k^{\top}$; the eigenvectors have unit empirical norm, $\|v_k\|_n^2 = 1$. The randomness thus all comes from the responses 
\begin{equation}
\label{eqn:fixed_graph_regression_model}
Y_i = f_{0}(x_i) + w_i
\end{equation}
where the noise variables $w_i$ are independent $N(0,1)$. In the rest of this section, we will mildly abuse notation and write $f_0 = (f_0(x_1),\ldots,f_0(x_n)) \in \Reals^n$. We will also write ${\bf Y} = (Y_1,\ldots,Y_n)$.

\subsection{Upper bound on Estimation Error of PCR-LE}

\begin{lemma}
	\label{lem:fixed_graph_estimation}
	For any integer $s > 0$, and any integer $0 \leq K \leq n$, the PCR-LE estimator $\wh{f}$ of~\eqref{eqn:laplacian_eigenmaps_estimator} satisfies
	\begin{equation}
	\label{eqn:fixed_graph_estimation}
	\|\wh{f} - f_0\|_n^2 \leq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} + \frac{5K}{n};
	\end{equation}
	this is guaranteed if $K = 0$, and otherwise holds with probability at least $1 - \exp(-K)$ if $1 \leq K \leq n$. 
\end{lemma}
\paragraph{Proof (of Lemma~\ref{lem:fixed_graph_estimation}).}
	By the triangle inequality,
	\begin{equation}
	\label{pf:fixed_graph_estimation_1}
	\|\wh{f} - f_0\|_n^2 \leq 2\Bigl(\|\mathbb{E}\wh{f} - f_0\|_n^2 + \|\wh{f} - \mathbb{E}\wh{f}\|_n^2\Bigr).
	\end{equation}
	The first term in~\eqref{pf:fixed_graph_estimation_1} (approximation error) is non-random, since the design is fixed. The expectation $\mathbb{E}\wh{f} = \sum_{k = 1}^{K} \dotp{v_k}{f_0}_n v_k$, so that
	\begin{equation*}
	\|\mathbb{E}\wh{f} - f_0\|_n^2 = \Bigl\|\sum_{k = K + 1}^{n} \dotp{v_k}{f_0}_n v_k\Bigr\|_n^2 = \sum_{k = K + 1}^n \dotp{v_k}{f_0}_n^2.
	\end{equation*}
	In the above, the last equality relies on the fact that $v_k$ are orthonormal in $L^2(P_n)$. Using the fact that the eigenvalues are in increasing order, we obtain
	\begin{equation*}
	\sum_{k = K + 1}^n \dotp{v_k}{f_0}_n^2 \leq \frac{1}{\lambda_{K + 1}^s} \sum_{k = K + 1}^n \lambda_k^s \dotp{v_k}{f_0}_n^2 \leq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s}.
	\end{equation*}

	If $K = 0$, $\wh{f} = \Ebb{\wh{f}} = 0$, and the second term in~\eqref{pf:fixed_graph_estimation_1} is $0$. Otherwise the second   in~\eqref{pf:fixed_graph_estimation_1} (estimation error) is random. Observe that $\dotp{v_k}{\varepsilon}_n \overset{d}{=} Z_k/\sqrt{n}$, where $(Z_1,\ldots,Z_n) \sim N(0,I_{n \times n})$. Again using the orthonormality of the eigenvectors $v_k$, we have
	\begin{equation*}
	\|\wh{f} - \mathbb{E}\wh{f}\|_n^2 = \sum_{k = 1}^{K} \dotp{v_k}{\varepsilon}_n^2 \overset{d}{=} \frac{1}{n}\sum_{k = 1}^{K} Z_k^2.
	\end{equation*}
	Thus $\|\wh{f} - \mathbb{E}\wh{f}\|_n^2$ is equal to $1/n$ times a $\chi^2$ distribution with $K$ degrees of freedom. Consequently, it follows from a result of \citep{laurent00} that
	\begin{equation*}
	\Pbb\biggl(\|\wh{f} - \mathbb{E}\wh{f}\|_n^2 \geq \frac{K}{n} + 2\frac{\sqrt{K}}{n}\sqrt{t} + \frac{2t}{n}\biggr) \leq \exp(-t).
	\end{equation*}
	Setting $t = K$ completes the proof of the lemma.

\subsection{Upper bound on Testing Error of PCR-LE}

Let $\wh{T} = \sum_{k = 1}^{K} \dotp{{\bf Y}}{v_k}_n^2$, and let $\varphi = \1\{\wh{T} \geq t_a\}$. In the following Lemma, we upper bound the Type I and Type II error of the test $\varphi$.

\begin{lemma}
	\label{lem:fixed_graph_testing}
	Suppose we observe $(Y_1,x_1),\ldots,(Y_n,x_n)$ according to~\eqref{eqn:fixed_graph_regression_model}.
	\begin{itemize}
		\item If $f_0 = 0$, then $\Ebb_0[\varphi] \leq a$.
		\item Suppose $f_0 \neq 0$ satisfies
		\begin{equation}
		\label{eqn:fixed_graph_testing_critical_radius}
		\|f_0\|_n^2 \geq \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} + \frac{\sqrt{2K}}{n}\biggl[2\sqrt{\frac{1}{a}} + \sqrt{\frac{2}{b}} + \frac{32}{bn}\biggr],
		\end{equation}
		for some $s \in \mathbb{N}\setminus \{0\}$. Then $\Ebb_{f_0}[1 - \phi] \leq b$.
	\end{itemize}
\end{lemma}
\paragraph{Proof (of Lemma~\ref{lem:fixed_graph_testing}).}
We first compute the expectation and variance of $\wh{T}$, then apply Chebyshev's inequality to upper bound the Type I and Type II error.

\underline{\emph{Expectation}.}
Recall that $\wh{T} = \sum_{k = 1}^{K} \dotp{Y}{v_k}_n^2$. Expanding the square gives
\begin{equation*}
\Ebb[\wh{T}] = \sum_{k = 1}^{K} \Ebb[\dotp{Y}{v_k}_n^2] = \sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 + \Ebb[2\dotp{f_0}{v_k}_n\dotp{\varepsilon}{v_k}_n + \dotp{\varepsilon}{v_k}_n^2] = \frac{K}{n} + \sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2.
\end{equation*}
Thus $\Ebb[\wh{T}] - t_a = \sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 - \sqrt{2K}/n \cdot \sqrt{1/a}$. Furthermore, it is a consequence of~\eqref{eqn:fixed_graph_testing_critical_radius} that 
\begin{equation}
\label{pf:fixed_graph_testing_1}
\sum_{k = 1}^{K} \dotp{f_0}{v_k}_n^2 - \frac{\sqrt{2K}}{n}\sqrt{1/a} \geq \|f_0\|_n^2 - \frac{\dotp{L^sf_0}{f_0}_n}{\lambda_{K + 1}^s} - \frac{\sqrt{2K}}{n}\sqrt{1/a} \geq \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \sqrt{\frac{2}{b}} + \frac{32}{bn}\biggr].
\end{equation}

\underline{\emph{Variance}.}
Recall from the proof of Lemma~\ref{lem:fixed_graph_estimation} that $\dotp{\varepsilon}{v_k}_n \overset{d}{=} Z_k/\sqrt{n}$ for $(Z_1,\ldots,Z_n) \sim N(0,I_{n \times n})$. Expanding the square, and recalling that $\Cov[Z,Z^2] = 0$ for Gaussian random variables, we have that
\begin{equation*}
\Var\bigl[\dotp{{\bf Y}}{v_k}_n^2\bigr] = \Var\biggl[\frac{2}{n}\dotp{f_0}{v_k}_nZ_k + \frac{2}{n^2}Z_k^2\biggr] = \frac{4\dotp{f_0}{v_k}_n^2}{n} + \frac{2}{n^2}.
\end{equation*}
Moreover, since $\Cov[Z_k^2,Z_{\ell}^2] = 0$ for each $k = 1,\ldots,K$, we see that
\begin{equation*}
\Var\bigl[\wh{T}\bigr] = \sum_{k = 1}^{K} \Var\bigl[\dotp{{\bf Y}}{v_k}_n^2\bigr] = \frac{2K}{n^2} + \sum_{k = 1}^{K}\frac{4\dotp{f_0}{v_k}_n^2}{n}.
\end{equation*}

\underline{\emph{Bounds on Type I and Type II error}.}
The upper bound on Type I error follows immediately from Chebyshev's inequality.

The upper bound on Type II error also follows from Chebyshev's inequality. We observe that~\eqref{eqn:fixed_graph_testing_critical_radius} implies $\Ebb_{f_0}[\wh{T}] = t_a$, and apply Chebyshev's inequality to deduce
\begin{equation*}
\Pbb_{f_0}\bigl(\wh{T} < t_a\bigr) \leq \Pbb_{f_0}\Bigl(|\wh{T} - \Ebb_{f_0}[\wh{T}]|^2 > |\Ebb_{f_0}[\wh{T}] - t_a|^2\Bigr) \leq \frac{\Var\bigl[\wh{T}\bigr]}{\bigl[\Ebb_{f_0}[\wh{T}] - t_a\bigr]^2} = \frac{2K/n^2 + 4/n\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2}{\bigl[\Ebb_{f_0}[\wh{T}] - t_a\bigr]^2}.
\end{equation*}
Thus we have upper bounded the Type II error by the sum of two terms, each of which are no more than $1/(2b)$, as we now show. For the first term, after noting that~\eqref{pf:fixed_graph_testing_1} implies $\Ebb_{f_0}[\wh{T}] - t_a \geq \sqrt{2K}/n \cdot \sqrt{2/b}$, the upper bound follows:
\begin{equation*}
\frac{2K/n^2}{\bigl[\Ebb_{f_0}[\wh{T}] - t_a\bigr]^2} \leq \frac{b}{2}.
\end{equation*}
On the other hand, for the second term we use~\eqref{pf:fixed_graph_testing_1} in two ways: first to conclude that $\Ebb_{f_0}[\wh{T}] - t_a \geq 1/2 \cdot \sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2$, and second to obtain
\begin{equation*}
\frac{4\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2}{n\bigl[\Ebb_{f_0}[\wh{T}] - t_a\bigr]^2} \leq \frac{4\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2}{n\bigl(\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2/2\bigr)^2} \leq \frac{16}{n\sum_{k = 1}^{K}\dotp{f_0}{v_k}_n^2} \leq \frac{b}{2}.
\end{equation*}

\section{Graph Sobolev semi-norm, flat Euclidean domain}
\label{sec:graph_quadratic_form_euclidean}
In this section we prove Proposition~\ref{prop:graph_seminorm_ho}. The proposition will follow from several intermediate results.
\begin{enumerate}
	\item~In Section~\ref{subsec:decomposition_graph_seminorm}, we show that
	\begin{equation}
	\label{pf:graph_seminorm_ho_1}
	\dotp{L_{n,\varepsilon}^sf}{f}_n \leq \frac{1}{\delta} \dotp{L_{P,\varepsilon}^sf}{f}_{P} + \frac{C\varepsilon^2}{\delta n\varepsilon^{2 + d}}M^2.
	\end{equation}
	with probability at least $1 - 2\delta$.

	We term the first term on the right hand side the \emph{non-local Sobolev semi-norm}, as it is a kernelized approximation to the Sobolev semi-norm $\dotp{\Delta_P^sf}{f}_{P}$. The second term on the right hand side is a pure bias term, which as we will see is negligible compared to the non-local Sobolev semi-norm as long as $\varepsilon \ll n^{-1/(2(s -1 + d))}$. 
	\item~In Section~\ref{subsec:approximation_error_nonlocal_laplacian}, we show that when $x$ is sufficiently in the interior of $\mc{X}$, then $L_{P,\varepsilon}^kf(x)$ is a good approximation to $\Delta_P^kf(x)$, as long as $f \in H^{s}(\mc{X})$ and $p \in C^{s - 1}(\mc{X})$ for some $s \geq 2k + 1$. 
	\item~In Section~\ref{subsec:boundary_behavior_nonlocal_laplacian}, we show that when $x$ is sufficiently near the boundary of $\mc{X}$, then $L_{P,\varepsilon}^kf(x)$ is close to $0$, as long as $f \in H_0^{s}(\mc{X})$ for some $s > 2k$.
	\item~In Section~\ref{subsec:estimate_nonlocal_seminorm}, we use the results of the preceding two sections to show that if $f \in H_0^s(\mc{X};M)$ and $p \in C^{s - 1}(\mc{X})$, there exists a constant $C$ which does not depend on $f$ such that
	\begin{equation}
	\label{pf:graph_seminorm_ho_2}
	\dotp{L_{P,\varepsilon}^sf}{f}_{P} \leq CM^2.
	\end{equation}
\end{enumerate}
Finally, in Section~\ref{subsec:integrals} we provide some assorted estimates used in Sections~\ref{subsec:decomposition_graph_seminorm}.

\paragraph{Proof (of Proposition~\ref{prop:graph_seminorm_ho}).}
Proposition~\ref{prop:graph_seminorm_ho} follows immediately from~\eqref{pf:graph_seminorm_ho_1} and~\eqref{pf:graph_seminorm_ho_2}. \qed

One note regarding notation: suppose a function $g \in H^{\ell}(U)$, where $\ell \in \mathbb{N}$ and $U$ is an open set. Let $V$ be another open set, compactly contained within $U$. Then we will use the notation $g \in H^{\ell}(V)$ to mean that the restriction $\restr{g}{V}$ of $g$ to $V$ belongs to $H^{\ell}(V)$.

\subsection{Decomposition of graph Sobolev semi-norm}
\label{subsec:decomposition_graph_seminorm}

In Lemma~\ref{lem:graph_seminorm_bias}, we decompose the graph Sobolev semi-norm (a V-statistic) into an unbiased estimate of the non-local Sobolev semi-norm (a U-statistic), and a pure bias term. We establish that the pure bias term will be small (in expectation) relative to the U-statistic whenever $\varepsilon$ is sufficiently small.
\begin{lemma}
	\label{lem:graph_seminorm_bias}
	For any $f \in L^2(\mc{X})$, the graph Sobolev semi-norm satisfies
	\begin{equation}
	\label{eqn:graph_seminorm_bias_1}
	\dotp{L_{n,\varepsilon}^sf}{f}_{n} = U_{n,\varepsilon}^{(s)}(f) + B_{n,\varepsilon}^{(s)}(f),
	\end{equation}
	such that $\mathbb{E}[U_{n,\varepsilon}^{(s)}(f)] = (n - s - 1)!/n! \cdot \dotp{L_{P,\varepsilon}^sf}{f}_P$. If additionally $f \in H^1(\mc{X};M)$ and $\varepsilon \geq n^{-1/d}$, then the bias term $B_{n,\varepsilon}^{(s)}(f)$ satisfies
	\begin{equation}
	\label{eqn:graph_seminorm_bias_2}
	\mathbb{E}\bigl[|B_{n,\varepsilon}^{(s)}(f)|\bigr] \leq \frac{C\varepsilon^2}{\delta n\varepsilon^{2 + d}}M^2.
	\end{equation}
\end{lemma}
Then~\ref{pf:graph_seminorm_ho_1} follows immediately from Lemma~\ref{lem:graph_seminorm_bias}, by Markov's inequality.
\paragraph{Proof (of Lemma~\ref{lem:graph_seminorm_bias}).}
We begin by introducing some notation. We will use bold notation $\bj = (j_1,\ldots,j_s)$ for a vector of indices where $j_i \in [n]$ for each $i$. We write $[n]^s$ for the collection of all such vectors, and $(n)^s$ for the subset of such vectors with no repeated indices. Finally, we write $D_if$ for a kernelized difference operator,
\begin{equation*}
D_if(x) := \bigl(f(x) - f(X_i)\bigr) \eta\biggl(\frac{\|X_i - x\|}{\varepsilon}\biggr),
\end{equation*}
and we let $D_{\bj}f(x) := \bigl(D_{j_1}\circ \cdots \circ D_{j_s}f\bigr)(x)$.

With this notation in hand, it is easy to represent $\dotp{L_{n,\varepsilon}^sf}{f}_{n}$ as the sum of a U-statistic and a bias term,
\begin{align*}
\dotp{L_{n,\varepsilon}^sf}{f}_{n} & = \frac{1}{n} \sum_{i = 1}^{n} L_{n,\varepsilon}^sf(X_i) \cdot f(X_i) \\
& = \underbrace{\frac{1}{n^{s + 1}\varepsilon^{s(d + 2)}} \sum_{{i\bf j} \in (n)^{s + 1}}D_{\bj}f(X_i) \cdot f(X_i)}_{=:U_{n,\varepsilon}^{(s)}(f)} + \underbrace{\frac{1}{n^{s + 1}\varepsilon^{s(d + 2)}} \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot f(X_i)}_{=:B_{n,\varepsilon}^{(s)}(f)}
\end{align*}
When the indices of $i\bj$ are all distinct, it follows straightforwardly from the law of iterated expectation that
\begin{equation*}
\mathbb{E}[ D_{\bj}f(X_i) \cdot f(X_i)] = \varepsilon^{s(d + 2)}\mathbb{E}[L_{P,\varepsilon}^sf(X_i) \cdot f(X_i)] = \dotp{L_{P,\varepsilon}^sf}{f}_{P}, 
\end{equation*}
which in turn implies $\mathbb{E}[U_{n,\varepsilon}^{(s)}(f)] = (n - s - 1)!/n! \cdot \dotp{L_{P,\varepsilon}^sf}{f}_P$.

It remains to show~\eqref{eqn:graph_seminorm_bias_2}. By adding and subtracting $f(X_{\bj_1})$, we obtain by symmetry that
\begin{equation*}
\sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot f(X_i) = \frac{1}{2} \cdot \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot \bigl(f(X_i) - f(X_{\bj_1})\bigr),
\end{equation*}
and consequently
\begin{equation*}
\Ebb\Bigl[\sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} D_{\bj}f(X_i) \cdot f(X_i)\Bigr] \leq \frac{1}{2} \cdot \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} \Ebb\Bigl[\bigl|D_{\bj}f(X_i)\bigr| \cdot \bigl|f(X_i) - f(X_{\bj_1})\bigr|\Bigr].
\end{equation*}
In Lemma~\ref{lem:graph_seminorm_bias2}, we show that if $f \in H^1(\mc{X};M)$, then for any $i\bj \in [n]^{s + 1}$ which contains a total of $k + 1$ distinct indices, 
\begin{equation*}
\Ebb\Bigl[\bigl|D_{\bj}f(X_i)\bigr| \cdot \bigl|f(X_i) - f(X_{\bj_1})\bigr|\Bigr] \leq C_1 \varepsilon^{2 + kd} M^2.
\end{equation*}
This shows us that the expectation of $|B_{n,\varepsilon}^s(f)|$ can bounded from above by the sum over several different terms, as follows:
\begin{align*}
\Ebb\Bigl[|B_{n,\varepsilon}^s(f)|\Bigr] & \leq C_1\frac{\varepsilon^2}{n\varepsilon^{2s}}M^2 \sum_{\substack{i\bj \in \\ [n]^{s + 1}\setminus (n)^{s + 1}}} \frac{1}{(n\varepsilon^d)^s}  \varepsilon^{(|i\bj| - 1)d} \\
& \leq C_1\frac{\varepsilon^2}{n\varepsilon^{2s}}M^2  \sum_{k = 1}^{s - 1} \frac{(n\varepsilon^d)^k}{(n\varepsilon^d)^s}n.
\end{align*}
Finally, we note that by assumption $n\varepsilon^d \geq 1$, so that in the above sum the factor of $(n\varepsilon^d)^k$ is largest when $k = s- 1$. We conclude that
\begin{equation*}
\Ebb\Bigl[|B_{n,\varepsilon}^s(f)|\Bigr] \leq C_1 (s - 1) \frac{\varepsilon^2}{n\varepsilon^{2s + d}}M^2,
\end{equation*}
which is the desired result.

\subsection{Approximation error of non-local Laplacian}
\label{subsec:approximation_error_nonlocal_laplacian}

In this section, we establish the convergence $L_{P,\varepsilon}^kf \to \sigma_{\eta}^k\Delta_P^kf$ as $\varepsilon \to 0$. More precisely, we give an upper bound on the squared difference between $L_{P,\varepsilon}^kf$ and  $\sigma_{\eta}^k\Delta_P^kf$ as a function of $\varepsilon$. The bound holds for all $x \in \mc{X}_{k\varepsilon}$, and $f \in H^{s}(\mc{X})$, as long as $s \geq 2k + 1$. 
\begin{lemma}
	\label{lem:approximation_error_nonlocal_laplacian}
	Assume Model~\ref{def:model_flat_euclidean}. Let $s \in \mathbb{N} \setminus \{0,1\}$, suppose that $f \in H^s(\mc{X};M)$, and if $s > 1$ suppose that $p \in C^{s - 1}(\mc{X})$. Let $L_{P,\varepsilon}$ be define with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_flat_euclidean}. Then there exist constants $C_1$ and $C_2$ that do not depend on $f$, such that each of the following statements hold.
	\begin{itemize}
		\item If $s$ is odd and $k = (s - 1)/2$, then
		\begin{equation}
		\label{eqn:approximation_error_nonlocal_laplacian_1}
		\|L_{P,\varepsilon}^kf - \Delta_P^kf\|_{L^2(\mc{X}_{k\varepsilon})} \leq C_1 M \varepsilon
		\end{equation}
		\item If $s$ is even and $k = (s - 2)/2$, then
		\begin{equation}
		\label{eqn:approximation_error_nonlocal_laplacian_2}
		\|L_{P,\varepsilon}^kf - \Delta_P^kf\|_{L^2(\mc{X}_{k\varepsilon})} \leq C_2 M \varepsilon^2.
		\end{equation}
	\end{itemize}
\end{lemma}
We remark that when $k = 1$ and $f \in C^3(\mc{X})$ or $C^4(\mc{X})$, statements of this kind are well known, and indeed stronger results---with $L^{\infty}(\mc{X})$ norm replacing $L^2(\mc{X})$ norm---hold. When dealing with the iterated Laplacian, and functions $f$ which are regular only in the Sobolev sense, the proof is somewhat more lengthy, but in result is similar in spirit.

\paragraph{Proof (of Lemma~\ref{lem:approximation_error_nonlocal_laplacian}).}
Throughout this proof, we shall assume that $f$ and $p$ are smooth functions, meaning they belong to $C^{\infty}(\mc{X})$. This is without loss of generality, since $C^{\infty}(\mc{X})$ is dense in both $H^s(\mc{X})$ and $C^{s - 1}(\mc{X})$, and since both sides of the inequalities~\eqref{eqn:approximation_error_nonlocal_laplacian_1} and~\eqref{eqn:approximation_error_nonlocal_laplacian_2} are continuous with respect to $\|\cdot\|_{H^s(\mc{X})}$ and $\|\cdot\|_{C^{s - 1}(\mc{X})}$ norms.

We will actually prove a more general set of statements than contained in Lemma~\ref{lem:approximation_error_nonlocal_laplacian}, more general in the sense that they give estimates for all $k$, rather than simply the particular choices of $k$ given above. In particular, we will prove that the following two statements hold for any $s \in \mathbb{N}$ and any $k \in \mathbb{N} \setminus \{0\}$. 
\begin{itemize}
	\item If $k \geq s/2$, then for every $x \in \mc{X}_{k\varepsilon}$, 
	\begin{equation}
	\label{pf:approximation_error_nonlocal_laplacian_0}
	L_{P,\varepsilon}^kf(x) = g_s(x) \varepsilon^{s - 2k}
	\end{equation}
	for a function $g_s$ that satisfies
	\begin{equation}
	\label{pf:approximation_error_nonlocal_laplacian_0.5}
	\|g_s\|_{L^2(\mc{X}_{k\varepsilon})} \leq C \|p\|_{C^{q}(\mc{X})}^k M 
	\end{equation}
	where $q = 1$ if $s =0$ or $s = 1$, and otherwise $q = s - 1$. 
	\item If $k < s/2$, then for every $x \in \mc{X}_{k\varepsilon}$,
	\begin{equation}
	\label{pf:approximation_error_nonlocal_laplacian_1}
	L_{P,\varepsilon}^kf(x) = \sigma_{\eta}^k \cdot \Delta_{P}^kf(x) + \sum_{j = 1}^{\floor{(s - 1)/2} - k} g_{2(j + k)}(x)\varepsilon^{2j} + g_{s}(x) \varepsilon^{s - 2k}.
	\end{equation}
	for functions $g_j$ that satisfy
	\begin{equation}
	\label{pf:approximation_error_nonlocal_laplacian_1.5}
	\|g_j\|_{H^{s - j}(\mc{X}_{k\varepsilon})} \leq C \|p\|_{C^{s - 1}(\mc{X})}^k M.
	\end{equation}
\end{itemize}
In the statement above, recall that $H^0(\mc{X}_{k\varepsilon}) = L^2(\mc{X}_{k\varepsilon})$. Additionally, note that we may speak of the pointwise behavior of derivatives of $f$ because we have assumed that $f$ is a smooth function. Observe that~\eqref{eqn:approximation_error_nonlocal_laplacian_1} follows upon taking $k = \floor{(s - 1)/2}$ in~\eqref{pf:approximation_error_nonlocal_laplacian_1}, whence we have
\begin{equation*}
\bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_{P}^kf(x)\bigr)^2 = \varepsilon^2 \bigl(g_s(x)\bigr)^2
\end{equation*}
for some $g_s \in L^2(\mc{X}_{k\varepsilon},C \cdot M \cdot \|p\|_{C^{s - 1}(\mc{X})})$, and integrating over $\mc{X}_{k\varepsilon}$ gives the desired result. \eqref{eqn:approximation_error_nonlocal_laplacian_2} follows from~\eqref{pf:approximation_error_nonlocal_laplacian_1} in an identical fashion.

It thus remains establish~\eqref{pf:approximation_error_nonlocal_laplacian_1}, and~\eqref{pf:approximation_error_nonlocal_laplacian_0} which is an important part of proving~\eqref{pf:approximation_error_nonlocal_laplacian_1}. We will do so by induction on $k$. Note that throughout, we will let $g_j$ refer to functions which may change from line to line, but which always satisfy~\eqref{pf:approximation_error_nonlocal_laplacian_1.5}.

\underline{\textit{Proof of~\eqref{pf:approximation_error_nonlocal_laplacian_0} and~\eqref{pf:approximation_error_nonlocal_laplacian_1}, base case.}}

We begin with the base case, where $k = 1$. Again, we point out that although desired result is known when $s = 3$ or $s = 4$, and $f$ is regular in the H\"{o}lder sense, we require estimates for all $s \in \mathbb{N}$ when $f$ is regular in the Sobolev sense.

When $s = 0$, the inequality~\eqref{pf:approximation_error_nonlocal_laplacian_0} is implied by Lemma~\ref{lem:l2estimate_nonlocal_laplacian}.  When $s \geq 1$, we proceed using Taylor expansion. For any $x \in \mc{X}_{\varepsilon}$, we have that $B(x,\varepsilon) \subseteq \mc{X}$. Thus for any $x' \in B(x,\varepsilon)$, we may take an order $s$ Taylor expansion of $f$ around $x' = x$, and an order $q$ Taylor expansion of $p$ around $x' = x$, where $q = 1$ if $s = 1$, and otherwise $q = s - 1$. (See Section~\ref{subsec:taylor_expansion} for a review of the notation we use for Taylor expansions, as well as some properties that we make use of shortly.) This allows us to express $L_{P,\varepsilon}f(x)$ as the sum of three terms,
\begin{align*}
L_{P,\varepsilon}f(x) & = \frac{1}{\varepsilon^{d + 2}}\sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{q - 1}\frac{1}{j_1!j_2!}  \int_{\mc{X}} \bigl(d_x^{j_1}f\bigr)(x' - x) \bigl(d_x^{j_2}p\bigr)(x' - x) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \quad + \\
& \quad \frac{1}{\varepsilon^{d + 2}}\sum_{j = 1}^{s - 1} \frac{1}{j!} \int_{\mc{X}} \bigl(d_x^jf\bigr)(x' - x)  r_{x'}^{q}(x;p) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \quad  + \\
& \quad \frac{1}{\varepsilon^{d + 2}} \int_{\mc{X}} r_{x'}^j(x;f) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x').
\end{align*}
Here we have adopted the convention that $\sum_{j = 1}^{0} = 0$.

Changing variables to $z = (x' - x)/\varepsilon$, we can rewrite the above expression as 
\begin{align*}
L_{P,\varepsilon}f(x) & = \frac{1}{\varepsilon^{2}}\sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{q - 1}\frac{\varepsilon^{j_1 + j_2}}{j_1!j_2!}  \int d_x^{j_1}f(z) d_x^{j_2}p(z) \eta\bigl(\|z\|\bigr) \,dz \quad + \\
& \quad \frac{1}{\varepsilon^{2}} \sum_{j = 1}^{s - 1} \frac{\varepsilon^j}{j!} \int d_x^jf(z)  r_{zh + x}^{q}(x;p) \eta\bigl(\|z\|\bigr) \,dz \quad  + \\
& \quad \frac{1}{\varepsilon^{2}} \int r_{zh + x}^j(x;f) \eta\bigl(\|z\|\bigr) p(zh + x)\,dz \\
& := G_1(x) + G_2(x) + G_3(x).
\end{align*}
We now separately consider each of $G_1(x),G_2(x)$ and $G_3(x)$. We will establish that if $s = 1$ or $s = 2$, then $G_1(x) = 0$, and otherwise if $s \geq 3$ that
\begin{equation*}
G_1(x) = \sigma_{\eta}\Delta_Pf(x) + \sum_{j = 1}^{\floor{(s - 1)/2} - 1}g_{2(j + 1)}(x)\varepsilon^{2j} + g_{s}(x)\varepsilon^{s - 2}.
\end{equation*}
On the other hand, we will establish that if $s = 1$ then $G_2(x) = 0$, and otherwise for $s \geq 2$
\begin{equation}
\label{pf:approximation_error_nonlocal_laplacian_2}
\|G_2\|_{L^2(\mc{X}_{\varepsilon})} \leq C \varepsilon^{s - 2} M \|p\|_{C^{s - 1}(\mc{X})};
\end{equation}
this same estimate will hold for $G_3$ for all $s \geq 1$. Together these will imply~\eqref{pf:approximation_error_nonlocal_laplacian_0} and~\eqref{pf:approximation_error_nonlocal_laplacian_1}.

\emph{Estimate on $G_1(x)$.}
If $s = 1$, then $s - 1 = 0$, and so $G_1(x) = 0$. We may therefore suppose $s \geq 2$. Recall that
\begin{equation}
G_1(x) = \sum_{j_1 = 1}^{s - 1} \sum_{j_2 = 0}^{q - 1} \frac{\varepsilon^{j_1 + j_2 - 2}}{j_1!j_2!}  \underbrace{\int_{B(0,1)} d_x^{j_1}f(z) d_x^{j_2}p(z) \eta(\|z\|) \,dz}_{:= g_{j_1,j_2}(x)} \label{pf:approximation_error_nonlocal_laplacian_3}
\end{equation}
The nature of $g_{j_1,j_2}(x)$ depends on the sum $j_1 + j_2$. Since $d_x^{j_1}f d_x^{j_2}$ is an order $j_1 + j_2$ (multivariate) monomial, we have (see Section~\ref{subsec:taylor_expansion}) that whenever $j_1 + j_2$ is odd,
\begin{equation*}
g_{j_1,j_2}(x) = \int_{\mc{X}} d_x^{j_1}f(z) d_x^{j_2}p(z) \eta(\|z\|) \,dz = 0.
\end{equation*}
In particular this is the case when $j_1 = 1$ and $j_2 = 0$. Thus when $s = 2$,  $G_1(x) = g_{1,0}(x) = 0$. On the other hand if $s \geq 3$, then the lowest order terms in~\eqref{pf:approximation_error_nonlocal_laplacian_3} are those where $j_1 + j_2 = 2$, so that either $j_1 = 1$ and $j_2 = 1$, or $j_1 = 2$ and $j_2 = 0$. We have that
\begin{align*}
g_{1,1}(x) + \frac{1}{2}g_{2,0}(x) & = \int_{\mc{X}} d_x^{1}f(z) d_x^{1}p(z) \eta(\|z\|) \,dz + \frac{p(x)}{2} \int_{\mc{X}} d_x^{2}f(z) \eta(\|z\|) \,dz \\
& = \sum_{i_1 = 1}^{d} \sum_{i_2 = 1}^{d} D^{e_{i_1}}f(x) D^{e_{i_2}}p(x) \int_{\mc{X}} z^{e_{i_1} + e_{i_2}} \eta(\|z\|) \,dz + \frac{p(x)}{2} \sum_{i_1 = 1}^{d} \sum_{i_2 = 1}^{d}D^{e_{i_2}+e_{i_2}}f(x)\int_{\mc{X}} z^{e_{i_1} + e_{i_2}} \eta(\|z\|) \,dz\\
& = \sum_{i = 1}^{d} D^{e_{i}}f(x) D^{e_{i}}p(x) \int_{\mc{X}} z^2 \eta(\|z\|) \,dz + \frac{p(x)}{2} \sum_{i = 1}^{d} D^{2e_{i}}f(x)\int_{\mc{X}} z^2 \eta(\|z\|) \,dz\\ 
& = \sigma_{\eta}\Delta_Pf(x),
\end{align*}
which is the leading term order term. Now it remains only to deal with the higher-order terms, where $j_1 + j_2 > 2$, and where it suffices to show that each function $g_{j_1,j_2}$ satisfies~\eqref{pf:approximation_error_nonlocal_laplacian_1.5} for $j = \min\{j_1 + j_2 - 2,s - 2\}$. It is helpful to write $g_{j_1,j_2}$ using multi-index notation, 
\begin{align*}
g_{j_1,j_2}(x) = \sum_{|\alpha_1| = j_1} \sum_{|\alpha_2| = j_2} D^{\alpha_1}f(x) D^{\alpha_2}p(x) \int_{B(0,1)} z^{\alpha_1 + \alpha_2} \eta(\|z\|) \,dz,
\end{align*}
where we note that $|\int_{B(0,1)} z^{\alpha_1 + \alpha_2} \eta(\|z\|) \,dz| < \infty$ for all $\alpha_1, \alpha_2$, by the assumption that $\eta$ is Lipschitz on its support. Finally, by H\"{o}lder's inequality we have that
\begin{align*}
\|D^{\alpha_1}f D^{\alpha_2}p\|_{H^{s - (j + 2)}(\mc{X})} & \leq \|D^{\alpha_1}f\|_{H^{s - (j + 2)}(\mc{X})} \|D^{\alpha_2}p\|_{C^{s - (j + 2)}(\mc{X})} \\
& \leq \|D^{\alpha_1}f\|_{H^{s - j_1}(\mc{X})} \|D^{\alpha_2}p\|_{C^{s - (j_2 + 1)}(\mc{X})} \\
& \leq M \cdot \|p\|_{C^{s - 1}(\mc{X})},
\end{align*}
and summing over all $|\alpha_1| = j_1$ and $|\alpha_2| = j_2$ establishes that $g_{j_1,j_2}$ satisfies~\eqref{pf:approximation_error_nonlocal_laplacian_1.5}.

\emph{Estimate on $G_2(x)$.}
Note immediately that $G_2(x) = 0$ if $s = 1$. Otherwise if $s \geq 2$, then $q = s - 1$. Recalling that $|r_{x + z\varepsilon}^{s - 1}(x; p)| \leq C\varepsilon^{s - 1}\|p\|_{C^{s - 1}(\mc{X})}$ for any $z \in B(0,1)$, and that $d_x^jf(\cdot)$ is a $j$-homogeneous function, we have that
\begin{align}
|G_2(x)| & \leq \sum_{j = 1}^{s - 1} \frac{\varepsilon^{j - 2}}{j!}\int_{B(0,1)} \Bigl|\bigl(d_x^{j}f\bigr)(z)\Bigr| \cdot |r_{x + z\varepsilon}^{s - 1}(x;p)| \cdot \eta(\|z\|) \,dz \nonumber \\
& \leq C\varepsilon^{s - 2}\|p\|_{C^{s - 1}(\mc{X})} \sum_{j = 1}^{s - 1} \frac{1}{j!} \int_{B(0,1)} \Bigl|\bigl(d_x^{j}f\bigr)(z)\Bigr| \cdot \eta(\|z\|) \,dz \label{pf:approximation_error_nonlocal_laplacian_4}.
\end{align}
Furthermore, for each $j = 1,\ldots,s - 1$ convolution of $d_x^jf$ with $\eta$ only decreases the $L^2(\mc{X}_{\varepsilon})$ norm, meaning
\begin{equation}
\label{pf:approximation_error_nonlocal_laplacian_5}
\begin{aligned}
\int_{\mc{X}_{\varepsilon}} \biggl(\int_{B(0,1)} \Bigl|\bigl(d_x^{j}f\bigr)(z)\Bigr| \cdot \eta(\|z\|) \,dz\biggr)^2 \,dx & \leq \int_{\mc{X}_{\varepsilon}} \biggl(\int_{B(0,1)} \Bigl|\bigl(d_x^jf\bigr)(z)\Bigr|^2 \eta(\|z\|)\,dz \biggr) \cdot \biggl(\int_{B(0,1)} \eta(\|z\|) \,dz \biggr) \,dx \\
& \leq \int_{B(0,1)} \int_{\mc{X}_{\varepsilon}} \Bigl[\bigl(d^jf\bigr)(x)\Bigr]^2 \eta(\|z\|) \,dx  \,dz \\
& \leq \|d^jf\|_{L^2(\mc{X_{\varepsilon}})}^2.
\end{aligned}
\end{equation}
In the above, we have used both that $|d_x^jf(z)| \leq |d^jf(x)|$ for all $z \in B(0,1)$, and that the kernel is normalized so that $\int \eta(\|z\|) \,dz = 1$. 
Combining this with~\eqref{pf:approximation_error_nonlocal_laplacian_4}, we conclude that
\begin{align*}
\int_{\mc{X}_{\varepsilon}} |G_2(x)|^2 \,dx & \leq C \Bigl(\varepsilon^{s - 2}\|p\|_{C^{s - 1}(\mc{X})}\Bigr)^2 \sum_{j = 1}^{s - 1} \int_{\mc{X}_{\varepsilon}}\biggl(\frac{1}{j!} \int_{B(0,1)} \Bigl|\bigl(d_x^{j}f\bigr)(z)\Bigr| \cdot \Bigl|\eta(\|z\|)\Bigr| \,dz\biggr)^2 \,dx \\
& \leq C \Bigl(\varepsilon^{s - 2}\|p\|_{C^{s - 1}(\mc{X})}\Bigr)^2 \sum_{j = 1}^{s - 1} \|d^ju\|_{L^2(\mc{X_{\varepsilon}})}^2,
\end{align*}
establishing the desired estimate.

\emph{Estimate on $G_3(x)$.}
Applying the Cauchy-Schwarz inequality, we deduce a pointwise upper bound on $|G_3(x)|^2$,
\begin{align*}
|G_3(x)|^2 & \leq \biggl(\frac{p_{\max}}{\varepsilon^2}\biggr)^2 \cdot \biggl(\int_{B(0,1)} \bigl|r_{x + \varepsilon z}^s(x;u)\bigr|^2 \eta(\|z\|)\,dz\biggr) \cdot \biggl(\int_{B(0,1)} \eta(\|z\|) \,dz\biggr) \\
& \leq \biggl(\frac{p_{\max}}{\varepsilon^2}\biggr)^2 \int_{B(0,1)} \bigl|r_{x + \varepsilon z}^s(x;u)\bigr|^2 \eta(\|z\|) \,dz.
\end{align*}
Applying this pointwise over all $x \in \mc{X}_{\varepsilon}$ and integrating, we obtain
\begin{align*}
\int_{\mc{X}_{\varepsilon}} |G_3(x)|^2 \,dx & \leq \biggl(\frac{p_{\max}}{\varepsilon^2}\biggr)^2 \int_{\mc{X}_{\varepsilon}} \int_{B(0,1)} \bigl|r_{x + \varepsilon z}^s(x;f)\bigr|^2 \eta(\|z\|) \,dz \,dx \\
& = \biggl(\frac{p_{\max}}{\varepsilon^2}\biggr)^2 \int_{B(0,1)} \int_{\mc{X}_{\varepsilon}} \bigl|r_{x + \varepsilon z}^s(x;f)\bigr|^2 \eta(\|z\|) \,dx \,dz \\
& \leq \biggl(\frac{p_{\max}\varepsilon^s}{\varepsilon^2}\biggr)^2  \|d^sf\|_{L^2(\mc{X}_{\varepsilon})}^2,
\end{align*}
with the last inequality following from~\eqref{eqn:sobolev_remainder_term}. Noting that $p_{\max} = \|p\|_{C^0(\mc{X})} \leq \|p\|_{C^{s - 1}(\mc{X})}$, we see that this is a sufficient bound on $\|G_3\|_{L^2(\mc{X}_{\varepsilon})}$.

\underline{\textit{Proof of~\eqref{pf:approximation_error_nonlocal_laplacian_0} and~\eqref{pf:approximation_error_nonlocal_laplacian_1}, induction step.}}
We now assume that~\eqref{pf:approximation_error_nonlocal_laplacian_0} and~\eqref{pf:approximation_error_nonlocal_laplacian_1} hold for all order up to some $k$, and show that they then hold for order $k + 1$ as well. The proof is relatively straightforward, once we introduce a bit of notation. Namely, for any $\ell,j \in \mathbb{N}$ such that $1 \leq j \leq \ell \leq$, we will use $g_j^{\ell}$ to refer to a function satisfying
\begin{equation}
\label{pf:approximation_error_nonlocal_laplacian_6}
\|g_j^{\ell}\|_{H^{\ell - j}(\mc{X}_{(k + 1)\varepsilon})} \leq C \|p\|_{C^{q}(\mc{X})}^{k + 1} M.
\end{equation}
Note that $g_j^{\ell}(x) = g_{(s - \ell) + j}(x)$, so that $g_j^{s}(x) = g_j(x)$. As before, the functions $g_j^{\ell}$ may change from line to line, but will always satisfy~\eqref{pf:approximation_error_nonlocal_laplacian_6}. We immediately illustrate the purpose of this notation. Suppose $g \in H^{\ell}(\mc{X}_{k\varepsilon}; C \|p\|_{C^{q}(\mc{X})}^k M)$ for some $\ell \leq s$. If $\ell \leq 2$, then by the inductive hypothesis, it follows that for any $x \in \mc{X}_{(k + 1)\varepsilon}$
\begin{equation}
\label{pf:approximation_error_nonlocal_laplacian_7}
L_{P,\varepsilon}g(x) = g_{\ell}^{\ell}(x) \varepsilon^{\ell - 2}.
\end{equation} 
On the other hand if $2 < \ell \leq s$, then by the inductive hypothesis, it follows that for any $x \in \mc{X}_{(k + 1)\varepsilon}$,
\begin{equation}
\label{pf:approximation_error_nonlocal_laplacian_8}
L_{P,\varepsilon}g(x) = \sigma_{\eta} \Delta_Pg(x) + \sum_{j = 1}^{\floor{(\ell - 1)/2} - 1} g_{2j + 2}^{\ell}(x) \varepsilon^{2j} + g_{\ell}^{\ell}(x) \varepsilon^{\ell - 2}.
\end{equation}

\emph{Proof of \eqref{pf:approximation_error_nonlocal_laplacian_0}.} If $s \leq 2(k + 1)$, then by the inductive hypothesis it follows that for all $x \in \mc{X}_{k\varepsilon}$, we have $L_{P,\varepsilon}^kf(x) = g_{s}(x) \cdot \varepsilon^{s - 2k}$, for some $g_s \in L^2(\mc{X}_{k\varepsilon}, C\|p\|_{C^{s - 1}(\mc{X})}^k M)$. Note that we may know more about $L_P^kf(x)$ than simply that it is bounded in $L^2$-norm, but a bound in $L^2$-norm suffices. In particular, from such a bound along with~\eqref{pf:approximation_error_nonlocal_laplacian_7} we deduce that for any $x \in \mc{X}_{(k + 1)\varepsilon}$,
\begin{equation}
\label{pf:approximation_error_nonlocal_laplacian_8.5}
L_{P,\varepsilon}^{k + 1}f(x) = \bigl(L_{P,\varepsilon} \circ L_{P,\varepsilon}^k f)(x)= L_{P,\varepsilon} g_s(x)\varepsilon^{s - 2k} = g_{s}^{s}(x) \varepsilon^{s - 2(k + 1)},
\end{equation}
establishing~\eqref{pf:approximation_error_nonlocal_laplacian_0}.

\emph{Proof of \eqref{pf:approximation_error_nonlocal_laplacian_1}.} If $s > 2(k + 1)$, then by the inductive hypothesis we have that for all $x \in \mc{X}_{k\varepsilon}$, 
\begin{equation*}
L_{P,\varepsilon}^kf(x) = \sigma_{\eta}^k \Delta_P^kf(x) + \sum_{j = 1}^{\floor{(s - 1)/2} - k} g_{2(j + k)}(x) \varepsilon^{2j} + g_s(x) \varepsilon^{s - 2k}.
\end{equation*}
Thus for any $x \in \mc{X}_{(k + 1)\varepsilon}$, 
\begin{equation*}
L_{P,\varepsilon}^{k + 1}f(x) = \bigl(L_{P,\varepsilon} \circ L_{P,\varepsilon}^k f\bigr)(x) = \sigma_{\eta}^k L_{P,\varepsilon}\Delta_P^kf(x) + \sum_{j = 1}^{\floor{(s - 1)/2} - k} L_{P,\varepsilon}g_{2(j + k)}(x) \varepsilon^{2j} + L_{P,\varepsilon}g_s(x) \varepsilon^{s - 2k}
\end{equation*}
There are three terms on the right hand side of this equality, and we now analyze each separately.
\begin{enumerate}
	\item Noting that $\Delta_P^kf \in H^{s - 2k}(\mc{X}; C\|p\|_{C^{s - 1}(\mc{X})}^kM)$, we use~\eqref{pf:approximation_error_nonlocal_laplacian_8} to derive that
	\begin{align}
	L_{P,\varepsilon}\Delta_P^kf(x) & = \sigma_{\eta} \Delta_P^{k + 1}f(x) + \sum_{j = 1}^{(s - 2k - 1)/2 - } g_{2j + 2}^{s - 2k}(x)\varepsilon^{2j} + g_{s - 2k}^{s - 2k}(x) \varepsilon^{s - 2k - 2} \nonumber \\
	& = \sigma_{\eta} \Delta_P^{k + 1}f(x) + \sum_{j = 1}^{(s - 1)/2 - (k + 1)} g_{2(k + 1 + j)}(x)\varepsilon^{2j} + g_{s}(x) \varepsilon^{s - 2(k + 1)}, \label{pf:approximation_error_nonlocal_laplacian_9}
	\end{align}
	where in the second equality we have simply used the fact $g_j^{\ell}(x) = g_{(s - \ell) + j}(x)$ to rewrite the equation.
	\item Suppose $j < \floor{(s - 1)/2} - k$. Then we use~\eqref{pf:approximation_error_nonlocal_laplacian_8} to derive that
	\begin{align*}
	L_{P,\varepsilon}g_{2(j + k)}(x) & = \sigma_{\eta}\Delta_P g_{2(j + k)}(x) + \sum_{i = 1}^{\floor{(s - 2j - 2k - 1)/2} - 1} g_{2(i + 1)}^{s - 2(j + k)}(x)\varepsilon^{2i} + g_{s - 2(j + k)}^{s - 2(j + k)}(x) \varepsilon^{s - 2(j + k + 1)} \\
	& = g_{2(j + k + 1)}(x) + \sum_{i = 1}^{\floor{(s - 1)/2} - (j + k + 1)} g_{2(i + j + k + 1)}(x)\varepsilon^{2i} + g_{s}(x) \varepsilon^{s - 2(j + k + 1)},
	\end{align*}
	where in the second equality we have again used $g_j^{\ell}(x) = g_{(s - \ell) + j}(x)$, and also written $\sigma_{\eta} \Delta_Pf = g_{2}^{s - 2(j + k)} = g_{2(j + k + 1)}$, since the particular dependence on the Laplacian $\Delta_P$ will not matter. From here, multiplying by $\varepsilon^{2j}$, we conclude that
	\begin{align}
	\varepsilon^{2j} L_{P,\varepsilon}g_{2(j + k)}(x) & = g_{2(j + k + 1)}(x) \varepsilon^{2j} + \sum_{i = 1}^{\floor{(s - 1)/2} - (j + k + 1)} g_{2(i + j + k + 1)}(x)\varepsilon^{2(i + j)} + g_{s}(x) \varepsilon^{s - 2(k + 1)} \nonumber \\ 
	& = g_{2(j + k + 1)}(x) \varepsilon^{2j} + \sum_{m = 1}^{\floor{(s - 1)/2} - (k + 1)}  g_{2(m + k + 1)}(x)\varepsilon^{2m} + g_{s}(x) \varepsilon^{s - 2(k + 1)} \label{pf:approximation_error_nonlocal_laplacian_10},
	\end{align}
	with the second equality following upon changing variables to $m = i + j$.

	On the other hand if $j = \floor{(s - 1)/2} - k$, then the calculation is much simpler,
	\begin{equation}
	\label{pf:approximation_error_nonlocal_laplacian_11}
	\varepsilon^{2j} L_{P,\varepsilon}g_{2(j + k)}(x) = g_{s - 2(j + k)}^{s - 2(j + k)}(x) \varepsilon^{2j} \varepsilon^{s - 2(j + k) - 2} = g_s(x) \varepsilon^{s - 2(k + 1)}.
	\end{equation}
	\item Finally, it follows immediately from~\eqref{pf:approximation_error_nonlocal_laplacian_8} that
	\begin{equation}
	\label{pf:approximation_error_nonlocal_laplacian_12}
	L_{P,\varepsilon}g_s(x) \varepsilon^{s - 2k} = g_{s}(x) \varepsilon^{s - 2(k + 1)}.
	\end{equation}
\end{enumerate}
Plugging~\eqref{pf:approximation_error_nonlocal_laplacian_9}-\eqref{pf:approximation_error_nonlocal_laplacian_12} back into~\eqref{pf:approximation_error_nonlocal_laplacian_8.5} proves the claim.

\subsection{Boundary behavior of non-local Laplacian}
\label{subsec:boundary_behavior_nonlocal_laplacian}

In Lemma~\ref{lem:approximation_error_nonlocal_laplacian_boundary}, we establish that if $f$ is Sobolev smooth of order $s > 2k$ and zero-trace, then near the boundary of $\mc{X}$ the non-local Laplacian $L_{P,\varepsilon}^kf$ is close to $0$ in the $L^2$-sense.
\begin{lemma}
	\label{lem:approximation_error_nonlocal_laplacian_boundary}
	Assume Model~\ref{def:model_flat_euclidean}. Let $s,k \in \mathbb{N}$. Suppose that $f \in H_0^{s}(\mc{X};M)$. Then there exist numbers $c,C > 0$ that do not depend on $M$, such that for all $\varepsilon < c$, 
	\begin{equation*}
	\|L_{P,\varepsilon}^kf\|_{L^2(\partial_{k\varepsilon}\mc{X})}^2 \leq C \varepsilon^{2(s - 2k)}M^2.
	\end{equation*}
\end{lemma}

\paragraph{Proof (of Lemma~\ref{lem:approximation_error_nonlocal_laplacian_boundary})}
Applying Lemma~\ref{lem:l2estimate_nonlocal_laplacian}, we have that
\begin{equation*}
\|L_{P,\varepsilon}^kf\|_{L^2(\partial_{k\varepsilon}(\mc{X}))}^2 \leq \frac{(Cp_{\max})^{2}}{\varepsilon^4} \|L_{P,\varepsilon}^{k - 1}f\|_{L^2(\partial_{k\varepsilon}(\mc{X}))}^2 \leq \cdots \leq \frac{(Cp_{\max})^{2}}{\varepsilon^{4k}} \|f\|_{L^2(\partial_{k\varepsilon}(\mc{X}))}^2   
\end{equation*}
Thus it remains to show that for all $\varepsilon < c$,
\begin{equation}
\label{pf:approximation_error_nonlocal_laplacian_boundary_0}
\|f\|_{L^2(\partial_{k\varepsilon}(\mc{X}))}^2 = \int_{\partial_{k\varepsilon}(\mc{X})} \bigl(f(x)\bigr)^2 \,dx \leq C_1 \varepsilon^{2s} \|f\|_{H^s(\mc{X})}^2.
\end{equation}
We will build to~\eqref{pf:approximation_error_nonlocal_laplacian_boundary_0} by a series of intermediate steps, following the same rough structure as the proof of Theorem 18.1 in \citet{leoni2017}. For simplicity, we will take $k = 1$; the exact same proof applies to the general case upon assuming $\varepsilon < c/k$.

\underline{\textit{Step 1: Local Patch.}}
To begin, we assume that for some $c_0 > 0$ and a Lipschitz mapping $\phi: \Reals^{d - 1} \to [-c_0,c_0]$, we have that $f \in C_c^{\infty}(U_{\phi}(c_0))$, where 
\begin{equation*}
U_{\phi}(c_0) = \Bigl\{y \in Q(0,c_0): \phi(y_{-d}) \leq y_d\Bigr\}, 
\end{equation*}
and here $Q(0,c_0)$ is the $d$-dimensional cube of side length $c_0$, centered at $0$. We will show that for all $0 < \varepsilon < c_0$, and for the tubular neighborhood $V_{\phi}(\varepsilon) = \{y \in Q(0,c_0): \phi(y_{-d}) \leq y_d \leq \phi(y_{-d}) + \varepsilon\}$, we have that
\begin{equation*}
\int_{V_{\phi}(\varepsilon)} |f(x)|^2 \,dx \leq C\varepsilon^{2s} \|f\|_{H^s(U_{\phi}(c_0))}^2.
\end{equation*}
For a given $y = (y',y_d) \in V_{\phi}(\varepsilon)$, let $y_0 = (y',\phi(y'))$. Taking the Taylor expansion of $f(y)$ around $y = y_0$ because $u$ is compactly supported in $V_{\phi}$ it follows that,
\begin{align*}
f(y) & = f(y_0) + \sum_{j = 1}^{s - 1} \frac{1}{j!} D^{je_d}f(y_0) \bigl(y_d - \phi(y')\bigr)^j + \frac{1}{(s - 1)!}\int_{\phi(y')}^{y_d} (1 - t)^{s - 1} D^{se_d}f(y',z) \bigl(y_d - z\bigr)^{s - 1} \,dz \Longrightarrow\\
|f(y)| & \leq C\varepsilon^{s - 1}\int_{\phi(y')}^{y_d} \bigl|D^{se_d}f(y',z)\bigr| \,dz. 
\end{align*}
Consequently, by squaring both sides and applying Cauchy-Schwarz, we have that
\begin{equation*}
|f(y)|^2 \leq C\varepsilon^{2(s - 1)} \biggl(\int_{\phi(y')}^{y_d} \bigl|D^{se_d}f(y',z)\bigr| \,dz\biggr)^2 \leq C\varepsilon^{2s - 1} \int_{\phi(y')}^{y_d} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz.
\end{equation*}
Applying this bound for each $y \in V_{\phi}(\varepsilon)$, and then integrating, we obtain
\begin{align}
\int_{V_{\phi}(\varepsilon)} |f(y)|^2 \,dy & \leq \int_{Q_{d - 1}(c_0)} \int_{\phi(y')}^{\phi(y') + \varepsilon} |f(y',y_d)|^2 \,dy_d \,dy' \nonumber \\
& \leq C\varepsilon^{2s - 1}\int_{Q_{d - 1}(c_0)}  \int_{\phi(y')}^{\phi(y') + \varepsilon} \int_{\phi(y')}^{y_d} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz \,dy_d \,dy' \label{pf:approximation_error_nonlocal_laplacian_boundary_1}
\end{align}
where we have written $Q_{d - 1}(0,c_0)$ for the $d - 1$ dimensional cube of side length $c_0$, centered at $0$. Exchanging the order of the inner two integrals then gives
\begin{align*}
\int_{\phi(y')}^{\phi(y') + \varepsilon} \int_{\phi(y')}^{y_d} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz \,dy_d & = \int_{\phi(y')}^{\phi(y') + \varepsilon} \int_{z}^{\varepsilon} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dy_d \,dz \\
& \leq C \varepsilon \int_{\phi(y')}^{\phi(y') + \varepsilon} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz \\
& \leq C \varepsilon \int_{\phi(y')}^{c_0} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz.
\end{align*}
Finally, plugging back into~\eqref{pf:approximation_error_nonlocal_laplacian_boundary_1}, we conclude that
\begin{equation*}
\int_{V_{\phi}(\varepsilon)} |f(y)|^2 \,dy \leq C \varepsilon^{2s} \int_{Q_{d - 1}(0,c_0)} \int_{\phi(y')}^{c_0} \bigl|D^{se_d}f(y',z)\bigr|^2 \,dz \,dy' \leq C \varepsilon^{2s} |u|_{H^s(U_{\phi}(c_0))}^2.
\end{equation*}

\underline{\textit{Step 2: Rigid motion of local patch.}} Now, suppose that at a point $x_0 \in \partial \mc{X}$, there exists a rigid motion $T: \Rd \to \Rd$ for which $T(x_0) = 0$, and a number $C_0$ such that for all $\varepsilon \cdot C_0 \leq c_0$, 
\begin{equation*}
T\bigl(Q_{T}(x_0,c_0) \cap \partial_{\varepsilon}\mc{X}\bigr) \subseteq V_{\phi}\bigl(C_0\varepsilon\bigr) \quad\textrm{and}\quad T\bigl(Q_T(x_0,c_0) \cap \mc{X}\bigr) = U_{\phi}(c_0).
\end{equation*}
Here $Q_{T}(x_0,c_0))$ is a (not necessarily coordinate-axis-aligned) cube of side length $c_0)$, centered at $x_0$. Define $v(y) := f(T^{-1}(y))$ for $y \in U_{\phi}(c_0)$. If $u \in C_c^{\infty}(\mc{X})$, then $v \in C_c^{\infty}(U_{\phi}(c_0))$, and moreover $\|v\|_{H^s(U_{\phi}(c_0))}^2 = \|f\|_{H^s(Q_{T}(x_0,c_0) \cap \mc{X})}^2$. Therefore, using the upper bound that we derived in Step 1,
\begin{equation*}
\int_{V_{\phi}(C_0 \cdot \varepsilon)} |v(y)|^2 \,dy \leq C \varepsilon^{2s} \|v\|_{H^s(U_{\phi}(c_0))}^2,
\end{equation*}
we conclude that
\begin{align*}
\int_{Q_{T}(x_0,c_0) \cap \partial_{\varepsilon}\mc{X}} |f(x)|^2 \,dx & = \int_{T(Q_T(x_0,c_0)) \cap \partial_{\varepsilon}\mc{X})} |v(y)|^2 \,dy \\
& \leq \int_{V_{\phi}(C_0 \cdot \varepsilon)} |v(y)|^2 \,dy \\
& \leq C \varepsilon^{2s} \|v\|_{H^s(U_{\phi}(c_0))}^2 = C \varepsilon^{2s} \|f\|_{H^s(Q_{T}(x_0,c_0)) \cap \mc{X})}^2 \leq C \varepsilon^{2s} \|f\|_{H^s(\mc{X})}^2.
\end{align*}

\underline{\textit{Step 3: Lipschitz domain}}.
Finally, we deal with the case where $\mc{X}$ is assumed to be an open, bounded subset of $\Rd$, with Lipschitz boundary. In this case, at every $x_0 \in \partial \mc{X}$, there exists a rigid motion $T_{x_0}: \Rd \to \Rd$ such that $T_{x_0}(x_0) = 0$, a number $c_0(x_0)$, a Lipschitz function $\phi_{x_0}:\Reals^{d - 1} \to [-c_0,c_0]$, and a number $C_0(x_0)$, such that for all $\varepsilon \cdot C_0(x_0) \leq c_0(x_0)$,
\begin{equation*}
T\bigl(Q_{T}(x_0,c_0(x_0)) \cap \partial_{\varepsilon}\mc{X}\bigr) \subseteq V_{\phi}\bigl(C_0(x_0) \cdot \varepsilon\bigr) \quad\textrm{and}\quad T\bigl(Q_T(x_0,c_0(x_0)) \cap \mc{X}\bigr) = U_{\phi}(c_0(x_0)).
\end{equation*}
Therefore for every $x_0 \in \partial \mc{X}$, it follows from the previous step that
\begin{equation*}
\int_{Q_{T_{x_0}}(x_0,c_0(x_0)) \cap \partial_{\varepsilon}\mc{X}} |f(x)|^2 \,dx \leq C(x_0) \varepsilon^{2s} \|f\|_{H^s(\mc{X})}^2,
\end{equation*}
where on the right hand side $C(x_0)$ is a constant that may depend on $x_0$, but not on $u$ or $\varepsilon$.

We conclude by taking a collection of cubes that covers $\partial_{\varepsilon}\mc{X}$ for all $\epsilon$ sufficiently small. First, we note that by a compactness argument there exists a finite subset of the collection of cubes $\{Q_{T_{x_0}}(x_0,c_0(x_0)/2): x_0 \in \partial\mc{X} \}$ which covers $\partial \mc{X}$, say $Q_{T_{x_1}}(x_1,c_0(x_1)/2),\ldots, Q_{T_{x_N}}(x_N,c_0(x_N)/2)$. Then, for any $\varepsilon \leq \min_{i = 1,\ldots,N} c_0(x_i)/2$, it follows from the triangle inequality that
\begin{equation*}
\partial_{\varepsilon}\mc{X} \subseteq \bigcup_{i = 1}^{N} Q_{T_{x_i}}(x_i, c_0(x_i)).
\end{equation*}
As a result,
\begin{equation*}
\int_{\partial_{\varepsilon}\mc{X}} |f(x)|^2 \leq \sum_{i = 1}^{N} \int_{Q_{T_{x_i}}(x_i, c_0(x_i)) \cap \partial_{\varepsilon}(\mc{X})} |f(x)|^2 \leq  \varepsilon^{2s} \|f\|_{H^s(\mc{X})}^2 \sum_{i = 1}^{N}C_0(x_i),
\end{equation*}
which proves the claim of~\eqref{pf:approximation_error_nonlocal_laplacian_boundary_0}.

\subsection{Estimate of non-local Sobolev seminorm}
\label{subsec:estimate_nonlocal_seminorm}

Now, we use the results of the preceding two sections to prove~\eqref{pf:graph_seminorm_ho_2}. We will divide our analysis in two cases, depending on whether $s$ is odd or even, but before we do this we state some facts that will be applicable to both cases. First, we  recall that $L_{P,\varepsilon}$ is self-adjoint in $L^2(P)$, meaning $\dotp{L_{P,\varepsilon}f}{g}_{P} = \dotp{f}{L_{P,\varepsilon}g}_{P}$ for all $f, g \in L^2(P)$. We also recall the definition of the Dirichlet energy $E_{P,\varepsilon}(f;\mc{X})$,
\begin{equation}
\label{eqn:dirichlet_energy}
\dotp{L_{P,\varepsilon}f}{f}_{P} = \frac{1}{\varepsilon^{d + 2}}\int_{\mc{X}} \int_{\mc{X}} \bigl(f(x) - f(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) =: E_{P,\varepsilon}(f;\mc{X}).
\end{equation}
Finally, we recall a result of~\cite{green2021}: there exist constants $c_0$ and $C_0$ which do not depend on $M$, such that for all $\varepsilon < c_0$ and for any $f \in H^1(\mc{X};M)$,
\begin{equation}
\label{pf:estimate_nonlocal_seminorm_1}
E_{P,\varepsilon}(f;\mc{X}) \leq C_0 M^2.
\end{equation}
\paragraph{Case 1: $s$ odd.}
Suppose $s$ is odd, so that $s \geq 3$. Taking $k = (s - 1)/2$, we use the self-adjointness of $L_{P,\varepsilon}$ to relate the non-local semi-norm $\dotp{L_{P,\varepsilon}^sf}{f}_{P}$ to a non-local Dirichlet energy,
\begin{equation*}
\dotp{L_{P,\varepsilon}^sf}{f}_P = \dotp{L_{P,\varepsilon}^{k + 1}f}{L_{P,\varepsilon}^{k}f}_P = E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\mc{X}).
\end{equation*}
We now separate this energy into integrals over $\mc{X}_{k\varepsilon}$ and $\partial_{k\varepsilon}(\mc{X})$,
\begin{align}
E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\mc{X}) & = \frac{1}{\varepsilon^{d + 2}}\Biggl\{\int_{\mc{X}_{k\varepsilon}} \int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - L_{P,\varepsilon}^kf(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) \nonumber \\
& \quad + \int_{\partial_{k\varepsilon}\mc{X}} \int_{\partial_{k\varepsilon}\mc{X}} \bigl(L_{P,\varepsilon}^kf(x) - L_{P,\varepsilon}^kf(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x)\Biggr\} \nonumber \\
& := E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\mc{X}_{k\varepsilon}) + E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\partial_{k\varepsilon}\mc{X}) \label{pf:estimate_nonlocal_seminorm_1.5}
\end{align}
and upper bound each energy separately. For the first term, we add and substract $\sigma_{\eta}^k\Delta_P^kf(x)$ and $\sigma_{\eta}^k\Delta_P^kf(x')$ within the integrand, then use the triangle inequality and the symmetry between $x$ and $x'$ to deduce that
\begin{equation}
\label{pf:estimate_nonlocal_seminorm_2}
E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\mc{X}_{k\varepsilon}) \leq 3 \sigma_{\eta}^{2k} E_{P,\varepsilon}(\Delta_P^kf;\mc{X}_{k\varepsilon}) + \frac{2}{\varepsilon^{d + 2}}\int_{\mc{X}_{k\varepsilon}} \int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_P^kf(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x).
\end{equation}
Noticing that $\Delta_P^kf \in H^1(\mc{X};\|p\|_{C^{s - 1}(\mc{X})}^kM)$, we use~\eqref{pf:estimate_nonlocal_seminorm_1} to conclude that $E_{P,\varepsilon}(\Delta_P^kf;\mc{X}_{k\varepsilon}) \leq C_0M^2$. On the other hand, it follows from Assumption~\ref{asmp:kernel_flat_euclidean} and~\eqref{eqn:approximation_error_nonlocal_laplacian_1} that
\begin{align*}
\frac{2}{\varepsilon^{d + 2}}\int_{\mc{X}_{k\varepsilon}} \int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_P^kf(x)\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) & \leq \frac{2p_{\max}}{\varepsilon^{2}}\int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^kf(x) - \sigma_{\eta}^k \Delta_P^kf(x)\bigr)^2 \,dP(x) \\
& \leq C_1M^2.
\end{align*}
Plugging these two bounds into~\eqref{pf:estimate_nonlocal_seminorm_2} gives the desired upper bound on $E_{P,\varepsilon}(L_{P,\varepsilon}^{k};\mc{X}_{k\varepsilon})$.

For the second term in~\eqref{pf:estimate_nonlocal_seminorm_1.5}, we apply Lemmas~\ref{lem:dirichlet_estimate_nonlocal_laplacian} and~\ref{lem:approximation_error_nonlocal_laplacian_boundary} and conclude that,
\begin{equation*}
E_{P,\varepsilon}(L_{P,\varepsilon}^{k}f;\partial_{k\varepsilon}\mc{X}) \leq \frac{4p_{\max}^2}{\varepsilon^2}\|L_{P,\varepsilon}^kf\|_{L^2(\partial_{k\varepsilon}\mc{X})} \leq C M^2.
\end{equation*}
\paragraph{Case 2: $s$ even.}
If $s \in \mathbb{N}$ is even, $s \geq 2$, then letting $k = (s - 2)/2$, the  self-adjointness of $L_{P,\varepsilon}$ implies
\begin{equation*}
\dotp{L_{P,\varepsilon}^sf}{f}_P = \|L_{P,\varepsilon}^{k + 1}f\|_P^2.
\end{equation*}
As in the first case, we divide the integral up into the interior region $\mc{X}_{k\varepsilon}$ and the boundary region $\partial_{k\varepsilon}\mc{X}$,
\begin{equation}
\label{pf:estimate_nonlocal_seminorm_3}
\|L_{P,\varepsilon}^{k + 1}f\|_P^2 \leq p_{\max} \|L_{P,\varepsilon}^{k + 1}f\|_{L^2(\mc{X})}^2 \leq p_{\max}\biggl\{\int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^{k + 1}f(x)\bigr)^2 \,dP(x) + \int_{\partial_{k\varepsilon}\mc{X}} \bigl(L_{P,\varepsilon}^{k + 1}f(x)\bigr)^2 \,dP(x)\biggr\},
\end{equation}
and upper bound each term separately. For the first term, adding and subtracting $\sigma_{\eta}^k \Delta_P^kf(x)$ gives
\begin{align*}
\int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}^{k + 1}f(x)\bigr)^2 \,dP(x) & \leq 2\int_{\mc{X}_{k\varepsilon}} \bigl(L_{P,\varepsilon}\Delta_P^kf(x)\bigr)^2 \,dP(x)  + 2 \int_{\mc{X}_{k\varepsilon}} \Bigl(L_{P,\varepsilon}\bigl(L_{P,\varepsilon}^kf- \sigma_{\eta}\Delta_P^kf\bigr)(x)\Bigr)^2 \,dP(x) \\
& \overset{(i)}{\leq} CM^2  + 2 \int_{\mc{X}_{k\varepsilon}} \Bigl(L_{P,\varepsilon}\bigl(L_{P,\varepsilon}^kf- \sigma_{\eta}\Delta_P^kf\bigr)(x)\Bigr)^2 \,dP(x) \\
& \overset{(ii)}{\leq} CM^2  + \frac{Cp_{\max}^2}{\varepsilon^{2}} \|L_{P,\varepsilon}^kf- \sigma_{\eta}\Delta_P^kf\|_{L^2(\mc{X}_{k\varepsilon})}^2 \\
& \overset{(iii)}{\leq} CM^2,
\end{align*}
with $(i)$ following from~\eqref{pf:approximation_error_nonlocal_laplacian_0} since $\Delta_P^kf \in H^2(\mc{X};M\|p\|_{C^{s - 1}(\mc{X})}^l)$, $(ii)$ following from Lemma~\ref{lem:l2estimate_nonlocal_laplacian}, and $(iii)$ following from~\eqref{eqn:approximation_error_nonlocal_laplacian_2}.

Then Lemma~\ref{lem:approximation_error_nonlocal_laplacian_boundary} shows that the second term in~\eqref{pf:estimate_nonlocal_seminorm_3} satisfies
\begin{equation*}
\int_{\partial_{k\varepsilon}\mc{X}} \bigl(L_{P,\varepsilon}^{k + 1}f(x)\bigr)^2 \,dP(x) \leq CM^2.
\end{equation*}

\subsection{Assorted integrals}
\label{subsec:integrals}

\begin{lemma}
	\label{lem:l2estimate_nonlocal_laplacian}
	Assume Model~\ref{def:model_flat_euclidean}. Suppose $f \in L^2(U;M)$ for a Borel set $U \subseteq \mc{X}$, and let $L_{P,\varepsilon}$ be defined with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_flat_euclidean}. Then there exists a constant $C$ which does not depend on $f$ or $M$ such that
	\begin{equation}
	\label{eqn:l2estimate_nonlocal_laplacian}
	\|L_{P,\varepsilon}f\|_{L^2(U)} \leq \frac{2 p_{\max}}{\varepsilon^2} \|f\|_{L^2(U)}
	\end{equation}
\end{lemma}

\begin{lemma}
	\label{lem:dirichlet_estimate_nonlocal_laplacian}
	Assume Model~\ref{def:model_flat_euclidean}. Suppose $f \in L^2(U;M)$ for a Borel set $U \subseteq \mc{X}$, and let $L_{P,\varepsilon}$ be defined with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_flat_euclidean}. Then there exists a constant $C$ which does not depend on $f$ or $M$ such that
	\begin{equation}
	\label{eqn:dirichlet_estimate_nonlocal_laplacian}
	E_{P,\varepsilon}(f;U) \leq \frac{4 p_{\max}^2}{\varepsilon^2} \|f\|_{L^2(U)}^2
	\end{equation}
\end{lemma}

\begin{lemma}
	\label{lem:graph_seminorm_bias2}
	Assume Model~\ref{def:model_flat_euclidean}. Suppose $f \in H^1(\mc{X};M)$, and let $D_if$ be defined with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_flat_euclidean}. Then there exists a constant $C$ which does not depend on $f$ or $M$, such that for any $i \in [n]$ and $\bj \in [n]^s$,
	\begin{equation*}
	\Ebb\Bigl[|D_{\bj}f(X_i)| \cdot |f(X_i) - f(X_{\bj_1})|\Bigr] \leq C \varepsilon^{2 + dk}M^2,
	\end{equation*}
	where $k + 1$ is the number of distinct indices in $i\bj$. 
\end{lemma}

\paragraph{Proof (of Lemma~\ref{lem:l2estimate_nonlocal_laplacian}).}
We fix a version of $f \in L^2(U)$, so that we may speak of its pointwise values.

At a given point $x \in U$, we can upper bound $|L_{P,\varepsilon}f(x)|^2$ using the Cauchy-Schwarz inequality as follows,
\begin{align*}
|L_{P,\varepsilon}f(x)|^2 & \leq \biggl(\frac{p_{\max}}{\varepsilon^{2 + d}}\biggr)^2 \Biggl(\int_U \bigl(|f(x')| + |f(x)|\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx'\Biggr)^2 \\
& \leq \biggl(\frac{p_{\max}}{\varepsilon^{2 + d}}\biggr)^2 \Biggl(\int_U \bigl(|f(x')| + |f(x)|\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \cdot \int \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \Biggr) \\
& = \frac{p_{\max}^2}{\varepsilon^{4 + d}} \int_U \bigl(|f(x')| + |f(x)|\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx'.
\end{align*}
The equality follows by the assumption $\int_{\Rd} \eta(\|z\|)\,dx = 1$ in~\ref{asmp:kernel_flat_euclidean}. Integrating over all $x \in U$, it follows from the triangle inequality that
\begin{align}
\|L_{P,\varepsilon}\|_{L^2(U)}^2 & \leq \frac{2 p_{\max}^2}{\varepsilon^{4 + d}} \int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx \nonumber \\
& \leq \frac{2 p_{\max}^2}{\varepsilon^{4 + d}} \int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx. \label{pf:l2estimate_nonlocal_laplacian_1}
\end{align}
Finally, using Fubini's Theorem we determine that
\begin{equation}
\int_{U} \int_{U} \bigl(|f(x')|^2 + |f(x)|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx = 2 \int_{U} \int_{U} |f(x)|^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx \leq 2 \varepsilon^d \int_{U} |f(x)|^2 \,dx = 2\varepsilon^d \|f\|_{L^2(U)}^2,
\label{pf:l2estimate_nonlocal_laplacian_2}
\end{equation}
and by combining~\eqref{pf:l2estimate_nonlocal_laplacian_1} and~\eqref{pf:l2estimate_nonlocal_laplacian_2} we conclude that
\begin{equation*}
\|L_{P,\varepsilon}\|_{L^2(U)}^2 \leq \frac{4p_{\max}^2}{\varepsilon^4} \|f\|_{L^2(U)}^2. 
\end{equation*}

\paragraph{Proof (of Lemma~\ref{lem:dirichlet_estimate_nonlocal_laplacian}).}
We have
\begin{equation*}
E_{P,\varepsilon}(f) = \frac{1}{\varepsilon^{2 + d}} \int_{U} \int_{U} \bigl(f(x) - f(x')\bigr)^2 \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dP(x') \,dP(x) \leq \frac{2p_{\max}^2}{\varepsilon^{2 + d}} \int_{U} \int_{U} \bigl(|f(x)|^2 + |f(x')|^2\bigr) \eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) \,dx' \,dx,
\end{equation*}
and the claim follows from~\eqref{pf:l2estimate_nonlocal_laplacian_2}.

\paragraph{Proof (of Lemma~\ref{lem:graph_seminorm_bias2}).}
Let $G_{n,\varepsilon}[X_{i\bj}]$ be the subgraph induced by vertices $X_i, X_{\bj_1},\ldots,X_{\bj_s}$. We make two observations. First, in order for $|D_{\bj}f(X_i)| \cdot |f(X_i)  - f(X_j)|$ to be non-zero, it must be the case that the subgraph $G_{n,\varepsilon}[X_{i\bj}]$ is connected. Second, noting that for any indices $i$ and $j$,
\begin{equation*}
|D_{ij}f(x)| \leq \Bigl(|D_jf(X_i)| + |D_jf(x)|\Bigr)\|\eta\|_{\infty}, 
\end{equation*}
a straightforward inductive argument implies that 
\begin{equation*}
|D_{\bj}f(X_i)| \leq s\|\eta\|_{\infty}^s \sum_{j \in i\bj} |D_{\bj_s}f(X_j)|.
\end{equation*}
Combining these two observations, we reduce the task to upper bounding the product of two (first-order) differences,
\begin{align*}
\mathbb{E}\Bigl[ |D_{\bj}f(X_i)| |f(X_i) - f(X_{\bj_1})| \Bigr] & = \mathbb{E}\Bigl[ |D_{\bj}f(X_i)| |f(X_i) - f(X_{\bj_1})| \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\} \Bigr] \\
& \leq s\|\eta\|_{\infty}^s \sum_{j \in i\bj} \mathbb{E}\Bigl[ |D_{\bj_s}f(X_j)| \cdot |f(X_i) - f(X_{\bj_1})| \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\}   \Bigr] \\
& \leq s\|\eta\|_{\infty}^s \sum_{j \in i\bj} \mathbb{E}\Bigl[ |f(X_j) - f(X_{\bj_s})| \cdot |f(X_i) - f(X_{\bj_1})| \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\}   \Bigr] 
\end{align*}
Next, from the Cauchy-Schwarz inequality we have that for any $j \in \bj$,
\begin{align*}
& \mathbb{E}\Bigl[ |f(X_j) - f(X_{\bj_s})| \cdot |f(X_i) - f(X_{\bj_1})| \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\}   \Bigr] \\
& \quad \leq \sqrt{\mathbb{E}\Bigl[ |f(X_j) - f(X_{\bj_s})|^2 \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\} \Bigr]} \cdot \sqrt{\mathbb{E}\Bigl[ |f(X_j) - f(X_{\bj_s})|^2 \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\} \Bigr]} \\
& \quad = \mathbb{E}\Bigl[ |f(X_j) - f(X_{i})|^2 \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\} \Bigr],
\end{align*}
with the equality following since each $X_i$ are identically distributed. Marginalizing out the contribution of all indices in $\bj$ not equal to $i$ or $j$ gives
\begin{align}
\mathbb{E}\Bigl[ |f(X_j) - f(X_{i})|^2 \cdot \1\bigl\{G_{n,\varepsilon}[X_{i\bj}]~~\textrm{is connected.} \bigr\} \Bigr] & \leq \bigl((s + 1)p_{\max}\nu_d\varepsilon^d\bigr)^{|i\bj\setminus\{j \cup i\}|} \cdot \mathbb{E}\Bigl[ |f(X_j) - f(X_{i})|^2 \1\{\|X_i - X_j\| \leq \varepsilon\}\Bigr] \nonumber \\
& \leq \bigl((s + 1)p_{\max}\nu_d\varepsilon^d\bigr)^{|i\bj\setminus\{j \cup i\}|} \cdot p_{\max}^2 \nu_d \varepsilon^{2 + d} M^2 \label{pf:graph_seminorm_bias2_1}
\end{align}
with the second inequality following from the proof of Lemma~1 in~\cite{green2021}. Finally, we notice that $|i\bj\setminus \{i \cup j\}| + 1 = k$, so that~\eqref{pf:graph_seminorm_bias2_1} gives the desired result.

\section{Graph Sobolev semi-norm, manifold domain}
\label{sec:graph_quadratic_form_manifold}
In this section we prove Proposition~\ref{prop:graph_seminorm_manifold}. Note that when $s = 1$, the upper bound~\eqref{eqn:graph_seminorm_manifold} follows immediately from Lemma~\ref{lem:dirichlet_energy_sobolev} and Markov's inequality.

On the other hand when $s = 2$ or $s = 3$, we prove Proposition~\ref{prop:graph_seminorm_manifold} by first establishing some intermediate results, many of which are analogous to results we have already shown in the flat Euclidean case. Indeed, in some ways the proof will be simpler in the manifold setting than in the flat Euclidean case: there is no boundary, and we do not need to analyze the iterated nonlocal Laplacian $L_{P,\varepsilon}^j$ for $j > 1$.

That being said, as mentioned in our main text, in the manifold setting there is some extra error induced by using Euclidean rather than geodesic distance. We upper bound this error by comparing $L_{P,\varepsilon}$ to an alternative nonlocal Laplacian $\wt{L}_{P,\varepsilon}$, which is defined with respect to geodesic distance. Precisely, let $d_{\mc{X}}(x,x')$ denote the geodesic distance between $x,x' \in \mc{X}$, and define
\begin{equation*}
\wt{L}_{P,\varepsilon}f(x) := \int_{\mc{X}} \bigl(f(x') - f(x)\bigr) \eta \biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr) p(x') \,dx'.
\end{equation*}

We show the following results, each of which hold under the same assumptions as Proposition~\ref{prop:graph_seminorm_manifold}.
\begin{itemize}
	\item In Section~\ref{subsec:manifold_decomposition_graph_seminorm} we show that the graph Sobolev seminorm $\dotp{L_{n,\varepsilon}^sf}{f}_n$ is upper bounded by the sum of a nonlocal seminorm and a pure bias term: specifically, with probability at least $1 - 2\delta$,
	\begin{equation}
	\label{pf:graph_seminorm_manifold_1}
	\dotp{L_{n,\varepsilon}^sf}{f}_n \leq \frac{\dotp{L_{P,\varepsilon}^sf}{f}_P}{\delta} + C_1\frac{\varepsilon^2}{n\varepsilon^{2s + m}}M^2.
	\end{equation}
	This upper bound is essentially the same as~\eqref{pf:graph_seminorm_ho_1}, but with the intrinsic dimension $m$ taking the place of the ambient dimension $d$. The pure bias term will be of at most constant order when $\varepsilon \gtrsim n^{-1/(2(s-1) + m)}$. 
	\item In Section~\ref{subsec:error_euclidean_distance}, we show that the error incurred by using the ``wrong'' metric is negligible. Precisely, we find that
	\begin{equation}
	\label{eqn:nonlocal_laplacian_geodesic_error}
	\|L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f\|_{L^2(\mc{X})}^2 \leq C_2 \varepsilon^2 |f|_{H^1(\mc{X})}^2.
	\end{equation}
	\item In Section~\ref{subsec:manifold_approximation_error_nonlocal_laplacian}, we analyze the approximation error of $\wt{L}_{P,\varepsilon}$. We show that when $f \in H^2(\mc{X})$ and $p \in C^1(\mc{X})$, 
	\begin{equation}
	\label{eqn:nonlocal_laplacian_approximation_error_manifold_l2}
	\|\wt{L}_{P,\varepsilon}f\|_{L^2(\mc{X})}^2 \leq C_3 \|f\|_{H^2(\mc{X})}^2,
	\end{equation}
	whereas if $f \in H^3(\mc{X})$ and $p \in C^2(\mc{X})$, 
	\begin{equation}
	\label{eqn:nonlocal_laplacian_approximation_error_manifold_sobolev}
	\|\wt{L}_{P,\varepsilon}f - \sigma_{\eta}\Delta_Pf\|_{L^2(\mc{X})}^2 \leq C_3 \varepsilon^2 \|f\|_{H^3(\mc{X})}^2.
	\end{equation}
	\item In Section~\ref{subsec:manifold_estimate_nonlocal_seminorm}, we use the results of the preceding two sections to show that if $f \in H^s(\mc{X})$ and $p \in C^{s - 1}(\mc{X})$, then 
	\begin{equation}
	\label{eqn:manifold_nonlocal_seminorm}
	\dotp{L_{P,\varepsilon}^sf}{f}_P \leq C_4\|f\|_{H^s(\mc{X})}^2.
	\end{equation}
	\item In Section~\ref{subsec:manifold_integrals} we state some technical results used in the previous sections.
\end{itemize}
We point out that when $f$ is H\"{o}lder smooth, results analogous to~\eqref{eqn:nonlocal_laplacian_approximation_error_manifold_sobolev} have been established in \citet{calder2019}. When $f$ is Sobolev smooth, our analysis (which relies heavily on Taylor expansions) is largely similar, except that the remainder term in the relevant Taylor expansion will be bounded in $L^2(\mc{X})$ norm rather than $L^{\infty}(\mc{X})$ norm. This is analogous to the situation in the flat Euclidean model.

In the proof of~\eqref{pf:graph_seminorm_manifold_1}-\eqref{eqn:manifold_nonlocal_seminorm}, we recall the following estimates from differential geometry: (i) letting $K_0$ be an upper bound on the absolute value of the sectional curvatures of $\mc{X}$, $K_0 \leq 2R$, and  letting (ii) $i_0$ be a lower bound on the injectivity radius of $\mc{X}$, $i_0 \geq \pi R$; see Proposition~1 of \cite{aamari2019}. Additionally, recall that for all $\delta < i_0$, the exponential map $\exp_x: B_m(0,\delta) \subset T_x(\mc{X}) \to B_{\mc{X}}(x,\delta) \subset \mc{X}$ is a diffeomorphism for all $x \in \mc{X}$. We shall therefore always assume $\varepsilon < i_0$.

\paragraph{Proof (of Proposition~\ref{prop:graph_seminorm_manifold}).} Follows immediately from~\eqref{pf:graph_seminorm_manifold_1} and~\eqref{eqn:manifold_nonlocal_seminorm}. \qed

\subsection{Decomposition of graph Sobolev seminorm}
\label{subsec:manifold_decomposition_graph_seminorm}
The proof of~\eqref{pf:graph_seminorm_manifold_1} is identical to the proof of~\eqref{pf:graph_seminorm_ho_1}, except substituting the intrinsic dimension $m$ for ambient dimension $d$, and using Lemma~\ref{lem:manifold_graph_seminorm_bias2} rather than Lemma~\ref{lem:graph_seminorm_bias2}.

\subsection{Error due to Euclidean Distance}
\label{subsec:error_euclidean_distance}

In this section, we prove~\eqref{eqn:nonlocal_laplacian_geodesic_error}. By applying Cauchy-Schwarz we obtain an upper bound on $|L_{P,\varepsilon}f(x) - \wt{L}_{P,\varepsilon}f(x)|^2$:
\begin{align}
\bigl[L_{P,\varepsilon}f(x) - \wt{L}_{P,\varepsilon}f(x)\bigr]^2 & \leq \frac{p_{\max}^2}{\varepsilon^{2(2 + m)}} \int_{\mc{X}} \bigl[f(x') - f(x)\bigr]^2 \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') \nonumber \\
& \quad \cdot \int_{\mc{X}} \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') \nonumber \\
& = \frac{1}{\varepsilon^{2(2 + m)}} A_1(x) \cdot A_2(x) \label{pf:error_euclidean_distance_0}
\end{align} 
Thus we have upper bounded $|L_{P,\varepsilon}f(x) - \wt{L}_{P,\varepsilon}f(x)|^2$ by the product of two terms, each of which we now suitably bound. To do so, we will use the following estimate, from Proposition 4 of \cite{trillos2019}: for all $\|x' - x\| \leq R/2$,
\begin{equation}
\label{eqn:distance_error}
\|x' - x\| \leq d_{\mc{X}}(x',x) \leq \|x' - x\| + \frac{8}{R^2} \|x' - x\|^3.
\end{equation}
From here forward we will assume $\varepsilon < R/2$. 
\paragraph{Upper bound on $A_1(x)$.}
Consequently $\eta(\|x' - x\|/\varepsilon) \geq \eta(d_{\mc{X}}(x',x)/\varepsilon)$. Furthermore, letting $L_{\eta}$ denote the Lipschitz constant of $\eta$, and setting $\wt{\varepsilon} := (1 + 27\varepsilon^2/R^2)\varepsilon$ we have that
\begin{equation*}
\biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \leq \frac{L_{\eta} 8 \varepsilon^2}{R^2} \cdot \1\bigl\{d_{\mc{X}}(x',x) \leq \varepsilon\bigr\} + \|\eta\|_{\infty} \cdot 1\{\varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon}\}.
\end{equation*}
Thus,
\begin{equation*}
A_1(x) \leq \frac{8L_{\eta}\varepsilon^2}{R^2}\int_{\mc{X}}\bigl[f(x') - f(x)\bigr]^2 \1\{\|x' - x\| \leq \varepsilon\} \,d\mu(x') + \|\eta\|_{\infty} \int_{\mc{X}}\bigl[f(x') - f(x)\bigr]^2 \1\bigl\{\varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon}\bigr\} \,d\mu(x') \\
\end{equation*}
Integrating over $\mc{X}$, we conclude from Lemma~\ref{lem:dirichlet_energy_remainder} and Lemma~3.3 of \citep{burago2014} and  that
\begin{equation*}
\int_{\mc{X}} A_1(x) \,d\mu(x) \leq \frac{8L_{\eta}\nu_m\varepsilon^2}{R^2(m + 2)} \Bigl(1 + CmK_0R^2\Bigr) \varepsilon^{m + 2} |f|_{H^1(\mc{X})}^2 + C\|\eta\|_{\infty}\varepsilon^{m + 4} |f|_{H^1(\mc{X})}^2 =: C_5 \varepsilon^{m + 4}|f|_{H^1(\mc{X})}.
\end{equation*}

\paragraph{Upper bound on $A_2(x)$.}
Integrating over $x' \in \mc{X}$, we see that
\begin{align}
\int_{\mc{X}} \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') & \leq \frac{8L_{\eta}\varepsilon^2}{R^2} \int_{\mc{X}} \1\bigl\{d_{\mc{X}}(x',x)\bigr\} \,d\mu(x') + p_{\max}\|\eta\|_{\infty} \int_{\mc{X}} \1\bigl\{ \varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon} \bigr\} \,d\mu(x') \nonumber \\
& = \frac{8L_{\eta}\varepsilon^2}{R^2} \cdot \mu\bigl(B(x,\varepsilon)\bigr) +  p_{\max}\|\eta\|_{\infty} \Bigl[\mu\bigl(B(x,\wt{\varepsilon})\bigr) - \mu\bigl(B(x,\varepsilon)\bigr) \Bigr]. \label{pf:error_euclidean_distance_1}
\end{align}
Equation (1.36) in \cite{trillos2019} states that
\begin{equation*}
\bigl|\mu(B_\mc{X}(x,\varepsilon))  - \omega_m \varepsilon^m\bigr|  \leq CmK_0\varepsilon^{m + 2},
\end{equation*}
where we recall $K_0$ is an upper bound on the sectional curvature of $\mc{X}$. Plugging this back into~\eqref{pf:error_euclidean_distance_1}, we conclude that
\begin{align*}
\int_{\mc{X}} \biggl|\eta\biggl(\frac{\|x' - x\|}{\varepsilon}\biggr) - \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr)\biggr| \,d\mu(x') & \leq  \frac{8L_{\eta}\varepsilon^2}{R^2}\Bigl[\omega_m\varepsilon^{m} + CmK_0\varepsilon^{m + 2}\Bigr] + \|\eta\|_{\infty} \Bigl[\omega_m(\wt{\varepsilon}^m - {\varepsilon}^m) + 2CmK_0\varepsilon^{m + 2}\Bigr] \\
& \leq \frac{8L_{\eta}\varepsilon^2}{R^2}\Bigl[\omega_m\varepsilon^{m} + R^2CmK_0\varepsilon^{m}\Bigr] + \|\eta\|_{\infty} \varepsilon^{m + 2}\Bigl[\frac{27\omega_m}{R^2} + 2CmK_0\Bigr] \\
& =: C_6 \varepsilon^{m + 2}.
\end{align*}

\paragraph{Putting together the pieces.}
Plugging our upper bounds on $A_1(x)$ and $A_2(x)$ back into~\eqref{pf:error_euclidean_distance_0}, we deduce that
\begin{align*}
\|\wt{L}_{P,\varepsilon}f - L_{P,\varepsilon}f\|_{L^2(\mc{X})}^2 & \leq \frac{1}{\varepsilon^{2(2 + m)}} \int_{\mc{X}} A_1(x) \cdot A_2(x) \,d\mu(x) \\
& \leq \frac{C_6}{\varepsilon^{(2 + m)}} \int_{\mc{X}} A_1(x) \,d\mu(x) \\
& \leq C_5 C_6 \varepsilon^2 |f|_{H^1(\mc{X})}^2,
\end{align*}
thus proving the claimed result.

\subsection{Approximation Error of non-local Laplacian}
\label{subsec:manifold_approximation_error_nonlocal_laplacian}
Fix $x \in \mc{X}$. We begin with a pointwise estimate of $\wt{L}_{P,\varepsilon}f$, facilitated by expressing $w(v) = f(\exp_x(v))$ and $q(v) = p(\exp_x(v))$ in normal coordinates, as in \citep{calder2019}. Let $J_x(\cdot)$ be the Jacobian of the exponential map $\exp_x$, we have
\begin{align*}
\wt{L}_{P,\varepsilon}f(x) & = \frac{1}{\varepsilon^{m + 2}} \int_{\mc{X}} \bigl(f(x') - f(x)\bigr) \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr) \,dP(x') \\
& = \frac{1}{\varepsilon^{m + 2}} \int_{B(0,\varepsilon) \subset T_x(\mc{X})} \bigl(w(v) - w(0)\bigr) \eta\biggl(\frac{\|v\|}{\varepsilon}\biggr) J_x(v) q(v) \,dv \\
& = \frac{1}{\varepsilon^{2}}\biggl\{\int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr) \eta(\|v\|) q(\varepsilon v) \,dv + \int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr) \eta(\|v\|) q(\varepsilon v) \bigl(J_x(\varepsilon v) - 1\bigr) \,dv \biggr\} \\
& = A_1(x) + A_2(x)
\end{align*}
Note that $w$ and $q$ have the same smoothness properties as $f$ and $p$. Moreover, arguing exactly as we did in the flat Euclidean case, we can show that when $f \in H^2(\mc{X})$ and $p \in C^1(\mc{X})$, then
\begin{equation*}
\|A_1\|_{L^2(\mc{X})}^2 \leq C \|f\|_{H^2(\mc{X})}^2
\end{equation*}
whereas if $f \in H^3(\mc{X})$ and $p \in C^2(\mc{X})$ then 
\begin{equation*}
\|A_1 - \sigma_{\eta} \Delta_Pf\|_{L^2(\mc{X})}^2 \leq C\|f\|_{H^3(\mc{X})}^2 \varepsilon^2.
\end{equation*}

Therefore it remains only to upper bound $A_2$ in $L^2(\mc{X})$ norm. To do so, we recall (1.34) of \cite{trillos2019}: for any $\varepsilon < i_0$ and all $x \in \mc{X}$, the Jacobian $J_x(v)$ satisfies the upper bound
\begin{equation*}
|J_x(v) - 1| \leq CmK_0\varepsilon^2, \quad  \textrm{for all} ~~ v \in B(0,\varepsilon) \subseteq T_x(\mc{X}).
\end{equation*}
Combining this estimate with the Cauchy-Schwarz inequality, we conclude that
\begin{align*}
\|A_2\|_{L^2(\mc{X})}^2 & \leq Cm^2K_0^2 \biggl[\int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr)^2 \eta(\|v\|) q(\varepsilon v) \,dv\biggr] \cdot \biggl[\int_{B(0,1)} \eta(\|v\|) q(\varepsilon v) \,dv\biggr] \\
& \leq Cm^2K_0^2 \sigma_{\eta} (1 + L_q\varepsilon)  \int_{B(0,1)} \bigl(w(\varepsilon v) - w(0)\bigr)^2 \eta(\|v\|) q(\varepsilon v) \,dv \\
& \leq Cm^2K_0^2 \sigma_{\eta}^2 (1 + L_q\varepsilon) p_{\max}  \varepsilon^2 |f|_{H^1(\mc{X})}^2,
\end{align*}
with the final inequality following from (3.2) of~\cite{burago2014}. Combining our estimates on $A_1$ and $A_2$ yields the claim.

\subsection{Estimate of non-local Sobolev seminorm}
\label{subsec:manifold_estimate_nonlocal_seminorm}
In this subsection we establish that the upper bound~\eqref{eqn:manifold_nonlocal_seminorm} holds when $f \in H^s(\mc{X})$ and $p \in C^{s - 1}(\mc{X})$. We first consider $s = 2$, and then $s = 3$.

\paragraph{Case 1: $s = 2$.}
When $s = 2$, the triangle inequality implies that
\begin{equation*}
\dotp{L_{P,\varepsilon}^sf}{f}_P \leq 2p_{\max}\Bigl(\|L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}\|_{L^2(\mc{X})}^2 + \|\wt{L}_{P,\varepsilon}f\|_{L^2(\mc{X})}^2\Bigr)
\end{equation*}
The first term on the right hand side is upper bounded in~\eqref{eqn:nonlocal_laplacian_geodesic_error}, and the second term is upper bounded in~\eqref{eqn:nonlocal_laplacian_approximation_error_manifold_l2}. Together these estimates imply the claim.

\paragraph{Case 2: $s = 3$.}
When $s = 3$, the triangle inequality implies that
\begin{equation*}
\dotp{L_{P,\varepsilon}^sf}{f}_P = E_{P,\varepsilon}(L_{P,\varepsilon}f;\mc{X}) \leq 3\Bigl( E_{P,\varepsilon}(L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f;\mc{X}) +  E_{P,\varepsilon}(\wt{L}_{P,\varepsilon}f - \sigma_{\eta} \Delta_Pf;\mc{X}) + \sigma_{\eta}^2 E_{P,\varepsilon}(\Delta_Pf;\mc{X})\Bigr)
\end{equation*}
We now upper bound each of the three terms on the right hand side of the above inequality. First, we note that by Lemma~\ref{lem:dirichlet_energy_l2} and~\eqref{eqn:nonlocal_laplacian_geodesic_error}, 
\begin{equation*}
 E_{P,\varepsilon}(L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f;\mc{X}) \leq  \frac{C}{\varepsilon^2}\|L_{P,\varepsilon}f - \wt{L}_{P,\varepsilon}f\|_{L^2(\mc{X})}^2 \leq C |f|_{H^1(\mc{X})}^2.
\end{equation*}
An equivalent upper bound on $E_{P,\varepsilon}(\wt{L}_{P,\varepsilon}f - \sigma_{\eta} \Delta_Pf;\mc{X})$ follows from Lemma~\ref{lem:dirichlet_energy_l2} and~\eqref{eqn:nonlocal_laplacian_approximation_error_manifold_sobolev}. Finally, we notice that $f \in H^3(\mc{X})$ and $p \in C^2(\mc{X})$ implies $\Delta_Pf \in H^1(\mc{X})$, and furthermore $|\Delta_Pf|_{H^1(\mc{X})} \leq \|p\|_{C^2(\mc{X})} \cdot \|f\|_{H^3(\mc{X})}$. We conclude from Lemma~\ref{lem:dirichlet_energy_sobolev} that
\begin{equation*}
E_{P,\varepsilon}(\Delta_Pf;\mc{X}) \leq C |\Delta_Pf|_{H^1(\mc{X})}^2 \leq C \|f\|_{H^3(\mc{X})}^2,
\end{equation*}
where in the final inequality we have absorbed $\|p\|_{C^2(\mc{X})}$ into the constant $C$. Together, these upper bounds prove the claim.

\subsection{Integrals}
\label{subsec:manifold_integrals}
Recall the Dirichlet energy $E_{P,\varepsilon}(f;\mc{X}) = \dotp{L_{P,\varepsilon}f}{f}_P$, defined in~\eqref{eqn:dirichlet_energy}. Now we establish some estimates on $E_{P,\varepsilon}(f;\mc{X})$ under Model~\ref{def:model_manifold}, and under various assumptions regarding the regularity of $f$.
\begin{lemma}
	\label{lem:dirichlet_energy_l2}
	Suppose Model~\ref{def:model_manifold}, and additionally that $f \in L^2(\mc{X})$. Then there exists a constant $C$ such that
	\begin{equation}
	\label{eqn:dirichlet_energy_l2}
	E_{P,\varepsilon}(f;\mc{X}) \leq \frac{C}{\varepsilon^2} \|f\|_{L^2(\mc{X})}^2.
	\end{equation}
\end{lemma}
\begin{lemma}
	\label{lem:dirichlet_energy_sobolev}
	Suppose Model~\ref{def:model_manifold}, and additionally that $f \in H^1(\mc{X})$. Then there exist constants $c$ and $C$ which do not depend on $f$ such that for any $0 < \varepsilon < c$,
	\begin{equation}
	\label{eqn:dirichlet_energy_sobolev}
	E_{P,\varepsilon}(f;\mc{X}) \leq C |f|_{H^1(\mc{X})}^2.
	\end{equation}
\end{lemma}

We use Lemma~\ref{lem:dirichlet_energy_remainder} to help upper bound the error incurred by using $\|\cdot\|$ rather than $d_{\mc{X}}(\cdot,\cdot)$. Recall the notation $\wt{\varepsilon} = (1 + 27\varepsilon^2/R^2)\varepsilon$, where $R$ is the reach of $\mc{X}$.
\begin{lemma}
	\label{lem:dirichlet_energy_remainder}
	Suppose Model~\ref{def:model_manifold}, and additionally that $f \in H^1(\mc{X})$. There exist constants $c$ and $C$ such that for any $\varepsilon < c$,
	\begin{equation}
	\label{eqn:dirichlet_energy_remainder}
	\int_{\mc{X}} \int_{\mc{X}} \bigl(f(x') - f(x)\bigr)^2 \1\{\varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon}\} \,d\mu(x') \,d\mu(x) \leq C \varepsilon^{4 + m} \|f\|_{H^1(\mc{X})}^2
	\end{equation}
\end{lemma}

Finally, we use Lemma~\ref{lem:manifold_graph_seminorm_bias2} to show that the pure bias component of $\dotp{L_n^sf,f}_n$ is small in expectation. This is analogous to Lemma~\ref{lem:graph_seminorm_bias2}, except assuming Model~\ref{def:model_manifold} rather than Model~\ref{def:model_flat_euclidean}.
\begin{lemma}
	\label{lem:manifold_graph_seminorm_bias2}
	Assume Model~\ref{def:model_manifold}. Suppose $f \in H^1(\mc{X})$, and let $D_if$ be defined with respect to a kernel $\eta$ that satisfies~\ref{asmp:kernel_manifold}. Then there exists a constant $C$ which does not depend on $f$ or $n$, such that for any $i \in [n]$ and $\bj \in [n]^s$,
	\begin{equation*}
	\Ebb\Bigl[|D_{\bj}f(X_i)| \cdot |f(X_i) - f(X_{\bj_1})|\Bigr] \leq C \varepsilon^{2 + mk} \cdot \|f\|_{H^1(\mc{X})}^2,
	\end{equation*}
	where $k + 1$ is the number of distinct indices in $i\bj$. 
\end{lemma}

\paragraph{Proof (of Lemmas~\ref{lem:dirichlet_energy_l2} and~\ref{lem:dirichlet_energy_sobolev}).}
Define the non-local energy $\wt{E}_{P,\varepsilon}$ with respect to geodesic distance,
\begin{equation*}
\wt{E}_{P,\varepsilon}(f;{\mc{X}}) := \dotp{\wt{L}_{P,\varepsilon}f}{f}_{P} = \int_{\mc{X}} \int_{\mc{X}} \bigl(f(x') - f(x)\bigr)^2 \eta\biggl(\frac{d_{\mc{X}}(x',x)}{\varepsilon}\biggr) \,dP(x') \,dP(x).
\end{equation*}
From the lower bound in~\eqref{eqn:distance_error}, it follows that $E_{P,\varepsilon}(f;X) \leq \wt{E}_{P,\varepsilon}(f;{\mc{X}})$, and from the upper bounds $p(x) \leq p_{\max}$ and $\eta(|x|) \leq \|\eta\|_{\infty} \cdot \1\{x \in [-1,1]\}$ we further have
\begin{equation*}
\wt{E}_{P,\varepsilon}(f;{\mc{X}}) \leq p_{\max}^2 \|\eta\|_{\infty} \cdot \int_{\mc{X}} \int_{B_{\mc{X}}(\varepsilon)} \bigl(f(x') - f(x)\bigr)^2 \,d\mu(x') \,d\mu(x).
\end{equation*}
The estimates~\eqref{eqn:dirichlet_energy_l2} and~\eqref{eqn:dirichlet_energy_sobolev} then respectively follow from (3.1) and Lemma~3.3 of \cite{burago2014}.

\paragraph{Proof (of Lemma~\ref{lem:dirichlet_energy_remainder}).}
Following exactly the steps of the proof of Lemma~3.3 of \citet{burago2014}, but replacing all references to a ball of radius $r$ by references to the set difference between balls of radius $\wt{\varepsilon}$ and $\varepsilon$, we obtain that
\begin{equation*}
\int_{\mc{X}} \int_{\mc{X}} \bigl(f(x') - f(x)\bigr)^2 \1\{\varepsilon < d_{\mc{X}}(x',x) \leq \wt{\varepsilon}\} \,d\mu(x') \,d\mu(x) \leq (1 + CmK_0\varepsilon^2) \cdot \int_{\mc{X}} \int_{B_{m}(0,\wt{\varepsilon})} |d_x^{1}f(v)|^2 \,dv \,d\mu(x).
\end{equation*}
From (2.7) of~\citet{burago2014}, we further have
\begin{equation*}
\int_{\mc{X}} \int_{B_{m}(0,\wt{\varepsilon})} |d_x^{1}f(v)|^2 \,dv \,d\mu(x)  = \frac{\nu_m}{2 + m} (\wt{\varepsilon}^{2 + m} - \varepsilon^{2 + m}) \int_{\mc{X}} |d_x^1f|^2 \,d\mu(x) = 27\frac{\nu_m}{(2 + m)R^2} \varepsilon^{4 + m} \|d^1f\|_{L^2(\mc{X})}^2. 
\end{equation*}
Recalling that $\|d^1f\|_{L^2(\mc{X})}^2 \leq \|f\|_{H^1(\mc{X})}^2$, we see that this implies the claim of Lemma~\ref{lem:dirichlet_energy_remainder}.

\paragraph{Proof (of Lemma~\ref{lem:manifold_graph_seminorm_bias2}).}
The proof of Lemma~\ref{lem:manifold_graph_seminorm_bias2} is identical to the proof of Lemma~\ref{lem:graph_seminorm_bias2}, upon substituting the ambient dimension $m$ for the intrinsic dimension $d$, and using Lemma~\ref{lem:dirichlet_energy_sobolev} rather than Lemma~\ref{lem:dirichlet_estimate_nonlocal_laplacian} to establish~\eqref{pf:graph_seminorm_bias2_1}.

\section{Lower bound on empirical norm}
\label{sec:empirical_norm}
In this Section we prove Proposition~\ref{prop:empirical_norm_sobolev} (in Section~\ref{subsec:empirical_norm_sobolev}). We also prove an analogous result when $\mc{X}$ is a manifold as in Model~\ref{def:model_manifold} (in Section~\ref{subsec:empirical_norm_sobolev_manifold}).

\subsection{Proof of Proposition~\ref{prop:empirical_norm_sobolev}}
\label{subsec:empirical_norm_sobolev}
In this section we establish Proposition~\ref{prop:empirical_norm_sobolev}. As mentioned, the proof of this Proposition follows from the Gagliardo-Nirenberg interpolation inequality, and a one-sided Bernstein's inequality (Lemma~\ref{lem:one_sided_bernstein}).

\begin{lemma}[Gagliardo-Nirenberg interpolation inequality]
	\label{lem:gagliardo_nirenberg}
	Suppose Model~\ref{def:model_flat_euclidean}, and that $f \in H^s(\mc{X})$ for some $s \geq d/4$. Then there exist constants $C_1$ and $C_2$ that do not depend on $f$, such that
	\begin{equation}
	\label{eqn:gagliardo_nirenberg}
	\|f\|_{L^4(\mc{X})} \leq C_1 |f|_{H^s(\mc{X})}^{d/4s} \|f\|_{L^2(\mc{X})}^{1 - d/(4s)} + C_2 \|f\|_{L^2(\mc{X})}
	\end{equation}
\end{lemma}

\paragraph{Proof (of Proposition~\ref{prop:empirical_norm_sobolev}).}
Rearranging~\eqref{eqn:gagliardo_nirenberg} and raising both sides to the $4$th power, we see that
\begin{equation*}
\frac{\Ebb[f^4(X)]}{\|f\|_P^4} \leq C \biggl(\frac{\|f\|_{L^4(\mc{X})}}{\|f\|_{L^2(\mc{X})}}\biggr)^4 \leq C_1\biggl(\frac{|f|_{H^s(\mc{X})}}{\|f\|_{L^2(\mc{X})}}\biggr)^{d/s} + C_2,
\end{equation*}
here the constants $C_1,C_2$ are not the same as in~\eqref{eqn:gagliardo_nirenberg}. Therefore taking the constant $C$ in assumption~\eqref{eqn:empirical_norm_sobolev_1} to be sufficiently large relative to $C_1$ and $C_2$, we have that
\begin{equation*}
C_1\biggl(\frac{|f|_{H^s(\mc{X})}}{\|f\|_{L^2(\mc{X})}}\biggr)^{d/s} \leq \frac{\delta n}{64},
\end{equation*} 
and consequently 
\begin{equation*}
\frac{\Ebb[f^4(X)]}{\|f\|_P^4} \leq \frac{\delta n}{8} + 8C_2^3.
\end{equation*}
The claim then follows from Lemma~\ref{lem:one_sided_bernstein}, upon taking $c = 1/(64C_2^3)$ in the statement of Proposition~\ref{prop:empirical_norm_sobolev}.

\subsection{Proof of Proposition~\ref{prop:empirical_norm_sobolev_manifold}}
\label{subsec:empirical_norm_sobolev_manifold}
The proof of Proposition~\ref{prop:empirical_norm_sobolev_manifold} follows exactly the same steps as the proof of Proposition~\ref{prop:empirical_norm_sobolev}, upon replacing Lemma~\ref{lem:gagliardo_nirenberg} by Lemma~\ref{lem:gagliardo_nirenberg_manifold}.
\begin{lemma}[(c.f Theorem~3.70 of~\citet{aubin2012})]
	\label{lem:gagliardo_nirenberg_manifold}
	Suppose Model~\ref{def:model_manifold}, and that $f \in H^s(\mc{X})$ for some $s \geq m/4$. Then there exist constants $C_1$ and $C_2$ that do not depend on $f$, such that
	\begin{equation}
	\label{eqn:gagliardo_nirenberg_manifold}
	\|f\|_{L^4(\mc{X})} \leq C_1 |f|_{H^s(\mc{X})}^{m/4s} \|f\|_{L^2(\mc{X})}^{1 - m/(4s)} + C_2 \|f\|_{L^2(\mc{X})}.
	\end{equation}
\end{lemma}

\section{Proof of Main Results}
\label{sec:proofs_main_results}

\subsection{Estimation Results}

\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}.}
We condition on the event that the design points $X_1,\ldots,X_n$ satisfy
\begin{equation}
\label{pf:laplacian_eigenmaps_estimation_fo_1}
\dotp{L_{n,\varepsilon}f_0}{f_0}_n \leq \frac{C}{\delta}M^2 \quad \textrm{and} \quad \lambda_k \geq \min\{\lambda_k(\Delta_P), \varepsilon^{-2}\}~~\textrm{for all $2 \leq k \leq n$.}
\end{equation}
Note that by Propositions~\ref{prop:graph_seminorm_fo} and~\ref{prop:graph_eigenvalue}, these statements are both satisfied with probability at least $1 - \delta - Cn\exp\{-cn\varepsilon^d\}$.

Conditional on~\eqref{pf:laplacian_eigenmaps_estimation_fo_1}, we have from Lemma~\ref{lem:fixed_graph_estimation} that for any $0 \leq K \leq n$,
\begin{equation*}
\|\wh{f} - f_0\|_n^2 \leq C\biggl\{\frac{M^2}{\delta (\lambda_{K + 1}(\Delta_P) \wedge \varepsilon^{-2})} + \frac{K}{n}\biggr\},
\end{equation*}
either deterministically (when $K = 0$), or with probability at least $1 - \exp(-K)$ (when $K \geq 1$). Further, from the bounds $\varepsilon \leq c_0 K^{-1/d}$ (Assumption~\ref{asmp:parameters_estimation_fo}) and $\lambda_{K + 1}(\Delta_P) \geq c (K + 1)^{2/d}$ (Weyl's Law) we can simply the above expression to the following,
\begin{equation}
\label{pf:laplacian_eigenmaps_estimation_fo_2}
\|\wh{f} - f_0\|_n^2 \leq C\biggl\{\frac{M^2}{\delta}(K + 1)^{-2/d} + \frac{K}{n}\biggr\}.
\end{equation}
We now upper bound the right hand side of~\eqref{pf:laplacian_eigenmaps_estimation_fo_2}, based on the value of $K$ chosen in~\ref{asmp:parameters_estimation_fo}.  When possible we choose $K = \floor{M^2n}^{d/(2 + d)}$ to balance bias and variance, in which case~\eqref{pf:laplacian_eigenmaps_estimation_fo_2} implies
\begin{equation*}
\|\wh{f} - f_0\|_n^2 \leq \frac{C}{\delta} M^2 (M^2n)^{-2/(2 + d)}.
\end{equation*}
If $M^2 < n^{-1}$, then we take $K = 1$, and from~\eqref{pf:laplacian_eigenmaps_estimation_fo_2} we get
\begin{equation*}
\|\wh{f} - f_0\|_n^2 \leq \frac{C}{n\delta}.
\end{equation*}
Finally if $M > n^{1/d}$, we take $K = n$. In this case, we note that $\wh{f}(X_i) = Y_i$ for all $i = 1,\ldots,n$, and it immediately follows that
\begin{equation*}
\|\wh{f} - f_0\|_n^2 = \frac{1}{n}\sum_{i = 1}^{n} w_i^2 \leq 5,
\end{equation*}
with probability at least $1 - \exp(-n)$. Combining these three separate cases yields the conclusion of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}.

\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}.}
Follows identically to the proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~\ref{prop:graph_seminorm_ho} rather than Proposition~\ref{prop:graph_seminorm_fo} and Assumption~\ref{asmp:parameters_estimation_ho} rather than Assumption~\ref{asmp:parameters_estimation_fo}.

\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_manifold}.}
Follows identically to the proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~\ref{prop:graph_seminorm_manifold} rather than Proposition~\ref{prop:graph_seminorm_fo}, Proposition~\ref{prop:graph_eigenvalue_manifold} rather than Proposition~\ref{prop:graph_eigenvalue}, and Assumption~\ref{asmp:parameters_estimation_manifold} rather than Assumption~\ref{asmp:parameters_testing_fo}.

\subsection{Testing Results}

\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}.}
We have already upper bounded the Type I error of $\varphi$ in Lemma~\ref{lem:fixed_graph_testing}, and it remains to upper bound the Type II error. To do so, we condition on the event that the design points $X_1,\ldots,X_n$ satisfy,
\begin{equation}
\label{pf:laplacian_eigenmaps_testing_fo_1}
\dotp{L_{n,\varepsilon}f_0}{f_0}_n \leq \frac{C}{\delta}M^2,\quad \textrm{and} \quad \lambda_k \geq \min\{\lambda_k(\Delta_P), \varepsilon^{-2}\}~~\textrm{for all $2 \leq k \leq n$,}
\end{equation}
as well as that
\begin{equation}
\label{pf:laplacian_eigenmaps_testing_fo_2}
\|f_0\|_n^2 \geq \frac{1}{2}\|f_0\|_P^2.
\end{equation}
Note that by Propositions~\ref{prop:graph_seminorm_fo} and~\ref{prop:graph_eigenvalue}, both statements in~\eqref{pf:laplacian_eigenmaps_testing_fo_1} are satisfied with probability at least $1 - \delta - Cn\exp\{-cn\varepsilon^d\}$. Additionally, by Proposition~\ref{prop:empirical_norm_sobolev} and the assumption in~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} that $\|f_0\|_P^2 \geq CM^2/(bn^{2/d})$, the one-sided inequality~\eqref{pf:laplacian_eigenmaps_testing_fo_2} follows with probability at least $1 - \exp\{-(cn \wedge 1/b)\}$. Setting $\delta = b/3$ and taking $n \geq N$ to be sufficiently large, the bottom line is that both~\eqref{pf:laplacian_eigenmaps_testing_fo_1} and~\eqref{pf:laplacian_eigenmaps_testing_fo_2} are together satisfied with probability at least $1 - b/2$.

Now, to complete the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}, we would like to invoke Lemma~\ref{lem:fixed_graph_testing}, and conclude that conditional on $X_1,\ldots,X_n$ satisfying~\eqref{pf:laplacian_eigenmaps_testing_fo_1} and~\eqref{pf:laplacian_eigenmaps_testing_fo_2}, our test $\varphi$ will equal $1$ with probability at least $1 - b/2$. To use Lemma~\ref{lem:fixed_graph_testing}, we will need to establish that~\eqref{eqn:fixed_graph_testing_critical_radius} is satisfied, which we now show.

On the one hand, we have that the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius} is upper bounded, 
\begin{align*}
\frac{\dotp{L_{n,\varepsilon}f_0}{f_0}_n}{\lambda_{K + 1}} + \frac{\sqrt{2K}}{n}\biggl[2\sqrt{\frac{1}{a}} + \sqrt{\frac{2}{b}} + \frac{32}{bn}\biggr] & \leq C\biggl(\frac{M^2}{b \min\{\lambda_{K+1}(\Delta_P), \varepsilon^{-2}\}} + \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr) \\
& \leq C\biggl(\frac{M^2}{b}K^{-2/d} + \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr)
\end{align*}
with the second inequality following by the assumption $\varepsilon \leq K^{-1/d}$ and Weyl's Law. On the other hand, we have that $\|f_0\|_n^2 \geq \|f_0\|_P^2/2$. Consequently, to prove Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}, it remains only to verify that
\begin{equation}
\label{pf:laplacian_eigenmaps_testing_fo_3}
\|f_0\|_P^2 \geq C\biggl(\frac{M^2}{b}K^{-2/d} + \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr).
\end{equation}
As in the estimation case, we can further upper bound the right hand side of~\eqref{pf:laplacian_eigenmaps_testing_fo_3}, depending on the value of $K$ chosen in~\ref{asmp:parameters_testing_fo}. The classical case is $K = (M^2n)^{d/(2 + d)}$, in which case~\eqref{pf:laplacian_eigenmaps_testing_fo_3} is satisfied as long as
\begin{equation*}
\|f_0\|_P^2 \geq CM^2(M^2n)^{-4/(4 + d)}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]
\end{equation*}
If $M^2 < n^{-1}$, then we take $K = 1$, and~\eqref{pf:laplacian_eigenmaps_testing_fo_3} is satisfied whenever
\begin{equation*}
\|f_0\|_P^2 \geq \frac{C}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr].
\end{equation*}
Finally if $M > n^{1/d}$, we take $K = n$, and~\eqref{pf:laplacian_eigenmaps_testing_fo_3} is satisfied if
\begin{equation*}
\|f_0\|_P^2 \geq C\biggl(\frac{M^2}{n^{2/d}b} + n^{-1/2}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr).
\end{equation*}
We conclude by observing that~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} implies each of these three inequalities, and thus implies~\eqref{pf:laplacian_eigenmaps_testing_fo_3}.

\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_ho}.}
Follows identically to the proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~\ref{prop:graph_seminorm_ho} rather than Proposition~\ref{prop:graph_seminorm_fo} and Assumption~\ref{asmp:parameters_testing_ho} rather than Assumption~\ref{asmp:parameters_testing_fo}.

\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_manifold}.}
Follows identically to the proof of Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo}, except substituting $L_{n,\varepsilon}^s$ for $L_{n,\varepsilon}$, $\lambda_k^s$ for $\lambda_k$, and using Proposition~\ref{prop:graph_seminorm_manifold} rather than Proposition~\ref{prop:graph_seminorm_fo}, Proposition~\ref{prop:graph_eigenvalue_manifold} rather than Proposition~\ref{prop:graph_eigenvalue}, Proposition~\ref{prop:empirical_norm_sobolev_manifold} rather than Proposition~\ref{prop:empirical_norm_sobolev}, and Assumption~\ref{asmp:parameters_testing_manifold} rather than Assumption~\ref{asmp:parameters_testing_fo}.

\paragraph{Proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal}.}
Note that our choices of $K$ and $\varepsilon$ ensure  that~\eqref{pf:laplacian_eigenmaps_testing_fo_1} (with $L_{n,\varepsilon}^s$ replacing $L_{n,\varepsilon}$) and~\eqref{pf:laplacian_eigenmaps_testing_fo_2} are satisfied with probability at least $1 - b/2$. Proceeding as in the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}, we upper bound the right hand side of~\eqref{eqn:fixed_graph_testing_critical_radius},
\begin{align*}
\frac{\dotp{L_{n,\varepsilon}f_0}{f_0}_n}{\lambda_{K + 1}} + \frac{\sqrt{2K}}{n}\biggl[2\sqrt{\frac{1}{a}} + \sqrt{\frac{2}{b}} + \frac{32}{bn}\biggr] & \leq C\biggl(\frac{M^2}{b \min\{\lambda_{K+1}(\Delta_P), \varepsilon^{-2}\}} + \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr) \\
& \leq C\biggl(\frac{M^2}{b}\varepsilon^2 + \frac{\sqrt{2K}}{n}\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr]\biggr).
\end{align*}
Unlike in the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}, we note that in this case $\varepsilon^2 \leq C\lambda_K(\Delta_P)$ rather than vice versa. From here, proceeding as in the proof of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} gives the claimed result.

\section{Graph Laplacian methods and the cluster assumption}
\label{subsec:eigenmaps_beats_kernel_smoothing}

A main conclusion of our paper is that PCR-LE is minimax optimal for nonparametric regression over certain Sobolev classes. It is not the only optimal method. For instance, kernel smoothing and least squares using an appropriate set of basis functions as features are two other minimax optimal methods over these Sobolev classes. We now give an example where PCR-LE is better than these two alternatives, in the sense of having (much) smaller risk. This is possible because PCR-LE performs remarkably well when the regression function $f_0$ and design distribution $P$ satisfy a \emph{cluster assumption}: that is, when the regression function is (approximately) piecewise constant over high-density clusters of the design distribution $P$. On the other hand, kernel smoothing (with Euclidean distance) and least squares (using eigenfunctions of an unweighted Laplace operator) cannot take advantage of the cluster assumption. We call this property of PCR-LE \emph{density adaptivity}.

\subsection{Setup}
We begin by specifying a sequence of design densities and regression functions $\{(p^{(n)}, f_0^{(n)}): n \in \mathbb{N}\}$. These distributions will all be chosen to satisfy the cluster assumption. To that end, we define two clusters $Q_1,Q_2 \subset \Reals$ using a cluster separation parameter $r$, as
\begin{equation*}
Q_1 := [0,1/2 - r], \quad Q_2 := [1/2 + r,1],
\end{equation*}
and take the domain $\mc{X}^{(n)} := Q_1 \cup Q_2$. We then take the design density to be uniform over $\mc{X}^{(n)}$ and the regression function to be a piecewise constant function over $Q_1$ and $Q_2$ of height $\theta$,
\begin{equation}
\label{def:model_cluster_assumption}
p^{(n)}(x) := \frac{1}{1 - 2r}\1\bigl\{x \in Q_1 \cup Q_2\bigr\}, \quad f_0^{(n)}(x) := \theta \cdot \Bigl(\1\bigl\{x \in Q_1\bigr\} - \1\bigl\{x \in Q_2\bigr\}\Bigr).
\end{equation}
Thus $p^{(n)}$ and $f_0^{(n)}$ belong to a two-parameter family, where the parameters are the cluster separation $r$ and height $\theta$. Generally speaking, the smaller the separation $r$, and the larger the height $\theta$, the more graph Laplacian methods will outperform both kernel smoothing and linear regression using eigenfunctions of the unweighted Laplace operator as features.

We now define kernel smoothing and least squares using eigenfunction of an unweighted Laplace operator For a kernel function $\psi$ and bandwidth parameter $h$, the kernel smoothing estimator $\wt{f}_{\mathrm{KS}}$ is defined at a point $x \in \mc{X}$ as 
\begin{equation}
\label{eqn:kernel_smoothing}
\wt{f}_{\mathrm{KS}}(x) := 
\begin{dcases*}
0, & \quad \textrm{if $d_{n,h}(x) = 0$,} \\
\frac{1}{d_{n,h}(x)} \sum_{i = 1}^{n} Y_i \psi\biggl(\frac{\|X_i - x\|}{h}\biggr), & \quad \textrm{otherwise.}
\end{dcases*}
\end{equation}

Let $(\lambda_1,\phi_1), (\lambda_2,\phi_2),\ldots$ be eigenpairs of the unweighted Laplace operator $\Delta$ on $[0,1]$, meaning
\begin{equation}
\label{eqn:laplace_beltrami_eigenpairs}
\Delta \phi_k = \lambda_k \phi_k, ~~ \|\phi_k\|_{L^2([0,1])} = 1,~~\frac{d}{dx}\phi_k(0) = \frac{d}{dx}\phi_k(1) = 0. 
\end{equation}
In this case the eigenfunctions $\phi_k$ of $\Delta$ are simply cosine functions, with eigenvalues proportional to their squared frequency. Noting that $\phi_1(x) = 1$ and $\lambda_1 = 0$, for $k = 2,3,\ldots$ we have
\begin{equation*}
\phi_k(x) = \sqrt{2}\cdot\cos(2\pi k x),~~\lambda_k(\Delta) = \pi^2 k^2.
\end{equation*}
The least squares estimator using $\phi_1,\ldots,\phi_K$ ($1 \leq K \leq n$) eigenfunctions as features is simply\footnote{For convenience, we will assume $\Phi \in \Reals^{n \times K}$ is full rank. If this is not the case, the least squares estimator $\wt{f}_K$ is not uniquely defined, but any solution will equal ${\bf Y}$ in-sample, and will satisfy $\|\wt{f}_K - f_0\|_n^2 \geq 1/2$ with high probability.}
\begin{equation}
\label{eqn:least_squares}
\wt{f}_{K} := \argmin_{f \in \mathrm{span}\{\phi_1,\ldots,\phi_K\}} \|Y - f\|_n^2 = \Phi (\Phi^{\top} \Phi)^{-1} \Phi^{\top} Y.
\end{equation}
Hereafter, we will refer to $\wt{f}_K$ as the \emph{uniform least squares} estimator.

\subsection{Upper bounds on risk of PCR-LE}
Now we are in a position to state our results. Both PCR-LE and kernel smoothing depend in part on the choice of kernel. For simplicity, in our analysis we only consider the boxcar kernel,
\begin{equation}
\label{asmp:boxcar_kernel}
\eta(z) = \psi(z) = \1\{z \leq 1\}.
\end{equation}
This is strictly for convenience, and the following results will also hold for any kernel that satisfies~\ref{asmp:kernel_flat_euclidean}.

\begin{proposition}
	\label{prop:cluster_assumption_ub}
	Suppose $(X_1,Y_1),\ldots(X_n,Y_n)$ are sampled according to~\eqref{def:model_cluster_assumption}. Compute the PCR-LE estimator $\wh{f}$ using a kernel $\eta$ which satisfies~\eqref{asmp:boxcar_kernel}, number of eigenvectors $K = 2$, and radius $\varepsilon = r/2$. Then,
	\begin{equation}
	\label{eqn:cluster_assumption_le_ub}
	\Ebb\Bigl[\|\wh{f} - f_0^{(n)}\|_{n}^2\Bigr] \leq \biggl(6\theta^2 + \frac{1}{n}\biggr) \cdot \frac{8}{r} \exp(-nr/8) + \frac{1}{n}.
	\end{equation}
\end{proposition}
\paragraph{Proof of Proposition~\ref{prop:cluster_assumption_ub}.}

We begin by showing that, with high probability, the eigenvectors $v_1,v_2$ respect the cluster structure of $p^{(n)}$. Denote $u_1 = (\1\{X_i \in Q_1\})_{i \in [n]}$, and likewise $u_2 = (\1\{X_i \in Q_2\})_{i \in [n]}$. We make the following two observations:
\begin{enumerate}
	\item Because $\varepsilon < r$ and the kernel $\eta$ is compactly supported on $[0,1]$, for each $X_i \in Q_1$ and $X_j \in Q_2$, it must be the case that $\eta(\|X_i - X_j\|/\varepsilon) = 0$.
	\item Using an elementary concentration argument (stated in Lemma~\ref{lem:balls_in_bins}) and the triangle inequality, we deduce that with probability at least $1 - 4/\varepsilon\exp(-n\varepsilon/4)$ there exists a path in $G_{n,\varepsilon}$ between each $X_i, X_j \in Q_1$, and likewise between each $X_i,X_j \in Q_2$.
\end{enumerate}
Together these observations imply that with high probability the neighborhood graph $G_{n,\varepsilon}$ consists of exactly two connected components: one consisting of all design points $X_i \in Q_1$, and the other consisting of all design points $X_i \in Q_2$. In other words, 
\begin{equation}
\label{pf:eigenmaps_beats_kernel_smoothing_1}
\Pbb\Bigl(\mathrm{span}\{v_1,v_2\} = \mathrm{span}\{u_1,u_2\}\Bigr) \geq 1 - 4/\varepsilon\exp(-n\varepsilon/4).
\end{equation}

Let us condition on the ``good'' event $\mc{E}$ that the design points $X_1,\ldots,X_n$ satisfy~\eqref{eqn:balls_in_bins}, and therefore that $\mathrm{span}\{v_1,v_2\} = \mathrm{span}\{u_1,u_2\}$. Consider the empirical mean $\wb{Y}_Q := \frac{1}{\sharp\{Q \cup {\bf X}\}} \sum_{i: X_i \in Q} Y_i$. Since $\mathrm{span}\{v_1,v_2\} = \mathrm{span}\{u_1,u_2\}$, the estimator $\wh{f} = \wh{f}_{\mathrm{LE}}$ will be piecewise constant on $Q_1$ and $Q_2$, and in fact we have that
\begin{equation}
\label{pf:eigenmaps_beats_kernel_smoothing_-1}
\wh{f} = \wb{Y}_{Q_1}u_1 + \wb{Y}_{Q_2}u_2.
\end{equation}
Therefore conditional on $\mc{E}$,
\begin{equation*}
\|\wh{f} - f_0^{(n)}\|_{n}^2 = P_n(Q_1) \cdot (\wb{Y}_{Q_1} - \theta)^2 + P_n(Q_2) \cdot (\wb{Y}_{Q_2} + \theta)^2
\end{equation*}
and consequently,
\begin{equation}
\label{pf:eigenmaps_beats_kernel_smoothing_0}
\Ebb\Bigl[\|\wh{f} - f_0^{(n)}\|_{n}^2 \Big|\mc{E}\Bigr] = \Ebb\Bigl[\Ebb\bigl[\|\wh{f} - f_0^{(n)}\|_{n}^2\big|X_1,\ldots,X_n\bigr] \Bigr|\mc{E}\Bigr] = \frac{1}{n}.
\end{equation}

Now we derive a crude upper bound on $\|\wh{f} - f_0^{(n)}\|_{n}$ that will suffice to control the error conditional on $\mc{E}^c$. We observe that the empirical norm of $\wh{f}$ is bounded,
\begin{equation*}
\|\wh{f}\|_n^2 \leq \frac{2}{n}\sum_{i = 1}^{n}\dotp{Y}{v_1}_n^2v_{1,i}^2 + \dotp{{\bf Y}}{v_2}_n^2v_{2,i}^2 \leq 2\bigl(\dotp{{\bf Y}}{v_1}_n^2 + \dotp{{\bf Y}}{v_2}_n^2\bigr) \leq 4 \|{\bf Y}\|_n^2. 
\end{equation*}
Noting that $\Ebb[\|{\bf Y}\|_n^2|X_1,\ldots,X_n] = \|f_0\|_n^2 + 1/n  = \theta^2 + 1/n$, we conclude that
\begin{equation*}
\Ebb\Bigl[\|\wh{f} - f_0\|_{n}^2 \cdot \1\{\mc{E}^c\}\Bigr] \leq \Ebb\Bigl[\Bigl(2 \|f_0\|_{n}^2 + 4(\theta^2 + 1/n) \cdot \1\{\mc{E}^c\} \Bigr] \leq (6\theta^2 + n^{-1}) \cdot 4\varepsilon^{-1} \exp(-n\varepsilon/4).
\end{equation*}
Combining this with~\eqref{pf:eigenmaps_beats_kernel_smoothing_0} implies~\eqref{eqn:cluster_assumption_le_ub}.

\subsection{Lower bounds on risk of kernel smoothing and least squares}

\begin{proposition}
	\label{prop:cluster_assumption_lb}
	Suppose $(X_1,Y_1),\ldots,(X_n,Y_n)$ are sampled according to~\eqref{def:model_cluster_assumption}. Suppose $(\log n)^2/n \leq r \leq c$, where $c$ is a universal constant. 
	\begin{itemize}
		\item Compute the kernel smoothing estimator $\wt{f} = \wt{f}_{\mathrm{KS}}$ as in~\eqref{eqn:kernel_smoothing}, using a kernel $\psi$ which satisfies~\eqref{asmp:boxcar_kernel}. Then there exist universal constants $c, N > 0$ such that for all $n > N$,
		\begin{equation}
		\label{eqn:cluster_assumption_ks_lb}
		\inf_{h' > 0} \Ebb\Bigl[\|\wt{f} - f_0^{(n)}\|_{n}^2\Bigr] \geq c\min\biggl\{ \frac{r^{-1}}{n}, \frac{\theta}{\sqrt{n}} \biggr\}.
		\end{equation}
		\item Compute the least squares estimator $\wt{f} = \wt{f}_{\mathrm{SP}}$ as in~\eqref{eqn:least_squares}. Then there exist universal constants $c, N > 0$ such that for all $n > N$,
		\begin{equation}
		\label{eqn:cluster_assumption_sp_lb}
		\inf_{1 \leq K \leq n} \Ebb \Bigl[\|\wt{f} - f_0^{(n)}\|_n^2\Bigr] \geq c\min\biggl\{\frac{r^{-1}}{n}, \frac{1}{\log(n)}, \frac{r^{-2/3}}{n}, \frac{\sqrt{\theta}}{n^{3/4}}\biggr\}.
		\end{equation}
	\end{itemize}
\end{proposition}
The proof of Proposition~\ref{prop:cluster_assumption_lb} is long, and we defer it until after some discussion of the implications of the proposition.

Together, Propositions~\ref{prop:cluster_assumption_ub} and~\ref{prop:cluster_assumption_lb} illustrate that the risk of PCR-LE can be dramatically smaller than that of kernel smoothing or uniform least squares. For instance, taking $\theta = n^{1/2}$ and $r = n^{-3/4}$, when appropriately tuned, $\wh{f}$ satisfies
\begin{equation*}
\Ebb\Bigl[\|\wh{f} - f_0^{(n)}\|_{n}^2\Bigr] \leq C\biggl(n^{7/4}\exp(-n^{1/4}/8)) + \frac{1}{n}\biggr) \leq \frac{C}{n},
\end{equation*}
for a universal constant $C$ and all $n$ larger than some universal constant $N$, whereas for $\wt{f} = \wt{f}_{\mathrm{KS}}$,
\begin{equation*}
\inf_{h' > 0} \Ebb\Bigl[\|\wt{f} - f_0^{(n)}\|_{n}^2\Bigr] \geq \frac{c}{n^{1/4}},
\end{equation*}
and for $\wt{f} = \wt{f}_{\mathrm{SP}}$,
\begin{equation*}
\inf_{1 \leq K \leq n} \Ebb\Bigl[\|\wt{f} - f_0^{(n)}\|_{n}^2\Bigr] \geq \frac{c}{n^{1/2}}.
\end{equation*}
Other choices of $\theta$ and $r$ lead to even more dramatic gaps between the risk of PCR-LE, and the risk of kernel smoothing and least squares. The overall takeaway is that under Model~\ref{def:model_cluster_assumption}, estimators that use the graph Laplacian can converge to the true regression function $f_0^{(n)}$ at fast rates---parametric rates that do not depend on the $L^2$ norm of $f_0^{(n)}$---whereas other estimators, optimal for estimation over Sobolev spaces, converge to $f_0^{(n)}$ at slow rates---nonparametric rates that deteriorate as the $L^2$ norm of $f_0^{(n)}$ grows.

Some remarks:
\begin{itemize}
	\item The lower bound on the in-sample risk of $\wt{f}_{\mathrm{KS}}$  given by \eqref{eqn:cluster_assumption_ks_lb} is larger than that of $\wt{f}_{\mathrm{SP}}$ given by \eqref{eqn:cluster_assumption_sp_lb}. This does not mean that kernel smoothing exhibits less adaptivity to the cluster assumption than uniform least squares. Instead, we suspect it is due to looseness in our lower bounds: we are able to tightly control the bias of kernel smoothing, whereas we must use a potentially loose bound on the bias of uniform least squares. Experimentally, it appears that kernel smoothing usually outperforms uniform least squares, under various instantiations of the cluster assumption.
	\item The cluster assumption---in which the regression function is piecewise constant and $p$ consists of multiple connected components---is a very strong assumption. The \emph{low-density separation} condition is a related but weaker assumption, in which the regression function is assumed to be smoother (but not constant) in regions of higher density. This is a rather general hypothesis which can formalized in a number of different ways. For instance, one could insist that the regression function $f_0$ belong to a normed ball in a \emph{weighted Sobolev space}, with semi-norm given by
	\begin{equation*}
	|f_0|_{H^s(P)} := \dotp{\Delta_P^s f_0}{f_0}_P.
	\end{equation*}
	Intuitively, when $\|f_0\|_{H^s(P)}$ is much smaller than $\|f_0\|_{H^s(\mc{X})}$, density-adaptive learners such as PCR-LE should have the advantage on non-density adaptive linear smoothers, such as kernel smoothing or uniform least squares. Indeed, in the case of Model~\ref{def:model_cluster_assumption} we see that
	\begin{equation*}
	\|f_0^{(n)}\|_{H^s(P^{(n)})} = 0~~\textrm{for all $s \in \mathbb{N}$, and all $r, \theta > 0$,}
	\end{equation*}
	whereas $f_0^{(n)}$ does not even belong to the first-order Sobolev space $H^1([0,1])$. In words, this shows the cluster assumption is an extreme case of the low-density separation condition. Unfortunately, it is quite difficult to analyze graph-based estimators under the general low-density separation condition, without making strong assumptions on $P$. 
	\item Finally, we note that either changing the graph or the normalization of the Laplacian fundamentally alters the type of density adaptivity displayed by graph-Laplacian-based estimators; see~\citet{hoffmann2019} for an extensive discussion.
\end{itemize}

\subsection{Proof of Proposition~\ref{prop:cluster_assumption_lb}}
First we show~\eqref{eqn:cluster_assumption_ks_lb}, then~\eqref{eqn:cluster_assumption_sp_lb}.

\subsubsection{Proof of~\eqref{eqn:cluster_assumption_ks_lb}}
A standard argument using the law of iterated expectation implies the following lower bound on the pointwise risk in terms of squared-bias and variance-like quantities,
\begin{equation*}
\Ebb\Bigl[\Bigl(\wt{f}(X_i) - f_0(X_i)\Bigr)^2|X_i = x\Bigr] \geq \frac{(n - 1)}{n}\Ebb\biggl[\Bigl(f_0(X) - f_0(x)\Bigr)^2|X \in B(x,h')\biggr] + \Ebb\biggl[\frac{1}{d_{n,h'}(x)}\biggr].
\end{equation*}
The variance term can be lower bounded quite simply for any $x \in \mc{X}^{(n)}$; noting that $\sup_{x} p^{(n)}(x) < 2$ and $\nu(B(x,h') \cap \mc{X}^{(n)}) \leq 2h'$, it follows by Jensen's inequality that 
\begin{equation*}
\Ebb\biggl[\frac{1}{d_{n,h'}(x)}\biggr] \geq \frac{1}{\Ebb[d_{n,h'}(x)]} \geq \frac{1}{4nh'}.
\end{equation*}
On the other hand the squared bias term is quite large for $x$ close to $1/2$. Precisely, if $h' \geq 4r$ then a simple calculation implies
\begin{equation*}
\Ebb[(f_0(X) - f_0(x))^2|X \in B(x,h')] \geq \frac{\theta^2}{8} \quad \textrm{for all}~x \in [(1 - h'/2)_{+}, 1/2 - r].
\end{equation*}
Combining these lower bounds on variance and squared bias terms and summing over $X_1,\ldots,X_n$, we arrive at the following: if $h' \leq 4r$, then
\begin{equation*}
\Ebb\Bigl[\|\wt{f} - f_0^{(n)}\|_{n}^2\Bigr] = \frac{1}{n}\sum_{i = 1}^{n} \Ebb\biggl[\Ebb\Bigl[\Bigl(\wt{f}(X_i) - f_0(X_i)\Bigr)^2|X_i\Bigr]\biggr]\geq \frac{1}{16rn},
\end{equation*}
whereas if $h' > 4r$  then
\begin{align*}
\Ebb\Bigl[\|\wt{f} - f_0^{(n)}\|_{n}^2\Bigr] & = \frac{1}{n}\sum_{i = 1}^{n} \Ebb\biggl[\Ebb\Bigl[\Bigl(\wt{f}(X_i) - f_0(X_i)\Bigr)^2|X_i\Bigr]\biggr] \\
& \geq \frac{1}{4nh'} + \frac{\theta^2}{8}\frac{(n - 1)}{n} P^{(n)}\Bigl([(1 - h'/2)_{+}, 1/2 - r]\Bigr) \\
&\geq \frac{1}{4nh'} + \frac{\theta^2h'}{64}.
\end{align*}
In the latter case, setting the derivative equal to $0$ shows that the right hand side is always at least $\theta/\sqrt{64 n}$, and taking the minimum over the two cases then yields~\eqref{eqn:cluster_assumption_ks_lb}.

\subsubsection{Proof of~\eqref{eqn:cluster_assumption_sp_lb}}
We begin by decomposing the risk into conditional bias and variance terms. Let $\mathbb{E}_n = \Ebb[\cdot|X_1,\ldots,X_n]$ denote expectation conditional on the design points $X_1,\ldots,X_n$.  Then by the law of iterated expectation, and the fact that $\Ebb_n[w] = 0$,
\begin{equation*}
\Ebb\bigl[\|\wt{f}_{\mathrm{SP}} - f_0\|_n^2\bigr] = \Ebb\bigl[\|\Ebb_n\wt{f}_{\mathrm{SP}} -  f_0\|_n^2\bigr] + \Ebb\bigl[\|\wt{f}_{\mathrm{SP}} - \Ebb_n\wt{f}_{\mathrm{SP}}\|_n^2\bigr]. 
\end{equation*}
We separately lower bound the expected conditional squared bias and variance terms. To anticipate what is to come: we will show that the expected conditional variance is equal to $K/n$; on the other hand we will show that the expected conditional squared bias is lower bounded,
\begin{equation}
\label{pf:cluster_assumption_sp_lb_1}
\Ebb\bigl[\|\Ebb_n\wt{f}_{\mathrm{SP}} -  f_0\|_n^2\bigr] = \frac{K}{n}\quad\textrm{and}\quad\Ebb\bigl[\|\Ebb_n\wt{f}_{\mathrm{SP}} -  f_0\|_n^2\bigr] \geq \frac{\theta^2}{2601\pi^2 K^{3}},
\end{equation}
with the lower bound holding so long as $K \leq \min\{1/(16r),n/(8 \log(8 n)), (\sqrt{160} \pi/r)^{2/3}\}$. If $K$ is larger than this, then the expected conditional variance is lower bounded,
\begin{equation}
\label{pf:cluster_assumption_sp_lb_2}
\Ebb\bigl[\|\wt{f}_{\mathrm{SP}} - \Ebb_n\wt{f}_{\mathrm{SP}}\|_n^2\bigr] \geq \min\biggl\{\frac{1}{16rn}, \frac{1}{8 \log(8n)}, \frac{(\sqrt{160} \pi)^{2/3}}{r^{2/3}n} \biggr\}
\end{equation}
Otherwise~\eqref{pf:cluster_assumption_sp_lb_1} implies that the in-sample risk is always at least 
\begin{equation*}
\Ebb\bigl[\|\wt{f}_{\mathrm{SP}} - f_0\|_n^2\bigr] \geq \frac{\theta^2}{2601\pi^2 K^{3}} + \frac{K}{n} \geq 2 \frac{\theta^{1/2}}{n^{3/4}} \frac{1}{(2601 \pi^2)^{1/4}}.
\end{equation*}
Along with~\eqref{pf:cluster_assumption_sp_lb_2}, this implies the claim. It remains to show the bounds on conditional bias and variance.

\paragraph{Conditional variance.}
The expected conditional variance is exactly equal to $K/n$, a standard fact that is verified by the following calculations: first,
\begin{align*}
\|\wt{f}_{\mathrm{SP}} - \Ebb_n\wt{f}_{\mathrm{SP}}\|_n^2 = \|\Phi(\Phi^{\top}\Phi)^{-1}\Phi^{\top}w\|_n^2 = \frac{1}{n}w^{\top}\Phi(\Phi^{\top}\Phi)^{-1}\Phi^{\top}w;
\end{align*}
thus standard properties of the Gaussian distribution and the trace trick imply
\begin{equation*}
\Ebb_n\bigl[\|\wt{f}_{\mathrm{SP}} - \Ebb_n\wt{f}_{\mathrm{SP}}\|_n^2 \bigr] = \frac{1}{n} \mathrm{tr} (\Phi(\Phi^{\top}\Phi)^{-1}\Phi^{\top}) = \frac{K}{n};
\end{equation*}
and finally by the law of iterated expectation and the independence of the noise $(w_1,\ldots,w_n)$ and the design points $X_1,\ldots,X_n$,
\begin{equation*}
\Ebb\Bigl[\Ebb_n\bigl[\|\wt{f}_{\mathrm{SP}} - \Ebb_n\wt{f}_{\mathrm{SP}}\|_n^2\Bigr] = K/n.
\end{equation*}

\paragraph{Conditional bias.}
It takes more work to lower bound the conditional bias. We will first upper bound the Lipschitz constant of $\Ebb_n\wt{f}_{\mathrm{SP}}$ in terms of the empirical norm $\|\Ebb_n\wt{f}_{\mathrm{SP}}\|_n$. Then we will use this upper bound to argue that either $\Ebb_n\wt{f}_{\mathrm{SP}}$ has empirical norm much larger than that of $f_0$, or $\Ebb_n\wt{f}_{\mathrm{SP}}$ is a smooth function, in the sense of having a small Lipschitz constant. In the former case, the triangle inequality will then imply that $\|\Ebb_n\wt{f}_{\mathrm{SP}} -  f_0\|_n$ must be large. In the latter case, the smoothness of $\Ebb_n\wt{f}_{\mathrm{SP}}$ will imply that $\Ebb_n\wt{f}_{\mathrm{SP}}$ must be far from $f_0$ at many points $X_i$ close to $x = 1/2$.

The following Lemma gives our upper bound on the Lipschitz constant of $\|\Ebb_n\wt{f}_{\mathrm{SP}}\|_n$. Here we treat $\Ebb_n\wt{f}_{\mathrm{SP}} = \sum_{k = 1}^{K} \wt{\beta}_k \phi_k$ as a function defined at all $x \in [0,1]$ by extending it in the canonical way. As a function over $[0,1]$, clearly $\Ebb_n\wt{f}_{\mathrm{SP}} \in C^\infty([0,1])$. Let $\Sigma \in \Reals^{K \times K}$ be the covariance matrix of $(\phi_1,\ldots,\phi_K)$, i.e. the matrix with entries $\Sigma_{k\ell} = \dotp{\phi_k,\phi_{\ell}}_{P^{(n)}}$. Let $\wh{\Sigma} := (\Phi^{\top}\Phi)/n$ be the empirical covariance matrix. Let $I_K \in \Reals^{K \times K}$ be the identity matrix.
\begin{lemma}[Lipschitz regularity of $\Ebb_n\wt{f}_{\mathrm{SP}}$.]
	\label{lem:lipschitz_regularity_sp}
	Let $\wt{f}_n = \Ebb_n\wt{f}_{\mathrm{SP}}$. Then 
	\begin{equation}
	\label{eqn:lipschitz_regularity_sp1}
	\|\wt{f}_n\|_{C^1(\mc{X})}^2 \leq \pi^2 \frac{K^{3} \cdot \|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\|_{\mathrm{op}}}{(1 - \|I_K - \Sigma\|_F)} \cdot \|\wt{f}_n\|_n^2.
	\end{equation}
	Moreover, suppose $K \leq 1/(16r)$ and $r \leq (1 - 2^{-1/2})/2$.
	\begin{itemize}
		\item {\bf(Matrix perturbation)} Then
		\begin{equation}
		\label{eqn:lipschitz_regularity_matrix_perturbation}
		\|\Sigma - I_K\|_F \leq \frac{1}{2}.
		\end{equation}
		\item {\bf(Matrix concentration, cf. \cite{hsu2012})} If additionally $n \geq 8K \log(K/\delta)$ for some $\delta \in (0,1/2)$, then with probability at least $1 - 2\delta$,
		\begin{equation}
		\label{eqn:lipschitz_regularity_matrix_concentration}
		\|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\|_{\mathrm{op}} \leq 5.
		\end{equation}
	\end{itemize}
	Therefore, if $K \leq \min\{1/(16r),n/(8 \log(K/\delta))\}$, then with probability at least $1 - 2\delta$,
	\begin{equation}
	\label{eqn:lipschitz_regularity_sp2}
	\|\wt{f}_n\|_{C^1(\mc{X})}^2 \leq 10 \pi^2 K^3 \|\wt{f}_n\|_n^2.
	\end{equation}
\end{lemma}
We defer the proof of Lemma~\ref{lem:lipschitz_regularity_sp} until after we complete the proof of~\eqref{eqn:cluster_assumption_sp_lb}.

Now, if $\|\wt{f}_n\|_n^2 \geq \frac{3}{2}\|f_0\|_n^2$, then by the triangle inequality
\begin{equation*}
\|\wt{f}_n - f_0\|_n \geq \|\wt{f}_n\|_n - \|f_0\|_n \geq \sqrt{\frac{3}{2}} \cdot \|f_0\|_n  = \sqrt{\frac{3}{2}} \cdot \theta.
\end{equation*}
Otherwise $\|\wt{f}_n\|_n^2 \geq \frac{3}{2}\|f_0\|_n^2$. In this case, we show that $|\wt{f}_n(X_i) - f_0(X_i)|$ must be large (on the order of $\theta$) for many points $X_i$ which are close to $x = 1/2$. Let us suppose without loss of generality that $\wt{f}_n(1/2) \leq \theta/2$ and consider points $X_i \in Q_1$ close to $x = 1/2$; otherwise if $\wt{f}_n(1/2) > \theta/2$ we could obtain the exact same bound by considering $X_i \in Q_2$. For each point $X_i \in Q_1$, by Lemma~\ref{lem:lipschitz_regularity_sp} we have that with probability at least $1 - 2\delta$,
\begin{equation*}
|\wt{f}_n(X_i) - \wt{f}_n(1/2)| \leq CK^{3/2} \|\wt{f}_n\|_n \cdot |X_i - 1/2| \leq \sqrt{10}\pi K^{3/2} \theta \cdot |X_i - 1/2|.
\end{equation*}
Since $\wt{f}_n(1/2) \leq \theta/2$ and $f_0(X_i) = \theta/2$ for all $X_i \in Q_1$ it follows that
\begin{equation*}
|\wt{f}_n(X_i) - f_0(X_i)| \geq \theta - \sqrt{10}\pi K^{3/2} \theta \cdot |X_i - 1/2|,
\end{equation*}
and consequently
\begin{equation*}
|\wt{f}_n(X_i) - f_0(X_i)| \geq \theta/2,\quad\textrm{for any $X_i \in Q_1$ such that $|X_i - 1/2| \leq 1/(\sqrt{40}\pi K^{3/2})$.}
\end{equation*}
This yields a lower bound on $\|\wt{f}_n - f_0\|_n$; letting $Q_K := \Bigl[\frac{1}{2} - \frac{1}{\sqrt{40}\pi K^{3/2}},\frac{1}{2} - r\Bigr]$, we have that 
\begin{equation*}
\|\wt{f}_n - f_0\|_n \geq \frac{\theta}{2} \cdot P_n\Bigl(Q_k\Bigr).
\end{equation*}
Then as long as $K^{-3/2} \geq \sqrt{160} \pi r$, from the multiplicative form of Hoeffding's inequality (Lemma~\ref{lem:hoeffding_smoothing})
\begin{equation*}
P^{(n)}(Q_K) \geq \frac{1}{\sqrt{160} \pi K^{3/2}} \geq 2r \Longrightarrow \Pbb\biggl(P_n(Q_K) \geq \frac{1}{\sqrt{640} \pi K^{3/2}}\biggr) \geq 1 - \exp(-nr/4) \geq 1 - \frac{4}{n^2}.
\end{equation*}
Putting the pieces together, we conclude that if $K \leq \min\{1/(8r),n/(8 \log(K/\delta)), (\sqrt{160} \pi/r)^{2/3}\}$, then
\begin{equation*}
\|\wt{f}_n - f_0\|_n \geq \frac{\theta}{51 \pi K^{3/2}},
\end{equation*}
with probability at least $1 - 2\delta - {4}{n^2}$. Taking $\delta = 1/8$ then implies the claim.

\paragraph{Proof of Lemma~\ref{lem:lipschitz_regularity_sp}.}

\underline{\it{Proof of~\eqref{eqn:lipschitz_regularity_sp1}}.}
Recall that $\wt{f}_n = \sum_{k = 1}^{K} \wt{\beta}_k \phi_k$. Exchanging sum with derivative, we have that
\begin{equation*}
\frac{d}{dx} \wt{f}_n(x) = -\pi \sum_{k = 1}^{K}\wt{\beta}_k k \sin(k\pi x).
\end{equation*}
Thus taking absolute value and applying the Cauchy-Schwarz inequality gives
\begin{equation*}
|\wt{f}_n'(x)|^2 \leq \pi^2 K^2 \sum_{k = 1}^{K} \bigl(\sin(k\pi x)\bigr)^2 \|\beta\|_2^2 \leq \pi^2 K^3 \|\beta\|_2^2.
\end{equation*}
On the other hand, we can also relate the empirical norm $\|\wt{f}_n\|_n^2$ to the $\ell^2$ norm of $\beta$. Specifically,
\begin{equation*}
\|\wt{f}_n\|_n^2 = \beta^{\top} \wh{\Sigma} \beta \geq \frac{\|\beta\|_2^2}{\|\wh{\Sigma}^{-1}\|_{\mathrm{op}}} \geq \frac{\|\beta\|_2^2}{\|\Sigma^{-1}\|_{\mathrm{op}} \cdot \|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\|_{\mathrm{op}}} = \frac{\|\beta\|_2^2 \|\Sigma\|_{\mathrm{op}} }{\|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\|_{\mathrm{op}}}
\end{equation*}
Rearranging, we see that 
\begin{equation*}
\sup_{x \in [0,1]} |\wt{f}_n'(x)|^2 \leq \frac{\pi^2 K^3}{\|\Sigma\|_{\mathrm{op}}} \|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\|_{\mathrm{op}} \leq \frac{\pi^2 K^3}{1 - \|I_K - \Sigma\|_F} \|\Sigma^{1/2} \wh{\Sigma}^{-1} \Sigma^{1/2}\|_{\mathrm{op}}
\end{equation*}
with the latter inequality following since $\|\Sigma\|_{\mathrm{op}} \geq \|I_K\|_{\mathrm{op}} - \|I_K - \Sigma\|_{\mathrm{op}} \geq 1 - \|I_K - \Sigma\|_{F}$.

\underline{\it{Proof of~\eqref{eqn:lipschitz_regularity_matrix_perturbation}}.}
We will show that for all $1 \leq k < \ell \leq K$, 
\begin{equation}
\label{pf:lipshcitz_regularity_matrix_perturbation_1}
(1 - \dotp{\phi_k}{\phi_k}_{P^{(n)}})^2 \leq 32r^2, \quad\textrm{and}\quad|\dotp{\phi_k}{\phi_{\ell}}_{P^{(n)}}| \leq 64r^2.
\end{equation}
This implies $\|I - \Sigma\|_F^2 \leq 32K^2r^2$, so that $\|I - \Sigma\|_F \leq 1/2$ so long as $K \leq 1/(16r)$.

The proof of~\eqref{pf:lipshcitz_regularity_matrix_perturbation_1} follows from computing some standard integrals. We separate the computation based on whether $k = 1$ or $k > 1$. \newline

\underline{\it{Case 1: $k = 1$.}}
When $k = 1$, $\dotp{\phi_1}{\phi_1}_{P^{(n)}} = 1$ and $(1 - \dotp{\phi_1}{\phi_1}_{P^{(n)}})^2 = 0$. Additionally, by symbolic integration we find that
\begin{equation*}
\dotp{\phi_k}{\phi_{\ell}}_{P^{(n)}} = \frac{-2\sqrt{2}}{(1 - 2r)} \cdot \frac{\cos(\ell\pi/2) \sin(\ell\pi r)}{\ell \pi},
\end{equation*}
and therefore
\begin{equation*}
\bigl[\dotp{\phi_k}{\phi_{\ell}}_{P^{(n)}}\bigr]^2 \leq \frac{8}{(1 - 2r)^2} \cdot \biggl(\frac{\sin(\ell\pi r)}{\ell \pi}\biggr)^2 \leq \frac{8}{(1 - 2r)^2} r^2 \leq 16r^2,
\end{equation*}
where in the second-to-last inequality follows because $\sin(x)/x \leq 1$, and the last inequality follows by our assumed upper bound on $r$. \newline

\underline{\it{Case 2: $k > 1$.}}
When $k > 1$,
\begin{equation*}
\dotp{\phi_k}{\phi_k}_{P^{(n)}} = 1 - \frac{2}{(1 - 2r)} \frac{\cos(k\pi) \sin(2k\pi r)}{k\pi} \Longrightarrow \bigl[1 - \dotp{\phi_k}{\phi_k}_{P^{(n)}}\bigr]^2 \leq \frac{4}{(1 - 2r)^2} \cdot \biggl(\frac{\sin(2k\pi r)}{k\pi}\biggr)^2 \leq 32r^2.
\end{equation*}
Similarly,
\begin{equation*}
\dotp{\phi_k}{\phi_\ell}_{P^{(n)}} = -\frac{4}{(1 - 2r)}\biggl[\frac{\cos((k + \ell)\pi) \sin((k + \ell)\pi r)}{(k + \ell)\pi} + \frac{\cos((k - \ell)\pi) \sin((k - \ell)\pi r)}{(k - \ell)\pi}\biggr]
\end{equation*}
and therefore
\begin{equation*}
\bigl[\dotp{\phi_k}{\phi_\ell}_{P^{(n)}}\bigr]^2 \leq \frac{16}{(1 - 2r)^2} \biggl(\biggl[\frac{\sin((k + \ell)\pi r)}{(k + \ell)\pi}\biggr]^2 + \biggl[\frac{\sin((k - \ell)\pi r)}{(k - \ell)\pi}\biggr]^2\biggr) \leq 64 r^2.
\end{equation*}

\underline{\it{Proof of~\eqref{eqn:lipschitz_regularity_matrix_concentration}}}
Denote $\Phi(x) = (\phi_1,\ldots,\phi_K(x)) \in \Reals^K$ for any $x \in [0,1]$. Then for any $x \in [0,1]$,
\begin{equation*}
\|\Sigma^{-1/2} \Phi(x)\| \leq \|\Sigma^{-1}\|_{\mathrm{op}}^{1/2} \|\Phi(x)\|_2 \leq \|\Sigma^{-1}\|_{\mathrm{op}}^{1/2} \sqrt{2K} \leq 2 \sqrt{K}
\end{equation*}
with the second-to-last inequality following from~\eqref{eqn:lipschitz_regularity_matrix_perturbation}, and the last inequality following since $|\phi_k(x)| \leq \sqrt{2}$ for all $k$. Thus $\|\Sigma^{-1/2} \Phi(x)\|/\sqrt{K} \leq 2$, and~\eqref{eqn:lipschitz_regularity_matrix_concentration} follows from Theorem~1 of \citet{hsu2012}.

\underline{\it{Proof of~\eqref{eqn:lipschitz_regularity_sp2}}.}
Follows immediately.

\section{Miscellaneous}
Here we give assorted helpful Lemmas used at various points in the above proofs. We also review notation and relevant facts regarding Taylor expansion.

\subsection{Concentration Inequalities}
Lemma~\ref{lem:chi_square_bound} controls the deviation of a chi-squared random variable. It is from~\cite{laurent00}.
\begin{lemma}
	\label{lem:chi_square_bound}
	Let $\xi_1,\ldots,\xi_N$ be independent $N(0,1)$ random variables, and let $U := \sum_{k = 1}^{N} a_k(\xi_k^2 - 1)$.  Then for any $t > 0$,
	\begin{equation*}
	\Pbb\Bigl[U \geq 2 \|a\|_2 \sqrt{t} + 2 \|a\|_{\infty}t\Bigr] \leq \exp(-t).
	\end{equation*}
	In particular if $a_k = 1$ for each $k = 1,\ldots,N$, then
	\begin{equation*}
	\Pbb\Bigl[U\geq 2\sqrt{N t} + 2t\Bigr] \leq \exp(-t).
	\end{equation*}
\end{lemma}

Lemma~\ref{lem:one_sided_bernstein} is an immediate consequence of the one-sided Bernstein's inequality (14.23) in \cite{wainwright2019}.
\begin{lemma}[One-sided Bernstein's inequality]
	\label{lem:one_sided_bernstein}
	Let $X, X_1,\ldots,X_n \sim P$, and $f$ satisfy $\Ebb[f^4(X)] < \infty$. Then
	\begin{equation*}
	\|f\|_n^2 \geq \frac{1}{2}\|f\|_P^2,
	\end{equation*}
	with probability at least $1 - \exp\bigl(-n/8 \cdot \|f\|_P^4 /\Ebb[f^4(X)]\bigr)$.
\end{lemma}

Lemma~\ref{lem:hoeffding_smoothing} is a multiplicative form of Hoeffding's inequality.
\begin{lemma}[Hoeffding's Inequality, multiplicative form]
	\label{lem:hoeffding_smoothing}
	Suppose $Z_i$ are independent random variables, which satisfy $Z_i \in [0,B]$ for $i = 1,\ldots,n$. For any $0 < \delta < 1$, it holds that
	\begin{equation*}
	\Pbb\biggl(\Bigl|S_n - \mu\Bigr| \geq \delta \mu\biggr) \leq 2\exp\biggl(-\frac{\delta^2\mu}{3B^2}\biggr).
	\end{equation*}
\end{lemma}

The following Lemma gives a ``balls-in-bins'' result. More precisely, it gives a lower bound on the probability that every bin
\begin{equation*}
Q_{i1} = [i/m,(i + 1)/m] \cdot (1/2 - r), \quad Q_{i2} = 1/2 + [i/m,(i + 1)/m] \cdot (1/2 - r).
\end{equation*}
will contain at least one ball.
\begin{lemma}
	\label{lem:balls_in_bins}
	Suppose $(X_1,Y_1),\ldots(X_n,Y_n)$ are sampled according to~\eqref{def:model_cluster_assumption}, and suppose $r \leq 1/4$. We have that 
	\begin{equation}
	\label{eqn:balls_in_bins}
	\Pbb\Bigl(\sharp\{Q_{ij} \cup {\bf X}\} > 0 ~~\textrm{for all $i = 1,\ldots,m - 1$ and $j = 1,2$} \Bigr) \geq 1 - 2m\exp\{-n/2m\}.
	\end{equation}
\end{lemma}
\paragraph{Proof (of Lemma~\ref{lem:balls_in_bins}).}
For each $Q_{ij}$, we have that $P(Q_{ij}) = (1/2 - r)/m \geq 1/(2m)$. Therefore
\begin{equation*}
\Pbb\bigl(\sharp\{Q_{ij} \cup {\bf X}\} = 0\bigr) = (1 - 1/(2m))^{n} \leq \exp\{-n/2m\}.
\end{equation*}
By a union bound,
\begin{equation*}
\Pbb\Bigl(\sharp\{Q_{ij} \cup {\bf X}\} = 0 ~~\textrm{for any $i = 1,\ldots,m - 1$ and $j = 1,2$} \Bigr) \leq 2m\exp\{-n/2m\}. \qed
\end{equation*}
Let $\varepsilon = 2/m$. Note that by construction,~\eqref{eqn:balls_in_bins} implies that any points $x$ and $x'$ in adjacent intervals $Q_{ij}$ and $Q_{i'j}$ must be connected in $G_{n,\varepsilon}$. Likewise, it implies that for $h = 1/m$ the degree $d_{n,h}(x) > 0$ for every $x \in Q_1 \cup Q_2$.

\subsection{Taylor expansion}
\label{subsec:taylor_expansion}
We begin with some notation that allows us to concisely derivatives. For a given $z \in \Rd$ and $s$-times differentiable function $f: \mc{X} \to \Reals$, we denote $\bigl(d_x^sf\bigr)(z) := \sum_{|\alpha| = s} D^{\alpha}f(x) z^{\alpha}$. We also write $d^sf := \sum_{|\alpha| = j} D^{\alpha}f$. We point out that in the first-order case $d_x^1f$ is the differential of $f$ at $x \in \mc{X}$, while $d^1f$ is the divergence of $f$.

Let $u$ be a function which is $s$ times continuously differentiable at all $x \in \mc{X}$, for $k \in \mathbb{N}\setminus\{0\}$. Suppose that for some $h > 0$, $x \in \mc{X}_{h}$ and $x' \in B(x,h)$. We write the order-$s$ Taylor expansion of $u(x')$ around $x' = x$ as
\begin{equation*}
u(x') = u(x) + \sum_{j = 1}^{s - 1} \frac{1}{j!}\bigl(d_x^{j}u\bigr)(x' - x) + r_{x'}^{s}(x;u)
\end{equation*}
For notational convenience we have adopted the convention that $\sum_{j = 1}^{0} a_j = 0$.  Thus $\bigl(d_x^{j}f\bigr)(z)$ is a degree-$j$ polynomial---and so a $j$-homogeneous function---in $z$, meaning for any $t \in \Reals$,
\begin{equation*}
\bigl(d_x^{j}f\bigr)(tz) = t^{j} \cdot \bigl(d_x^{j}f\bigr)(z).
\end{equation*}
The remainder term $r_{x'}$ is given by
\begin{equation*}
r_{x'}^s(x;f) = \frac{1}{(j - 1)!} \int_{0}^{1}(1 - t)^{j - 1} \bigl(d_{x + t(x' - x)}^{s}f\bigr)(x' - x) \,dt,
\end{equation*}
where we point out that the integral makes sense because $x + t(x' - x) \in B(x,h) \subseteq \mc{X}$. We now give estimates on the remainder term in both sup-norm and $L^2(\mc{X}_{h})$ norm, each of which hold for any $z \in B(0,1)$. In sup-norm, we have that 
\begin{equation*}
\sup_{x \in \mc{X}_{h}}|r_{x + hz}^j(x;f)| \leq C h^j \|f\|_{C^j(\mc{X})},
\end{equation*}
whereas in $L^2(\mc{X}_{h})$ norm we have,
\begin{equation}
\label{eqn:sobolev_remainder_term}
\int_{\mc{X}_{h}} \bigl|r_{x + thz}^j(x;f)\bigr|^2 \,dx \leq h^{2j} \int_{\mc{X}_{h}} \int_{0}^{1} |d_{x + thz}^jf(z)|^2 \,dt \,dx \leq h^{2j} \|d^jf\|_{L^2(\mc{X})}^2.
\end{equation}
In the last inequality

Finally, we recall some facts regarding the interaction between smoothing kernels and polynomials.  Let $q_j(z)$ be an arbitrary degree-$j$ (multivariate) polynomial. If $\eta$ is a radially symmetric kernel and $j$ is odd, then by symmetry it follows that
\begin{equation*}
\int_{B(0,1)} q_j(z) \eta(\|z\|) \,dz = 0.
\end{equation*}
On the other hand, if $\psi$ is an order-$s$ kernel for some $s > j$, then by converting to polar coordinates we can verify that
\begin{equation*}
\int_{B(0,1)} q_j(z) \eta(\|z\|) \,dz = 0.
\end{equation*}

\section{Computational considerations}
\label{sec:computational_considerations}
Recall that when $s = 1$, we have shown that PCR-LE is optimal when $\varepsilon \asymp (\log n/n)^{1/d}$ is (up to a constant) as small as possible while still ensuring the graph $G$ is connected. On the other hand, when $s > 1$, we can show PCR-LE is optimal only when $\varepsilon = \omega(n^{-c})$ for some $c < 1/d$. For such a choice of $\varepsilon$, the average degree in $G$ will grow polynomially in $n$ as $n \to \infty$, and computing eigenvectors of the Laplacian of a graph will be more computationally intensive than if the graph were sparse. 
In this dense-graph setting, we now discuss a procedure to more efficiently compute an approximation to the PCR-LE estimate: \emph{edge sparsification}.

By now there exist various methods see (e.g., the seminal papers of \citet{spielman2011,spielman2013,spielman2014}, or the overview by \citet{vishnoi2012} and references therein) to efficiently remove many edges from the graph $G$ while only slightly perturbing the spectrum of the Laplacian. Specifically such algorithms take as input a parameter $\sigma \geq 1$, and return a sparser graph $\wc{G}$, $E(\wc{G}) \subseteq E(G)$, with a Laplacian $\wc{L}_{n,\varepsilon}$ satisfying
\begin{equation*}
\frac{1}{\sigma} \cdot u^{\top} \wc{L}_{n,\varepsilon} u \leq u^{\top} L_{n,\varepsilon} u \leq \sigma \cdot u^{\top} \wc{L}_{n,\varepsilon}u \quad \textrm{for all $u \in \Reals^n$.}
\end{equation*}
Let $\wc{f}$ be the PCR-LE estimator computed using the eigenvectors of the sparsified graph Laplacian $\wc{L}_{n,\varepsilon}$ . Because $\wc{G}$ is sparser than $G$, it can be (much) faster to compute the eigenvectors of $\wc{L}_{n,\varepsilon}$ than the eigenvectors of $L_{n,\varepsilon}$, and consequently much faster to compute $\wc{f}$ than $\wh{f}$. Statistically speaking, letting $\wc{\lambda}_k$ be the $k$th eigenvalue of $\wc{L}_{n,\varepsilon}$, we have that conditional on $\{X_1,\ldots,X_n\}$,
\begin{equation*}
\|\wc{f} - f_0\|_n^2 \leq \frac{\dotp{\wc{L}_{n,\varepsilon}^s f_0}{f_0}_n}{\wc{\lambda}_{K + 1}^s} + \frac{5K}{n} \leq \sigma^{2s} \frac{\dotp{\wc{L}_{n,\varepsilon}^s f_0}{f_0}_n}{\wc{\lambda}_{K + 1}^s} + \frac{5K}{n},
\end{equation*}
with probability at least $1 - \exp(-K)$. Consequently $\wt{f}$ has $L^2(P_n)$-error of at most $\sigma^{2s}$ times our upper bound on the error of $\wh{f}$, and for any choice of $\sigma$ that is constant in $n$ the estimator $\wc{f}$ will also be rate-optimal.

In fact the aforementioned edge sparsification algorithms are overkill for our needs. For one thing, they are designed to work when $\sigma$ is very close to $1$, whereas in order for $\wc{f}$ to be rate-optimal, setting $\sigma$ to be any constant greater than $1$, say $\sigma = 2$, is sufficient. Additionally, edge sparsification algorithms are traditionally designed to work in the worst-case, where no assumptions are made on the structure of the graph $G$. But the geometric graphs we consider in this paper exhibit a special structure, in which very roughly speaking no single edge is a bottleneck. As pointed out by~\citet{sadhanala16b}, in this special case there are far simpler and faster methods for sparsification, which at least empirically seem to do the job.