### 2021-12-22 manuscripts/laplacian_eigenmaps/information_and_inference/cover_letter.tex
Added cover letter for Inf and Inf.

\documentclass{article}
\usepackage{enumerate}

\begin{document}

Dear Editors,

\vspace{.2 in}

Please find our attached submission titled: Minimax Optimal Regression over Sobolev Spaces via Laplacian Eigenmaps on Neighborhood Graphs. We analyze Principal Components Regression with Laplacian Eigenmaps, a spectral projection approach to nonparametric regression that projects observed responses onto the span of eigenvectors of a neighborhood graph Laplacian. We analyze this method in a classical random design regression model; in this model, we establish upper bounds showing that the method can optimally take advantage of Sobolev smoothness in the underlying regression function, and low intrinsic dimension in the design data manifold. Interestingly, these upper bounds imply that our estimator converges at rates far faster than the rates at which graph Laplacian eigenvectors converge to their continuum limits.

For completeness, we also derive some minimax rates for Sobolev classes, in the less well-studied subcritical regime where such classes contain discontinuous functions. However, we stress that our main contribution is the analysis of Principal Components Regression with Laplacian Eigenmaps, rather than novel minimax theory. Thank you for your consideration of this manuscript.
\vspace{.2 in}

Sincerely,

\vspace{.1 in}
Alden Green, Sivaraman Balakrishnan, and Ryan Tibshirani

\end{document}